### 2022-04-23 write-up/results.Rmd

---
title: "Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: preliminary report"
output:
  html_document:
    self_contained: yes
  pdf_document:
    toc: no
  bookdown::html_document2:
    number_sections: no
    toc: no
  rtf_document:
    toc: no
  word_document: default
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r setup, include=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')

# remotes::install_github("https://github.com/wjchulme/osutils")
# library('osutils')

# where are the outputs (ie the inputs for this manuscript!) saved?
output_dir_os <- here("output", "release-objects")
#output_dir_os <- here("released_output", "release-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("write-up", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

## Import design elements
source(here("lib", "design", "design.R"))

matchset <- "A"

match_flowchart <-  read_csv(here("output", "release-objects", matchset, "match_flowchart.csv"))
match_coverage <- read_csv(here("output", "release-objects", matchset, "match_coverage.csv"))

match_table1 <- read_csv(here("output", "release-objects", matchset, "match_table1.csv"))
match_table1by <- read_csv(here("output", "release-objects", matchset, "match_table1by.csv"))
match_smd <-  read_csv(here("output", "release-objects", matchset, "match_smd.csv"))

km_estimates <- read_csv(here("output", "release-objects", matchset, "km_estimates.csv")) %>% 
  mutate(
    outcome_descr = fct_recode(as.character(outcome),  !!!recoder$outcome),
    subgroup_descr = fct_recode(subgroup,  !!!recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
km_contrasts_daily <- read_csv(here("output", "release-objects", matchset, "km_contrasts_daily.csv")) %>% 
  mutate(
    outcome_descr = fct_recode(as.character(outcome),  !!!recoder$outcome),
    subgroup_descr = fct_recode(subgroup,  !!!recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
km_contrasts_overall <- read_csv(here("output", "release-objects", matchset, "km_contrasts_overall.csv")) %>% 
  mutate(
    outcome_descr = fct_recode(as.character(outcome),  !!!recoder$outcome),
    subgroup_descr = fct_recode(subgroup,  !!!recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidcc",
      "coviddeath",
      "noncoviddeath"
    ) %>% 
    set_names(., .)

```

\newpage

## Overview

The UK COVID-19 vaccination programme delivered its first "booster" doses in September 2021, initially in groups at high risk of severe disease then across the adult population. The BNT162b2 Pfizer-BioNTech vaccine was used initially, with Moderna mRNA-1273 subsequently also used. The concurrent administration BNT162b2 and Moderna vaccines for boosting in England enables a direct comparison of their effectiveness. We used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to compare the effectiveness of BNT162b2 versus Moderna in boosted adults between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`, during which time both vaccine brands were being used.

## Methods

## Data source

All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>. With the approval of NHS England, primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS).

## Eligibility criteria

We considered all adults aged $\ge$ 18 years who received a booster dose of BNT162b2 or Moderna between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyrend_date), "%e %B %Y")` inclusive (the "recruitment period"). People were eligible for matching if: they were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.

## Matching

BNT162b2 booster recipients were matched with MOderna booster recipients on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation. 
These matching criteria ensure that treatment groups are also matched within JCVI priority group.

## Outcomes

The following outcomes were considered: 
* Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection. 
* COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
* COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes. 
* COVID-19 death. Deaths were classified as from COVID-19 if deaths with the U07.1 or U07.2 ICD-10 codes were mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
* Non-COVID-19 death.

<!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->

<!--#

## Follow-up

Each person was followed from the day of boosting (i.e., time zero) until the earliest of outcome, death, practice de-registration, 10 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.

## Statistical Analysis

We estimated Kaplan-Meier cumulative incidence curves separately by vaccine brand, and compared groups using the 10-week risk differences and risk ratios.

<!-- [The primary estimand is the average treatment effect in the recruited population. To estimate this, expected cumulative incidence rates are estimated for each participant assuming treatment was received by everyone in both treatment groups at time zero. These curves are then averaged over all participants to give the marginal cumulative incidence. This is repeated assuming nobody is treated in either group, and their difference taken. Calculating standard errors for these quantities is challenging, whether approached analytically (due to complex variance components) or computationally (via bootstrapping, due to the large sample sizes involved) and are therefore not reported. ] -->

## Subgroup analyses

We also estimated comparative effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.

## Software, code, and reproducibility

Data management and analyses were conducted in Python version 3.8.10 and R version 4.0.2. All code is shared openly for review and re-use under MIT open license at <https://github.com/opensafely/comparative-booster>. The supplementary materials provide further details of the codelists and data sources used for all variables in the study. Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared. Any reported figures based on counts below 6 are redacted or rounded for disclosure control.

This study will follow the STROBE-RECORD reporting guidelines.

# Results

## Study population and matching

`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="moderna") %>% pull(pct_all))`) eligible for matching. `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).

#### Table 1: inclusion criteria

```{r table-flowchart, echo=FALSE, warning=FALSE, message=FALSE}

match_flowchart %>%
  select(-crit) %>%
   pivot_wider(
    id_cols = c(criteria, vax3_type),
    names_from = vax3_type,
    values_from = c(n, n_exclude, pct_all)
  ) %>%
  gt() %>%
  cols_label(
    criteria = "",

    n_pfizer = "N",
    n_moderna = "N",

    n_exclude_pfizer = "N excluded",
    n_exclude_moderna = "N excluded",

    pct_all_pfizer = "% remaining",
    pct_all_moderna = "% remaining",
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("pfizer")
  ) %>%
   tab_spanner(
    label = "Moderna mRNA-1273",
    columns = ends_with("moderna")
  ) %>%
  # tab_spanner(
  #   label = "mRNA-1273",
  #   columns = ends_with("moderna")
  # ) %>%
  fmt_percent(
    columns = starts_with(c("pct_")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("n_")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "criteria")
  )
```

\newpage 
#### Figure 1: Cumulative matching success

```{r, figure-matchingsuccess, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))

xmin <- min(match_coverage$vax3_date )
xmax <- max(match_coverage$vax3_date )+1

plot_coverage_cumuln <-
  match_coverage %>%
  mutate(
    treatment_descr = fct_recode(as.character(treatment), !!!recoder$treatment),
    status_descr = fct_recode(as.character(status), !!!recoder$status),
    cumuln = cumuln*((treatment*2) - 1)
  ) %>%
  ggplot()+
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=paste0(treatment,status_descr),
      fill=treatment_descr,
      alpha=status_descr,
      colour=NULL
    ),
    position=position_stack(reverse=TRUE),
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(match_coverage$vax3_date, "1 month")),
    limits = c(xmin-1, NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = ~scales::label_number(accuracy = 1, big.mark=",")(abs(.)),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2")+
  scale_colour_brewer(type="qual", palette="Set2")+
  scale_alpha_discrete(range= c(0.8,0.4))+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "bottom"
  )+
  NULL

plot_coverage_cumuln
cat("  \n\n")
```

\newpage

#### Table 2: baseline characteristics

```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
table1 <- 
  match_table1 %>%
  #filter(outcome %in% outcomes) %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(match_table1by$by_col, match_table1by$by_chr))

table1 %>%
  gt(groupname_col="treatment_descr") %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
     label=md("Characteristic")
  ) %>%
  cols_hide(c("variable", "var_label", "order"))
```

## Estimated comparative effectiveness

### Main analysis

#### Table 3: Kaplan-Meier cumulative incidence estimates

```{r table-km, echo=FALSE, warning=FALSE, message=FALSE}

km_contrasts_overall %>%
  transmute(
    outcome_descr,
    subgroup_descr,
    subgroup_level_descr,
    n.atrisk=n.atrisk_0,
    cml.atrisk_0=label_number(1)(cml.atrisk_0/7), 
    cml.atrisk_1=label_number(1)(cml.atrisk_1/7),
    n.event_0, 
    n.event_1,
    riskCI_0 = paste0(label_number(0.01)(risk_0), " (", label_number(0.01)(risk.ll_0), "-", label_number(0.01)(risk.ul_0), ")"),
    riskCI_1 = paste0(label_number(0.01)(risk_1), " (", label_number(0.01)(risk.ll_1), "-", label_number(0.01)(risk.ul_1), ")"),
    kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
    kmrr = paste0(label_number(0.01)(kmrr)),
  ) %>%
  gt(groupname_col="outcome_descr") %>%
  cols_label(

    outcome_descr = "Outcome",
    subgroup_descr = "Subgroup",
    subgroup_level_descr = "Subgroup",

    n.atrisk = "N per group",

    cml.atrisk_0 = "Person-weeks", 
    cml.atrisk_1 = "Person-weeks",

    n.event_0 = "Events",
    n.event_1 = "Events",

    riskCI_0 = "KM risk (95% CI)",
    riskCI_1 = "KM risk (95% CI)",

    kmrdCI = "Risk-difference (95% CI)",
    kmrr = "Relative-risk"
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("_0")
  ) %>%
   tab_spanner(
    label = "Moderna",
    columns = ends_with("_1")
  )  %>%
   tab_spanner(
    label = "Comparison",
    columns = starts_with("km")
  )

```

#### Figure 3: Kaplan-Meier cumulative incidence estimates

```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
plot_km <- km_estimates %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
  geom_step(aes(x=time, y=1-surv))+
  geom_rect(aes(xmin=time, xmax=leadtime, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
  facet_grid(rows=vars(outcome_descr))+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y="Cumulative incidence",
    colour=NULL,
    title=NULL
  )+
  theme_minimal()+
  theme(
    axis.line.x = element_line(colour = "black"),
    panel.grid.minor.x = element_blank(),
    legend.position=c(.05,.95),
    legend.justification = c(0,1),
  )

plot_km
```

\newpage

## References

### 2022-04-23 write-up/report.Rmd
move report

--- 2022-04-23 write-up/results.Rmd
@@
 
-BNT162b2 booster recipients were matched with MOderna booster recipients on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation.
+BNT162b2 booster recipients were matched with Moderna booster recipients on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation.
 These matching criteria ensure that treatment groups are also matched within JCVI priority group.
@@
 
+# Supplement
+
+## JCVI vaccine priority groups
+
+The Joint Committee on Vaccine and Immunisation (JCVI) priority groups were defined as follows:
+
+```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE}
+
+tibble(
+  `Priority group` = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10a", "10b"),
+  `Description` = c(
+    "Residents in a care home for older adults\nStaff working in care homes for older adults",
+    "All those 80 years of age and over\nFrontline health and social care workers",
+    "All those 75-79 years of age",
+    "All those 70-74 years of age\nIndividuals aged 16-69 in a high risk group",
+    "All those 65-69 years of age",
+    "Adults aged 16-64 years in an at-risk group",
+    "All those 60-64 years of age",
+    "All those 55-59 years of age",
+    "All those 50-54 years of age",
+    "All those 40-49 years of age",
+    "All those 18-39 years of age"
+  )
+) %>%
+mutate(Description = str_replace_all(Description, "\n", "<br>")) %>% 
+gt() %>% 
+fmt_markdown(columns = TRUE)
+
+```
+
+See original priority groups here: https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1007737/Greenbook_chapter_14a_30July2021.pdf#page=15
+See revised priority groups here: https://www.england.nhs.uk/coronavirus/wp-content/uploads/sites/52/2021/07/C1327-covid-19-vaccination-autumn-winter-phase-3-planning.pdf
+
+Note that we excluded care home residents and health care workers in our analysis, so members of JCVI group 1 are not included and JCVI group 2 includes only those aged 80 and over.
+The original priority group list has 9 groups, with a 10th group implicitly defined as "everybody else".
+Here we explicitly define this group, and split into two (10a and 10b) because of the earlier booster eligibility in the 40-49 group from 15 November 2021 onwards.
+(https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds)
+
 \newpage

### 2022-04-23 analysis/study_definition.py
move local scripts

@@ -972,6 +972,17 @@ study = StudyDefinition(
     date_format="YYYY-MM-DD",
     find_first_match_in_period=True,
   ),
+  
+  covidadmitted_ccdays = patients.admitted_to_hospital(
+        returning = "days_in_critical_care",
+        find_first_match_in_period = True,
+        with_admission_method=["21", "22", "23", "24", "25", "2A", "2B", "2C", "2D", "28"],
+        with_these_diagnoses=codelists.covid_icd10,
+        on_or_after="covid_vax_disease_3_date",
+        return_expectations={
+        "category": {"ratios": {"0": 0.75, "1": 0.20,  "2": 0.05}},
+        "incidence": 0.5,
+      },
 
   
   # Covid-related death

### 2022-04-23 lib/dummydata/dummydata.R

@@ -58,7 +58,7 @@ sim_list = lst(
   ),
 
   age = bn_node(
-    ~as.integer(rnorm(n=..n, mean=60, sd=15))
+    ~as.integer(rnorm(n=..n, mean=60, sd=14))
   ),
 
   age_august2021 = bn_node(~age),
@@ -306,6 +306,11 @@ sim_list = lst(
   covidadmitted_day = bn_node(
     ~as.integer(runif(n=..n, covid_vax_disease_3_day, covid_vax_disease_3_day+100)),
     missing_rate = ~0.7
+  ),
+
+  covidadmitted_ccdays = bn_node(
+    ~as.integer(rpois(n=..n, 0.5)),
+    missing_rate = ~is.na(covidadmitted_day)
   ),
 
   admitted_unplanned_day = bn_node(

### 2022-04-23 analysis/data_process.R
include covid admissions in attendance definition; add placeholder for ICU admissions

@@ -229,10 +229,7 @@ data_processed <- data_extract %>%
 
     prior_covid_infection = !is.na(postest_0_date)  | !is.na(covidemergency_0_date) | !is.na(covidadmitted_0_date) | !is.na(primary_care_covid_case_0_date),
 
-
-    #covidemergency_1_date = pmin(covidemergency_1_date, covidadmitted_1_date, na.rm=TRUE),
-
-
+    covidemergency_1_date = pmin(covidemergency_1_date, covidadmitted_1_date, na.rm=TRUE),
 
     # latest covid event before study start
     anycovid_0_date = pmax(postest_0_date, covidemergency_0_date, covidadmitted_0_date, na.rm=TRUE),
@@ -243,6 +240,9 @@ data_processed <- data_extract %>%
 
     # earliest covid event after study start
     anycovid_date = pmin(postest_date, covidemergency_date, covidadmitted_date, coviddeath_date, na.rm=TRUE),
+
+    #FIXME ignores any covid-icu episodes subsequent to the first known covid admission, so we need to extract multiple covid admissions to capture them
+    #covidicu_date = if_else(covidadmitted_ccdays>0, covidadmitted_date, Date(NA)),
 
     noncoviddeath_date = if_else(!is.na(death_date) & is.na(coviddeath_date), death_date, as.Date(NA_character_)),
 

### 2022-04-23 analysis/study_definition.py
bug

@@ -974,16 +974,16 @@ study = StudyDefinition(
   ),
   
   covidadmitted_ccdays = patients.admitted_to_hospital(
-        returning = "days_in_critical_care",
-        find_first_match_in_period = True,
-        with_admission_method=["21", "22", "23", "24", "25", "2A", "2B", "2C", "2D", "28"],
-        with_these_diagnoses=codelists.covid_icd10,
-        on_or_after="covid_vax_disease_3_date",
-        return_expectations={
-        "category": {"ratios": {"0": 0.75, "1": 0.20,  "2": 0.05}},
-        "incidence": 0.5,
-      },
-
+    returning = "days_in_critical_care",
+    find_first_match_in_period = True,
+    with_admission_method=["21", "22", "23", "24", "25", "2A", "2B", "2C", "2D", "28"],
+    with_these_diagnoses=codelists.covid_icd10,
+    on_or_after="covid_vax_disease_3_date",
+    return_expectations={
+      "category": {"ratios": {"0": 0.75, "1": 0.20,  "2": 0.05}},
+      "incidence": 0.5,
+    },
+  ),
   
   # Covid-related death
   coviddeath_date=patients.with_these_codes_on_death_certificate(

### 2022-04-23 write-up/report.Rmd
use new fct_Recdoerelevel function

@@
 
+## Import custom user functions from lib
+source(here("lib", "functions", "utility.R"))
+
 ## Import design elements
@@
   mutate(
-    outcome_descr = fct_recode(as.character(outcome),  !!!recoder$outcome),
-    subgroup_descr = fct_recode(subgroup,  !!!recoder$subgroup),
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
     subgroup_level_descr = replace_na(subgroup_level_descr, "")
@@
 
-`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="moderna") %>% pull(pct_all))`) eligible for matching.
-`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
+`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type=="moderna") %>% pull(pct_all))`) eligible for matching.
+`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
 
@@
   mutate(
-    treatment_descr = fct_recode(as.character(treatment), !!!recoder$treatment),
-    status_descr = fct_recode(as.character(status), !!!recoder$status),
+    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
+    status_descr = fct_recoderelevel(as.character(status), recoder$status),
     cumuln = cumuln*((treatment*2) - 1)

### 2022-04-23 analysis/km_combine.R

@@ -49,27 +49,24 @@ metaparams <-
   expand_grid(
     outcome = factor(c("postest", "covidemergency", "covidadmittedproxy1", "covidadmitted", "coviddeath", "noncoviddeath")),
     subgroup = factor(recoder$subgroup),
-    #treatment = recoder$treatment
   ) %>%
   mutate(
-    #treatment_descr = fct_recode(as.character(treatment),  !!!recoder$treatment),
-    outcome_descr = fct_recode(as.character(outcome),  !!!recoder$outcome),
-    subgroup_descr = fct_recode(subgroup,  !!!recoder$subgroup),
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
   )
 
-
-# create lookup -- quasiquotation not working as expected inside pmap, so can't use fct_recode(var, !!!recoder$var) as anticipated
-subgroup_level_lookup <- map_dfr(recoder, ~ tibble(subgroup_level=., subgroup_level_descr=names(.)), .id="variable")
 
 km_estimates <- metaparams %>%
   mutate(
     data = pmap(list(matchset, subgroup, outcome), function(matchset, subgroup, outcome) {
       dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("km_estimates.csv")), na="NA")
-      subgroup <- as.character(subgroup)
-      dat[[subgroup]] <- as.character(dat[[subgroup]])
-      lookup <- subgroup_level_lookup[subgroup_level_lookup$variable==subgroup,]
-      right_join(lookup, dat, by=c("subgroup_level"=subgroup)) %>%
-        select(-variable)
+      dat %>%
+      add_column(
+        subgroup_level = as.character(.[[subgroup]]),
+        subgroup_level_descr = fct_recoderelevel(.[[subgroup]], recoder[[subgroup]]),
+        .before=1
+      ) %>%
+      select(-all_of(subgroup))
     })
   ) %>%
   unnest(data)
@@ -81,10 +78,13 @@ km_contrasts_daily <- metaparams %>%
   mutate(
     data = pmap(list(matchset, outcome, subgroup), function(matchset, outcome, subgroup){
         dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("km_contrasts_daily.csv")), na="NA")
-        subgroup <- as.character(subgroup)
[…]

### 2022-04-23 analysis/match_report.R

@@ -100,7 +100,7 @@ data_coverage <-
   ungroup() %>%
   mutate(
     status = factor(status, levels=c("unmatched", "matched")),
-    status_descr = fct_recode(status, !!!recoder$status)
+    status_descr = fct_recoderelevel(status, recoder$status)
   ) %>%
   arrange(treatment, status_descr, vax3_date)
 
@@ -118,7 +118,7 @@ xmax <- max(data_coverage$vax3_date )+1
 plot_coverage_n <-
   data_coverage %>%
   mutate(
-    treatment_descr = fct_recode(as.character(treatment), !!!recoder$treatment),
+    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
     n=n*((treatment*2) - 1)
   ) %>%
   ggplot()+
@@ -174,7 +174,7 @@ ggsave(plot_coverage_n, filename="coverage_count.png", path=output_dir)
 plot_coverage_cumuln <-
   data_coverage %>%
   mutate(
-    treatment_descr = fct_recode(as.character(treatment), !!!recoder$treatment),
+    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
     cumuln=cumuln*((treatment*2) - 1)
   ) %>%
   ggplot()+
@@ -286,7 +286,7 @@ tab_summary_baseline <-
   data_matched_baseline %>%
   mutate(
     N = 1L,
-    treatment_descr = fct_recode(as.character(treatment), !!!recoder$treatment),
+    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
   ) %>%
   select(
     treatment_descr,

### 2022-04-23 lib/functions/utility.R

@@ -6,8 +6,6 @@ ceiling_any <- function(x, to=1){
   ceiling(plyr::round_any(x/to, 1/100000000))*to
 }
 
-
-
 fct_case_when <- function(...) {
   # uses dplyr::case_when but converts the output to a factor,
   # with factors ordered as they appear in the case_when's  ... argument
@@ -16,4 +14,11 @@ fct_case_when <- function(...) {
   levels <- levels[!is.na(levels)]
   factor(dplyr::case_when(...), levels=levels)
 }
+
+fct_recoderelevel <- function(x, lookup){
+  stopifnot(!is.na(names(lookup)))
+  factor(x, levels=lookup, labels=names(lookup))
+}
+
+
 

### 2022-04-23 analysis/model.R
use new function

@@ -376,7 +376,7 @@ foreach::foreach(sub = subgroup) %dopar% {
       summary3$glance,
     ) %>%
     mutate(
-      model_descr = fct_recode(as.character(model), !!!model_descr)
+      model_descr = fct_recoderelevel(as.character(model), model_descr)
     )
   write_csv(model_glance, fs::path(output_dir, "glance.csv"))
 
@@ -388,7 +388,7 @@ foreach::foreach(sub = subgroup) %dopar% {
       summary3$tidy,
     ) %>%
     mutate(
-      model_descr = fct_recode(as.character(model), !!!model_descr)
+      model_descr = fct_recoderelevel(as.character(model), model_descr)
     )
   write_csv(model_tidy, fs::path(output_dir, "tidy.csv"))
 

### 2022-04-23 analysis/data_process.R
use correct types for critical care days

@@ -45,7 +45,7 @@ if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
 
   data_custom_dummy <- read_feather(here("lib", "dummydata", "dummyinput.feather")) %>%
     mutate(
-     msoa = sample(factor(c("1", "2")), size=n(), replace=TRUE) # override msoa so matching success more likely
+      msoa = sample(factor(c("1", "2")), size=n(), replace=TRUE), # override msoa so matching success more likely
     )
 
 
@@ -241,6 +241,7 @@ data_processed <- data_extract %>%
     # earliest covid event after study start
     anycovid_date = pmin(postest_date, covidemergency_date, covidadmitted_date, coviddeath_date, na.rm=TRUE),
 
+    covidadmitted_ccdays = as.integer(as.character(covidadmitted_ccdays)),
     #FIXME ignores any covid-icu episodes subsequent to the first known covid admission, so we need to extract multiple covid admissions to capture them
     #covidicu_date = if_else(covidadmitted_ccdays>0, covidadmitted_date, Date(NA)),
 

### 2022-04-23 lib/dummydata/dummydata.R

@@ -309,7 +309,7 @@ sim_list = lst(
   ),
 
   covidadmitted_ccdays = bn_node(
-    ~as.integer(rpois(n=..n, 0.5)),
+    ~as.factor(as.integer(rpois(n=..n, 0.5))),
     missing_rate = ~is.na(covidadmitted_day)
   ),
 

### 2022-04-23 analysis/data_process.R
use correct var names

@@ -229,7 +229,7 @@ data_processed <- data_extract %>%
 
     prior_covid_infection = !is.na(postest_0_date)  | !is.na(covidemergency_0_date) | !is.na(covidadmitted_0_date) | !is.na(primary_care_covid_case_0_date),
 
-    covidemergency_1_date = pmin(covidemergency_1_date, covidadmitted_1_date, na.rm=TRUE),
+    covidemergency_date = pmin(covidemergency_date, covidadmitted_date, na.rm=TRUE),
 
     # latest covid event before study start
     anycovid_0_date = pmax(postest_0_date, covidemergency_0_date, covidadmitted_0_date, na.rm=TRUE),

### 2022-04-23 write-up/protocol.Rmd
add draft protocol; add styles; add bibliography

--- 2022-04-23 write-up/report.Rmd
@@
 ---
-title: "Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: preliminary report"
+title: "Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: protocol"
 output:
@@
 
-matchset <- "A"
-
-match_flowchart <-  read_csv(here("output", "release-objects", matchset, "match_flowchart.csv"))
-match_coverage <- read_csv(here("output", "release-objects", matchset, "match_coverage.csv"))
-
-match_table1 <- read_csv(here("output", "release-objects", matchset, "match_table1.csv"))
-match_table1by <- read_csv(here("output", "release-objects", matchset, "match_table1by.csv"))
-match_smd <-  read_csv(here("output", "release-objects", matchset, "match_smd.csv"))
-
-km_estimates <- read_csv(here("output", "release-objects", matchset, "km_estimates.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-km_contrasts_daily <- read_csv(here("output", "release-objects", matchset, "km_contrasts_daily.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-km_contrasts_overall <- read_csv(here("output", "release-objects", matchset, "km_contrasts_overall.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-
-  outcomes <- 
-    c(
-      "postest",
-      #"covidemergency",
-      "covidadmitted",
-      #"covidcc",
-      "coviddeath",
-      "noncoviddeath"
-    ) %>% 
-    set_names(., .)
-
 ```
@@
 \newpage
+
+## Objectives
+
+To assess the compare the two most widely used COVID-19 booster vaccinations in England, Pfizer-BioNTech (BNT162b2) and Moderna (mRNA-1273) in their effectiveness against COVID-19-related outcomes.
 
@@
 
-All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>.
-With the approval of NHS England, primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
+This study will use the [OpenSAFELY platform](https://opensafely.org/), providing secure access to the OpenSAFELY-TPP database containing primary care records managed by the GP software provider TPP, roughly 40% of English primary care practices in England.
+These recorded are linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
 COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS).
 
-## Eligibility criteria
+## Study population
 
-We considered all adults aged $\ge$ 18 years who received a booster dose of BNT162b2 or Moderna between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyrend_date), "%e %B %Y")` inclusive (the "recruitment period").
-People were eligible for matching if: they were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.
+We will consider all adults aged $\ge$ 18 years who received a booster dose of BNT162b2 or Moderna between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyrend_date), "%e %B %Y")` inclusive (the "recruitment period").
+People will be eligible for matching if:
 
-## Matching
+* they are registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting;
+* they received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers);
+* they are not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care;
+* there is no evidence of SARS-CoV-2 infection or COVID-19 disease within the 90 days prior to booster dose;
+* they are not undergoing an unplanned hospital admission at the time of boosting;
+* they have complete information on sex, ethnicity, deprivation, and NHS region.
 
-BNT162b2 booster recipients were matched with Moderna booster recipients on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation.
-These matching criteria ensure that treatment groups are also matched within JCVI priority group.
+## Exposure
+
+The exposure of interest is the brand of COVID-19 vaccine administered for the third or booster dose; either BNT162b2 or Moderna.
+Other vaccine brands were used much less frequently &mdash; for example if there were contraindications or previous reactions to BNT162b2 or Moderna brands &mdash; and so will not be considered in this study.
 
@@
 
-The following outcomes were considered: 
+The following primary outcomes were considered: 
 * Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date.
@@
 * COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
+This outcome also includes COVID-19 admissions (see below) to pick up more severe disease episodes that may have bypassed A&E.
 * COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes. 
@@
 
-<!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->
+Additional secondary outcomes:
+* COVID-19 A&E attendances that resulted in a hospital admission, according to the discharge location. This can be used as a surrogate for hospital admission 
+* COVID-19 ICU / critical care admission
 
-<!--#
+ICU admission and mortality are expected to be low in this group of healthy vaccine recipients, and there may be insufficient statistical power to assess these outcomes reliably.
 
-## Follow-up
+## Study duration
 
-Each person was followed from the day of boosting (i.e., time zero) until the earliest of outcome, death, practice de-registration, 10 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.
+Time-zero is the day of booster dose receipt. Booster recipients will be followed from the this day until the first occurrence of:
+
+* The outcome of interest
+* death
+* practice de-registration
+* 10 weeks since boosting
+* fourth vaccine dose
+* `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.
+
+Outcome events occurring on the day of boosting will be considered to have occurred _after_ boosting, one day after time-zero (i.e., boosting is assumed to have occurred at the start of the day, and all other events are assumed to have occurred at the end of the day).
+This is necessarily true for death and likely true for hospitalisation outcomes.
+For positive SARS-CoV-2 tests the order of events is less clear, but given the policy of a recent positive test precluding vaccination, this is a reasonable assumption.
+
+## Matching
+
+BNT162b2 booster recipients will be matched with Moderna booster recipients on one or more of the following characteristics, calculated as at the day of boosting:
+
+Essential matching criteria:
+* date of booster dose
+* primary course vaccine brand
+* date of second vaccine dose, within 7 days
+* evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission)
+* clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither)
+* age, within 3 years, and age groups used by JCVI for prioritisation
+* Geographical region, one of the below in decreasing order of precision:
+  * Lower Layer Super Output Area (LSOA)
+  * Middle Layer Super Output Area (MSOA)
+  * NHS Strategic Health Partnership (STP)
+  * NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West)
+
+Additional matching criteria if possible:
+* English Index of Multiple Deprivation (IMD), within X days (tbd)
+* ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census);
+* existing morbidity count: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease
+* learning disabilities
+* severe mental illness
+* the number of SARS-CoV-2 tests in the 6 months prior to boosting
+* elective in-hospital episode at the time of boosting
+
+The final specification of matching characteristics will depend on the time required to complete the matching run (matching complexity will be reduced if it is infeasibly long), and number of candidates successfully matched (matching complexity will be reduced if success rate is too low).
+
+We will use one-to-one matching without replacement initially, but may revise to include many-to-many matching to increase power if required.
+However, many-to-many matching is only straight-forward with pure exact matching.
+If any variables are matched using calipers, then nearest-neighbour or optimal matching will be required instead.
+
+Deprivation (IMD) is defined within LSOAs (lower-layer super output areas).
+Each MSOA (middle-layer super output area) is composed of a small number of LSOAs.
+MSOA is the smallest geographical unit available within OpenSAFELY-TPP.
+If MSOA is used as a matching variable then deprivation will add little additional information, and may be poorly balanced, and so should not be used.
+If MSOA is not used, IMD should be used if possible.
 
@@
 
-We estimated Kaplan-Meier cumulative incidence curves separately by vaccine brand, and compared groups using the 10-week risk differences and risk ratios.
+We estimated Kaplan-Meier (KM) cumulative incidence curves separately by vaccine brand, and compared brands using the 10-week risk differences and risk ratios.
+
+Confidence limits for the KM estimate will be derived from complimentary log-log transformation.
+Confidence limits for the risk difference will be derived from the square root of the sum of squares the treatment-specific KM standard errors. The KM standard errors will use Greenwood's formula.
+Confidence limits for the relative risk??
+
+Alternatively, we could boostrap.
 
@@
 
-Data management and analyses were conducted in Python version 3.8.10 and R version 4.0.2.
-All code is shared openly for review and re-use under MIT open license at <https://github.com/opensafely/comparative-booster>.
-The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
-Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.
-Any reported figures based on counts below 6 are redacted or rounded for disclosure control.
+Data management and analyses will be conducted in Python version 3.8.10 and R version 4.0.2.
+All code will be shared openly for review and re-use under MIT open license at <https://github.com/opensafely/comparative-booster>.
 
 This study will follow the STROBE-RECORD reporting guidelines.
+
+## Disclosure control
+
+Any reported figures based on counts below 6 must be redacted or rounded for disclosure control.
+To reduce secondary disclosure risk (e.g., by using two unredacted values to derive a redacted value), rounding is preferred.
+
+Kaplan-Meier survival estimates are rounded upwards by deferring event-times until there are at least 7 events, so that each KM "step" is based on at least 7 events.
+The same principle is applied to KM confidence limits.
 
@@
 
-`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type=="moderna") %>% pull(pct_all))`) eligible for matching.
-`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
+`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="moderna") %>% pull(pct_all))`) eligible for matching.
+`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
 

### 2022-04-25 analysis/study_definition.py
used fixed_value

@@ -14,7 +14,7 @@ import codelists
 # import json module
 import json
 
-# import study dates defined in "design.R" script
+# import study dates defined in "./lib/design/study-dates.R" script
 with open("./lib/design/study-dates.json") as f:
   study_dates = json.load(f)
 
@@ -91,9 +91,9 @@ study = StudyDefinition(
       AND
       NOT has_died
       AND 
-      covid_vax_disease_3_date >= '{studystart_date}'
+      covid_vax_disease_3_date >= startdate
       AND
-      covid_vax_disease_3_date <= '{studyend_date}'
+      covid_vax_disease_3_date <= enddate
     """,
     # we define baseline variables on the day _before_ the study date (start date = day of first possible booster vaccination)
     registered=patients.registered_as_of(
@@ -102,10 +102,13 @@ study = StudyDefinition(
     has_died=patients.died_from_any_cause(
       on_or_before="covid_vax_disease_3_date - 1 day",
       returning="binary_flag",
-    ),
+    ), 
+   
+    startdate = patients.fixed_value(studystart_date),
+    enddate = patients.fixed_value(studyend_date),
+  
     
   ),
-  
   
   
   #################################################################

### 2022-04-25 analysis/km.R
Update km.R

@@ -328,7 +328,7 @@ kmcontrast <- function(data, cuts=NULL){
 
       # survival ratio, standard error, and confidence limits
       kmsr = surv_1 / surv_0,
-      kmsr.ln = log(kmrs),
+      kmsr.ln = log(kmsr),
       kmsr.ln.se = (surv.se_0/surv_0) + (surv.se_1/surv_1), #because cmlhaz = -log(surv) and cmlhaz.se = surv.se/surv
       kmsr.ll = exp(kmsr.ln + qnorm(0.025)*kmsr.ln.se),
       kmsr.ul = exp(kmsr.ln + qnorm(0.975)*kmsr.ln.se),

### 2022-04-25 analysis/km.R
add placeholder for hazard ratio

@@ -279,7 +279,11 @@ km_incidence <-
     !!subgroup_sym,
     treatment,
     time, interval,
-    surv, surv.se, surv.ll, surv.ul, risk, risk.ll, risk.ul, n.atrisk, n.event, n.censor, sumerand, rate, cml.atrisk, cml.event, cml.censor, cml.sumerand, cml.rate
+    surv, surv.se, surv.ll, surv.ul,
+    risk, risk.ll, risk.ul,
+    haz=haz_km, haz.se=haz_km.se,
+    n.atrisk, n.event, n.censor, sumerand, rate,
+    cml.atrisk, cml.event, cml.censor, cml.sumerand, cml.haz = cml.haz_km
   )
 
 kmcontrast <- function(data, cuts=NULL){
@@ -343,6 +347,13 @@ kmcontrast <- function(data, cuts=NULL){
       kmrd.se = sqrt( (surv.se_0^2) + (surv.se_1^2) ), # combining SEs from greenwood's formula
       kmrd.ll = kmrd + qnorm(0.025)*kmrd.se,
       kmrd.ul = kmrd + qnorm(0.975)*kmrd.se,
+
+      # hazard ratio, standard error and confidence limits
+      # kmhr = haz_1 / haz_0,
+      # kmhr.se = sqrt( (haz.se_0^2) + (haz.se_1^2) ),
+      # kmhr.ll = kmrd + qnorm(0.025)*kmhr.se,
+      # kmhr.ul = kmrd + qnorm(0.975)*kmhr.se,
+
 
       # incidence rate ratio
       irr = rate_1 / rate_0,

### 2022-04-25 protocol/protocol.Rmd

--- 2022-04-23 write-up/protocol.Rmd
@@
 output:
-  html_document:
-    self_contained: yes
+  word_document: default
   pdf_document:
@@
     toc: no
-  word_document: default
-bibliography: references.bib
-csl: vancouver.csl
-link-citations: yes
-zotero: yes
+  html_document:
+    self_contained: yes
+#bibliography: references.bib
+#csl: vancouver.csl
+#link-citations: yes
+#zotero: yes
 header-includes:
@@
 Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection.
-* COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
+* COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes.
 This outcome also includes COVID-19 admissions (see below) to pick up more severe disease episodes that may have bypassed A&E.
@@
 Essential matching criteria:
+
 * date of booster dose
@@
 * Geographical region, one of the below in decreasing order of precision:
+
   * Lower Layer Super Output Area (LSOA)
@@
 Additional matching criteria if possible:
+
 * English Index of Multiple Deprivation (IMD), within X days (tbd)
@@
 Confidence limits for the risk difference will be derived from the square root of the sum of squares the treatment-specific KM standard errors. The KM standard errors will use Greenwood's formula.
-Confidence limits for the relative risk??
+Confidence limits for the survival ratio are straight-forward, and can be based on the standard error of the cumulative hazard.
+But confidence limits for the risk ratio (or relative risk) are more complicated.
+TBD.
 
-Alternatively, we could boostrap.
+Alternatively, we could bootstrap.
 
@@
 The same principle is applied to KM confidence limits.
-
-# Results
-
-## Study population and matching
-
-`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="moderna") %>% pull(pct_all))`) eligible for matching.
-`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
-
-#### Table 1: inclusion criteria
-
-```{r table-flowchart, echo=FALSE, warning=FALSE, message=FALSE}
-
-match_flowchart %>%
-  select(-crit) %>%
-   pivot_wider(
-    id_cols = c(criteria, vax3_type),
-    names_from = vax3_type,
-    values_from = c(n, n_exclude, pct_all)
-  ) %>%
-  gt() %>%
-  cols_label(
-    criteria = "",
-
-    n_pfizer = "N",
-    n_moderna = "N",
-
-    n_exclude_pfizer = "N excluded",
-    n_exclude_moderna = "N excluded",
-
-    pct_all_pfizer = "% remaining",
-    pct_all_moderna = "% remaining",
-  ) %>%
-  tab_spanner(
-    label = "BNT162b2",
-    columns = ends_with("pfizer")
-  ) %>%
-   tab_spanner(
-    label = "Moderna mRNA-1273",
-    columns = ends_with("moderna")
-  ) %>%
-  # tab_spanner(
-  #   label = "mRNA-1273",
-  #   columns = ends_with("moderna")
-  # ) %>%
-  fmt_percent(
-    columns = starts_with(c("pct_")),
-    decimals = 1
-  ) %>%
-  fmt_number(
-    columns = starts_with(c("n_")),
-    decimals= 0
-  ) %>%
-  fmt_missing(
-    everything(),
-    missing_text="-"
-  ) %>%
-  tab_style(
-    style=cell_text(whitespace ="pre-wrap"),
-    #style=cell_text(style ="italic"),
-    locations = cells_body(columns = "criteria")
-  )
-```
-
-\newpage 
-#### Figure 1: Cumulative matching success
-
-```{r, figure-matchingsuccess, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
-#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))
-
-xmin <- min(match_coverage$vax3_date )
-xmax <- max(match_coverage$vax3_date )+1
-
-plot_coverage_cumuln <-
-  match_coverage %>%
-  mutate(
-    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
-    status_descr = fct_recoderelevel(as.character(status), recoder$status),
-    cumuln = cumuln*((treatment*2) - 1)
-  ) %>%
-  ggplot()+
-  geom_col(
-    aes(
-      x=vax3_date+0.5,
-      y=cumuln,
-      group=paste0(treatment,status_descr),
-      fill=treatment_descr,
-      alpha=status_descr,
-      colour=NULL
-    ),
-    position=position_stack(reverse=TRUE),
-    width=1
-  )+
-  geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
-  scale_x_date(
-    breaks = unique(lubridate::ceiling_date(match_coverage$vax3_date, "1 month")),
-    limits = c(xmin-1, NA),
-    labels = scales::label_date("%d/%m"),
-    expand = expansion(add=1),
-  )+
-  scale_y_continuous(
-    labels = ~scales::label_number(accuracy = 1, big.mark=",")(abs(.)),
-    expand = expansion(c(0, NA))
-  )+
-  scale_fill_brewer(type="qual", palette="Set2")+
-  scale_colour_brewer(type="qual", palette="Set2")+
-  scale_alpha_discrete(range= c(0.8,0.4))+
-  labs(
-    x="Date",
-    y="Cumulative booster vaccines",
-    colour=NULL,
-    fill=NULL,
-    alpha=NULL
-  ) +
-  theme_minimal()+
-  theme(
-    axis.line.x.bottom = element_line(),
-    axis.text.x.top=element_text(hjust=0),
-    strip.text.y.right = element_text(angle = 0),
-    axis.ticks.x=element_line(),
-    legend.position = "bottom"
-  )+
-  NULL
-
-plot_coverage_cumuln
-cat("  \n\n")
-```
-
-\newpage
-
-#### Table 2: baseline characteristics
-
-```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
-table1 <- 
-  match_table1 %>%
-  #filter(outcome %in% outcomes) %>%
-  group_by(variable) %>%
-  mutate(
-    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
-  ) %>%
-  mutate(
-    across(
-      starts_with("stat_"),
-      ~if_else(is.na(.x) & row_number()==1, "", .x)
-    )
-  ) %>%
-  ungroup() %>%
-  mutate(
-    order = cumsum(lag(variable, 1, first(variable))!=variable)
-  ) %>%
-  select(
-    variable, var_label, label, starts_with("stat_"), order
-  ) %>%
-  rename(!!!set_names(match_table1by$by_col, match_table1by$by_chr))
-
-table1 %>%
-  gt(groupname_col="treatment_descr") %>%
-  fmt_markdown(columns = "label") %>%
-  tab_style(
-    style = list(
-      cell_fill(color = "gray96")
-    ),
-    locations = cells_body(
-      rows = order%%2 == 0
-    )
-  ) %>%
-  tab_style(
-    style = list(
-      cell_text(weight = "bold")
-    ),
-    locations = cells_column_labels(everything())
-  ) %>%
-  cols_label(
-     label=md("Characteristic")
-  ) %>%
-  cols_hide(c("variable", "var_label", "order"))
-```
-
-## Estimated comparative effectiveness
-
-### Main analysis
-
-#### Table 3: Kaplan-Meier cumulative incidence estimates
-
-```{r table-km, echo=FALSE, warning=FALSE, message=FALSE}
-
-km_contrasts_overall %>%
-  transmute(
-    outcome_descr,
-    subgroup_descr,
-    subgroup_level_descr,
-    n.atrisk=n.atrisk_0,
-    cml.atrisk_0=label_number(1)(cml.atrisk_0/7), 
-    cml.atrisk_1=label_number(1)(cml.atrisk_1/7),
-    n.event_0, 
-    n.event_1,
-    riskCI_0 = paste0(label_number(0.01)(risk_0), " (", label_number(0.01)(risk.ll_0), "-", label_number(0.01)(risk.ul_0), ")"),
-    riskCI_1 = paste0(label_number(0.01)(risk_1), " (", label_number(0.01)(risk.ll_1), "-", label_number(0.01)(risk.ul_1), ")"),
-    kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
-    kmrr = paste0(label_number(0.01)(kmrr)),
-  ) %>%
-  gt(groupname_col="outcome_descr") %>%
-  cols_label(
-
-    outcome_descr = "Outcome",
-    subgroup_descr = "Subgroup",
-    subgroup_level_descr = "Subgroup",
-
-    n.atrisk = "N per group",
-
-    cml.atrisk_0 = "Person-weeks", 
-    cml.atrisk_1 = "Person-weeks",
-
-    n.event_0 = "Events",
-    n.event_1 = "Events",
-
-    riskCI_0 = "KM risk (95% CI)",
-    riskCI_1 = "KM risk (95% CI)",
-
-    kmrdCI = "Risk-difference (95% CI)",
-    kmrr = "Relative-risk"
-  ) %>%
-  tab_spanner(
-    label = "BNT162b2",
-    columns = ends_with("_0")
-  ) %>%
-   tab_spanner(
-    label = "Moderna",
-    columns = ends_with("_1")
-  )  %>%
-   tab_spanner(
-    label = "Comparison",
-    columns = starts_with("km")
-  )
-
-```
-
-#### Figure 3: Kaplan-Meier cumulative incidence estimates
-
-```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
-plot_km <- km_estimates %>%
-  filter(
-    subgroup=="all",
-    outcome %in% outcomes
-  ) %>%
-  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
-  geom_step(aes(x=time, y=1-surv))+
-  geom_rect(aes(xmin=time, xmax=leadtime, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
-  facet_grid(rows=vars(outcome_descr))+
-  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
-  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-  scale_x_continuous(breaks = seq(0,600,14))+
-  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
-  coord_cartesian(xlim=c(0, NA))+
-  labs(
-    x="Days",
-    y="Cumulative incidence",
-    colour=NULL,
-    title=NULL
-  )+
-  theme_minimal()+
-  theme(
-    axis.line.x = element_line(colour = "black"),
-    panel.grid.minor.x = element_blank(),
-    legend.position=c(.05,.95),
-    legend.justification = c(0,1),
-  )
-
-plot_km
-```
 
@@
 (https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds)
-
-\newpage
-
-## References

### 2022-04-25 write-up/report.Rmd

@@
   ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
-  geom_step(aes(x=time, y=1-surv))+
+  geom_step(aes(x=period_end, y=1-surv))+
   geom_rect(aes(xmin=time, xmax=leadtime, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
-  facet_grid(rows=vars(outcome_descr))+
+  facet_grid(rows=vars(outcome_descr), scales="free_y")+
   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
@@
   )+
-  theme_minimal()+
+  theme_minimal(base_size = 9)+
     theme(
-    axis.line.x = element_line(colour = "black"),
+      legend.position=c(0.05,0.1),
+      legend.justification = c(0,0),
+      legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+
       panel.grid.minor.x = element_blank(),
-    legend.position=c(.05,.95),
-    legend.justification = c(0,1),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
+
+plot_km
+
+plot_kmrd <- km_contrasts_daily %>%
+  filter(
+    subgroup=="all",
+    outcome %in% outcomes
+  ) %>%
+  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+  geom_step(aes(x=period_start, y=kmrd))+
+  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
+  facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  scale_x_continuous(breaks = seq(0,600,14))+
+  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
+  coord_cartesian(xlim=c(0, NA))+
+  labs(
+    x="Days",
+    y="Cumulative risk difference",
+    colour=NULL,
+    title=NULL
+  )+
+  theme_minimal(base_size = 9)+
+    theme(
+      legend.position="none",
+      legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
+
+plot_kmrd
+
+  plot_combined <- patchwork::wrap_plots(plot_km, plot_kmrd, plotkmrr, ncol=1)
+  ggsave(
+    filename="km.png",
+    path=output_dir_rmd,
+    plot=plot_combined,
   )
 
-plot_km
+plot_combined
+
 ```

### 2022-04-25 analysis/km.R
Update km.R

@@ -283,7 +283,7 @@ km_incidence <-
     risk, risk.ll, risk.ul,
     haz=haz_km, haz.se=haz_km.se,
     n.atrisk, n.event, n.censor, sumerand, rate,
-    cml.atrisk, cml.event, cml.censor, cml.sumerand, cml.haz = cml.haz_km
+    cml.atrisk, cml.event, cml.censor, cml.sumerand, cml.haz = cml.haz_km, cml.rate
   )
 
 kmcontrast <- function(data, cuts=NULL){

### 2022-04-25 analysis/match.R
add matching variables; increase number of threads

@@ -96,7 +96,7 @@ data_matchingcandidates <-
 library("doParallel")
 
 parallel::detectCores() # how many cores available?
-n_threads <- 2
+n_threads <- 8
 
 cluster <- parallel::makeCluster(
   n_threads,
@@ -106,14 +106,6 @@ print(cluster)
 #register it to be used by %dopar%
 doParallel::registerDoParallel(cl = cluster)
 
-
-
-foreach(
-  i = 1:10,
-  .combine = 'c'
-) %dopar% {
-  sqrt(i)
-}
 
 # create function that catches errors in case no matches are found within a thread
 safely_matchit <- purrr::safely(matchit)

### 2022-04-25 lib/design/design.R

@@ -167,10 +167,12 @@ if(exists("matchset")){
       "cev_cv",
       "vax12_type",
       #"vax2_week",
-      "region",
+      "stp",
       "sex",
+      "ethnicity",
 
-      #"multimorb",
+
+      "multimorb",
       "prior_covid_infection",
       #"immunosuppressed",
       #"status_hospplanned"
@@ -180,6 +182,7 @@ if(exists("matchset")){
     caliper <- c(
       age = 3,
       vax2_day = 7,
+      imd = 1000,
       NULL
     )
 

### 2022-05-05 analysis/data_process.R
add variant era subgroup

@@ -224,7 +224,12 @@ data_processed <- data_extract %>%
 
     # subgroups
     all=factor("all"),
-    age65plus=age>=65,
+    age65plus = age>=65,
+    variantera = fct_case_when(
+      covid_vax_disease_3_date <= as.Date("2021-12-31") ~ "Delta (29 Nov - 31 Dec)",
+      covid_vax_disease_3_date >= as.Date("2022-01-01") ~ "Omicron (1 Jan onwards)",
+      TRUE ~ NA_character_
+    ),
 
     prior_tests_cat = cut(prior_covid_test_frequency, breaks=c(0, 1, 2, 3, Inf), labels=c("0", "1", "2", "3+"), right=FALSE),
 

### 2022-05-05 create-project.R

@@ -384,11 +384,23 @@ actions_list <- splice(
   action_km("A", "age65plus", "coviddeath"),
   action_km("A", "age65plus", "noncoviddeath"),
 
+  comment("### Models by variant era ('variantera')"),
+
+  action_km("A", "variantera", "postest"),
+  action_km("A", "variantera", "covidemergency"),
+  action_km("A", "variantera", "covidadmittedproxy1"),
+  action_km("A", "variantera", "covidadmitted"),
+  #action_km("A", "variantera", "noncovidadmitted"),
+  #action_km("A", "variantera", "covidcc"),
+  action_km("A", "variantera", "coviddeath"),
+  action_km("A", "variantera", "noncoviddeath"),
+
+
   comment("# # # # # # # # # # # # # # # # # # #", "Combine KM estimates across outcomes and subgroups", "# # # # # # # # # # # # # # # # # # #"),
 
   action_km_combine(
     "A",
-    subgroups = c("all", "vax12_type", "prior_covid_infection", "age65plus", "cev_cv"),
+    subgroups = c("all", "vax12_type", "prior_covid_infection", "age65plus", "cev_cv", "variantera"),
     outcomes=c("postest", "covidemergency", "covidadmittedproxy1", "covidadmitted", "coviddeath", "noncoviddeath")
   ),
 

### 2022-05-05 lib/design/design.R

@@ -109,6 +109,10 @@ recoder <-
     prior_covid_infection = c(
       `No prior SARS-CoV-2 infection` = "FALSE",
       `Prior SARS-CoV-2 infection` = "TRUE"
+    ),
+    variantera = c(
+      `Delta (29 Nov - 31 Dec)` = "Delta (29 Nov - 31 Dec)",
+      `Omicron (1 Jan onwards)` = "Omicron (1 Jan onwards)"
     )
   )
 

### 2022-05-05 analysis/data_selection.R
headings

@@ -101,6 +101,9 @@ data_inclusioncriteria <- data_criteria %>%
 
 write_rds(data_inclusioncriteria, here("output", "data", "data_inclusioncriteria.rds"), compress="gz")
 
+
+## flowchart -- rounded so disclosure-safe ----
+
 data_flowchart <-
   data_inclusioncriteria %>%
   select(-patient_id) %>%
@@ -135,7 +138,7 @@ data_flowchart <-
 
 
 
-## flowchart -- rounded -- disclosure-safe
+## flowchart -- rounded so disclosure-safe ----
 data_flowchart_rounded <-
   data_inclusioncriteria %>%
   select(-patient_id) %>%

### 2022-05-05 analysis/match_report.R

@@ -416,6 +416,8 @@ data_flowchart_match <-
   )
 
 
+# flowchart -- rounded so disclosure-safe ----
+
 data_flowchart_match_rounded <-
   read_rds(here("output", "data", "data_inclusioncriteria.rds")) %>%
   left_join(

### 2022-05-05 analysis/km.R
combine cox and KM contrasts

@@ -278,8 +278,8 @@ kmcontrast <- function(data, cuts=NULL){
       treatment,
 
       time, lagtime, interval,
-      period_start = cut(time, cuts, right=TRUE, label=cuts[-length(cuts)]),
-      period_end = cut(time, cuts, right=TRUE, label=cuts[-1]),
+      period_start = as.integer(as.character(cut(time, cuts, right=TRUE, label=cuts[-length(cuts)]))),
+      period_end = as.integer(as.character(cut(time, cuts, right=TRUE, label=cuts[-1]))),
       period = cut(time, cuts, right=TRUE, label=paste0(cuts[-length(cuts)]+1, " - ", cuts[-1])),
 
       n.atrisk = n.risk,
@@ -370,9 +370,6 @@ kmcontrast <- function(data, cuts=NULL){
 
       # hazard ratio, standard error and confidence limits
 
-      #hr_cml = cml.haz_1 / cml.haz_0, # CHECK
-      #hr_cmlac = cml.haz_ac_1 / cml.haz_ac_0, # CHECK
-
       # incidence rate ratio
       irr = rate_1 / rate_0,
       irr.ln.se = sqrt((1/n.event_0) + (1/n.event_1)),
@@ -420,8 +417,6 @@ km_contrasts_daily <- kmcontrast(data_surv)
 km_contrasts_cuts <- kmcontrast(data_surv, postbaselinecuts)
 km_contrasts_overall <- kmcontrast(data_surv, c(0,maxfup))
 
-write_csv(km_contrasts_daily, fs::path(output_dir, "km_contrasts_daily.csv"))
-write_csv(km_contrasts_overall, fs::path(output_dir, "km_contrasts_overall.csv"))
 
 
 ## Cox models ----
@@ -476,8 +471,12 @@ coxcontrast <- function(data, cuts=NULL){
     ) %>%
     select(!!subgroup_sym, period_start, period_end, cox_obj_tidy) %>%
     unnest(cox_obj_tidy) %>%
-    mutate(
+    transmute(
+      !!subgroup_sym,
+      period_start,
+      period_end,
       coxhaz = exp(estimate),
+      coxhaz.se = robust.se,
       coxhaz.ll = exp(estimate + qnorm(0.025)*robust.se),
       coxhaz.ul = exp(estimate + qnorm(0.975)*robust.se),
     )
@@ -488,3 +487,11 @@ coxcontrast <- function(data, cuts=NULL){
[…]

### 2022-05-05 analysis/km_combine.R

@@ -57,7 +57,7 @@ metaparams <-
 km_estimates <- metaparams %>%
   mutate(
     data = pmap(list(matchset, subgroup, outcome), function(matchset, subgroup, outcome) {
-      dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("km_estimates.csv")), na="NA")
+      dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("km_estimates.csv")), na="NA", col_types = cols())
       dat %>%
       add_column(
         subgroup_level = as.character(.[[subgroup]]),
@@ -72,10 +72,10 @@ km_estimates <- metaparams %>%
 write_csv(km_estimates, fs::path(output_dir, "km_estimates.csv"))
 
 
-km_contrasts_daily <- metaparams %>%
+contrasts_daily <- metaparams %>%
   mutate(
     data = pmap(list(matchset, outcome, subgroup), function(matchset, outcome, subgroup){
-        dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("km_contrasts_daily.csv")), na="NA")
+        dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("contrasts_daily.csv")), na="NA", col_types = cols())
         dat %>%
           add_column(
             subgroup_level = as.character(.[[subgroup]]),
@@ -88,14 +88,14 @@ km_contrasts_daily <- metaparams %>%
   ) %>%
   unnest(data)
 
-write_csv(km_contrasts_daily, fs::path(output_dir, "km_contrasts_daily.csv"))
+write_csv(contrasts_daily, fs::path(output_dir, "contrasts_daily.csv"))
 
 
 
-km_contrasts_overall <- metaparams %>%
+contrasts_overall <- metaparams %>%
   mutate(
     data = pmap(list(matchset, outcome, subgroup), function(matchset, outcome, subgroup) {
-        dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("km_contrasts_overall.csv")), na="NA")
+        dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("contrasts_overall.csv")), na="NA", col_types = cols())
         dat %>%
           add_column(
             subgroup_level = as.character(.[[subgroup]]),
@@ -108,4 +108,4 @@ km_contrasts_overall <- metaparams %>%
   ) %>%
[…]

### 2022-05-05 analysis/release_objects.R

@@ -27,8 +27,8 @@ for(matchset in c("A")){
     ## KM ----
 
   fs::file_copy(here("output", "match", "A", "km", "combined", "km_estimates.csv"), here("output", "release-objects", matchset, "km_estimates.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "match", "A", "km", "combined", "km_contrasts_daily.csv"), here("output", "release-objects", matchset, "km_contrasts_daily.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "match", "A", "km", "combined", "km_contrasts_overall.csv"), here("output", "release-objects", matchset, "km_contrasts_overall.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", "A", "km", "combined", "contrasts_daily.csv"), here("output", "release-objects", matchset, "contrasts_daily.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", "A", "km", "combined", "contrasts_overall.csv"), here("output", "release-objects", matchset, "contrasts_overall.csv"), overwrite = TRUE)
 
 }
 

### 2022-05-05 run-local.R

@@ -95,6 +95,15 @@ action_km("A", "age65plus", "covidadmitted")
 action_km("A", "age65plus", "coviddeath")
 action_km("A", "age65plus", "noncoviddeath")
 
+action_km("A", "variantera", "postest")
+action_km("A", "variantera", "covidemergency")
+action_km("A", "variantera", "covidadmittedproxy1")
+action_km("A", "variantera", "covidadmitted")
+#action_km("A", "variantera", "noncovidadmitted")
+#action_km("A", "variantera", "covidcc")
+action_km("A", "variantera", "coviddeath")
+action_km("A", "variantera", "noncoviddeath")
+
 action_km_combine("A")
 
 action_ls(

### 2022-05-05 lib/functions/survival.R
use complimentary log-log for surv confidence limits

@@ -66,8 +66,8 @@ tidy_surv <-
           summand = n.event / ((n.risk - n.event) * n.risk),
 
           surv=cumprod(1 - n.event / n.risk),
-          surv.ll = conf.low,
-          surv.ul = conf.high,
+          #surv.ll = conf.low,
+          #surv.ul = conf.high,
           surv.se = surv * sqrt(cumsum(summand)), #greenwood's formula
 
           lsurv = log(surv),
@@ -89,6 +89,9 @@ tidy_surv <-
 
           llsurv = log(-log(surv)),
           llsurv.se = sqrt((1 / log(surv)^2) * cumsum(summand)),
+
+          surv.ll = exp(-exp(llsurv + qnorm(0.025)*llsurv.se)),
+          surv.ul = exp(-exp(llsurv + qnorm(0.975)*llsurv.se)),
 
       )
     }
@@ -117,8 +120,8 @@ tidy_surv <-
 
           surv=cumprod(1 - n.event / n.risk),
 
-          surv.ll = conf.low,
-          surv.ul = conf.high,
+          #surv.ll = conf.low,
+          #surv.ul = conf.high,
           surv.se = surv * sqrt(cumsum(summand)), #greenwood's formula
 
           lsurv = log(surv),
@@ -139,6 +142,10 @@ tidy_surv <-
 
           llsurv = log(-log(surv)),
           llsurv.se = sqrt((1 / log(surv)^2) * cumsum(summand)),
+
+
+          surv.ll = exp(-exp(llsurv + qnorm(0.025)*llsurv.se)),
+          surv.ul = exp(-exp(llsurv + qnorm(0.975)*llsurv.se)),
         )
     }
 

### 2022-05-05 write-up/report.Rmd
update report

--- 2022-04-23 write-up/report.Rmd
@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, message=FALSE}
 
@@
   )
-km_contrasts_daily <- read_csv(here("output", "release-objects", matchset, "km_contrasts_daily.csv")) %>% 
+contrasts_daily <- read_csv(here("output", "release-objects", matchset, "contrasts_daily.csv")) %>% 
   mutate(
@@
   )
-km_contrasts_overall <- read_csv(here("output", "release-objects", matchset, "km_contrasts_overall.csv")) %>% 
+contrasts_overall <- read_csv(here("output", "release-objects", matchset, "contrasts_overall.csv")) %>% 
   mutate(
@@
 
-km_contrasts_overall %>%
+contrasts_overall %>%
   transmute(
@@
     n.atrisk=n.atrisk_0,
-    cml.atrisk_0=label_number(1)(cml.atrisk_0/7), 
-    cml.atrisk_1=label_number(1)(cml.atrisk_1/7),
+    persontime_0=label_number(1)(persontime_0/7), 
+    persontime_1=label_number(1)(persontime_1/7),
     n.event_0, 
@@
 
-    cml.atrisk_0 = "Person-weeks", 
-    cml.atrisk_1 = "Person-weeks",
+    persontime_0 = "Person-weeks", 
+    persontime_1 = "Person-weeks",
 
@@
   geom_step(aes(x=time, y=1-surv))+
-  geom_rect(aes(xmin=time, xmax=leadtime, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
-  facet_grid(rows=vars(outcome_descr))+
+  geom_rect(aes(xmin=lagtime, xmax=time, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
+  facet_grid(rows=vars(outcome_descr), scales="free_y")+
   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
@@
     x="Days",
-    y="Cumulative incidence",
-    colour=NULL,
-    title=NULL
+    y=NULL,
+    title="Cumulative incidence",
+    colour=NULL
   )+
-  theme_minimal()+
+  theme_minimal(base_size = 9)+
     theme(
-    axis.line.x = element_line(colour = "black"),
+      legend.position=c(0.05,0.1),
+      legend.justification = c(0,0),
+      legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+
       panel.grid.minor.x = element_blank(),
-    legend.position=c(.05,.95),
-    legend.justification = c(0,1),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
+
+plot_km
+
+plot_kmrd <- contrasts_daily %>%
+  filter(
+    subgroup=="all",
+    outcome %in% outcomes
+  ) %>% 
+  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+  geom_step(aes(x=period_start, y=kmrd))+
+  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
+  facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  scale_x_continuous(breaks = seq(0,600,14))+
+  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
+  coord_cartesian(xlim=c(0, NA))+
+  labs(
+    x="Days",
+    y=NULL,
+    title="Cumulative risk difference",
+    colour=NULL
+  )+
+  theme_minimal(base_size = 9)+
+    theme(
+      legend.position="none",
+      legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
+
+plot_kmrd
+
+plot_kmrr <- contrasts_daily %>%
+  filter(
+    subgroup=="all",
+    outcome %in% outcomes
+  ) %>%
+  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+  geom_step(aes(x=period_start, y=kmrr))+
+  #geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrr.ll, ymax=kmrr.ul), alpha=0.1, colour="transparent")+
+  facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  scale_x_continuous(breaks = seq(0,600,14))+
+  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
+  coord_cartesian(xlim=c(0, NA))+
+  labs(
+    x="Days",
+    y=NULL,
+    title="Cumulative risk ratio",
+    colour=NULL
+  )+
+  theme_minimal(base_size = 9)+
+    theme(
+      legend.position="none",
+      legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
+
+plot_kmrr
+
+  plot_combined <- patchwork::wrap_plots(plot_km, plot_kmrd, plot_kmrr, ncol=3)
+  ggsave(
+    filename="km.png",
+    path=output_dir_rmd,
+    plot=plot_combined,
   )
 
-plot_km
+plot_combined
+
 ```

### 2022-05-05 write-up/report.Rmd
minor plot changes

@@
 match_flowchart <-  read_csv(here("output", "release-objects", matchset, "match_flowchart.csv"))
-match_coverage <- read_csv(here("output", "release-objects", matchset, "match_coverage.csv"))
+match_coverage <- read_csv(here("output", "release-objects", matchset, "match_coverage.csv")) %>%
+   mutate(
+    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
+  )
 
@@
   mutate(
-    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
-    status_descr = fct_recoderelevel(as.character(status), recoder$status),
+   # treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
+    status_descr = fct_rev(fct_recoderelevel(as.character(status), recoder$status)),
     cumuln = cumuln*((treatment*2) - 1)
@@
   scale_x_date(
-    breaks = unique(lubridate::ceiling_date(match_coverage$vax3_date, "1 month")),
+    breaks = seq(xmin, xmax+1, by=14),
+    #breaks = unique(lubridate::ceiling_date(match_coverage$vax3_date, "1 month")),
     limits = c(xmin-1, NA),
@@
     expand = expansion(add=1),
+    sec.axis = sec_axis(
+      trans = ~as.Date(.),
+      breaks=as.Date(seq(floor_date(xmin, "month"), ceiling_date(xmax, "month"),by="month")),
+      labels = scales::label_date("%B %y")
+    )
   )+
@@
   geom_rect(aes(xmin=lagtime, xmax=time, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
-  facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
@@
       strip.background = element_blank(),
+      strip.placement="outside",
+      strip.text.y = element_text(angle=0),
 
@@
     NULL
-
-plot_km
 
@@
   geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
+  geom_hline(aes(yintercept=0), colour="black")+
   facet_grid(rows=vars(outcome_descr), scales="free_y")+
@@
       strip.background = element_blank(),
+      strip.text = element_blank(),
 
@@
     NULL
-
-plot_kmrd
 
@@
   #geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrr.ll, ymax=kmrr.ul), alpha=0.1, colour="transparent")+
+  geom_hline(aes(yintercept=0), colour="black")+
   facet_grid(rows=vars(outcome_descr), scales="free_y")+
@@
     NULL
-
-plot_kmrr
 

### 2022-05-05 analysis/km.R
add matching set B

@@ -475,10 +475,10 @@ coxcontrast <- function(data, cuts=NULL){
       !!subgroup_sym,
       period_start,
       period_end,
-      coxhaz = exp(estimate),
-      coxhaz.se = robust.se,
-      coxhaz.ll = exp(estimate + qnorm(0.025)*robust.se),
-      coxhaz.ul = exp(estimate + qnorm(0.975)*robust.se),
+      coxhazr = exp(estimate),
+      coxhr.se = robust.se,
+      coxhr.ll = exp(estimate + qnorm(0.025)*robust.se),
+      coxhr.ul = exp(estimate + qnorm(0.975)*robust.se),
     )
   data_cox
 

### 2022-05-05 analysis/release_objects.R

@@ -10,7 +10,7 @@ library('glue')
 library('survival')
 
 
-for(matchset in c("A")){
+for(matchset in c("A", "B")){
 
   output_dir <- here("output", "release-objects", matchset)
   fs::dir_create(output_dir)
@@ -18,18 +18,17 @@ for(matchset in c("A")){
 
   ## matching ----
 
-  fs::file_copy(here("output", "match", "A", "report", "table1.csv"), here("output", "release-objects", matchset, "match_table1.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "match", "A", "report", "table1by.csv"), here("output", "release-objects", matchset, "match_table1by.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "match", "A", "report", "data_coverage.csv"), here("output", "release-objects", matchset, "match_coverage.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "match", "A", "report", "data_smd.csv"), here("output", "release-objects", matchset, "match_smd.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "match", "A", "report", "flowchart.csv"), here("output", "release-objects", matchset, "match_flowchart.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", matchset, "report", "table1.csv"), here("output", "release-objects", matchset, "match_table1.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", matchset, "report", "table1by.csv"), here("output", "release-objects", matchset, "match_table1by.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", matchset, "report", "data_coverage.csv"), here("output", "release-objects", matchset, "match_coverage.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", matchset, "report", "data_smd.csv"), here("output", "release-objects", matchset, "match_smd.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", matchset, "report", "flowchart.csv"), here("output", "release-objects", matchset, "match_flowchart.csv"), overwrite = TRUE)
 
     ## KM ----
 
-  fs::file_copy(here("output", "match", "A", "km", "combined", "km_estimates.csv"), here("output", "release-objects", matchset, "km_estimates.csv"), overwrite = TRUE)
[…]

### 2022-05-05 create-project.R

@@ -318,12 +318,10 @@ actions_list <- splice(
   # ),
 
 
-  comment("# # # # # # # # # # # # # # # # # # #", "Matching", "# # # # # # # # # # # # # # # # # # #"),
-
-  action_match("A"),
 
   comment("# # # # # # # # # # # # # # # # # # #", "matching set A", "# # # # # # # # # # # # # # # # # # #"),
 
+  action_match("A"),
 
   comment("### Overall models ('all')"),
 
@@ -396,6 +394,81 @@ actions_list <- splice(
   action_km("A", "variantera", "noncoviddeath"),
 
 
+  comment("# # # # # # # # # # # # # # # # # # #", "matching set B", "# # # # # # # # # # # # # # # # # # #"),
+
+  action_match("B"),
+
+  comment("### Overall models ('all')"),
+
+  action_km("B", "all", "postest"),
+  action_km("B", "all", "covidemergency"),
+  action_km("B", "all", "covidadmittedproxy1"),
+  action_km("B", "all", "covidadmitted"),
+  #action_km("B", "all", "noncovidadmitted"),
+  #action_km("B", "all", "covidcc"),
+  action_km("B", "all", "coviddeath"),
+  action_km("B", "all", "noncoviddeath"),
+
+
+  comment("### Models by primary course ('vax12_type')"),
+
+  action_km("B", "vax12_type", "postest"),
+  action_km("B", "vax12_type", "covidemergency"),
+  action_km("B", "vax12_type", "covidadmittedproxy1"),
+  action_km("B", "vax12_type", "covidadmitted"),
+  #action_km("B", "vax12_type", "noncovidadmitted"),
+  #action_km("B", "vax12_type", "covidcc"),
+  action_km("B", "vax12_type", "coviddeath"),
+  action_km("B", "vax12_type", "noncoviddeath"),
+
+
+  comment("### Models by clinically vulnerable group ('cev_cv')"),
+
+  action_km("B", "cev_cv", "postest"),
+  action_km("B", "cev_cv", "covidemergency"),
+  action_km("B", "cev_cv", "covidadmittedproxy1"),
+  action_km("B", "cev_cv", "covidadmitted"),
+  #action_km("B", "cev_cv", "noncovidadmitted"),
+  #action_km("B", "cev_cv", "covidcc"),
+  action_km("B", "cev_cv", "coviddeath"),
+  action_km("B", "cev_cv", "noncoviddeath"),
+
+
+  comment("### Models by prior infection ('prior_covid_infection')"),
+
[…]

### 2022-05-05 lib/design/design.R

@@ -170,15 +170,8 @@ if(exists("matchset")){
       "jcvi_ageband",
       "cev_cv",
       "vax12_type",
-      "stp",
-      "sex",
-      #"ethnicity",
+      "region",
       "imd_Q5",
-
-      "multimorb",
-      "prior_covid_infection",
-      #"immunosuppressed",
-      #"status_hospplanned"
       NULL
     )
 
@@ -200,12 +193,14 @@ if(exists("matchset")){
       "vax3_date",
       "jcvi_ageband",
       "cev_cv",
+      "sex",
       "vax12_type",
       #"vax2_week",
-      "msoa",
-      "sex",
+      "stp",
+      "imd_Q5",
 
-      #"multimorb",
+
+      "multimorb",
       "prior_covid_infection",
       #"immunosuppressed",
       #"status_hospplanned"
@@ -219,7 +214,6 @@ if(exists("matchset")){
     )
 
     all <- c(exact, names(caliper))
-
 
     matching_variables$B = lst(exact, caliper, all)
 

### 2022-05-05 write-up/report.Rmd
Update report.Rmd

@@
 The following outcomes were considered:
+
 * Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date.
@@
 
-We also estimated comparative effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.
+We also estimated comparative effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely .vulnerable or not; evidence of prior SARS-CoV-2 infection or not; dominant SARS-CoV-2 variant era (Delta or Omicron).
 
@@
   )
+
+cat("  \n\n")
 ```
@@
 plot_coverage_cumuln
+
 cat("  \n\n")
@@
   cols_hide(c("variable", "var_label", "order"))
+
+cat("  \n\n")
 ```
@@
     kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
-    kmrr = paste0(label_number(0.01)(kmrr)),
+    irrCI =  paste0(label_number(0.01)(irr), " (", label_number(0.01)(irr.ll), "-", label_number(0.01)(irr.ul), ")"),
+    coxhrCI =  paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(coxhr.ll), "-", label_number(0.01)(coxhr.ul), ")"),
   ) %>%
@@
 
-    kmrdCI = "Risk-difference (95% CI)",
-    kmrr = "Relative-risk"
+    kmrdCI = "Risk difference (95% CI)",
+    irrCI = "Incidence rate ratio (95% CI)",
+    coxhrCI = "Hazard ratio (95% CI)"
   ) %>%
@@
 
+cat("  \n\n")
 ```
@@
   #geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrr.ll, ymax=kmrr.ul), alpha=0.1, colour="transparent")+
-  geom_hline(aes(yintercept=0), colour="black")+
+  geom_hline(aes(yintercept=1), colour="black")+
   facet_grid(rows=vars(outcome_descr), scales="free_y")+
@@
   scale_x_continuous(breaks = seq(0,600,14))+
-  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
+  scale_y_log10(expand = expansion(mult=c(0,0.01)))+
   coord_cartesian(xlim=c(0, NA))+
@@
 plot_combined
+cat("  \n\n")
 

### 2022-05-05 run-local.R
Update run-local.R

@@ -106,6 +106,71 @@ action_km("A", "variantera", "noncoviddeath")
 
 action_km_combine("A")
 
+
+
+
+action_match("B")
+action_km("B", "all", "postest")
+action_km("B", "all", "covidemergency")
+action_km("B", "all", "covidadmittedproxy1")
+action_km("B", "all", "covidadmitted")
+#action_km("B", "all", "noncovidadmitted")
+#action_km("B", "all", "covidcc")
+action_km("B", "all", "coviddeath")
+action_km("B", "all", "noncoviddeath")
+
+
+action_km("B", "vax12_type", "postest")
+action_km("B", "vax12_type", "covidemergency")
+action_km("B", "vax12_type", "covidadmittedproxy1")
+action_km("B", "vax12_type", "covidadmitted")
+#action_km("B", "vax12_type", "noncovidadmitted")
+#action_km("B", "vax12_type", "covidcc")
+action_km("B", "vax12_type", "coviddeath")
+action_km("B", "vax12_type", "noncoviddeath")
+
+
+
+action_km("B", "cev_cv", "postest")
+action_km("B", "cev_cv", "covidemergency")
+action_km("B", "cev_cv", "covidadmittedproxy1")
+action_km("B", "cev_cv", "covidadmitted")
+#action_km("B", "cev_cv", "noncovidadmitted")
+#action_km("B", "cev_cv", "covidcc")
+action_km("B", "cev_cv", "coviddeath")
+action_km("B", "cev_cv", "noncoviddeath")
+
+
+action_km("B", "prior_covid_infection", "postest")
+action_km("B", "prior_covid_infection", "covidemergency")
+action_km("B", "prior_covid_infection", "covidadmittedproxy1")
+action_km("B", "prior_covid_infection", "covidadmitted")
+#action_km("B", "prior_covid_infection", "noncovidadmitted")
+#action_km("B", "prior_covid_infection", "covidcc")
+action_km("B", "prior_covid_infection", "coviddeath")
+action_km("B", "prior_covid_infection", "noncoviddeath")
+
+
+action_km("B", "age65plus", "postest")
+action_km("B", "age65plus", "covidemergency")
+action_km("B", "age65plus", "covidadmittedproxy1")
+action_km("B", "age65plus", "covidadmitted")
+#action_km("B", "age65plus", "noncovidadmitted")
+#action_km("B", "age65plus", "covidcc")
+action_km("B", "age65plus", "coviddeath")
[…]

### 2022-05-06 analysis/match_report.R
Update match_report.R

@@ -462,12 +462,6 @@ write_csv(data_flowchart_match_rounded, fs::path(output_dir, "flowchart.csv"))
 
 
 
-
-
-
-
-
-
 ## matching summary ----
 # FIXME -- need to import baseline variabels here
 

### 2022-05-06 analysis/km_bootstrap.R
add bootstrapping for confidence limits

@@ -1,0 +1,521 @@
+
+# # # # # # # # # # # # # # # # # # # # #
+# Purpose: Get Kaplan-Meier estimates for specified outcome, and derive risk differences
+#  - import matched data
+#  - adds outcome variable and restricts follow-up
+#  - gets KM estimates
+#  - The script must be accompanied by three arguments:
+#    `matchset` - the matching set used for matching
+#    `subgroup` - the subgroup variable, which is used to stratify KM estimates
+#    `outcome` - the dependent variable
+
+# # # # # # # # # # # # # # # # # # # # #
+
+# Preliminaries ----
+
+
+# import command-line arguments ----
+
+args <- commandArgs(trailingOnly=TRUE)
+
+
+if(length(args)==0){
+  # use for interactive testing
+  removeobjects <- FALSE
+  matchset <- "A"
+  subgroup <- "all"
+  #subgroup <- "vax12_type"
+  outcome <- "covidadmitted"
+
+} else {
+  removeobjects <- TRUE
+  matchset <- args[[1]]
+  subgroup <- args[[2]]
+  outcome <- args[[3]]
+
+}
+
+
+## Import libraries ----
+library('tidyverse')
+library('here')
+library('glue')
+library('survival')
+library("furrr")
+
+## Import custom user functions from lib
+source(here("lib", "functions", "utility.R"))
+source(here("lib", "functions", "survival.R"))
+source(here("lib", "functions", "redaction.R"))
+
+## Import design elements
+source(here("lib", "design", "design.R"))
+
+
+# derive subgroup info
+
+subgroup_sym <- sym(subgroup)
+#subgroup_variable <-  str_split_fixed(subgroup,"-",2)[,1]
+#subgroup_level <- str_split_fixed(subgroup,"-",2)[,2]
+#subgroup_dummy <- paste(c(subgroup_variable,subgroup_level), collapse="_")
+
+# create output directories ----
+
+output_dir <- here("output", "match", matchset, "km", subgroup, outcome)
+fs::dir_create(output_dir)
+
+## create special log file ----
+cat(glue("## script info for {outcome} ##"), "  \n", file = fs::path(output_dir, glue("log.txt")), append = FALSE)
+
+## functions to pass additional log info to seperate file
+logoutput <- function(...){
[…]

### 2022-05-06 analysis/km_combine.R

@@ -49,14 +49,23 @@ metaparams <-
     subgroup = factor(recoder$subgroup),
   ) %>%
   mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    outcome_descr = fct_recoderelevel(outcome,  recoder$outcome),
     subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
+  )
+
+
+test <- read_csv(here("output", "match", "A", "km", "vax12_type", "postest", glue("km_estimates.csv")), na="NA", col_types = cols()) %>%
+  add_column(
+    subgroup_level = as.character(.[[factor("vax12_type")]]),
+    subgroup_level_descr = fct_recoderelevel((.[["vax12_type"]]), recoder[["vax12_type"]]),
+    .before=1
   )
 
 
 km_estimates <- metaparams %>%
   mutate(
     data = pmap(list(matchset, subgroup, outcome), function(matchset, subgroup, outcome) {
+      subgroup <- as.character(subgroup)
       dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("km_estimates.csv")), na="NA", col_types = cols())
       dat %>%
       add_column(
@@ -75,14 +84,15 @@ write_csv(km_estimates, fs::path(output_dir, "km_estimates.csv"))
 contrasts_daily <- metaparams %>%
   mutate(
     data = pmap(list(matchset, outcome, subgroup), function(matchset, outcome, subgroup){
-        dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("contrasts_daily.csv")), na="NA", col_types = cols())
-        dat %>%
-          add_column(
-            subgroup_level = as.character(.[[subgroup]]),
-            subgroup_level_descr = fct_recoderelevel(.[[subgroup]], recoder[[subgroup]]),
-            .before=1
-          ) %>%
-          select(-all_of(subgroup))
+      subgroup <- as.character(subgroup)
+      dat <- read_csv(here("output", "match", matchset, "km", subgroup, outcome, glue("contrasts_daily.csv")), na="NA", col_types = cols())
+      dat %>%
+        add_column(
+          subgroup_level = as.character(.[[subgroup]]),
+          subgroup_level_descr = fct_recoderelevel(.[[subgroup]], recoder[[subgroup]]),
+          .before=1
+        ) %>%
+        select(-all_of(subgroup))
       }
     )
   ) %>%
[…]

### 2022-05-06 analysis/match.R

@@ -31,11 +31,7 @@ library('MatchIt')
 
 ## Import custom user functions from lib
 source(here("lib", "functions", "utility.R"))
-source(here("lib", "functions", "survival.R"))
 source(here("lib", "functions", "redaction.R"))
-
-## Import custom user functions from lib
-source(here("lib", "functions", "utility.R"))
 
 ## Import design elements
 source(here("lib", "design", "design.R"))
@@ -90,8 +86,10 @@ data_matchingcandidates <-
   arrange(patient_id)
 
 
+# create function that catches errors in case no matches are found within a thread
+safely_matchit <- purrr::safely(matchit)
 
-## manage parallelisation
+## parallelisation preliminaries ----
 
 library("doParallel")
 
@@ -107,15 +105,12 @@ print(cluster)
 doParallel::registerDoParallel(cl = cluster)
 
 
-# create function that catches errors in case no matches are found within a thread
-safely_matchit <- purrr::safely(matchit)
-
-
-
 # create parallel matching streams
 matchthreads <- unique(as.character(data_matchingcandidates$thread_variable))
+
 table(data_matchingcandidates$thread_variable, useNA="ifany")
 
+## match in parallel ----
 data_matchstatus <-
   foreach(
     matchthread = matchthreads,
@@ -151,7 +146,7 @@ data_matchstatus <-
           patient_id = data_thread$patient_id,
           matched = FALSE,
           thread_id = data_thread$thread_id,
-          match_id = factor(NA),
+          threadmatch_id = NA_integer_,
           treatment = data_thread$treatment,
           weight = 0,
           vax3_date = data_thread$vax3_date
@@ -163,7 +158,7 @@ data_matchstatus <-
           patient_id = data_thread$patient_id,
           matched = !is.na(obj_matchit$subclass),
           thread_id = data_thread$thread_id,
-          match_id = obj_matchit$subclass,
+          threadmatch_id = as.integer(as.character(obj_matchit$subclass)),
           treatment = obj_matchit$treat,
           weight = obj_matchit$weights,
           .before = 1
@@ -175,6 +170,11 @@ data_matchstatus <-
 
 parallel::stopCluster(cl = cluster)
 
+data_matchstatus <- data_matchstatus %>%
+  arrange(threadmatch_id, thread_id) %>%
+  mutate(
[…]

### 2022-05-06 create-project.R

@@ -97,18 +97,35 @@ action_km <- function(
   matchset, subgroup, outcome
 ){
 
-  action(
-    name = glue("km_{matchset}_{subgroup}_{outcome}"),
-    run = glue("r:latest analysis/km.R"),
-    arguments = c(matchset, subgroup, outcome),
-    needs = list(
-      glue("match_{matchset}"),
-      "data_selection"
-    ),
-    moderately_sensitive = lst(
-      txt = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.txt"),
-      csv = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.csv"),
-      png = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.png"),
+  splice(
+    # action(
+    #   name = glue("km_{matchset}_{subgroup}_{outcome}"),
+    #   run = glue("r:latest analysis/km.R"),
+    #   arguments = c(matchset, subgroup, outcome),
+    #   needs = list(
+    #     glue("match_{matchset}"),
+    #     "data_selection"
+    #   ),
+    #   moderately_sensitive = lst(
+    #     txt = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.txt"),
+    #     csv = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.csv"),
+    #     png = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.png"),
+    #   )
+    # ),
+
+    action(
+      name = glue("km_{matchset}_{subgroup}_{outcome}"),
+      run = glue("r:latest analysis/km_bootstrap.R"),
+      arguments = c(matchset, subgroup, outcome),
+      needs = list(
+        glue("match_{matchset}"),
+        "data_selection"
+      ),
+      moderately_sensitive = lst(
+        txt = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.txt"),
+        csv = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.csv"),
+        png = glue("output/match/{matchset}/km/{subgroup}/{outcome}/*.png"),
+      )
     )
   )
 }

### 2022-05-06 lib/design/design.R

@@ -98,8 +98,8 @@ recoder <-
       `ChAdOx1-S` = "az-az"
     ),
     age65plus = c(
-      `Aged 18-64` = "FALSE",
-      `Aged 65 and over` = "TRUE"
+      `18-64` = "FALSE",
+      `65 and over` = "TRUE"
     ),
     cev_cv = c(
       "Clinically extremely vulnerable" = "Clinically extremely vulnerable",

### 2022-05-06 run-local.R

@@ -32,7 +32,7 @@ action_match <- compose_action(
 )
 
 action_km <- compose_action(
-    here("analysis", "km.R")
+    here("analysis", "km_bootstrap.R")
 )
 
 action_km_combine <- compose_action(

### 2022-05-06 write-up/report.Rmd
Update report.Rmd

@@
     locations = cells_body(columns = "criteria")
+  ) %>%
+  tab_options(
+    table.font.size = 8
   )
@@
 
-\newpage 
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
 #### Figure 1: Cumulative matching success
@@
 
-\newpage
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
 
@@
   ) %>%
-  cols_hide(c("variable", "var_label", "order"))
+  cols_hide(c("variable", "var_label", "order")) %>%
+  tab_options(
+    table.font.size = 8
+  )
 
@@
 ```
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
 
@@
 
-#### Table 3: Kaplan-Meier cumulative incidence estimates
+#### Table 3: Kaplan-Meier 10-week risk estimates
 
@@
     n.event_1,
-    riskCI_0 = paste0(label_number(0.01)(risk_0), " (", label_number(0.01)(risk.ll_0), "-", label_number(0.01)(risk.ul_0), ")"),
-    riskCI_1 = paste0(label_number(0.01)(risk_1), " (", label_number(0.01)(risk.ll_1), "-", label_number(0.01)(risk.ul_1), ")"),
+    riskCI_0 = paste0(label_number(0.01)(1-surv_0), " (", label_number(0.01)(1-surv.ul_0), "-", label_number(0.01)(1-surv.ll_0), ")"),
+    riskCI_1 = paste0(label_number(0.01)(1-surv_1), " (", label_number(0.01)(1-surv.ul_1), "-", label_number(0.01)(1-surv.ll_1), ")"),
     kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
+    kmrrCI = paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(kmrr.ll), "-", label_number(0.01)(kmrr.ul), ")"),
     irrCI =  paste0(label_number(0.01)(irr), " (", label_number(0.01)(irr.ll), "-", label_number(0.01)(irr.ul), ")"),
-    coxhrCI =  paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(coxhr.ll), "-", label_number(0.01)(coxhr.ul), ")"),
+    #coxhrCI =  paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(coxhr.ll), "-", label_number(0.01)(coxhr.ul), ")"),
   ) %>%
+  group_by(outcome_descr) %>%
+  mutate(
+    # add indicator for row shading
+    shade = cumsum((lag(subgroup_descr, 1, "")!=subgroup_descr)*1L),
+    # remove repeat values of subgroup_level
+    subgroup_descr = if_else(lag(subgroup_descr)==subgroup_descr & row_number()!=1, "", as.character(subgroup_descr)),
+  ) %>%
+  ungroup() %>%
   gt(groupname_col="outcome_descr") %>%
@@
     outcome_descr = "Outcome",
-    subgroup_descr = "Subgroup",
-    subgroup_level_descr = "Subgroup",
+    subgroup_descr = "Sub-group",
+    subgroup_level_descr = "",
 
@@
 
-    riskCI_0 = "KM risk (95% CI)",
-    riskCI_1 = "KM risk (95% CI)",
+    riskCI_0 = "Risk (95% CI)",
+    riskCI_1 = "Risk (95% CI)",
 
     kmrdCI = "Risk difference (95% CI)",
+    kmrrCI = "Risk ratio (95% CI)",
     irrCI = "Incidence rate ratio (95% CI)",
-    coxhrCI = "Hazard ratio (95% CI)"
+    #coxhrCI = "Hazard ratio (95% CI)"
   ) %>%
@@
     label = "Comparison",
-    columns = starts_with("km")
+    columns = c("kmrrCI", "irrCI")
+  ) %>%
+  cols_align(
+    align = c("right"),
+    columns = ends_with(c("_1", "_0", "CI"))
+  ) %>%
+  tab_options(
+    table.font.size = 8
+  ) %>%
+  tab_style(
+    style = list(
+      cell_fill(color = "gray96")
+    ),
+    locations = cells_body(
+      rows = shade%%2 != 0
     )
+  ) %>%
+  cols_hide(c("shade"))
 
@@
   geom_rect(aes(xmin=lagtime, xmax=time, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
+  geom_hline(aes(yintercept=0), colour="black")+
   facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
@@
   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-  scale_x_continuous(breaks = seq(0,600,14))+
-  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
+  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+  scale_y_continuous(expand = expansion(mult=c(0,0.01)), )+
   coord_cartesian(xlim=c(0, NA))+
@@
 
-plot_kmrd <- contrasts_daily %>%
-  filter(
-    subgroup=="all",
-    outcome %in% outcomes
-  ) %>% 
-  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
-  geom_step(aes(x=period_start, y=kmrd))+
-  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
-  geom_hline(aes(yintercept=0), colour="black")+
-  facet_grid(rows=vars(outcome_descr), scales="free_y")+
-  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
-  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-  scale_x_continuous(breaks = seq(0,600,14))+
-  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
-  coord_cartesian(xlim=c(0, NA))+
-  labs(
-    x="Days",
-    y=NULL,
-    title="Cumulative risk difference",
-    colour=NULL
-  )+
-  theme_minimal(base_size = 9)+
-    theme(
-      legend.position="none",
-      legend.direction = "vertical",
-      axis.text.x.top=element_text(hjust=0),
-
-      panel.border = element_blank(),
-      axis.line.y = element_line(colour = "black"),
-      axis.line.x.top = element_line(colour = "black"),
-
-      panel.grid.minor.x = element_blank(),
-      panel.grid.minor.y = element_blank(),
-      strip.background = element_blank(),
-      strip.text = element_blank(),
-
-      panel.spacing = unit(0.8, "lines"),
-
-    )+
-    NULL
+# 
+# plot_kmrd <- contrasts_daily %>%
+#   filter(
+#     subgroup=="all",
+#     outcome %in% outcomes
+#   ) %>% 
+#   ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+#   geom_step(aes(x=period_start, y=kmrd))+
+#   geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
+#   geom_hline(aes(yintercept=0), colour="black")+
+#   facet_grid(rows=vars(outcome_descr), scales="free_y")+
+#   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+#   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+#   scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+#   scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
+#   coord_cartesian(xlim=c(0, NA))+
+#   labs(
+#     x="Days",
+#     y=NULL,
+#     title="Cumulative risk difference",
+#     colour=NULL
+#   )+
+#   theme_minimal(base_size = 9)+
+#     theme(
+#       legend.position="none",
+#       legend.direction = "vertical",
+#       axis.text.x.top=element_text(hjust=0),
+# 
+#       panel.border = element_blank(),
+#       axis.line.y = element_line(colour = "black"),
+#       axis.line.x.top = element_line(colour = "black"),
+# 
+#       panel.grid.minor.x = element_blank(),
+#       panel.grid.minor.y = element_blank(),
+#       strip.background = element_blank(),
+#       strip.text = element_blank(),
+# 
+# 
+#       panel.spacing = unit(0.8, "lines"),
+# 
+#     )+
+#     NULL
+# 
+#
 
@@
   geom_step(aes(x=period_start, y=kmrr))+
-  #geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrr.ll, ymax=kmrr.ul), alpha=0.1, colour="transparent")+
+  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrr.ll, ymax=kmrr.ul), alpha=0.1, colour="transparent")+
   geom_hline(aes(yintercept=1), colour="black")+
@@
   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-  scale_x_continuous(breaks = seq(0,600,14))+
+  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
   scale_y_log10(expand = expansion(mult=c(0,0.01)))+
@@
 
-  plot_combined <- patchwork::wrap_plots(plot_km, plot_kmrd, plot_kmrr, ncol=3)
+  plot_combined <- patchwork::wrap_plots(
+    plot_km, 
+    #plot_kmrd, 
+    plot_kmrr, 
+    ncol=3
+  )
   ggsave(
@@
 
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+\newpage 
 # Supplement

### 2022-05-09 analysis/km_combine.R
clean up script

@@ -33,8 +33,6 @@ library('survival')
 
 ## Import custom user functions from lib
 source(here("lib", "functions", "utility.R"))
-source(here("lib", "functions", "survival.R"))
-source(here("lib", "functions", "redaction.R"))
 
 ## Import design elements
 source(here("lib", "design", "design.R"))
@@ -51,14 +49,6 @@ metaparams <-
   mutate(
     outcome_descr = fct_recoderelevel(outcome,  recoder$outcome),
     subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-  )
-
-
-test <- read_csv(here("output", "match", "A", "km", "vax12_type", "postest", glue("km_estimates.csv")), na="NA", col_types = cols()) %>%
-  add_column(
-    subgroup_level = as.character(.[[factor("vax12_type")]]),
-    subgroup_level_descr = fct_recoderelevel((.[["vax12_type"]]), recoder[["vax12_type"]]),
-    .before=1
   )
 
 

### 2022-05-09 analysis/km_combine.R
Update km_combine.R

@@ -1,6 +1,6 @@
 
 # # # # # # # # # # # # # # # # # # # # #
-# Purpose: Combine km estimates for different outcomes
+# Purpose: Combine km estimates from different outcomes
 #  - import matched data
 #  - adds outcome variable and restricts follow-up
 #  - gets KM estimates

### 2022-05-09 analysis/km_bootstrap.R
redo bootstrap to provide option of rounding survival curves prior to contrast calculations

@@ -146,7 +146,6 @@ threshold <- 7
 
 ## get bootstrap samples ----
 
-
 quantile_bs <- partial(quantile, na.rm = TRUE, names=FALSE)
 
 boot_n <-
@@ -156,45 +155,37 @@ boot_n <-
     500L
   }
 
-boot_samples <- read_rds(here("output", "match", matchset, "boot_samples.rds")) %>%
-  filter(boot_id <= boot_n)
+# look-up for bootstrap samples ----
+# boot_id = 0 is the unsampled (ie, sample everyone once) id list
 
-## kaplan meier cumulative risk differences ----
-data_surv0 <-
-  data_matched %>%
-  group_by(!!subgroup_sym, treatment) %>%
-  nest() %>%
-  mutate(
-    n_events = map_int(data, ~sum(.x$ind_outcome, na.rm=TRUE)),
-    surv_obj = map(data, ~{
-      survfit(Surv(tte_outcome, ind_outcome) ~ 1, data = .x, conf.type="log-log")
-    }),
-    surv_obj_tidy = map(surv_obj, ~tidy_surv(.x, times=seq_len(maxfup))), # return survival table for each day of follow up
-  ) %>%
-  select(!!subgroup_sym, treatment, n_events, surv_obj_tidy) %>%
-  unnest(surv_obj_tidy) %>%
-  mutate(
-    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment)
+boot_samples <- read_rds(here("output", "match", matchset, "boot_samples.rds")) %>%
+  filter(boot_id <= boot_n) %>%
+  bind_rows(
+    data_matched %>% transmute(boot_id=0L, match_id),
+    .
   )
 
+boot_ids <- unique(boot_samples$boot_id)
 
-## parallelisation preliminaries ----
+## calculate confidence limits with boot strapping ----
+
+### parallelisation preliminaries ----
 
 library("doParallel")
 
-parallel::detectCores() # how many cores available?
+# how many cores available?
+#parallel::detectCores()
+
 n_threads <- 8
 
 cluster <- parallel::makeCluster(
   n_threads,
   type = "PSOCK" # this should work across multi-core windows or linux machines
 )
-print(cluster)
+#print(cluster)
+
 #register it to be used by %dopar%
 doParallel::registerDoParallel(cl = cluster)
-
-
-boot_ids <- unique(as.character(boot_samples$boot_id))
 
 #### survival table in parallel ----
[…]

### 2022-05-09 write-up/report.Rmd
Update report.Rmd

@@
 
+Counts and survival estimates are based on values rounded up to the nearest 7, for disclosure control.
+
 ```{r table-km, echo=FALSE, warning=FALSE, message=FALSE}
@@
     n.event_1,
-    riskCI_0 = paste0(label_number(0.01)(1-surv_0), " (", label_number(0.01)(1-surv.ul_0), "-", label_number(0.01)(1-surv.ll_0), ")"),
-    riskCI_1 = paste0(label_number(0.01)(1-surv_1), " (", label_number(0.01)(1-surv.ul_1), "-", label_number(0.01)(1-surv.ll_1), ")"),
+    riskCI_0 = paste0(label_number(0.01)(risk_0), " (", label_number(0.01)(risk.ll_0), "-", label_number(0.01)(risk.ul_0), ")"),
+    riskCI_1 = paste0(label_number(0.01)(risk_1), " (", label_number(0.01)(risk.ll_1), "-", label_number(0.01)(risk.ul_1), ")"),
     kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
@@
 
-    riskCI_0 = "Risk (95% CI)",
-    riskCI_1 = "Risk (95% CI)",
+    riskCI_0 = "10-week Risk (95% CI)",
+    riskCI_1 = "10-week Risk (95% CI)",
 
@@
     label = "Comparison",
-    columns = c("kmrrCI", "irrCI")
+    columns = ends_with("CI")
   ) %>%
@@
 
+Counts and survival estimates are based on values rounded up to the nearest 7, for disclosure control.
+
 ```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
@@
     y=NULL,
-    title="Cumulative incidence",
+    title="Risk",
     colour=NULL
@@
       strip.placement="outside",
-      strip.text.y = element_text(angle=0),
+      strip.text.x = element_text(angle=0),
 
@@
 
-# 
-# plot_kmrd <- contrasts_daily %>%
-#   filter(
-#     subgroup=="all",
-#     outcome %in% outcomes
-#   ) %>% 
-#   ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
-#   geom_step(aes(x=period_start, y=kmrd))+
-#   geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
-#   geom_hline(aes(yintercept=0), colour="black")+
-#   facet_grid(rows=vars(outcome_descr), scales="free_y")+
-#   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
-#   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-#   scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
-#   scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
-#   coord_cartesian(xlim=c(0, NA))+
-#   labs(
-#     x="Days",
-#     y=NULL,
-#     title="Cumulative risk difference",
-#     colour=NULL
-#   )+
-#   theme_minimal(base_size = 9)+
-#     theme(
-#       legend.position="none",
-#       legend.direction = "vertical",
-#       axis.text.x.top=element_text(hjust=0),
-# 
-#       panel.border = element_blank(),
-#       axis.line.y = element_line(colour = "black"),
-#       axis.line.x.top = element_line(colour = "black"),
-# 
-#       panel.grid.minor.x = element_blank(),
-#       panel.grid.minor.y = element_blank(),
-#       strip.background = element_blank(),
-#       strip.text = element_blank(),
-# 
-# 
-#       panel.spacing = unit(0.8, "lines"),
-# 
-#     )+
-#     NULL
-# 
-#
+plot_kmrd <- contrasts_daily %>%
+  filter(
+    subgroup=="all",
+    outcome %in% outcomes
+  ) %>%
+  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+  geom_step(aes(x=period_start, y=kmrd))+
+  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
+  geom_hline(aes(yintercept=0), colour="black")+
+  facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
+  coord_cartesian(xlim=c(0, NA))+
+  labs(
+    x="Days",
+    y=NULL,
+    title="Risk difference",
+    colour=NULL
+  )+
+  theme_minimal(base_size = 9)+
+    theme(
+      legend.position="none",
+      legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+      strip.text = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
 
@@
     y=NULL,
-    title="Cumulative risk ratio",
+    title="Risk ratio",
+    colour=NULL
+  )+
+  theme_minimal(base_size = 9)+
+    theme(
+      legend.position="none",
+      legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+      strip.text = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
+
+plot_irr <- contrasts_daily %>%
+  # equivalent to the hazard
+  filter(
+    subgroup=="all",
+    outcome %in% outcomes
+  ) %>%
+  ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+  geom_step(aes(x=period_start, y=irr))+
+  geom_rect(aes(xmin=period_start, xmax=period_end, ymin=irr.ll, ymax=irr.ul), alpha=0.1, colour="transparent")+
+  geom_hline(aes(yintercept=0), colour="black")+
+  facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+  scale_y_log10(expand = expansion(mult=c(0,0.01)))+
+  coord_cartesian(xlim=c(0, NA))+
+  labs(
+    x="Days",
+    y=NULL,
+    title="Hazard ratio",
     colour=NULL
@@
   plot_km, 
-    #plot_kmrd, 
+  plot_kmrd, 
   plot_kmrr, 
-    ncol=3
+  #plot_irr,
+  ncol=4
 )
+
 ggsave(

### 2022-05-09 write-up/report.Rmd
Update report.Rmd

@@
     n.event_1,
-    riskCI_0 = paste0(label_number(0.01)(risk_0), " (", label_number(0.01)(risk.ll_0), "-", label_number(0.01)(risk.ul_0), ")"),
-    riskCI_1 = paste0(label_number(0.01)(risk_1), " (", label_number(0.01)(risk.ll_1), "-", label_number(0.01)(risk.ul_1), ")"),
+    riskCI_0 = paste0(label_number(0.01)(1-surv_0), " (", label_number(0.01)(1-surv.ul_0), "-", label_number(0.01)(1-surv.ll_0), ")"),
+    riskCI_1 = paste0(label_number(0.01)(1-surv_1), " (", label_number(0.01)(1-surv.ul_1), "-", label_number(0.01)(1-surv.ll_1), ")"),
     kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),

### 2022-05-09 create-project.R
add B actions to release action

@@ -510,7 +510,8 @@ actions_list <- splice(
       "data_selection",
       "match_report_A",
       "match_report_B",
-      "km_combine_A"
+      "km_combine_A",
+      "km_combine_B"
     ),
     moderately_sensitive = lst(
       txt = "output/files-for-release.txt",

### 2022-05-17 protocol/protocol.Rmd
update protocol / report

@@
   * Middle Layer Super Output Area (MSOA)
-  * NHS Strategic Health Partnership (STP)
+  * NHS Sustainability and Transformation Partnership (STP)
   * NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West)
@@
 We estimated Kaplan-Meier (KM) cumulative incidence curves separately by vaccine brand, and compared brands using the 10-week risk differences and risk ratios.
+The incidence rate ratio (events per person-time-at-risk) in time-period specific intervals will also be reported.
+This is equivalent to a period-specific hazard ratio if a constant hazard within each time period is assumed.
 
-Confidence limits for the KM estimate will be derived from complimentary log-log transformation.
-Confidence limits for the risk difference will be derived from the square root of the sum of squares the treatment-specific KM standard errors. The KM standard errors will use Greenwood's formula.
+95% confidence limits for the KM estimates, risk differences, risk ratios, and incidence rate ratios are derived by bootstrapping the matched pairs. 500 bootstrap replicates will be used.
+
+<!-- Alternatively, analytic confidence limits can be derived as follows: Confidence limits for the KM estimate can be derived from complementary log-log transformation.
+Confidence limits for the risk difference can be derived from the square root of the sum of squares the treatment-specific KM standard errors.
+The KM standard errors can use Greenwood’s formula.
 Confidence limits for the survival ratio are straight-forward, and can be based on the standard error of the cumulative hazard.
-But confidence limits for the risk ratio (or relative risk) are more complicated.
-TBD.
-
-Alternatively, we could bootstrap.
+Confidence limits for the risk ratio (or relative risk) are TBD.
+-->
 

### 2022-05-17 write-up/report.Rmd
update protocol / report

--- 2022-05-06 write-up/report.Rmd
@@
 # where are the outputs (ie the inputs for this manuscript!) saved?
-output_dir_os <- here("output", "release-objects")
-#output_dir_os <- here("released_output", "release-objects")
+#output_dir_os <- here("output", "release-objects")
+output_dir_os <- here("released-output", "release-objects")
 
@@
 
-matchset <- "A"
-
-match_flowchart <-  read_csv(here("output", "release-objects", matchset, "match_flowchart.csv"))
-match_coverage <- read_csv(here("output", "release-objects", matchset, "match_coverage.csv")) %>%
-   mutate(
-    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
-  )
-
-match_table1 <- read_csv(here("output", "release-objects", matchset, "match_table1.csv"))
-match_table1by <- read_csv(here("output", "release-objects", matchset, "match_table1by.csv"))
-match_smd <-  read_csv(here("output", "release-objects", matchset, "match_smd.csv"))
-
-km_estimates <- read_csv(here("output", "release-objects", matchset, "km_estimates.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-contrasts_daily <- read_csv(here("output", "release-objects", matchset, "contrasts_daily.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-contrasts_overall <- read_csv(here("output", "release-objects", matchset, "contrasts_overall.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-
   outcomes <- 
@@
       "covidadmitted",
+      "covidadmittedproxy1",
       #"covidcc",
@@
 
-BNT162b2 booster recipients were matched with Moderna booster recipients on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation.
-These matching criteria ensure that treatment groups are also matched within JCVI priority group.
+BNT162b2 booster recipients were matched with Moderna booster recipients using two different matching strategies.
+one simpler strategy that will identify more matches but the risk of greater imbalance between treatment groups, and another more complex strategy that will identify fewer matches but with less imbalance.
+
+#### Matching strategy A
+
+Booster recipients were matched on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years and age groups used by JCVI for prioritisation; English Index of Multiple Deprivation, grouped by quintile.
+
+#### Matching strategy B
+
+Booster recipients were matched on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; Sustainability and Transformation Partnership (STP) a geographical grouping of NHS and local authorities; evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation; English Index of Multiple Deprivation, grouped by quintile; sex; and morbidity count (0, 1, 2 or more).
+
+Morbidity count included the following conditions: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease.
 
@@
 
-<!--#
-
 ## Follow-up
@@
 
-We estimated Kaplan-Meier cumulative incidence curves separately by vaccine brand, and compared groups using the 10-week risk differences and risk ratios.
+We estimated Kaplan-Meier (KM) cumulative incidence curves separately by vaccine brand, and compared brands using the 10-week risk differences and risk ratios.
+The incidence rate ratio (events per person-time-at-risk) in time-period specific intervals will also be reported.
+This is equivalent to a period-specific hazard ratio if a constant hazard within each time period is assumed.
+
+<!-- 95% confidence limits for the KM estimates, risk differences, risk ratios, and incidence rate ratios are derived by bootstrapping the matched pairs. 500 bootstrap replicates will be used. -->
+
+95% confidence limits for the KM estimates were derived from the complementary log-log transformation.
+Confidence limits for the risk difference were derived from the sum of squares of the treatment-specific KM standard errors, derived from Greenwood’s formula.
+Confidence limits for the survival ratio were based on the standard error of the cumulative hazard.
+Confidence limits for the risk ratio were derived using the Delta method.
 
@@
 Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.
-Any reported figures based on counts below 6 are redacted or rounded for disclosure control.
 
-This study will follow the STROBE-RECORD reporting guidelines.
+## Disclosure control
+Any reported figures based on counts below 6 must be redacted or rounded for disclosure control.
+To reduce secondary disclosure risk (e.g., by using two unredacted values to derive a redacted value), rounding is preferred.
+
+In this study, we round counts up to the nearest multiple of 7.
+Kaplan-Meier survival estimates are rounded upwards by deferring event-times until there are at least 7 events, so that each KM “step” is based on at least 7 events.
+All other quantities will be derived from these rounded KM estimates.
 
 # Results
+
+```{r match-A, echo=FALSE, warning=FALSE, message=FALSE}
+
+matchset <- "A"
+
+match_flowchart <-  read_csv(fs::path(output_dir_os, matchset, "match_flowchart.csv"))
+match_coverage <- read_csv(fs::path(output_dir_os, matchset, "match_coverage.csv")) %>%
+   mutate(
+    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
+  )
+
+match_table1 <- read_csv(fs::path(output_dir_os, matchset, "match_table1.csv"))
+match_table1by <- read_csv(fs::path(output_dir_os, matchset, "match_table1by.csv"))
+match_smd <-  read_csv(fs::path(output_dir_os, matchset, "match_smd.csv"))
+
+km_estimates <- read_csv(fs::path(output_dir_os, matchset, "km_estimates.csv")) %>% 
+  mutate(
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
+    subgroup_level_descr = replace_na(subgroup_level_descr, "")
+  )
+contrasts_daily <- read_csv(fs::path(output_dir_os, matchset, "contrasts_daily.csv")) %>% 
+  mutate(
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
+    subgroup_level_descr = replace_na(subgroup_level_descr, "")
+  )
+contrasts_overall <- read_csv(fs::path(output_dir_os, matchset, "contrasts_overall.csv")) %>% 
+  mutate(
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
+    subgroup_level_descr = replace_na(subgroup_level_descr, "")
+  )
+
+```
 
@@
 `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type=="moderna") %>% pull(pct_all))`) eligible for matching.
-`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
+
+## Matching strategy A
+
+For the simpler matching criteria, `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
 
@@
     n.event_1,
-    riskCI_0 = paste0(label_number(0.01)(1-surv_0), " (", label_number(0.01)(1-surv.ul_0), "-", label_number(0.01)(1-surv.ll_0), ")"),
-    riskCI_1 = paste0(label_number(0.01)(1-surv_1), " (", label_number(0.01)(1-surv.ul_1), "-", label_number(0.01)(1-surv.ll_1), ")"),
-    kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
+    riskCI_0 = paste0(label_number(0.1, 10000)(1-surv_0), " (", label_number(0.1, 10000)(1-surv.ul_0), "-", label_number(0.1, 10000)(1-surv.ll_0), ")"),
+    riskCI_1 = paste0(label_number(0.1, 10000)(1-surv_1), " (", label_number(0.1, 10000)(1-surv.ul_1), "-", label_number(0.1, 10000)(1-surv.ll_1), ")"),
+    kmrdCI = paste0(label_number(0.1, 10000)(kmrd), " (", label_number(0.1, 10000)(kmrd.ll), "-", label_number(1, 10000)(kmrd.ul), ")"),
     kmrrCI = paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(kmrr.ll), "-", label_number(0.01)(kmrr.ul), ")"),
@@
 
-    riskCI_0 = "Risk (95% CI)",
-    riskCI_1 = "Risk (95% CI)",
+    riskCI_0 = "10-week Risk per 10,000 people (95% CI)",
+    riskCI_1 = "10-week Risk per 10,000 people (95% CI)",
 
-    kmrdCI = "Risk difference (95% CI)",
+    kmrdCI = "Risk difference per 10,000 people (95% CI)",
     kmrrCI = "Risk ratio (95% CI)",
@@
    tab_spanner(
-    label = "Comparison",
-    columns = c("kmrrCI", "irrCI")
+    label = "Comparison (reference=BNT162b2)",
+    columns = ends_with("CI")
   ) %>%
@@
 
-#### Figure 3: Kaplan-Meier cumulative incidence estimates
+```{r, function-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
 
-```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
+km_theme <- function(...) {
+  theme(
+      legend.position="none",
+      axis.text.x.top=element_text(hjust=0),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+      strip.text = element_blank(),
+
+      panel.border = element_blank(),
+      panel.spacing = unit(0.8, "lines"),
+      ...
+    )
+}
+
+colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
+
+plot_km_subgroup <- function(subgroup){
+
+  subgroup0 <- subgroup
+
+  if(subgroup=="all"){
+    colour_guide <- list(guides(colour="none"))
+  } else{
+    colour_guide <- NULL
+  }
+
   plot_km <- km_estimates %>%
     filter(
-    subgroup=="all",
+      subgroup==subgroup0,
       outcome %in% outcomes
     ) %>%
-  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
+    mutate(
+      colgroup = interaction(treatment_descr, subgroup_level_descr, sep=", "),
+      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10)
+    ) %>%
+    ggplot(aes(group=colgroup, colour=subgroup_level_descr, fill=subgroup_level_descr, alpha=treatment_descr)) +
     geom_step(aes(x=time, y=1-surv))+
@@
     scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+    #scale_color_manual(values=colourvalues, na.value="grey") +
+    #scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
     scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
-  scale_y_continuous(expand = expansion(mult=c(0,0.01)), )+
+    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, 10000))+
+    scale_alpha_discrete(range=c(0.3,0.9))+
     coord_cartesian(xlim=c(0, NA))+
@@
       y=NULL,
-    title="Cumulative incidence",
-    colour=NULL
+      title="Risk per 10,000",
+      colour=NULL,
+      alpha=NULL
     )+
+    colour_guide+
     theme_minimal(base_size = 9)+
       theme(
-      legend.position=c(0.05,0.1),
+        legend.position = c(0.05,0.1),
         legend.justification = c(0,0),
@@
         axis.text.x.top=element_text(hjust=0),
-
-      panel.border = element_blank(),
-      axis.line.y = element_line(colour = "black"),
-      axis.line.x.top = element_line(colour = "black"),
 
@@
         strip.placement="outside",
-      strip.text.y = element_text(angle=0),
+        strip.text.y.left = element_text(angle=0),
 
+        panel.border = element_blank(),
         panel.spacing = unit(0.8, "lines"),
@@
 
-# 
-# plot_kmrd <- contrasts_daily %>%
-#   filter(
-#     subgroup=="all",
-#     outcome %in% outcomes
-#   ) %>% 
-#   ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
-#   geom_step(aes(x=period_start, y=kmrd))+
-#   geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
-#   geom_hline(aes(yintercept=0), colour="black")+
-#   facet_grid(rows=vars(outcome_descr), scales="free_y")+
-#   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
-#   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-#   scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
-#   scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
-#   coord_cartesian(xlim=c(0, NA))+
-#   labs(
-#     x="Days",
-#     y=NULL,
-#     title="Cumulative risk difference",
-#     colour=NULL
-#   )+
-#   theme_minimal(base_size = 9)+
-#     theme(
-#       legend.position="none",
-#       legend.direction = "vertical",
-#       axis.text.x.top=element_text(hjust=0),
-# 
-#       panel.border = element_blank(),
-#       axis.line.y = element_line(colour = "black"),
-#       axis.line.x.top = element_line(colour = "black"),
-# 
-#       panel.grid.minor.x = element_blank(),
-#       panel.grid.minor.y = element_blank(),
-#       strip.background = element_blank(),
-#       strip.text = element_blank(),
-# 
-# 
-#       panel.spacing = unit(0.8, "lines"),
-# 
-#     )+
-#     NULL
-# 
-#
+  plot_kmrd <- contrasts_daily %>%
+    filter(
+      subgroup==subgroup0,
+      outcome %in% outcomes
+    ) %>%
+    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+    geom_step(aes(x=period_start, y=kmrd))+
+    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
+    geom_hline(aes(yintercept=0), colour="black")+
+    facet_grid(rows=vars(outcome_descr), scales="free_y")+
+    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, 10000))+
+    coord_cartesian(xlim=c(0, NA))+
+    labs(
+      x="Days",
+      y=NULL,
+      title="Risk difference",
+      colour=NULL
+    )+
+    theme_minimal(base_size = 9)+
+    km_theme()+
+    NULL
 
@@
     filter(
-    subgroup=="all",
+      subgroup==subgroup0,
       outcome %in% outcomes
@@
       y=NULL,
-    title="Cumulative risk ratio",
+      title="Risk ratio",
       colour=NULL
@@
     theme_minimal(base_size = 9)+
-    theme(
-      legend.position="none",
-      legend.direction = "vertical",
-      axis.text.x.top=element_text(hjust=0),
+    km_theme()+
+    NULL
 
-      panel.border = element_blank(),
-      axis.line.y = element_line(colour = "black"),
-      axis.line.x.top = element_line(colour = "black"),
-
-      panel.grid.minor.x = element_blank(),
-      panel.grid.minor.y = element_blank(),
-      strip.background = element_blank(),
-      strip.text = element_blank(),
-
-      panel.spacing = unit(0.8, "lines"),
-
+  plot_irr <- contrasts_daily %>%
+    # equivalent to the hazard
+    filter(
+      subgroup==subgroup0,
+      outcome %in% outcomes
+    ) %>%
+    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+    geom_step(aes(x=period_start, y=irr))+
+    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=irr.ll, ymax=irr.ul), alpha=0.1, colour="transparent")+
+    geom_hline(aes(yintercept=0), colour="black")+
+    facet_grid(rows=vars(outcome_descr), scales="free_y")+
+    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
+    coord_cartesian(xlim=c(0, NA))+
+    labs(
+      x="Days",
+      y=NULL,
+      title="Hazard ratio",
+      colour=NULL
     )+
+    theme_minimal(base_size = 9)+
+    km_theme()+
     NULL
@@
     plot_km, 
-    #plot_kmrd, 
+    plot_kmrd, 
     plot_kmrr, 
+    #plot_irr,
     ncol=3
   )
+
   ggsave(
-    filename="km.png",
+    filename=glue("km_{subgroup}.png"),
     path=output_dir_rmd,
@@
   plot_combined
-cat("  \n\n")
+}
 
+```
+
+#### Figure 3a: Kaplan-Meier cumulative incidence estimates
+
+```{r, figure-km-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_km_subgroup("all")
+```
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+#### Figure 3b: Kaplan-Meier cumulative incidence estimates, by primary course type
+
+```{r, figure-km-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_km_subgroup("vax12_type")
+```
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+#### Figure 3c: Kaplan-Meier cumulative incidence estimates, by age group
+
+```{r, figure-km-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_km_subgroup("age65plus")
+```
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+#### Figure 3d Kaplan-Meier cumulative incidence estimates, by prior infection status
+
+```{r, figure-km-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_km_subgroup("prior_covid_infection")
+```
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+#### Figure 3e Kaplan-Meier cumulative incidence estimates, by clinical vulnerability
+
+```{r, figure-km-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_km_subgroup("cev_cv")
 ```
@@
 \newpage 
-# Supplement
+<!-- # Supplement -->
 
-## JCVI vaccine priority groups
+<!-- ## JCVI vaccine priority groups -->
 
-The Joint Committee on Vaccine and Immunisation (JCVI) priority groups were defined as follows:
+<!-- The Joint Committee on Vaccine and Immunisation (JCVI) priority groups were defined as follows: -->
 
-```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE}
+<!-- ```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE} -->
 
-tibble(
-  `Priority group` = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10a", "10b"),
-  `Description` = c(
-    "Residents in a care home for older adults\nStaff working in care homes for older adults",
-    "All those 80 years of age and over\nFrontline health and social care workers",
-    "All those 75-79 years of age",
-    "All those 70-74 years of age\nIndividuals aged 16-69 in a high risk group",
-    "All those 65-69 years of age",
-    "Adults aged 16-64 years in an at-risk group",
-    "All those 60-64 years of age",
-    "All those 55-59 years of age",
-    "All those 50-54 years of age",
-    "All those 40-49 years of age",
-    "All those 18-39 years of age"
-  )
-) %>%
-mutate(Description = str_replace_all(Description, "\n", "<br>")) %>% 
-gt() %>% 
-fmt_markdown(columns = TRUE)
+<!-- tibble( -->
+<!--   `Priority group` = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10a", "10b"), -->
+<!--   `Description` = c( -->
+<!--     "Residents in a care home for older adults\nStaff working in care homes for older adults", -->
+<!--     "80 years of age and over\nFrontline health and social care workers", -->
+<!--     "75-79 years of age", -->
+<!--     "70-74 years of age\nIndividuals aged 16-69 in a high risk group", -->
+<!--     "65-69 years of age", -->
+<!--     "Adults aged 16-64 years in an at-risk group", -->
+<!--     "60-64 years of age", -->
+<!--     "55-59 years of age", -->
+<!--     "50-54 years of age", -->
+<!--     "40-49 years of age", -->
+<!--     "18-39 years of age" -->
+<!--   ) -->
+<!-- ) %>% -->
+<!-- mutate(Description = str_replace_all(Description, "\n", "<br>")) %>%  -->
+<!-- gt() %>%  -->
+<!-- fmt_markdown(columns = TRUE) -->
 
-cat("  \n\n")
-```
+<!-- cat("  \n\n") -->
+<!-- ``` -->
 
-See original priority groups here: https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1007737/Greenbook_chapter_14a_30July2021.pdf#page=15
-See revised priority groups here: https://www.england.nhs.uk/coronavirus/wp-content/uploads/sites/52/2021/07/C1327-covid-19-vaccination-autumn-winter-phase-3-planning.pdf
+<!-- See original priority groups here: https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1007737/Greenbook_chapter_14a_30July2021.pdf#page=15 -->
+<!-- See revised priority groups here: https://www.england.nhs.uk/coronavirus/wp-content/uploads/sites/52/2021/07/C1327-covid-19-vaccination-autumn-winter-phase-3-planning.pdf -->
 
-Note that we excluded care home residents and health care workers in our analysis, so members of JCVI group 1 are not included and JCVI group 2 includes only those aged 80 and over.
+<!-- Note that we excluded care home residents and health care workers in our analysis, so members of JCVI group 1 are not included and JCVI group 2 includes only those aged 80 and over.
 The original priority group list has 9 groups, with a 10th group implicitly defined as "everybody else".
 Here we explicitly define this group, and split into two (10a and 10b) because of the earlier booster eligibility in the 40-49 group from 15 November 2021 onwards.
-(https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds)
+(https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds) -->
 
-\newpage
+<!-- \newpage -->
 
-## References
+<!-- ## References -->

### 2022-05-17 analysis/study_definition.py
add covid critical care admissions as an outcome

@@ -68,6 +68,113 @@ def vaccination_date_X(name, index_date, n, product_name_matches=None, target_di
 
 
 
+def critcare_dates(name, on_or_after, n, with_these_diagnoses, with_admission_method):
+  
+  
+  def var_signature_date(
+    # variable signature for date of hosp admission
+    name,
+    on_or_after,
+    with_these_diagnoses,
+    with_admission_method
+  ):
+    return {
+      name: patients.admitted_to_hospital(
+        returning = "date_admitted",
+        with_these_diagnoses = with_these_diagnoses,
+        with_admission_method = with_admission_method,
+        on_or_after = on_or_after,
+        date_format = "YYYY-MM-DD",
+        find_first_match_in_period = True
+      )
+    }
+    
+  
+  def var_signature_ccdays(
+    # variable signature for days in critical care
+    name,
+    event_date,
+    with_these_diagnoses,
+    with_admission_method
+  ):
+    return {
+      name: patients.admitted_to_hospital(
+        returning = "days_in_critical_care",
+        with_these_diagnoses = with_these_diagnoses,
+        with_admission_method = with_admission_method,
+        between = [event_date, event_date],
+        find_first_match_in_period = True,
+        return_expectations = {
+          "category":{"ratios": {"0": 0.8, "1": 0.1, "2": 0.1}}
+        }
+      )
+    }
+    
+  # define a sequence of n variables for date of admission and associated number of days in critical care
+  
+  # initialise for first date
+  variables_date = var_signature_date(f"{name}_1_date", on_or_after, with_these_diagnoses, with_admission_method)
+  variables_ccdays = var_signature_ccdays(f"{name}_1_ccdays", f"{name}_1_date", with_these_diagnoses, with_admission_method)
+  #isadmission_cc = {"1" : f"{name}_1_date AND ({name}_1_ccdays > 0)"}
+  
+  # loop for subsequent dates 
+  for i in range(2, n+1):
+    variables_date.update(
+      var_signature_date(
+        name = f"{name}_{i}_date", 
+        on_or_after = f"{name}_{i-1}_date + 1 days",
+        with_these_diagnoses = with_these_diagnoses,
+        with_admission_method = with_admission_method
+      )
+    )
+    
+    variables_ccdays.update(
+      var_signature_ccdays(
+        name = f"{name}_{i}_ccdays", 
[…]

### 2022-05-17 create-project.R

@@ -348,7 +348,7 @@ actions_list <- splice(
   action_km("A", "all", "covidadmittedproxy1"),
   action_km("A", "all", "covidadmitted"),
   #action_km("A", "all", "noncovidadmitted"),
-  #action_km("A", "all", "covidcc"),
+  #action_km("A", "all", "covidcritcare"),
   action_km("A", "all", "coviddeath"),
   action_km("A", "all", "noncoviddeath"),
 
@@ -360,7 +360,7 @@ actions_list <- splice(
   action_km("A", "vax12_type", "covidadmittedproxy1"),
   action_km("A", "vax12_type", "covidadmitted"),
   #action_km("A", "vax12_type", "noncovidadmitted"),
-  #action_km("A", "vax12_type", "covidcc"),
+  #action_km("A", "vax12_type", "covidcritcare"),
   action_km("A", "vax12_type", "coviddeath"),
   action_km("A", "vax12_type", "noncoviddeath"),
 
@@ -372,7 +372,7 @@ actions_list <- splice(
   action_km("A", "cev_cv", "covidadmittedproxy1"),
   action_km("A", "cev_cv", "covidadmitted"),
   #action_km("A", "cev_cv", "noncovidadmitted"),
-  #action_km("A", "cev_cv", "covidcc"),
+  #action_km("A", "cev_cv", "covidcritcare"),
   action_km("A", "cev_cv", "coviddeath"),
   action_km("A", "cev_cv", "noncoviddeath"),
 
@@ -384,7 +384,7 @@ actions_list <- splice(
   action_km("A", "prior_covid_infection", "covidadmittedproxy1"),
   action_km("A", "prior_covid_infection", "covidadmitted"),
   #action_km("A", "prior_covid_infection", "noncovidadmitted"),
-  #action_km("A", "prior_covid_infection", "covidcc"),
+  #action_km("A", "prior_covid_infection", "covidcritcare"),
   action_km("A", "prior_covid_infection", "coviddeath"),
   action_km("A", "prior_covid_infection", "noncoviddeath"),
 
@@ -396,7 +396,7 @@ actions_list <- splice(
   action_km("A", "age65plus", "covidadmittedproxy1"),
   action_km("A", "age65plus", "covidadmitted"),
   #action_km("A", "age65plus", "noncovidadmitted"),
-  #action_km("A", "age65plus", "covidcc"),
+  #action_km("A", "age65plus", "covidcritcare"),
[…]

### 2022-05-17 lib/design/design.R

@@ -33,7 +33,7 @@ events_lookup <- tribble(
   "noncovidadmitted", "noncovidadmitted_date", "Non-COVID-19 hospitalisation",
   "covidadmittedproxy1", "covidadmittedproxy1_date", "COVID-19 hospitalisation (A&E proxy)",
   "covidadmittedproxy2", "covidadmittedproxy2_date", "COVID-19 hospitalisation (A&E proxy v2)",
-  "covidcc", "covidcc_date", "COVID-19 critical care",
+  "covidcritcare", "covidcritcare_date", "COVID-19 critical care",
   "coviddeath", "coviddeath_date", "COVID-19 death",
   "noncoviddeath", "noncoviddeath_date", "Non-COVID-19 death",
   "death", "death_date", "Any death",
@@ -87,7 +87,7 @@ recoder <-
       "COVID-19 hospitalisation" = "covidadmitted",
       "COVID-19 hospitalisation (A&E proxy)" = "covidadmittedproxy1",
       "COVID-19 hospitalisation (A&E proxy v2)" = "covidadmittedproxy2",
-      "COVID-19 critical care" = "covidcc",
+      "COVID-19 critical care" = "covidcritcare",
       "COVID-19 death" = "coviddeath",
       "Non-COVID-19 death" = "noncoviddeath",
       "All-cause death" = "death"

### 2022-05-17 lib/dummydata/dummydata.R

@@ -308,10 +308,42 @@ sim_list = lst(
     missing_rate = ~0.7
   ),
 
-  covidadmitted_ccdays = bn_node(
-    ~as.factor(as.integer(rpois(n=..n, 0.5))),
-    missing_rate = ~is.na(covidadmitted_day)
+  #placeholder for single criticalcare variable ---
+  # covidcritcare_day = bn_node(
+  #   ~as.integer(runif(n=..n, covid_vax_disease_3_day, covid_vax_disease_3_day+100)),
+  #   missing_rate = ~0.8
+  # ),
+
+  potentialcovidcritcare_1_day = bn_node(
+    ~as.integer(runif(n=..n, covid_vax_disease_3_day, covid_vax_disease_3_day+70)),
+    missing_rate = ~0.8
   ),
+
+  potentialcovidcritcare_2_day = bn_node(
+    ~as.integer(runif(n=..n, potentialcovidcritcare_1_day, potentialcovidcritcare_1_day+70)),
+    missing_rate = ~0.8
+  ),
+
+  potentialcovidcritcare_3_day = bn_node(
+    ~as.integer(runif(n=..n, potentialcovidcritcare_2_day, potentialcovidcritcare_2_day+70)),
+    missing_rate = ~0.8
+  ),
+
+  potentialcovidcritcare_1_ccdays = bn_node(
+    ~as.factor(as.integer(rexp(n=..n, 1))),
+    needs = "potentialcovidcritcare_1_day"
+  ),
+
+  potentialcovidcritcare_2_ccdays = bn_node(
+    ~as.factor(as.integer(rexp(n=..n, 1))),
+    needs = "potentialcovidcritcare_2_day"
+  ),
+
+  potentialcovidcritcare_3_ccdays = bn_node(
+    ~as.factor(as.integer(rexp(n=..n, 1))),
+    needs = "potentialcovidcritcare_3_day"
+  ),
+
 
   admitted_unplanned_day = bn_node(
     ~as.integer(runif(n=..n, covid_vax_disease_3_day, covid_vax_disease_3_day+100)),

### 2022-05-17 run-local.R

@@ -50,7 +50,7 @@ action_km("A", "all", "covidemergency")
 action_km("A", "all", "covidadmittedproxy1")
 action_km("A", "all", "covidadmitted")
 #action_km("A", "all", "noncovidadmitted")
-#action_km("A", "all", "covidcc")
+#action_km("A", "all", "covidcritcare")
 action_km("A", "all", "coviddeath")
 action_km("A", "all", "noncoviddeath")
 
@@ -60,7 +60,7 @@ action_km("A", "vax12_type", "covidemergency")
 action_km("A", "vax12_type", "covidadmittedproxy1")
 action_km("A", "vax12_type", "covidadmitted")
 #action_km("A", "vax12_type", "noncovidadmitted")
-#action_km("A", "vax12_type", "covidcc")
+#action_km("A", "vax12_type", "covidcritcare")
 action_km("A", "vax12_type", "coviddeath")
 action_km("A", "vax12_type", "noncoviddeath")
 
@@ -71,7 +71,7 @@ action_km("A", "cev_cv", "covidemergency")
 action_km("A", "cev_cv", "covidadmittedproxy1")
 action_km("A", "cev_cv", "covidadmitted")
 #action_km("A", "cev_cv", "noncovidadmitted")
-#action_km("A", "cev_cv", "covidcc")
+#action_km("A", "cev_cv", "covidcritcare")
 action_km("A", "cev_cv", "coviddeath")
 action_km("A", "cev_cv", "noncoviddeath")
 
@@ -81,7 +81,7 @@ action_km("A", "prior_covid_infection", "covidemergency")
 action_km("A", "prior_covid_infection", "covidadmittedproxy1")
 action_km("A", "prior_covid_infection", "covidadmitted")
 #action_km("A", "prior_covid_infection", "noncovidadmitted")
-#action_km("A", "prior_covid_infection", "covidcc")
+#action_km("A", "prior_covid_infection", "covidcritcare")
 action_km("A", "prior_covid_infection", "coviddeath")
 action_km("A", "prior_covid_infection", "noncoviddeath")
 
@@ -91,7 +91,7 @@ action_km("A", "age65plus", "covidemergency")
 action_km("A", "age65plus", "covidadmittedproxy1")
 action_km("A", "age65plus", "covidadmitted")
 #action_km("A", "age65plus", "noncovidadmitted")
-#action_km("A", "age65plus", "covidcc")
+#action_km("A", "age65plus", "covidcritcare")
[…]

### 2022-05-18 write-up/report.Rmd

@@
       "covidadmittedproxy1",
-      #"covidcc",
+      "covidcritcare",
       "coviddeath",

### 2022-05-23 analysis/ci_combine.R
use Aalen-Johansen estimates (with cause specific deaths as competing events)
* add scripts and actions for calculating cumulative incidence under competing risks
* add km estimates to tables, alongside cumulative incidence estimates
* remove KM scripts from running
* use midpoint of rounding points to remove bias, instead of ceiling
* improve plots for checking on server
* remove unused formulae
* change post baseline cuts

@@ -1,0 +1,131 @@
+
+# # # # # # # # # # # # # # # # # # # # #
+# Purpose: Combine ci estimates from different outcomes
+#  - import matched data
+#  - adds outcome variable and restricts follow-up
+#  - gets CI estimates
+#  - The script must be accompanied by two arguments:
+#    `matchset` - the matching set used for matching
+#    `subgroup` - the subgroup variable, which is used to stratify CI estimates
+# # # # # # # # # # # # # # # # # # # # #
+
+# Preliminaries ----
+
+## import command-line arguments ----
+
+args <- commandArgs(trailingOnly=TRUE)
+
+
+if(length(args)==0){
+  # use for interactive testing
+  matchset <- "A"
+
+} else {
+  matchset <- args[[1]]
+
+}
+
+## Import libraries ----
+library('tidyverse')
+library('here')
+library('glue')
+library('survival')
+
+## Import custom user functions from lib
+source(here("lib", "functions", "utility.R"))
+
+## Import design elements
+source(here("lib", "design", "design.R"))
+
+
+output_dir <- here("output", "match", matchset, "ci", "combined")
+fs::dir_create(output_dir)
+
+metaparams <-
+  expand_grid(
+    outcome = factor(c("postest", "covidemergency", "covidadmittedproxy1", "covidadmitted", "covidcritcare", "coviddeath", "noncoviddeath")),
+    subgroup = factor(recoder$subgroups),
+  ) %>%
+  mutate(
+    outcome_descr = fct_recoderelevel(outcome,  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroups),
+  )
+
+
+ci_estimates <- metaparams %>%
+  mutate(
+    data = pmap(list(matchset, subgroup, outcome), function(matchset, subgroup, outcome) {
+      subgroup <- as.character(subgroup)
+      dat <- read_csv(here("output", "match", matchset, "ci", subgroup, outcome, "ci_estimates.csv"), na="NA", col_types = cols())
+      dat %>%
+      add_column(
+        subgroup_level = as.character(.[[subgroup]]),
+        subgroup_level_descr = fct_recoderelevel(.[[subgroup]], recoder[[subgroup]]),
+        .before=1
+      ) %>%
+      select(-all_of(subgroup))
+    })
+  ) %>%
+  unnest(data)
[…]

### 2022-05-23 analysis/misc/testcif.R

@@ -1,0 +1,174 @@
+
+library('tidyverse')
+library('here')
+library('glue')
+library('survival')
+
+## Import custom user functions from lib
+source(here("lib", "functions", "utility.R"))
+source(here("lib", "functions", "survival.R"))
+source(here("lib", "functions", "redaction.R"))
+
+## Import design elements
+source(here("lib", "design", "design.R"))
+
+matchset <- "A"
+subgroup <- "all"
+#subgroup <- "vax12_type"
+outcome <- "covidadmitted"
+
+cif.se <- function(time, ci, n.risk, n.event, kmsurv, kmsummand){
+  # from https://ncss-wpengine.netdna-ssl.com/wp-content/themes/ncss/pdf/Procedures/NCSS/Cumulative_Incidence.pdf
+  # also here: https://onlinelibrary.wiley.com/doi/full/10.1002/bimj.200900039?saml_referrer
+  timeindex <- seq_along(time)
+  lapply(timeindex, function(i) {
+    cii <- ci[1:i]
+    bi <- ((n.risk - n.event)/n.risk)[1:i]
+    di <- (n.event/(n.risk^2))[1:i]
+    lagkmi <- lag(kmsurv,1,1)[1:i]
+    kmsummandi <- kmsummand[1:i]
+    vt <-
+      sum(((cii[i] - cii)^2) * kmsummandi) +
+      sum((lagkmi^2) * bi * di) +
+      -2* sum((cii[i] - cii) * lagkmi * di)
+
+    sqrt(vt)
+  })
+}
+
+
+output_dir <- here("output", "match", matchset, "ci", subgroup, outcome)
+
+## import match data
+data_matchstatus <- read_rds(here("output", "match", matchset, "data_matchstatus.rds"))
+
+## import baseline data, restrict to matched individuals and derive time-to-event variables
+data_matched <-
+  read_rds(here("output", "data", "data_cohort.rds")) %>%
+  select(
+    # select only variables needed for models to save space
+    patient_id, vax3_date,
+    all_of(subgroup),
+    all_of(paste0(c(outcome, "death", "dereg", "coviddeath", "noncoviddeath", "vax4"), "_date")),
+    all_of(matching_variables[[matchset]]$all), # model formula with all variables
+    all_of(all.vars(formula_allcovariates)), # model formula with all variables
+  ) %>%
[…]

### 2022-05-23 analysis/release_objects.R

@@ -26,9 +26,15 @@ for(matchset in c("A", "B")){
 
   ## KM ----
 
-  fs::file_copy(here("output", "match", matchset, "km", "combined", "km_estimates.csv"), fs::path(output_dir, "km_estimates.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "match", matchset, "km", "combined", "contrasts_daily.csv"), fs::path(output_dir, "contrasts_daily.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "match", matchset, "km", "combined", "contrasts_overall.csv"), fs::path(output_dir, "contrasts_overall.csv"), overwrite = TRUE)
+  #fs::file_copy(here("output", "match", matchset, "km", "combined", "km_estimates.csv"), fs::path(output_dir, "km_estimates.csv"), overwrite = TRUE)
+  #fs::file_copy(here("output", "match", matchset, "km", "combined", "contrasts_daily.csv"), fs::path(output_dir, "km_contrasts_daily.csv"), overwrite = TRUE)
+  #fs::file_copy(here("output", "match", matchset, "km", "combined", "contrasts_overall.csv"), fs::path(output_dir, "km_contrasts_overall.csv"), overwrite = TRUE)
+
+  ## CI ----
+
+  fs::file_copy(here("output", "match", matchset, "ci", "combined", "ci_estimates.csv"), fs::path(output_dir, "ci_estimates.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", matchset, "ci", "combined", "contrasts_daily.csv"), fs::path(output_dir, "ci_contrasts_daily.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "match", matchset, "ci", "combined", "contrasts_overall.csv"), fs::path(output_dir, "ci_contrasts_overall.csv"), overwrite = TRUE)
 
 }
 

### 2022-05-23 lib/design/design.R

@@ -55,7 +55,7 @@ treatement_lookup <-
   )
 
 # where to split follow-up time after recruitment
-postbaselinecuts <- c(0,7,14,28,42,70)
+postbaselinecuts <- c(0,7,14,21,28,42,56,70,84)
 
 # maximum follow-up duration
 
@@ -119,42 +119,6 @@ recoder <-
 ## model formulae ----
 
 treated_period_variables <- paste0("treatment_period_id", "_", seq_len(length(postbaselinecuts)-1))
-
-
-# unadjusted
-formula_vaxonly <- as.formula(
-  str_c(
-    "Surv(tstart, tstop, ind_outcome) ~ ",
-    str_c(treated_period_variables, collapse = " + ")
-  )
-)
-
-#formula_vaxonly <- Surv(tstart, tstop, ind_outcome) ~ treated:strata(fup_period)
-
-# cox stratification
-formula_strata <- . ~ . +
-  strata(vax3_date) +
-  strata(region) +
-  #strata(jcvi_group) +
-  strata(vax12_type)
-
-formula_demog <- . ~ . +
-  poly(age, degree=2, raw=TRUE) +
-  sex +
-  imd_Q5 +
-  ethnicity_combined
-
-formula_clinical <- . ~ . +
-  prior_tests_cat +
-  multimorb +
-  learndis +
-  sev_mental +
-  immunosuppressed +
-  asplenia
-
-formula_timedependent <- . ~ . +
-  prior_covid_infection +
-  inhospital_planned
 
 
 if(exists("matchset")){
@@ -222,21 +186,4 @@ if(exists("matchset")){
 
   })
 
-  # remove matching variables from formulae, as treatment groups are already balanced
-  formula_remove_matching <- as.formula(paste(". ~ . - ", paste(matching_variables[[matchset]]$all, collapse=" -"), "- poly(age, degree=2, raw=TRUE)"))
-
-  formula0_pw <- formula_vaxonly %>% update(formula_remove_matching)
-  formula1_pw <- formula_vaxonly %>% update(formula_strata) %>% update(formula_remove_matching)
-  formula2_pw <- formula_vaxonly %>% update(formula_strata) %>% update(formula_demog) %>% update(formula_remove_matching)
-  formula3_pw <- formula_vaxonly %>% update(formula_strata) %>% update(formula_demog) %>% update(formula_clinical) %>% update(formula_timedependent) %>% update(formula_remove_matching)
-
[…]

### 2022-05-23 lib/functions/utility.R

@@ -6,6 +6,12 @@ ceiling_any <- function(x, to=1){
   ceiling(plyr::round_any(x/to, 1/100000000))*to
 }
 
+roundmid_any <- function(x, to=1){
+  # like ceiling_any, but centers on (integer) midpoint of the rounding points
+  ceiling(plyr::round_any(x/to, 1/100000000))*to - floor(to/2)
+  #plyr::round_any(x+floor(to/2), to)-floor(to/2)
+}
+
 fct_case_when <- function(...) {
   # uses dplyr::case_when but converts the output to a factor,
   # with factors ordered as they appear in the case_when's  ... argument
@@ -19,6 +25,4 @@ fct_recoderelevel <- function(x, lookup){
   stopifnot(!is.na(names(lookup)))
   factor(x, levels=lookup, labels=names(lookup))
 }
-
-
 

### 2022-05-23 run-local.R

@@ -31,12 +31,14 @@ action_match <- compose_action(
   here("analysis", "match_report.R")
 )
 
-action_km <- compose_action(
-    here("analysis", "km.R")
+action_contrasts <- compose_action(
+    #here("analysis", "km.R"),
+    here("analysis", "ci.R")
 )
 
-action_km_combine <- compose_action(
-  here("analysis", "km_combine.R")
+action_combine <- compose_action(
+  #here("analysis", "km_combine.R"),
+  here("analysis", "ci_combine.R")
 )
 
 action_ls(
@@ -45,131 +47,131 @@ action_ls(
 )
 
 action_match("A")
-action_km("A", "all", "postest")
-action_km("A", "all", "covidemergency")
-action_km("A", "all", "covidadmittedproxy1")
-action_km("A", "all", "covidadmitted")
-#action_km("A", "all", "noncovidadmitted")
-action_km("A", "all", "covidcritcare")
-action_km("A", "all", "coviddeath")
-action_km("A", "all", "noncoviddeath")
+action_contrasts("A", "all", "postest")
+action_contrasts("A", "all", "covidemergency")
+action_contrasts("A", "all", "covidadmittedproxy1")
+action_contrasts("A", "all", "covidadmitted")
+#action_contrasts("A", "all", "noncovidadmitted")
+action_contrasts("A", "all", "covidcritcare")
+action_contrasts("A", "all", "coviddeath")
+action_contrasts("A", "all", "noncoviddeath")
 
 
-action_km("A", "vax12_type", "postest")
-action_km("A", "vax12_type", "covidemergency")
-action_km("A", "vax12_type", "covidadmittedproxy1")
-action_km("A", "vax12_type", "covidadmitted")
-#action_km("A", "vax12_type", "noncovidadmitted")
-action_km("A", "vax12_type", "covidcritcare")
-action_km("A", "vax12_type", "coviddeath")
-action_km("A", "vax12_type", "noncoviddeath")
+action_contrasts("A", "vax12_type", "postest")
+action_contrasts("A", "vax12_type", "covidemergency")
+action_contrasts("A", "vax12_type", "covidadmittedproxy1")
+action_contrasts("A", "vax12_type", "covidadmitted")
+#action_contrasts("A", "vax12_type", "noncovidadmitted")
+action_contrasts("A", "vax12_type", "covidcritcare")
+action_contrasts("A", "vax12_type", "coviddeath")
[…]

### 2022-05-23 write-up/report.Rmd
fizz comments

@@
 
-For the simpler matching criteria, `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
+For the simpler matching criteria, `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
 
@@
     names_from = vax3_type,
-    values_from = c(n, n_exclude, pct_all)
+    values_from = c(n, n_exclude, pct_all, pct_step)
   ) %>%
@@
 
-    pct_all_pfizer = "% remaining",
-    pct_all_moderna = "% remaining",
+    pct_all_pfizer = "% (of all)",
+    pct_all_moderna = "% (of all)",
+    pct_step_pfizer = "% (previous)",
+    pct_step_moderna = "% (previous)",
   ) %>%
@@
 
-#### Figure 1: Cumulative matching success
+#### Figure 1: Cumulative numbers included in the matched sample
 
-```{r, figure-matchingsuccess, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
+```{r, figure-matchingrates, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
 #read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))
@@
 
+label_number_n <- label_number(1, 1, big.mark=",")
+label_number_risk <- label_number(0.1, 10000, style_negative="minus", big.mark=",")
+label_number_rr <- label_number(0.01, 1, style_negative="minus")
+
 contrasts_overall %>%
@@
     subgroup_level_descr,
-    n.atrisk=n.atrisk_0,
-    persontime_0=label_number(1)(persontime_0/7), 
-    persontime_1=label_number(1)(persontime_1/7),
-    n.event_0, 
-    n.event_1,
-    riskCI_0 = paste0(label_number(0.1, 10000)(1-surv_0), " (", label_number(0.1, 10000)(1-surv.ul_0), "-", label_number(0.1, 10000)(1-surv.ll_0), ")"),
-    riskCI_1 = paste0(label_number(0.1, 10000)(1-surv_1), " (", label_number(0.1, 10000)(1-surv.ul_1), "-", label_number(0.1, 10000)(1-surv.ll_1), ")"),
-    kmrdCI = paste0(label_number(0.1, 10000)(kmrd), " (", label_number(0.1, 10000)(kmrd.ll), "-", label_number(1, 10000)(kmrd.ul), ")"),
-    kmrrCI = paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(kmrr.ll), "-", label_number(0.01)(kmrr.ul), ")"),
-    irrCI =  paste0(label_number(0.01)(irr), " (", label_number(0.01)(irr.ll), "-", label_number(0.01)(irr.ul), ")"),
+    n.atrisk=label_number_n(n.atrisk_0),
+    persontime_0=label_number_n(persontime_0/7), 
+    persontime_1=label_number_n(persontime_1/7),
+    n.event_0=label_number_n(n.event_0), 
+    n.event_1=label_number_n(n.event_1),
+    riskCI_0 = glue("{label_number_risk(1-surv_0)} ({label_number_risk(1-surv.ul_0)} - {label_number_risk(1-surv.ll_0)})"),
+    riskCI_1 = glue("{label_number_risk(1-surv_1)} ({label_number_risk(1-surv.ul_1)} - {label_number_risk(1-surv.ll_1)})"),
+    kmrdCI = glue("{label_number_risk(kmrd)} ({label_number_risk(kmrd.ll)} - {label_number_risk(kmrd.ul)})"),
+    kmrrCI = glue("{label_number_rr(kmrr)} ({label_number_rr(kmrr.ll)} - {label_number_rr(kmrr.ul)})"),
+    irrCI =   glue("{label_number_rr(irr)} ({label_number_rr(irr.ll)} - {label_number_rr(irr.ul)})"),
     #coxhrCI =  paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(coxhr.ll), "-", label_number(0.01)(coxhr.ul), ")"),

### 2022-07-06 write-up/manuscriptv2.Rmd
add manuscript draft

---
title: "Comparative effectiveness of BNT162b2 versus mRNA-1273 boosting in England: a cohort study in OpenSAFELY-TPP"
output:
  html_document:
    self_contained: yes
  pdf_document:
    toc: no
  word_document: default
  rtf_document:
    toc: no
  bookdown::html_document2:
    number_sections: no
    toc: no
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r setup, include=FALSE, message=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')

# remotes::install_github("https://github.com/wjchulme/osutils")
# library('osutils')

# where are the outputs (ie the inputs for this manuscript!) saved?
#output_dir_os <- here("output", "release-objects")
output_dir_os <- here("released-output", "release-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("write-up", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

## Import custom user functions from lib
source(here("lib", "functions", "utility.R"))

## Import design elements
source(here("lib", "design", "design.R"))

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidadmittedproxy1",
      #"covidcritcare",
      "coviddeath"
      #"noncoviddeath"
    ) %>% 
    set_names(., .)

```

```{r match-A, echo=FALSE, warning=FALSE, message=FALSE}

matchA_flowchart <-  read_csv(fs::path(output_dir_os, "A", "match_flowchart.csv"))
matchB_flowchart <-  read_csv(fs::path(output_dir_os, "B", "match_flowchart.csv"))

match_flowchart <-
  bind_rows(
    matchA_flowchart %>% 
      mutate(
        criteria = if_else(crit=="c7", "and matched with strategy A", criteria),
        crit=if_else(crit=="c7", "c7A", crit), 
      ),
    matchB_flowchart %>% 
      mutate(
        criteria = if_else(crit=="c7", "and matched with strategy B", criteria),
        crit=if_else(crit=="c7", "c7B", crit), 
      ) %>%
      filter(
        crit=="c7B"
      )
  ) %>%
  arrange(vax3_type, crit)

matchA_coverage <- read_csv(fs::path(output_dir_os, "A", "match_coverage.csv")) %>%
   mutate(
    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
    #status = if_else(status=="matched", "matchedA", status)
  ) %>%
  rename(nA=n, cumulnA=cumuln)

matchB_coverage <- read_csv(fs::path(output_dir_os, "B", "match_coverage.csv")) %>%
   mutate(
    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
    #status = if_else(status=="matched", "matchedB", status)
  )%>%
  rename(nB=n, cumulnB=cumuln)

match_coverage <-
  bind_rows(
    matchA_coverage %>% filter(status=="unmatched"),
    left_join(
      matchA_coverage, 
      matchB_coverage, 
      by=c("treatment", "treatment_descr", "status", "status_descr", "vax3_date")
    ) %>% 
      filter(status=="matched") %>% 
      mutate(strategy="A"),
    matchB_coverage %>% 
      filter(status=="matched") %>% 
      mutate(strategy="B"),
  ) %>%
  mutate(
    n = case_when(
      is.na(strategy) ~ nA,
      strategy=="A" ~ nA-nB,
      strategy=="B" ~ nB
    ),
    cumuln = case_when(
      is.na(strategy) ~ cumulnA,
      strategy=="A" ~ cumulnA-cumulnB,
      strategy=="B" ~ cumulnB
    ),
    status = case_when(
      is.na(strategy) ~ "unmatched",
      strategy=="A" ~ "matched (A only)",
      strategy=="B" ~ "matched (A and B)"
    ),
    status_descr = fct_case_when(
      is.na(strategy) ~ "Unmatched",
      strategy=="A" ~ "Matched (strategy A only)",
      strategy=="B" ~ "Matched (strategy A and B)"

    ),
  )

matchA_table1 <- read_csv(fs::path(output_dir_os, "A", "match_table1.csv"))
matchA_table1by <- read_csv(fs::path(output_dir_os, "A", "match_table1by.csv"))

matchB_table1 <- read_csv(fs::path(output_dir_os, "B", "match_table1.csv"))
matchB_table1by <- read_csv(fs::path(output_dir_os, "B", "match_table1by.csv"))

match_table1 <- 
  bind_rows(
    matchA_table1 %>% mutate(strategy="A"),
    matchB_table1 %>% mutate(strategy="B"),
  )

matchset <- "B"

ci_estimates <- read_csv(fs::path(output_dir_os, matchset, "ci_estimates.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrasts_daily <- read_csv(fs::path(output_dir_os, matchset, "ci_contrasts_daily.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrasts_cuts <- read_csv(fs::path(output_dir_os, matchset, "ci_contrasts_cuts.csv")) %>%
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsA_overall <- read_csv(fs::path(output_dir_os, "A", "ci_contrasts_overall.csv")) 
contrastsB_overall <- read_csv(fs::path(output_dir_os, "B", "ci_contrasts_overall.csv"))

contrasts_overall <- 
  bind_rows(
    contrastsA_overall %>% mutate(strategy="A"),
    contrastsB_overall %>% mutate(strategy="B"),
  ) %>%
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = fct_inorder(replace_na(subgroup_level_descr, "")),
  )

perN <- 1000
perN_format <- label_number(1, 1, big.mark=",")(perN)

label_number_n <- label_number(1, 1, big.mark=",")
label_number_risk <- label_number(0.01, perN, style_negative="minus", big.mark=",")
label_number_rr <- label_number(0.01, 1, style_negative="minus")

contrasts_overall_table <-
  contrasts_overall %>%
  mutate(
    outcome,
    outcome_descr,
    subgroup,
    subgroup_descr,
    subgroup_level_descr,
    n.event = label_number_n(n.event_0+n.event_1),
    persontime_weeks = label_number_n((persontime_0 + persontime_1) / 7),
    n.atrisk = label_number_n(n.atrisk_0),
    persontime_0=label_number_n(persontime_0/7), 
    persontime_1=label_number_n(persontime_1/7),
    n.event_0=label_number_n(n.event_0), 
    n.event_1=label_number_n(n.event_1),

    risk.ll_0 = 1-surv.ul_0,
    risk.ul_0 = 1-surv.ll_0,
    risk.ll_1 = 1-surv.ul_1,
    risk.ul_1 = 1-surv.ll_1,
    riskCI_0 = glue("{label_number_risk(risk_0)} ({label_number_risk(risk.ll_0)} to {label_number_risk(risk.ul_0)})"),
    riskCI_0_95 = glue("{label_number_risk(risk_0)} (95%CI {label_number_risk(risk.ll_0)} to {label_number_risk(risk.ul_0)})"),
    riskCI_1 = glue("{label_number_risk(risk_1)} ({label_number_risk(risk.ll_1)} to {label_number_risk(risk.ul_1)})"),
    cirdCI = glue("{label_number_risk(cird)} ({label_number_risk(cird.ll)} to {label_number_risk(cird.ul)})"),
    cirdCI_95 = glue("{label_number_risk(cird)} (95%CI {label_number_risk(cird.ll)} to {label_number_risk(cird.ul)})"),
    #cirrCI = glue("{label_number_rr(cirr)} ({label_number_rr(cirr.ll)}-{label_number_rr(cirr.ul)})"),
    #kmirrCI =   glue("{label_number_rr(kmirr)} ({label_number_rr(kmirr.ll)}-{label_number_rr(kmirr.ul)})"),
    coxhrCI =   glue("{label_number_rr(coxhr)} ({label_number_rr(coxhr.ll)} to {label_number_rr(coxhr.ul)})"),
    coxhrCI_95 =   glue("{label_number_rr(coxhr)} (95%CI {label_number_rr(coxhr.ll)} to {label_number_rr(coxhr.ul)})"),
  )

```

William J Hulme^1^, Elsie Horne^3,4^, Ruth Keogh^2^, Edward PK Parker^2^, Venexia Walker^3,5^, Elizabeth J Williamson^2^, Tom Palmer^3,5^, Linda Nab^1^, Helen J Curtis^1^, Ben Goldacre^1^, Miguel A Hernán^8,9^, Jonathan AC Sterne^3,4,7^, and the OpenSAFELY collaborative.

1\. The Bennett Institute for Applied Data Science, Nuffield Department of Primary Care Health Sciences, University of Oxford, OX26GG, UK

2\. London School of Hygiene and Tropical Medicine, Keppel Street, London, WC1E 7HT, UK

3\. Population Health Sciences, University of Bristol, Oakfield House, Oakfield Grove, Bristol, BS8 2BN, UK

4\. NIHR Bristol Biomedical Research Centre, Bristol, UK

5\. MRC Integrative Epidemiology Unit, Bristol Medical School, University of Bristol, Bristol, UK

6\. TPP, TPP House, 129 Low Lane, Horsforth, Leeds, LS18 5PX

7\. Health Data Research UK South West

8\. CAUSALab, Harvard T.H. Chan School of Public Health, Boston, MA 02115

9\. Departments of Epidemiology and Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA 02115

\newpage

## Abstract

*Introduction* The BNT162b2 and mRNA-1273 vaccines were both widely used for COVID-19 booster vaccination in England, and are both effectiveness against COVID-19 related outcomes, particularly severe outcomes. Direct comparisons of the effectiveness of these two vaccine types when used for boosting have not been made in trials or observational data.

*Methods* Using the OpenSAFELY-TPP linked electronic health record database, we matched adult recipients of each vaccine type on date of vaccination, primary vaccine course, prior SARS-CoV-2 infection, and other important demographic and clinical characteristics. Recipients were eligible for inclusion if boosted between 29 October 2021 to 31 January 2022, and following up for 12 weeks. Outcomes were positive SARS-CoV-2 test, COVID-19 hospitalisation, and COVID-19 death. We used the Aalen-johansen estimator to obtain the cumulative incidence of each outcome, and compared these using risk differences (RD) and hazard ratios (HRs).

*Results* `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(n))` people were matched in each group, contributing a total of `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up. The 12-week risk per `r perN_format` people of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 was `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`) respectively.

*Conclusion* COVID-19-related outcomes after booster vaccination with BNT162b2 or mRNA-1273 were rare, and the risk was slightly lower for mRNA-1273 than for BNT162b2. These finding were broadly consistent across subgroups.

\newpage

## Background

The UK COVID-19 vaccination programme delivered its first "booster" doses in September 2021 @NHSEnglandNHS. Following guidance from the Joint Committee for Vaccination and Immunisation (JCVI) expert advisory group@covid19 boosting followed a staged roll-out that prioritised groups at high risk of infection or severe disease, and was then progressively extended to the whole adult population by mid December 2021 @NHSEnglandNHSb @NHSEnglandNHSa1 @NHSEnglandNHSa2. The BNT162b2 Pfizer-BioNTech vaccine was used initially for boosting, with a half-dose of Moderna mRNA-1273 subsequently also used. The concurrent administration of the BNT162b2 and mRNA-1273 vaccines for boosting in England, receipt of which was largely determined by local availability rather than any clinical criteria, enables a direct comparison of their effectiveness. No randomised trials have made such a comparison for efficacy outcomes.

On behalf of NHS England, we used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to compare the effectiveness of BNT162b2 and mRNA-1273 boosting in adult recipients.

\newpage

## Methods

## Data source

All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>. With the approval of NHS England, primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS).

## Eligibility criteria

We considered all adults (aged 18 years and over) who received a booster dose of BNT162b2 or mRNA-1273 between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")` inclusive, during which time both vaccine brands were being used. People were eligible if they: were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and mRNA-1273 were not considered due to small numbers); were not a health or social care worker, not resident in a care or nursing home and not medically housebound or receiving end-of-life care; had no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; were not hospitalised at the time of boosting; and had complete information on sex, deprivation, and NHS region (defined using Sustainability and Transformation Partnerships (STP, a geographical grouping of NHS and local authorities).

## Matching

BNT162b2 booster recipients were matched with mRNA-1273 booster recipients in a one to one ratio without replacement, on the following characteristics: date of booster dose; primary vaccine course (BNT162b2 or ChadOx1-S); Date of second vaccine dose (7 day caliper); sex (male or female); age (3 year caliper and within age groups defined by JVCI risk groups); Clinical risk group defined by JCVI (clinically extremely vulnerable, clinically at-risk, neither); Index of Multiple Deprivation (IMD, grouped by quintile); STP as a surrogate for geographical region; evidence of prior SARS-CoV-2 infection (any of positive SARS-CoV-2 test, probable infection documented in primary care, or COVID-19 hospital attendance or admission); morbidity count (0, 1, or 2 or more of diabetes, BMI over 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease). The supplementary materials provide more information on how these characteristics were defined.

<!-- #### Matching strategies -->

<!-- ```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE} -->

<!-- tribble( -->
<!--   ~Factor, ~Description, ~A, ~B, -->

<!--   "Date of booster dose",  "Matched exactly",   -->
<!--     "X", "X", -->
<!--   "Primary course vaccine brand", "BNT162b2; ChadOx1-S",  -->
<!--     "X", "X", -->
<!--   "Date of second vaccine dose", "7 day caliper",  -->
<!--     "X", "X", -->
<!--   "Age",  "3 year caliper",  -->
<!--     "X", "X", -->
<!--   "JCVI age group", "18-39; 40-49; 50-54; 55-59; 60-64; 65-69; 70-75; 75-79; 80 and over", -->
<!--     "X", "X", -->
<!--   "JCVI Clinical risk group", "clinically extremely vulnerable; clinically at-risk; neither", -->
<!--     "X", "X", -->
<!--   "English Index of Multiple Deprivation", "Grouped by quintile", -->
<!--     "X", "X", -->
<!--   "NHS region", "East of England; Midlands; London; North East and Yorkshire; North West; South East; South West", -->
<!--     "X", "", -->
<!--   "Sustainability and Transformation Partnership (STP)", "a geographical grouping of NHS and local authorities", -->
<!--     "", "X", -->
<!--   "Sex", "Male; Female", -->
<!--     "", "X",  -->
<!--   "Evidence of prior SARS-CoV-2 infection", 'Any of: positive SARS-CoV-2 test; "probable" infection documented in primary care; COVID-19 hospital admission', -->
<!--     "", "X", -->
<!--   "Morbidity count", "0, 1, or 2 or more of the following conditions: diabetes; BMI over 40kg/m^2^; chronic heart disease; chronic kidney disease; chronic liver disease; chronic respiratory disease or severe asthma; chronic neurological disease",  -->
<!--     "", "X" -->
<!-- ) %>% -->
<!-- gt() %>% -->
<!-- tab_spanner( -->
<!--   label = "Matching strategy", -->
<!--   columns = all_of(c("A", "B")) -->
<!-- ) %>% -->
<!-- fmt_markdown(columns = TRUE) -->

<!-- cat("  \n\n") -->
<!-- ``` -->

## Outcomes and follow-up

We considered three effectiveness outcomes. Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection. COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes @emergenc. COVID-19 death, defined as deaths with U07.1 or U07.2 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).

<!-- We also considered the follows safety outcomes: -->

<!-- * Adverse allergic reaction / Anaphylaxis -->

<!-- * Myocarditis or Pericarditis -->

<!-- *  -->

<!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->

Each person was followed from the day of boosting (i.e., time zero) until the earliest of the outcome of interest, death, practice de-registration, 12 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.

## Statistical Analysis

We estimated the cumulative incidence in each treatment group using the Aalen-Johansen estimator. COVID-19 and non-COVID-19 deaths were treated as competing events. Risk differences (RD) and hazard ratios (RRs) comparing the cumulative incidences were derived at 12 weeks. We also estimated hazard ratios (HRs) comparing booster vaccine groups using Cox models.

<!-- 95% confidence limits for the AJ estimates were derived from a Greenwood-type formula. Confidence limits for the risk difference were derived from the sum of squares of the AJ standard errors, derived from Greenwood’s formula. Confidence limits for the survival ratio were based on the standard error of the cumulative hazard. Confidence limits for the risk ratio were derived using the Delta method. -->

<!-- 95% confidence limits for the KM estimates, risk differences, risk ratios, and incidence rate ratios are derived by bootstrapping the matched pairs. 500 bootstrap replicates will be used. -->

<!-- [The primary estimand is the average treatment effect in the recruited population. To estimate this, expected cumulative incidence rates are estimated for each participant assuming treatment was received by everyone in both treatment groups at time zero. These curves are then averaged over all participants to give the marginal cumulative incidence. This is repeated assuming nobody is treated in either group, and their difference taken. Calculating standard errors for these quantities is challenging, whether approached analytically (due to complex variance components) or computationally (via bootstrapping, due to the large sample sizes involved) and are therefore not reported. ] -->

## Subgroup analyses

We estimated comparative effectiveness separately in the following subgroups: primary vaccine course (ChAdOx1-S or BNT162b2); age (18-64 or $\ge$ 65 years); clinical vulnerability, as defined by JCVI (Not clinical at-risk, clinically at-risk, or clinically extremely vulnerable); evidence of prior SARS-CoV-2 infection or not.

## Software, code, and reproducibility

Data management and analyses were conducted in Python version 3.8.10 and R version 4.0.2. All code is shared openly for review and re-use under MIT open license at <https://github.com/opensafely/comparative-booster>. The supplementary materials provide further details of the codelists and data sources used for all variables in the study. Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.

## Disclosure control

To satisfy strict re-identification minimisation requirements for statistical outputs from OpenSAFELY's secure environment, all counts were reported to the nearest $6n-3$ ($3, 9, 15, 21, ...$). Zero counts were unchanged. Similarly, the Aalen-Johansen cumulative incidence estimates were rounded such that each "step" is based on at least 6 events.

\newpage

# Results

## Study population and matching

`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or mRNA-1273 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type=="moderna") %>% pull(pct_all))`) eligible for matching.

<!-- Baseline characteristics of the eligible groups before matching showed... (supplementary Table 1). -->

`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="moderna") %>% pull(pct_step))` respectively (Figure 1).

Characteristics were well-balanced between vaccine types at the start of follow up (Table 2), and the proportion of people with prior clinical conditions was generally similar between the groups, with the standardised mean difference consistently below 0.1 (supplementary Figure 2).

## Estimated comparative effectiveness

### Main analysis

There were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(n.event)` positive SARS-CoV-2 tests, `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths at 12-weeks, across `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 3).

Absolute risks at 12 weeks of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people respectively, representing a risk difference of `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` and a hazard ratio of `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)` (Table 3, Figure 3). Corresponding 12-week risks for COVID-19 hospitalisation were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(coxhrCI)`), and for COVID-19 death were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(coxhrCI)`).

### Subgroup analyses

Whilst absolute risks differed between subgroups, and therefore so did the absolute risk differences comparing vaccine types, subgroup-specific hazard ratios were broadly similar.

A notable exception was outcomes stratified by primary vaccination series: the estimated benefit of mRA-1273 over BNT162b2 was higher amongst those who received BNT162b2 for their first two vaccine doses (HR for positive SARS-CoV-2 `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="vax12_type", subgroup_level=="pfizer-pfizer") %>% pull(coxhrCI)`) compared with ChAdOx1 (HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhrCI)`).

## Discussion

This observational study in over 4 million adults compares the effectiveness of BNT162b2 and mRNA-1273 vaccines against COVID19-related outcomes in the national booster vaccination programme in England. We estimated that mRNA-1273 provided slightly more protection than BNT162b2 did against a positive SARS-CoV-2 test (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` per `r perN_format`) and hospitalisation (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI_95)`) in the first 12 weeks after boosting. For COVID-19 death, the difference was equivocal (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI_95)`) though there were only `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` events across both vaccine types, limiting statistical power.

Around 1 in 10 people had a positive SARS-CoV-2 test in the 12 weeks following booster vaccination reflecting both the high infection rates in the study period, particularly from December 2021 onward, and the imperfect protection booster vaccination offers against infection. However, rates of severe disease were rare: after 12-weeks, fewer than 1 in 1,000 booster recipients were hospitalised with COVID-19, and fewer than 1 in 20,000 died due to COVID-19 disease.

Comparative effectiveness against severe outcomes appeared similar regardless of age, clinically vulnerability, or whether there was evidence of prior infection or not. However, the relative advantage of mRNA-1273 against severe COVID-19 was even higher in people who primary vaccination course was BNT162b2 compared with ChAdOx1. That is, the relative benefit of BNT162b2-BNT162b2-mRNA1273 versus BNT162b2-BNT162b2-BNT162b2 was greater than that of ChAdOx1-ChAdOx1-mRNA1273 versus ChAdOx1-ChAdOx1-BNT162b2, at the point of a booster dose onwards. This suggests a benefit of heterologous boosting.

### Strengths and weaknesses

The OpenSAFELY-TPP database covers around 40% of the English population and contains rich clinical information which enabled us to closely match BNT162b2 and mRNA-1273 recipients to control for potential confounding, to study a range of clinical outcomes, including rare outcomes, and to investigate important clinical subgroups.

To make fair comparisons between the two vaccine types, we exploited the concurrent roll-out of both vaccines across the same eligible population in the same time period. The type of vaccine administered was based largely on local supply and availability, and clinical characteristics did not inform the type of vaccine offered on-site. Thus vaccine allocation was, to some degree, random. It remained important to control for any potential differences in the distribution of confounding characteristics between the two vaccine groups, and these were reasonably balanced after matching. We cannot rule out residual confounding, but crucially many unmeasured confounders, particularly those relating to health-seeking behaviours, are far less problematic in comparative effectiveness studies (compared with studies comparing vaccinated with unvaccinated people) as all people entering the study have sought and received a booster dose. Notably, the apparent benefit of mRNA-1273 versus BNT162b2 against positive SARS-CoV-2 test was less pronounced for matching strategy B versus A, suggesting some additional confounder control for this outcome with closer matching, but estimates were similar across strategies for more severe outcomes where outcome ascertainment is more reliable.

<!--# [People were usually not offered a choice of vaccine type (although they could refuse boosting if they were not offered their preference) and clinical characteristics did not inform the type of vaccine offered. Anecdotally, we understand that occasionally some people were offered a choice in centres using both vaccines, though we are not aware of any specific policy that allowed this.] -->

We excluded certain groups who were unlikely to undergo boosting such as those with a recent positive test and those in palliative care. We also excluded groups where we could not reliably capture high heterogeneity in outcome risk such as those living in care homes and health and social care workers. This may limit the generalisability of out findings.

To control for spatio-temporal heterogeneity in infection risk during the study period we matched on the date of booster receipt and local health administration (STP). However, some residual heterogeneity within STP regions is possible, though more precise geographical matching on middle layer super output areas {<https://www.ons.gov.uk/methodology/geography/ukgeographies/censusgeography>} was not feasible due to poor matching success at this level, partly due to time periods where only one vaccine type was available within a small geographical area.

Some outcomes may be under-ascertained. Positive SARS-CoV-2 tests only include those reported via the national COVID-19 surveillance system (SGSS), and so many asymptomatic and some symptomatic infections will be missed. Positive SARS-CoV-2 testing is therefore not a reliable surrogate for SARS-CoV-2 infection. Some COVID-19 hospitalisations resulting in longer hospital stays may have been missed as these events are only reported once the patient has been discharged.

### Findings in context

The COV-BOOST phase II trial assessed the safety and immunogenicity of seven COVID-19 vaccines for boosting, including BNT162b2 and mRNA-1273 against a menACWY control in people with no prior history of infection @munro2021. It found higher reactogenicity of mRNA-1273 compared with BNT162b2 in subgroups with a two-dose BNT162b2 primary course (geometric mean ratio (GMR) of SARS-CoV-2 anti-spike IgG versus control = 11.49 (95%CI 9.36 to 14.12) versus (6.78 (5.51 to 8.35)) and a two-dose ChAdOx1 primary course (GMR = 32.30 (24.84 to 42.01) versus 16.80 (12.97 to 21.76)). The trial is under-powered to study efficacy endpoints, and we are aware of no other RCTs where direct BNT162b2 versus mRNA-1273 comparisons for boosting are planned or published.

We believe this is the first study to compare BNT162b2 and mRNA-1273 in the context of booster vaccination in observational data, though comparisons have been made for primary vaccination. A study using data from the US Veteran Affairs health care system during the Alpha period using a similar matching approach to the present study @dickerman2022, identified an additional 1.23 (95% CI 0.72 to 1.81) documented infections per 1,000 people in those with a BNT162b2 first dose compared with a mRNA-1273 first dose at 24 weeks, 0.55 (95%CI 0.36 to 0.83) COVID-19 hospitalisations, 0.10 (95%CI 0.00 to 0.26) COVID-19 ICU admissions, and 0.02 (95%CI −0.06 to 0.12) COVID-19 deaths. Another study also using Veterans Affairs data with exact- and propensity-matching in a similar time period @ioannou2022a, unsurprisingly estimated similar results at 24 weeks for documented infection (1.73, 95%CI 1.50 to 1.96 additional events per 1,000 people in those with a BNT162b2 first dose compared with a mRNA-1273), COVID-19 hospitalisation (0.56, 95%CI 0.45 to 0.67) and COVID-19 death (0.03, 95%CI -0.00 to 0.07). Accounting for the longer follow-up period and statistical uncertainty, these estimates are consistent with our findings for comparative effectiveness of boosting.

A study in English data estimated the effectiveness of BNT162b2 boosting versus no boosting and mRNA-1273 boosting versus no boosting separately using a test-negative-control design @andrews2022, suggests marginally higher effectiveness against symptomatic disease in mRNA-1273 than in BNT162b2, but this assumes that the respective control populations in each analysis were similar.

The evidence therefore points strongly to a small benefit of mRNA-1273 over BNT162b2 for primary and subsequent booster doses. Though small, an additional 0.22 hospitalisations per 1,000 people becomes 100s or over 1,000 hospitalisations in England at a population level, depending on booster uptake and infection risk. This carries a substantial additional burden on healthcare resources both during the acute phase of disease and for those with longer-term consequences. However, [safety... better than no boosting....]

### Conclusions

COVID-19-related outcomes after booster vaccination with BNT162b2 or mRNA-1273 were rare, though the risk was slightly higher for BNT162b2 and for mRNA-1273. These finding were broadly consistent across subgroups. These are important differences at a population-level and may help inform vaccine procurement decisions for future booster programmes. Individuals should not defer booster vaccination by BNT162b2 in favour of waiting for mRNA-1273 if offered.

## IG / Funding / COI / etc

Statements to add

## Figures and tables

#### Table 1: inclusion criteria

```{r table-flowchart, echo=FALSE, warning=FALSE, message=FALSE}
match_flowchart %>%
  select(-crit) %>%
   pivot_wider(
    id_cols = c(criteria, vax3_type),
    names_from = vax3_type,
    values_from = c(n, n_exclude, pct_all, pct_step)
  ) %>%
  gt() %>%
  cols_label(
    criteria = "",

    n_pfizer = "N",
    n_moderna = "N",

    n_exclude_pfizer = "N excluded",
    n_exclude_moderna = "N excluded",

    pct_all_pfizer = "% (overall)",
    pct_all_moderna = "% (overall)",
    pct_step_pfizer = "% (incremental)",
    pct_step_moderna = "% (incremental)",
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("pfizer")
  ) %>%
   tab_spanner(
    label = "Moderna mRNA-1273",
    columns = ends_with("moderna")
  ) %>%
  fmt_percent(
    columns = starts_with(c("pct_")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("n_")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "criteria")
  ) %>%
  tab_options(
    table.font.size = 8
  )

cat("  \n\n")
```

#### Table 2: baseline characteristics

```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
table1A <- 
  matchA_table1 %>%
  filter(!is.na(`stat_1`)) %>%
  group_by(variable) %>%
  mutate(
    nlabel=n(),
    #label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
    var_label = if_else(row_number()==1, var_label, ""),
    label = if_else(nlabel!=1, label, ""),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(matchA_table1by$by_col, matchA_table1by$by_chr))

table1B <- 
  matchB_table1 %>%
  filter(!is.na(`stat_1`)) %>%
  group_by(variable) %>%
  mutate(
    nlabel=n(),
    #label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
    var_label = if_else(row_number()==1, var_label, ""),
    label = if_else(nlabel!=1, label, ""),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(matchB_table1by$by_col, matchB_table1by$by_chr))

table1 <- 
  left_join(
    table1A, table1B, by=c("variable", "var_label", "label", "order"),
    suffix=c("_A", "_B")
  )

table1B %>%
  gt() %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
    var_label = "Variable",
    #`BNT162b2_A` = "BNT162b2",
    #`Moderna mRNA-1273_A` = "mRNA-1273",
    `BNT162b2` = "BNT162b2",
    `Moderna mRNA-1273` = "mRNA-1273",
    label=md("Characteristic")
  ) %>%
  # tab_spanner(
  #   label = "Matching strategy A",
  #   columns = ends_with("_A")
  # ) %>%
  # tab_spanner(
  #   label = "Matching strategy B",
  #   columns = ends_with("_B")
  # ) %>%
  cols_align(
    align = c("right"),
    columns = (c("BNT162b2", "Moderna mRNA-1273"))
  ) %>%
  cols_hide(c("variable", "order")) %>%
  tab_options(
    table.font.size = 8
  )

cat("  \n\n")
```

#### Table 3: 12-week risk estimates

Counts and survival estimates are based on values rounded up to the nearest n\*6-3, for disclosure control.

```{r table-contrasts, echo=FALSE, warning=FALSE, message=FALSE}

contrasts_overall_table %>%
  filter(
    outcome %in% outcomes,
    strategy=="B"
  ) %>%
  group_by(outcome_descr) %>%
  select(
    outcome_descr,
    subgroup_descr,
    subgroup_level_descr,

    n.atrisk,

    persontime_0,
    persontime_1,

    n.event_0,
    n.event_1,

    riskCI_0,
    riskCI_1,

    cirdCI,
    #cirrCI,
    #kmirrCI,
    coxhrCI,
  ) %>%
  mutate(
    # add indicator for row shading
    shade = cumsum((lag(subgroup_descr, 1, "")!=subgroup_descr)*1L),
    # remove repeat values of subgroup_level
    subgroup_descr = if_else(lag(subgroup_descr)==subgroup_descr & row_number()!=1, "", as.character(subgroup_descr)),
  ) %>%
  ungroup() %>%
  gt(groupname_col="outcome_descr") %>%
  cols_label(

    outcome_descr = "Outcome",
    subgroup_descr = "Sub-group",
    subgroup_level_descr = "",

    n.atrisk = "N per group",

    persontime_0 = "Person-weeks", 
    persontime_1 = "Person-weeks",

    n.event_0 = "Events",
    n.event_1 = "Events",

    riskCI_0 = glue("12-week risk per {perN_format} people (95% CI)"),
    riskCI_1 = glue("12-week risk per {perN_format} people (95% CI)"),

    cirdCI = glue("Risk difference per {perN_format} people (95% CI)"),
    #cirrCI = "Risk ratio (95% CI)",
    #kmirrCI = "Incidence rate ratio (95% CI)",
    coxhrCI = "Hazard ratio (95% CI)"
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("_0")
  ) %>%
   tab_spanner(
    label = "Moderna",
    columns = ends_with("_1")
  )  %>%
   tab_spanner(
    label = "Comparison (reference=BNT162b2)",
    columns = ends_with("CI")
  ) %>%
  cols_align(
    align = c("right"),
    columns = ends_with(c("atrisk", "_1", "_0", "CI"))
  ) %>%
  tab_options(
    table.font.size = 8
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = shade%%2 != 0
    )
  ) %>%
  cols_hide(c("shade"))

cat("  \n\n")
```

#### Figure 1: Cumulative numbers included in the matched sample

```{r, figure-matchingrates, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))

xmin <- min(match_coverage$vax3_date )
xmax <- max(match_coverage$vax3_date )+1

data_annotate <- tibble(
    vax3_date = date(c("2021-11-26", "2021-11-26", "2022-01-07", "2022-01-07")),
    cumuln = c(1000000, -2900000, 0, -2900000),
    label = c("mRNA-1273", "BNT162b2", "Matched", "Unmatched"),
    colour = c(brewer_pal(type="qual", palette="Set2")(2)[c(2,1)], "darkgrey", "grey"),
    vjust = c(1.2,-0.1,1.2,-0.1)
  )

plot_coverage_cumuln <-
  matchB_coverage %>%
  mutate(
   # treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
    #status_descr = fct_rev(fct_recoderelevel(as.character(status), recoder$status)),
    #status_descr = fct_rev(status_descr),
    cumuln = cumulnB*((treatment*2) - 1)
  ) %>%
  ggplot()+
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=paste0(treatment,status_descr),
      fill=treatment_descr,
      alpha=status_descr,
      colour=NULL
    ),
    position=position_stack(reverse=TRUE),
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
  geom_text(data=data_annotate, aes(x=vax3_date, y=cumuln, label=label), colour=data_annotate$colour, vjust=data_annotate$vjust, hjust=1.05, fontface="bold")+
  scale_x_date(
    breaks = seq(xmin, xmax+1, by=14),
    #breaks = unique(lubridate::ceiling_date(match_coverage$vax3_date, "1 month")),
    limits = c(xmin-1, NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
    sec.axis = sec_axis(
      trans = ~as.Date(.),
      breaks=as.Date(seq(floor_date(xmin, "month"), ceiling_date(xmax, "month"),by="month")),
      labels = scales::label_date("%B %y")
    )
  )+
  scale_y_continuous(
    labels = ~scales::label_number(accuracy = 1, big.mark=",")(abs(.)),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2")+
  scale_colour_brewer(type="qual", palette="Set2")+
  scale_alpha_discrete(range= c(0.8,0.4))+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "none",
    legend.box = "horizontal"
  )+
  NULL

plot_coverage_cumuln

cat("  \n\n")
```

```{r, function-ci, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}

ci_theme <- function(...) {
  theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
      ...
    )
}

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_ci_subgroup <- function(subgroup){

  subgroup0 <- subgroup

  if(subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_ci <- ci_estimates %>%
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    mutate(
      colgroup = interaction(treatment_descr, subgroup_level_descr, sep=", "),
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10)
    ) %>%
    group_by(colgroup, outcome_descr) %>%
    mutate(
      lagrisk=lag(risk,1,0),
    ) %>%
    ggplot(aes(group=colgroup, colour=treatment_descr, fill=treatment_descr, alpha=treatment_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    #geom_step(aes(x=lagtime, y=risk, linetype=treatment_descr))+
    geom_segment(aes(x=lagtime, xend=time, y=risk, yend=risk))+
    geom_segment(aes(x=lagtime, xend=lagtime, y=lagrisk, yend=risk))+
    geom_rect(aes(xmin=lagtime, xmax=time, ymin=risk.ll, ymax=risk.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set2", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set2", guide="none", na.value="grey") +
    #scale_color_manual(values=colourvalues, na.value="grey") +
    #scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number_risk)+
    scale_alpha_discrete(range=c(0.3,0.9))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Days",
      y=NULL,
      title=glue("Cumulative incidence per {perN_format}"),
      colour=NULL,
      alpha=NULL,
      linetype=NULL
    )+
    guides(linetype="none", fill="none", alpha="none")+
    theme_minimal(base_size = 9)+
      theme(
        # legend.position = c(0.05,0.1),
        # legend.justification = c(0,0),
        legend.position = "bottom",
        #legend.direction = "vertical",
        axis.text.x.top=element_text(hjust=0),

        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),
        strip.placement="outside",
        strip.text.y.left = element_text(angle=0),

        panel.border = element_blank(),
        panel.spacing = unit(0.8, "lines"),

      )+
      NULL

  plot_contrasts_data <- contrasts_daily %>%
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    mutate(
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
      lagcird=lag(cird,1,0),
      lagcirr=lag(cirr,1,1)
    )

  plot_cirr <- 
    plot_contrasts_data %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    ggplot(aes(group=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    geom_segment(aes(x=period_start, xend=period_end, y=cirr, yend=cirr))+
    geom_segment(aes(x=period_start, xend=period_start, y=lagcirr, yend=cirr))+
    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=cirr.ll, ymax=cirr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    #scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    #scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Days",
      y=NULL,
      title="Risk ratio",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines")
    )+
    NULL

  plot_coxhr <- contrasts_cuts %>%
    # equivalent to the hazard
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    group_by(outcome_descr, subgroup_level_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    geom_segment(aes(x=period_start, xend=period_end, y=coxhr, yend=coxhr))+
    geom_segment(aes(x=period_start, xend=period_start, y=lag(coxhr), yend=coxhr))+
    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=coxhr.ll, ymax=coxhr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y")+
    # scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    # scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_color_manual(values=colourvalues, na.value="grey") +
    scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
    coord_cartesian(xlim=c(0, NA))+
    colour_guide+
    labs(
      x="Days",
      y="higher risk\n<- BNT162b2 | mRNA 1273 ->",
      title="Hazard ratio",
      colour=NULL
    )+
    theme_minimal(base_size = 9)+
    ci_theme()+
    NULL

  n_subgroups <-
    ci_estimates %>%
    filter(
      subgroup==subgroup0
    ) %>% 
    `[[`("subgroup_level_descr") %>%
    n_distinct()

  plot_combined <- patchwork::wrap_plots(
    plot_ci, 
    #plot_cird, 
    plot_cirr, 
    #plot_kmirr,
    #plot_coxhr,
    ncol=2,
    widths=c(n_subgroups,1)
  )

  ggsave(
    filename=glue("ciB_{subgroup}.png"),
    path=output_dir_rmd,
    plot=plot_combined,
  )

  plot_combined
}

```

#### Figure 2: Estimates of cumulative risk per `r perN_format` people

```{r, figure-ci-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("all")
```

\newpage

#### Figure 3a: 12-week risk difference by subgroups

```{r figure-contrasts-cird, echo=FALSE, warning=FALSE, message=FALSE}

contrasts_overall_plot <- 
  contrasts_overall %>%
  filter(
    outcome %in% outcomes,
    strategy=="B"
  ) %>%
  group_by(outcome_descr) %>%
  mutate(
    outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
    subgroup_level_descr = fct_rev(subgroup_level_descr),
  )

contrasts_overall_plot %>%
  ggplot(aes(y=subgroup_level_descr)) +
  geom_vline(aes(xintercept=0), linetype="dotted", colour="darkgrey")+
  geom_point(aes(x=cird), position=position_dodge(width=-0.3))+
  geom_linerange(aes(xmin=cird.ll, xmax=cird.ul), position=position_dodge(width=-0.3))+
  facet_grid(rows=vars(subgroup_descr), cols=vars(outcome_descr), scales="free", space="free_y", switch="y")+
  scale_x_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, perN, style_negative="minus", big.mark=","))+
  #scale_fill_brewer(type="qual", palette="Set2")+
  #scale_colour_brewer(type="qual", palette="Set2")+
  labs(
    x="12-week risk difference per 1,000 people\n <- favours mRNA-1273  |  favours BNT162b2 ->",
    y=NULL
  )+
  theme_minimal()+
  theme(
    legend.position="bottom",
    axis.text.x.top=element_text(hjust=0),

    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.background = element_blank(),
    strip.placement="outside",
    #strip.text.y.left = element_text(angle=0),
    strip.text.y.left = element_blank(),

    panel.border = element_blank(),
    panel.spacing = unit(0.3, "lines"),
  )

cat("  \n\n")
```

#### Figure 3b: 12-week risk ratio by subgroups

```{r figure-contrasts-cirr, echo=FALSE, warning=FALSE, message=FALSE}

contrasts_overall_plot %>%
  ggplot(aes(y=subgroup_level_descr)) +
  geom_vline(aes(xintercept=1), linetype="dotted", colour="darkgrey")+
  geom_point(aes(x=cirr), position=position_dodge(width=-0.3))+
  geom_linerange(aes(xmin=cirr.ll, xmax=cirr.ul), position=position_dodge(width=-0.3))+

  facet_grid(rows=vars(subgroup_descr), cols=vars(outcome_descr), scales="free", space="free_y", switch="y")+
  scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","), oob=oob_keep)+
  #scale_fill_brewer(type="qual", palette="Set2")+
  #scale_colour_brewer(type="qual", palette="Set2")+
  labs(
    x="12-week risk ratio\n <- favours mRNA-1273  |  favours BNT162b2 ->",
    y=NULL
  )+
  theme_minimal()+
  theme(
    legend.position="bottom",
    axis.text.x.top=element_text(hjust=0),

    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.background = element_blank(),
    strip.placement="outside",
    #strip.text.y.left = element_text(angle=0),
    strip.text.y.left = element_blank(),

    panel.border = element_blank(),
    panel.spacing = unit(0.3, "lines"),
  )

cat("  \n\n")
```

#### Figure 3c: 12-week hazard ratio by subgroups

```{r figure-contrasts-coxhr, echo=FALSE, warning=FALSE, message=FALSE}

contrasts_overall_plot %>%
  ggplot(aes(y=subgroup_level_descr)) +
  geom_vline(aes(xintercept=1), linetype="dotted", colour="darkgrey")+
  geom_point(aes(x=coxhr), position=position_dodge(width=-0.3))+
  geom_linerange(aes(xmin=coxhr.ll, xmax=coxhr.ul), position=position_dodge(width=-0.3))+
  # note this is very similar to using ciirr or kmirr instead of coxhr
  facet_grid(rows=vars(subgroup_descr), cols=vars(outcome_descr), scales="free", space="free_y", switch="y")+
  scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","), oob=oob_keep)+
  #scale_fill_brewer(type="qual", palette="Set2")+
  #scale_colour_brewer(type="qual", palette="Set2")+
  labs(
    x="12-week hazard ratio\n <- favours mRNA-1273  |  favours BNT162b2 ->",
    y=NULL
  )+
  theme_minimal()+
  theme(
    legend.position="bottom",
    axis.text.x.top=element_text(hjust=0),

    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    strip.background = element_blank(),
    strip.placement="outside",
    #strip.text.y.left = element_text(angle=0),
    strip.text.y.left = element_blank(),

    panel.border = element_blank(),
    panel.spacing = unit(0.3, "lines"),
  )

cat("  \n\n")
```

\newpage

## References

### 2022-07-06 write-up/supplement.Rmd
add manuscript draft

---
title: "Supplementary materials for Comparative effectiveness of BNT162b2 versus mRNA-1273 booster doses in England"
output:
  html_document:
    self_contained: yes
  pdf_document:
    toc: no
  bookdown::html_document2:
    number_sections: no
    toc: no
  rtf_document:
    toc: no
  word_document: default
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r setup, include=FALSE, message=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')

# remotes::install_github("https://github.com/wjchulme/osutils")
# library('osutils')

# where are the outputs (ie the inputs for this manuscript!) saved?
#output_dir_os <- here("output", "release-objects")
output_dir_os <- here("released-output", "release-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("write-up", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

## Import custom user functions from lib
source(here("lib", "functions", "utility.R"))

## Import design elements
source(here("lib", "design", "design.R"))

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidadmittedproxy1",
      #"covidcritcare",
      "coviddeath",
      #"noncoviddeath"
    ) %>% 
    set_names(., .)

```

```{r match-A, echo=FALSE, warning=FALSE, message=FALSE}

# 
# prematch_table1 <- read_csv(fs::path(output_dir_os, "prematch", "match_table1.csv"))
# prematch_table1by <- read_csv(fs::path(output_dir_os, "prematch", "match_table1by.csv"))

## matching A

matchA_flowchart <-  read_csv(fs::path(output_dir_os, "A", "match_flowchart.csv"))
matchA_coverage <- read_csv(fs::path(output_dir_os, "A", "match_coverage.csv")) %>%
   mutate(
    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
  )

matchA_table1 <- read_csv(fs::path(output_dir_os, "A", "match_table1.csv"))
matchA_table1by <- read_csv(fs::path(output_dir_os, "A", "match_table1by.csv"))
matchA_smd <-  read_csv(fs::path(output_dir_os, "A", "match_smd.csv"))

ciA_estimates <- read_csv(fs::path(output_dir_os, "A", "ci_estimates.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsA_daily <- read_csv(fs::path(output_dir_os, "A", "ci_contrasts_daily.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsA_cuts <- read_csv(fs::path(output_dir_os, "A", "ci_contrasts_cuts.csv")) %>%
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsA_overall <- read_csv(fs::path(output_dir_os, "A", "ci_contrasts_overall.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, ""),

  )

## matching B

matchB_flowchart <-  read_csv(fs::path(output_dir_os, "B", "match_flowchart.csv"))
matchB_coverage <- read_csv(fs::path(output_dir_os, "B", "match_coverage.csv")) %>%
   mutate(
    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
  )

matchB_table1 <- read_csv(fs::path(output_dir_os, "B", "match_table1.csv"))
matchB_table1by <- read_csv(fs::path(output_dir_os, "B", "match_table1by.csv"))
matchB_smd <-  read_csv(fs::path(output_dir_os, "B", "match_smd.csv"))

ciB_estimates <- read_csv(fs::path(output_dir_os, "B", "ci_estimates.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsB_daily <- read_csv(fs::path(output_dir_os, "B", "ci_contrasts_daily.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsB_cuts <- read_csv(fs::path(output_dir_os, "B", "ci_contrasts_cuts.csv")) %>%
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsB_overall <- read_csv(fs::path(output_dir_os, "B", "ci_contrasts_overall.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, ""),

  )

perN <- 1000
perN_format <- label_number(1, 1, big.mark=",")(perN)

label_number_n <- label_number(1, 1, big.mark=",")
label_number_risk <- label_number(0.01, perN, style_negative="minus", big.mark=",")
label_number_rr <- label_number(0.01, 1, style_negative="minus")

```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Table S1: baseline characteristics of booster vaccine recipients eligible for matching

```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
# table1 <- 
#   prematch_table1 %>%
#   group_by(variable) %>%
#   mutate(
#     label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
#   ) %>%
#   mutate(
#     across(
#       starts_with("stat_"),
#       ~if_else(is.na(.x) & row_number()==1, "", .x)
#     )
#   ) %>%
#   ungroup() %>%
#   mutate(
#     order = cumsum(lag(variable, 1, first(variable))!=variable)
#   ) %>%
#   select(
#     variable, var_label, label, starts_with("stat_"), order
#   ) %>%
#   rename(!!!set_names(match_table1by$by_col, match_table1by$by_chr))
# 
# 
# table1 %>%
#   gt(groupname_col="treatment_descr") %>%
#   fmt_markdown(columns = "label") %>%
#   tab_style(
#     style = list(
#       cell_fill(color = "gray96")
#     ),
#     locations = cells_body(
#       rows = order%%2 == 0
#     )
#   ) %>%
#   tab_style(
#     style = list(
#       cell_text(weight = "bold")
#     ),
#     locations = cells_column_labels(everything())
#   ) %>%
#   cols_label(
#      label=md("Characteristic")
#   ) %>%
#   cols_hide(c("variable", "var_label", "order")) %>%
#   tab_options(
#     table.font.size = 8
#   )
# 
# cat("  \n\n")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S1: Standardised Mean Differences between BNT12b2 and mRNA-1273 groups

<!-- For variables that were matched exactly the SMD is zero, and so are not shown. -->

```{r figure-smd, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

match_smd <-
  bind_rows(
    matchA_smd %>% mutate(strategy="A"), 
    matchB_smd %>% mutate(strategy="B"), 
  )

plot_smd <-
  match_smd %>%
  filter(
    !(var_label %in% c("Primary vaccine course", "Region", "Prior documented SARS-CoV-2 infection", "JCVI clinical risk group", "IMD"))
  ) %>%
  mutate(
    level=fct_rev(fct_inorder(level))
  ) %>%
  droplevels() %>%
  group_by(strategy) %>%
  mutate(
    variable = fct_inorder(variable),
    variable_card = as.numeric(variable)%%2,
    variable_levels = replace_na(as.character(variable_levels), ""),
  ) %>%
  arrange(strategy, variable) %>%
  mutate(
    level = fct_rev(fct_inorder(str_replace(paste(variable, variable_levels, sep=": "),  "\\:\\s$", ""))),
    cardn = row_number()
  ) %>%
  ggplot()+
  geom_point(aes(x=smd, y=level, group=strategy, colour=strategy), position=position_dodge(width=0.3))+
  geom_linerange(aes(xmax=smd, xmin=0, y=level, group=strategy, colour=strategy), position=position_dodge(width=0.3))+
  geom_rect(aes(alpha = variable_card, ymin = rev(cardn)-0.5, ymax =rev(cardn+0.5), group=strategy), xmin = -Inf, xmax = Inf, fill='grey', colour="transparent") +
  scale_alpha_continuous(range=c(0,0.1), guide=FALSE)+
  scale_colour_brewer(type="qual", palette="Set2")+
  labs(
    x="Standardised mean difference\n <- proportion higher in BNT162b2 | proportion higher in mRNA-1273 -> ",
    y=NULL,
    alpha=NULL,
    colour= "Matching strategy"
  )+
  theme_minimal(base_size=10) +
  theme(
    legend.position = "bottom",
    strip.placement = "outside",
    strip.background = element_rect(fill="transparent", colour="transparent"),
    strip.text.y.left = element_text(angle = 0, hjust=1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.spacing = unit(0, "lines")
  )

plot_smd

```

\newpage

## Comparative effectiveness estimates, matching strategy A

### Cumulative incidence

```{r, function-ci, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}

ci_theme <- function(...) {
  theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
      ...
    )
}

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_ci_subgroup <- function(subgroup){

  subgroup0 <- subgroup

  if(subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_ci <- ciB_estimates %>%
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    mutate(
      colgroup = interaction(treatment_descr, subgroup_level_descr, sep=", "),
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10)
    ) %>%
    #group_by(colgroup, outcome_descr) %>%
    ggplot(aes(group=colgroup, colour=treatment_descr, fill=treatment_descr, alpha=treatment_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    geom_step(aes(x=lagtime, y=risk, linetype=treatment_descr))+
    #geom_segment(aes(x=lagtime, xend=time, y=risk, yend=risk))+
    #geom_segment(aes(x=lagtime, xend=lagtime, y=lag(risk,1,0), yend=risk))+
    geom_rect(aes(xmin=lagtime, xmax=time, ymin=risk.ll, ymax=risk.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    #scale_color_manual(values=colourvalues, na.value="grey") +
    #scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number_risk)+
    scale_alpha_discrete(range=c(0.3,0.9))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Days",
      y=NULL,
      #title=glue("Cumulative incidence per {perN_format}"),
      colour=NULL,
      alpha=NULL,
      linetype=NULL
    )+
    guides(linetype="none", fill="none", alpha="none")+
    theme_minimal(base_size = 9)+
      theme(
        # legend.position = c(0.05,0.1),
        # legend.justification = c(0,0),
        legend.position = "bottom",
        #legend.direction = "vertical",
        axis.text.x.top=element_text(hjust=0),

        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),
        strip.placement="outside",
        strip.text.y.left = element_text(angle=0),

        panel.border = element_blank(),
        panel.spacing = unit(0.8, "lines"),

      )+
      NULL

  plot_coxhr <- contrastsB_cuts %>%
    # equivalent to the hazard
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    group_by(outcome_descr, subgroup_level_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    geom_segment(aes(x=period_start, xend=period_end, y=coxhr, yend=coxhr))+
    geom_segment(aes(x=period_start, xend=period_start, y=lag(coxhr), yend=coxhr))+
    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=coxhr.ll, ymax=coxhr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y")+
    # scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    # scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_color_manual(values=colourvalues, na.value="grey") +
    scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
    coord_cartesian(xlim=c(0, NA))+
    colour_guide+
    labs(
      x="Days",
      y="higher risk\n<- BNT162b2 | mRNA 1273 ->",
      title="Hazard ratio",
      colour=NULL
    )+
    theme_minimal(base_size = 9)+
    ci_theme()+
    NULL

  n_subgroups <-
    ciB_estimates %>%
    filter(
      subgroup==subgroup0
    ) %>% 
    `[[`("subgroup_level_descr") %>%
    n_distinct()

  plot_combined <- patchwork::wrap_plots(
    plot_ci, 
    #plot_cird, 
    #plot_cirr, 
    #plot_kmirr,
    plot_coxhr,
    ncol=2,
    widths=c(n_subgroups,1)
  )

  ggsave(
    filename=glue("ciB_{subgroup}.png"),
    path=output_dir_rmd,
    plot=plot_ci,
  )

  plot_ci
}

```

#### Figure S2a: Cumulative incidence estimates per `r perN_format` people

```{r, figure-ci-A-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("all")
```

#### Figure S2b: Cumulative incidence estimates per `r perN_format` people by primary course type

```{r, figure-ci-A-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("vax12_type")
```

#### Figure S2c: Cumulative incidence estimates per `r perN_format` people by age group

```{r, figure-ci-A-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("age65plus")
```

#### Figure S2d: Cumulative incidence estimates per `r perN_format` people by prior infection status

```{r, figure-ci-A-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("prior_covid_infection")
```

#### Figure S2e: Cumulative incidence estimates per `r perN_format` people by clinical vulnerability

```{r, figure-ci-A-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("cev_cv")
```

```{r, function-plot-contrasts, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_contrast_subgroup <- function(subgroup, matchset){

  subgroup0 <- subgroup

  if(subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_data <- get(glue("contrasts{matchset}_daily")) %>%
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    mutate(
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
      lagcird=lag(cird,1,0),
      lagcirr=lag(cirr,1,1)
    )

  plot_cird <- 
    plot_data %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    geom_segment(aes(x=period_start, xend=period_end, y=cird, yend=cird))+
    geom_segment(aes(x=period_start, xend=period_start, y=lagcird, yend=cird))+
    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=cird.ll, ymax=cird.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, 10000))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Days",
      y=NULL,
      title="Risk difference",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      # legend.position = c(0.05,0.1),
      # legend.justification = c(0,0),
      legend.position = "bottom",
      #legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement="outside",
      strip.text.y.left = element_text(angle=0),

      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
    )+
    NULL

  plot_cirr <- 
    plot_data %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    geom_segment(aes(x=period_start, xend=period_end, y=cirr, yend=cirr))+
    geom_segment(aes(x=period_start, xend=period_start, y=lagcirr, yend=cirr))+
    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=cirr.ll, ymax=cirr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Days",
      y=NULL,
      title="Risk ratio",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines")
    )+
    NULL

  plot_combined <- patchwork::wrap_plots(
    plot_cird, 
    plot_cirr, 
    ncol=2
  )

  ggsave(
    filename=glue("contrasts{matchset}_{subgroup}.png"),
    path=output_dir_rmd,
    plot=plot_combined,
  )

  plot_combined
}

```

### Cumulative contrasts

#### Figure S3a: Cumulative risk difference and risk ratio for matching strategy A

```{r, figure-contrasts-A-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("all","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S3b: Cumulative risk difference and risk ratio for matching strategy A by primary course type

```{r, figure-contrasts-A-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("vax12_type","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S3c: Cumulative risk difference and risk ratio for matching strategy A by age group

```{r, figure-contrasts-A-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("age65plus","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S2d: Cumulative risk difference and risk ratio for matching strategy A by prior infection status

```{r, figure-contrasts-A-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("prior_covid_infection","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S3e: Cumulative risk difference and risk ratio for matching strategy A by clinical vulnerability

```{r, figure-contrasts-A-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("cev_cv","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

## Comparative effectiveness estimates, matching strategy B

### Cumulative incidence

#### Figure S4a: Cumulative incidence estimates per `r perN_format` people for matching strategy B

```{r, figure-ci-B-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("all")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S4b: Cumulative incidence estimates per `r perN_format` people for matching strategy B by primary course type

```{r, figure-ci-B-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("vax12_type")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S4c: Cumulative incidence estimates per `r perN_format` people for matching strategy B by age group

```{r, figure-ci-B-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("age65plus")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S4d: Cumulative incidence estimates per `r perN_format` people for matching strategy B by prior infection status

```{r, figure-ci-B-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("prior_covid_infection")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S4e: Cumulative incidence estimates per `r perN_format` people for matching strategy B by clinical vulnerability

```{r, figure-ci-B-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("cev_cv")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

### Cumulative contrasts

#### Figure S5a: Cumulative risk difference and risk ratio for matching strategy B

```{r, figure-contrasts-B-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("all","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S5b: Cumulative risk difference and risk ratio for matching strategy B by primary course type

```{r, figure-contrasts-B-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("vax12_type","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S5c: Cumulative risk difference and risk ratio for matching strategy B by age group

```{r, figure-contrasts-B-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("age65plus","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S5d: Cumulative risk difference and risk ratio for matching strategy B by prior infection status

```{r, figure-contrasts-B-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("prior_covid_infection","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S5e: Cumulative risk difference and risk ratio for matching strategy B by clinical vulnerability

```{r, figure-contrasts-B-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("cev_cv","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

### 2022-07-06 analysis/study_definition.py
correct housebound definition (#11)
* correct housebound definition

This bug fix ensures that we no longer count people as being not housebound if they stopped being housebound at any point in time, _even_ if its after the baseline date.  h/t @evansd

* this is what code reviews are for

h/t @HelenCEBM

@@ -897,11 +897,11 @@ study = StudyDefinition(
     ),   
     no_longer_housebound=patients.with_these_clinical_events( 
       codelists.no_longer_housebound, 
-      on_or_after="housebound_date",
+      between=["housebound_date", "covid_vax_disease_3_date - 1 day"],
     ),
     moved_into_care_home=patients.with_these_clinical_events(
       codelists.carehome,
-      on_or_after="housebound_date",
+      between=["housebound_date", "covid_vax_disease_3_date - 1 day"],
     ),
   ),
   

### 2022-07-06 write-up/manuscript.Rmd
rmds

--- 2022-07-06 write-up/manuscriptv2.Rmd
@@
       "postest",
-      #"covidemergency",
+      "covidemergency",
       "covidadmitted",
@@
 
-matchset <- "B"
+matchset <- "A"
 
@@
 
-*Results* `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(n))` people were matched in each group, contributing a total of `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up.
-The 12-week risk per `r perN_format` people of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 was `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`) respectively.
+*Results* `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` people were matched in each group, contributing a total of `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up.
+The 12-week risk per `r perN_format` people of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 was `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` (RD `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)`; HR `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`) respectively.
 
@@
 
-The UK COVID-19 vaccination programme delivered its first "booster" doses in September 2021 @NHSEnglandNHS.
-Following guidance from the Joint Committee for Vaccination and Immunisation (JCVI) expert advisory group@covid19 boosting followed a staged roll-out that prioritised groups at high risk of infection or severe disease, and was then progressively extended to the whole adult population by mid December 2021 @NHSEnglandNHSb @NHSEnglandNHSa1 @NHSEnglandNHSa2.
-The BNT162b2 Pfizer-BioNTech vaccine was used initially for boosting, with a half-dose of Moderna mRNA-1273 subsequently also used.
+The UK COVID-19 vaccination programme delivered its first "booster" doses in September 2021, initially in groups at high risk of severe disease then across the adult population.
+A half-dose of the BNT162b2 Pfizer-BioNTech vaccine was used initially, with Moderna mRNA-1273 subsequently also used.
 The concurrent administration of the BNT162b2 and mRNA-1273 vaccines for boosting in England, receipt of which was largely determined by local availability rather than any clinical criteria, enables a direct comparison of their effectiveness.
-No randomised trials have made such a comparison for efficacy outcomes.
-
-On behalf of NHS England, we used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to compare the effectiveness of BNT162b2 and mRNA-1273 boosting in adult recipients.
-
-\newpage
+No randomised trials have made such a comparison.
+
+We used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to compare the effectiveness of BNT162b2 versus mRNA-1273 in boosted adults between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`, during which time both vaccine brands were being used.
 
@@
 
-We considered all adults (aged 18 years and over) who received a booster dose of BNT162b2 or mRNA-1273 between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")` inclusive, during which time both vaccine brands were being used.
-People were eligible if they: were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and mRNA-1273 were not considered due to small numbers); were not a health or social care worker, not resident in a care or nursing home and not medically housebound or receiving end-of-life care; had no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; were not hospitalised at the time of boosting; and had complete information on sex, deprivation, and NHS region (defined using Sustainability and Transformation Partnerships (STP, a geographical grouping of NHS and local authorities).
+We considered all adults aged $\ge$ 18 years who received a booster dose of BNT162b2 or mRNA-1273 between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")` inclusive (the "recruitment period").
+People were eligible if they: were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and mRNA-1273 were not considered due to small numbers); were not a health or social care worker, not resident in a care or nursing home and not medically housebound or receiving end-of-life care; had no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; were not undergoing an unplanned hospital admission; and had complete information on sex, ethnicity, deprivation, and NHS region.
 
@@
 
-BNT162b2 booster recipients were matched with mRNA-1273 booster recipients in a one to one ratio without replacement, on the following characteristics: date of booster dose; primary vaccine course (BNT162b2 or ChadOx1-S); Date of second vaccine dose (7 day caliper); sex (male or female); age (3 year caliper and within age groups defined by JVCI risk groups); Clinical risk group defined by JCVI (clinically extremely vulnerable, clinically at-risk, neither); Index of Multiple Deprivation (IMD, grouped by quintile); STP as a surrogate for geographical region; evidence of prior SARS-CoV-2 infection (any of positive SARS-CoV-2 test, probable infection documented in primary care, or COVID-19 hospital attendance or admission); morbidity count (0, 1, or 2 or more of diabetes, BMI over 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease).
-The supplementary materials provide more information on how these characteristics were defined.
-
-<!-- #### Matching strategies -->
-
-<!-- ```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE} -->
-
-<!-- tribble( -->
-<!--   ~Factor, ~Description, ~A, ~B, -->
-
-<!--   "Date of booster dose",  "Matched exactly",   -->
-<!--     "X", "X", -->
-<!--   "Primary course vaccine brand", "BNT162b2; ChadOx1-S",  -->
-<!--     "X", "X", -->
-<!--   "Date of second vaccine dose", "7 day caliper",  -->
-<!--     "X", "X", -->
-<!--   "Age",  "3 year caliper",  -->
-<!--     "X", "X", -->
-<!--   "JCVI age group", "18-39; 40-49; 50-54; 55-59; 60-64; 65-69; 70-75; 75-79; 80 and over", -->
-<!--     "X", "X", -->
-<!--   "JCVI Clinical risk group", "clinically extremely vulnerable; clinically at-risk; neither", -->
-<!--     "X", "X", -->
-<!--   "English Index of Multiple Deprivation", "Grouped by quintile", -->
-<!--     "X", "X", -->
-<!--   "NHS region", "East of England; Midlands; London; North East and Yorkshire; North West; South East; South West", -->
-<!--     "X", "", -->
-<!--   "Sustainability and Transformation Partnership (STP)", "a geographical grouping of NHS and local authorities", -->
-<!--     "", "X", -->
-<!--   "Sex", "Male; Female", -->
-<!--     "", "X",  -->
-<!--   "Evidence of prior SARS-CoV-2 infection", 'Any of: positive SARS-CoV-2 test; "probable" infection documented in primary care; COVID-19 hospital admission', -->
-<!--     "", "X", -->
-<!--   "Morbidity count", "0, 1, or 2 or more of the following conditions: diabetes; BMI over 40kg/m^2^; chronic heart disease; chronic kidney disease; chronic liver disease; chronic respiratory disease or severe asthma; chronic neurological disease",  -->
-<!--     "", "X" -->
-<!-- ) %>% -->
-<!-- gt() %>% -->
-<!-- tab_spanner( -->
-<!--   label = "Matching strategy", -->
-<!--   columns = all_of(c("A", "B")) -->
-<!-- ) %>% -->
-<!-- fmt_markdown(columns = TRUE) -->
-
-<!-- cat("  \n\n") -->
-<!-- ``` -->
+BNT162b2 booster recipients were matched 1:1 with mRNA-1273 booster recipients without replacement.
+Two different matching strategies were used: one simpler strategy resulting in more matches but with greater imbalance between booster vaccine groups, and another more complex strategy resulting in fewer matches but with less imbalance.
+
+#### Matching strategies
+
+```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE}
+
+tribble(
+  ~Factor, ~Description, ~A, ~B,
+
+  "Date of booster dose",  "Matched exactly",  
+    "X", "X",
+  "Primary course vaccine brand", "BNT162b2; ChadOx1-S", 
+    "X", "X",
+  "Date of second vaccine dose", "7 day caliper", 
+    "X", "X",
+  "Age",  "3 year caliper", 
+    "X", "X",
+  "JCVI age group", "18-39; 40-49; 50-54; 55-59; 60-64; 65-69; 70-75; 75-79; 80 and over",
+    "X", "X",
+  "JCVI Clinical risk group", "clinically extremely vulnerable; clinically at-risk; neither",
+    "X", "X",
+  "English Index of Multiple Deprivation", "Grouped by quintile",
+    "X", "X",
+  "NHS region", "East of England; Midlands; London; North East and Yorkshire; North West; South East; South West",
+    "X", "",
+  "Sustainability and Transformation Partnership (STP)", "a geographical grouping of NHS and local authorities",
+    "", "X",
+  "Sex", "Male; Female",
+    "", "X", 
+  "Evidence of prior SARS-CoV-2 infection", 'Any of: positive SARS-CoV-2 test; "probable" infection documented in primary care; COVID-19 hospital admission',
+    "", "X",
+  "Morbidity count", "0, 1, or 2 or more of the following conditions: diabetes; BMI over 40kg/m^2^; chronic heart disease; chronic kidney disease; chronic liver disease; chronic respiratory disease or severe asthma; chronic neurological disease",
+    "", "X"
+) %>%
+gt() %>%
+tab_spanner(
+  label = "Matching strategy",
+  columns = all_of(c("A", "B"))
+) %>%
+fmt_markdown(columns = TRUE)
+
+cat("  \n\n")
+```
+
+Morbidity count included the following conditions: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease"
 
@@
 
-We considered three effectiveness outcomes.
+We considered four effectiveness outcomes.
 Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date.
 Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection.
-COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes @emergenc.
-COVID-19 death, defined as deaths with U07.1 or U07.2 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
+COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
+COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes.
+COVID-19 death.
+Deaths were classified as from COVID-19 if deaths with the U07.1 or U07.2 ICD-10 codes were mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
 
@@
 
-Each person was followed from the day of boosting (i.e., time zero) until the earliest of the outcome of interest, death, practice de-registration, 12 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.
+Each person was followed from the day of boosting (i.e., time zero) until the earliest of outcome, death, practice de-registration, 12 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.
 
@@
 
-We estimated the cumulative incidence in each treatment group using the Aalen-Johansen estimator.
-COVID-19 and non-COVID-19 deaths were treated as competing events.
+We estimated the cumulative incidence in each treatment group using the Aalen-Johansen estimator, with COVID-19 and non-COVID-19 treated as competing events (when these were not the event of interest).
 Risk differences (RD) and hazard ratios (RRs) comparing the cumulative incidences were derived at 12 weeks.
@@
 
-We estimated comparative effectiveness separately in the following subgroups: primary vaccine course (ChAdOx1-S or BNT162b2); age (18-64 or $\ge$ 65 years); clinical vulnerability, as defined by JCVI (Not clinical at-risk, clinically at-risk, or clinically extremely vulnerable); evidence of prior SARS-CoV-2 infection or not.
+We estimated comparative effectiveness separately in the following subgroups: primary vaccine course (ChAdOx1-S or BNT162b2); age (18-64 or $\ge$ 65 years); clinical vulnerability, as defined by the UK Joint Committee for Vaccination and immunisation (JCVI) (Not clinical at-risk, clinically at-risk, or clinically extremely vulnerable); evidence of prior SARS-CoV-2 infection or not.
 
@@
 
-To satisfy strict re-identification minimisation requirements for statistical outputs from OpenSAFELY's secure environment, all counts were reported to the nearest $6n-3$ ($3, 9, 15, 21, ...$).
-Zero counts were unchanged.
-Similarly, the Aalen-Johansen cumulative incidence estimates were rounded such that each "step" is based on at least 6 events.
-
-\newpage
+To reduce re-identification risk, counts are reported to the nearest $6n-3$ ($3, 9, 15, 21, ...$).
+Event times were deferred until there were at least 6 events, so that each "step" in the Aalen-Johansen cumulative incidence estimates is based on at least 6 events.
+An offset of -3 is then applied to de-bias this rounding.
 
@@
 
-`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="moderna") %>% pull(pct_step))` respectively (Figure 1).
+For matching strategy A `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7A", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7A", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7A", vax3_type=="moderna") %>% pull(pct_step))` respectively.
+For matching strategy B `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7B", vax3_type=="moderna") %>% pull(pct_step))` respectively (Figure 1).
 
@@
 
-There were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(n.event)` positive SARS-CoV-2 tests, `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths at 12-weeks, across `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 3).
-
-Absolute risks at 12 weeks of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people respectively, representing a risk difference of `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` and a hazard ratio of `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)` (Table 3, Figure 3).
-Corresponding 12-week risks for COVID-19 hospitalisation were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(coxhrCI)`), and for COVID-19 death were `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(coxhrCI)`).
+For matching strategy A, there were `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(n.event)` positive SARS-CoV-2 tests, `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths at 12-weeks, across `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 3).
+
+Absolute risks at 12 weeks of positive SARS-CoV-2 test for those receiving BNT162b2 and mRNA-1273 were `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people respectively, representing a risk difference of `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` and a hazard ratio of `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)` (Table 3, Figure 3).
+Corresponding 12-week risks for COVID-19 hospitalisation were `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(coxhrCI)`), and for COVID-19 death were `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(coxhrCI)`).
+
+For severe outcomes, RDs and HRs were similar for matching strategy B where treatment groups were more closely matched.
+The risk of positive SARS-CoV-2 test was more similar between BNT162b2 and mRNA-1273 when estimated under matching strategy B, with `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` events per `r perN_format` people respectively, (RD `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)`; HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`).
 
@@
 
-A notable exception was outcomes stratified by primary vaccination series: the estimated benefit of mRA-1273 over BNT162b2 was higher amongst those who received BNT162b2 for their first two vaccine doses (HR for positive SARS-CoV-2 `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="vax12_type", subgroup_level=="pfizer-pfizer") %>% pull(coxhrCI)`) compared with ChAdOx1 (HR `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhrCI)`).
+A notable exception was outcomes stratified by primary vaccination series: the estimated benefit of mRA-1273 over BNT162b2 was higher amongst those who received BNT162b2 for their first two vaccine doses (HR for positive SARS-CoV-2 `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="vax12_type", subgroup_level=="pfizer-pfizer") %>% pull(coxhrCI)`) compared with ChAdOx1 (HR `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhrCI)`).
 
@@
 This observational study in over 4 million adults compares the effectiveness of BNT162b2 and mRNA-1273 vaccines against COVID19-related outcomes in the national booster vaccination programme in England.
-We estimated that mRNA-1273 provided slightly more protection than BNT162b2 did against a positive SARS-CoV-2 test (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` per `r perN_format`) and hospitalisation (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI_95)`) in the first 12 weeks after boosting.
-For COVID-19 death, the difference was equivocal (RD = `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI_95)`) though there were only `r contrasts_overall_table %>% filter(strategy=="B", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` events across both vaccine types, limiting statistical power.
+We estimated that mRNA-1273 provided slightly more protection than BNT162b2 did against a positive SARS-CoV-2 test (RD = `r contrasts_overall_table %>% filter(strategy=="A", outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` per `r perN_format`) and hospitalisation (RD = `r contrasts_overall_table %>% filter(strategy=="A", outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI_95)`) in the first 12 weeks after boosting.
+For COVID-19 death, the difference was equivocal (RD = `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI_95)`) though there were only `r contrasts_overall_table %>% filter(strategy=="A", outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` events across both vaccine types, limiting statistical power.
 
@@
   matchA_table1 %>%
-  filter(!is.na(`stat_1`)) %>%
   group_by(variable) %>%
   mutate(
-    nlabel=n(),
-    #label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
-    var_label = if_else(row_number()==1, var_label, ""),
-    label = if_else(nlabel!=1, label, ""),
+    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
   ) %>%
@@
   matchB_table1 %>%
-  filter(!is.na(`stat_1`)) %>%
   group_by(variable) %>%
   mutate(
-    nlabel=n(),
-    #label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
-    var_label = if_else(row_number()==1, var_label, ""),
-    label = if_else(nlabel!=1, label, ""),
+    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
   ) %>%
@@
 
-table1B %>%
+table1 %>%
   gt() %>%
@@
   cols_label(
-    var_label = "Variable",
-    #`BNT162b2_A` = "BNT162b2",
-    #`Moderna mRNA-1273_A` = "mRNA-1273",
-    `BNT162b2` = "BNT162b2",
-    `Moderna mRNA-1273` = "mRNA-1273",
+    `BNT162b2_A` = "BNT162b2",
+    `Moderna mRNA-1273_A` = "mRNA-1273",
+    `BNT162b2_B` = "BNT162b2",
+    `Moderna mRNA-1273_B` = "mRNA-1273",
     label=md("Characteristic")
   ) %>%
-  # tab_spanner(
-  #   label = "Matching strategy A",
-  #   columns = ends_with("_A")
-  # ) %>%
-  # tab_spanner(
-  #   label = "Matching strategy B",
-  #   columns = ends_with("_B")
-  # ) %>%
+  tab_spanner(
+    label = "Matching strategy A",
+    columns = ends_with("_A")
+  ) %>%
+  tab_spanner(
+    label = "Matching strategy B",
+    columns = ends_with("_B")
+  ) %>%
   cols_align(
     align = c("right"),
-    columns = (c("BNT162b2", "Moderna mRNA-1273"))
-  ) %>%
-  cols_hide(c("variable", "order")) %>%
+    columns = ends_with(c("_A", "_B"))
+  ) %>%
+  cols_hide(c("variable", "var_label", "order")) %>%
   tab_options(
@@
     outcome %in% outcomes,
-    strategy=="B"
+    strategy=="A"
   ) %>%
@@
 
-data_annotate <- tibble(
-    vax3_date = date(c("2021-11-26", "2021-11-26", "2022-01-07", "2022-01-07")),
-    cumuln = c(1000000, -2900000, 0, -2900000),
-    label = c("mRNA-1273", "BNT162b2", "Matched", "Unmatched"),
-    colour = c(brewer_pal(type="qual", palette="Set2")(2)[c(2,1)], "darkgrey", "grey"),
-    vjust = c(1.2,-0.1,1.2,-0.1)
-  )
-
 plot_coverage_cumuln <-
-  matchB_coverage %>%
+  match_coverage %>%
   mutate(
@@
     #status_descr = fct_rev(fct_recoderelevel(as.character(status), recoder$status)),
-    #status_descr = fct_rev(status_descr),
-    cumuln = cumulnB*((treatment*2) - 1)
+    status_descr = fct_rev(status_descr),
+    cumuln = cumuln*((treatment*2) - 1)
   ) %>%
@@
   geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
-  geom_text(data=data_annotate, aes(x=vax3_date, y=cumuln, label=label), colour=data_annotate$colour, vjust=data_annotate$vjust, hjust=1.05, fontface="bold")+
   scale_x_date(
@@
     axis.ticks.x=element_line(),
-    legend.position = "none",
-    legend.box = "horizontal"
+    legend.position = "bottom",
+    legend.box = "vertical"
   )+
@@
     mutate(
-      lagrisk=lag(risk,1,0),
+      lagrisk=lag(risk,1,0)
     ) %>%
@@
     facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), scales="free_y", switch="y")+
-    scale_color_brewer(type="qual", palette="Set2", na.value="grey") +
-    scale_fill_brewer(type="qual", palette="Set2", guide="none", na.value="grey") +
+    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
     #scale_color_manual(values=colourvalues, na.value="grey") +
@@
       y=NULL,
-      title=glue("Cumulative incidence per {perN_format}"),
+      #title=glue("Cumulative incidence per {perN_format}"),
       colour=NULL,
@@
 
-      )+
-      NULL
-
-  plot_contrasts_data <- contrasts_daily %>%
-    filter(
-      subgroup==subgroup0,
-      outcome %in% outcomes
-    ) %>%
-    group_by(subgroup_level_descr, outcome_descr) %>%
-    mutate(
-      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
-      lagcird=lag(cird,1,0),
-      lagcirr=lag(cirr,1,1)
-    )
-
-  plot_cirr <- 
-    plot_contrasts_data %>%
-    group_by(subgroup_level_descr, outcome_descr) %>%
-    ggplot(aes(group=subgroup_level_descr)) +
-    geom_hline(aes(yintercept=1), colour="black")+
-    geom_segment(aes(x=period_start, xend=period_end, y=cirr, yend=cirr))+
-    geom_segment(aes(x=period_start, xend=period_start, y=lagcirr, yend=cirr))+
-    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=cirr.ll, ymax=cirr.ul), alpha=0.1, colour="transparent")+
-    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
-    #scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
-    #scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
-    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
-    coord_cartesian(xlim=c(0, NA))+
-    labs(
-      x="Days",
-      y=NULL,
-      title="Risk ratio",
-      colour=NULL
-    )+
-    colour_guide+
-    theme_minimal(base_size = 9)+
-    theme(
-      legend.position="none",
-      axis.text.x.top=element_text(hjust=0),
-      axis.line.y = element_line(colour = "black"),
-      axis.line.x.top = element_line(colour = "black"),
-
-      panel.grid.minor.x = element_blank(),
-      panel.grid.minor.y = element_blank(),
-      strip.background = element_blank(),
-      strip.text = element_blank(),
-      strip.placement="outside",
-      panel.border = element_blank(),
-      panel.spacing = unit(0.8, "lines")
       )+
@@
     #plot_cird, 
-    plot_cirr, 
+    #plot_cirr, 
     #plot_kmirr,
-    #plot_coxhr,
+    plot_coxhr,
     ncol=2,
@@
   ggsave(
-    filename=glue("ciB_{subgroup}.png"),
+    filename=glue("ciA_{subgroup}.png"),
     path=output_dir_rmd,
-    plot=plot_combined,
-  )
-
-  plot_combined
+    plot=plot_ci,
+  )
+
+  plot_ci
 }
@@
   contrasts_overall %>%
-  filter(
-    outcome %in% outcomes,
-    strategy=="B"
-  ) %>%
+  filter(outcome %in% outcomes) %>%
   group_by(outcome_descr) %>%
@@
 contrasts_overall_plot %>%
-  ggplot(aes(y=subgroup_level_descr)) +
+  ggplot(aes(y=subgroup_level_descr, colour=strategy)) +
   geom_vline(aes(xintercept=0), linetype="dotted", colour="darkgrey")+
@@
   scale_x_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, perN, style_negative="minus", big.mark=","))+
-  #scale_fill_brewer(type="qual", palette="Set2")+
-  #scale_colour_brewer(type="qual", palette="Set2")+
+  scale_fill_brewer(type="qual", palette="Set2")+
+  scale_colour_brewer(type="qual", palette="Set2")+
   labs(
     x="12-week risk difference per 1,000 people\n <- favours mRNA-1273 | favours BNT162b2 ->",
-    y=NULL
+    y=NULL,
+    colour="Matching strategy"
   )+
@@
 contrasts_overall_plot %>%
-  ggplot(aes(y=subgroup_level_descr)) +
+  ggplot(aes(y=subgroup_level_descr, colour=strategy)) +
   geom_vline(aes(xintercept=1), linetype="dotted", colour="darkgrey")+
@@
   scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","), oob=oob_keep)+
-  #scale_fill_brewer(type="qual", palette="Set2")+
-  #scale_colour_brewer(type="qual", palette="Set2")+
+  scale_fill_brewer(type="qual", palette="Set2")+
+  scale_colour_brewer(type="qual", palette="Set2")+
   labs(
     x="12-week risk ratio\n <- favours mRNA-1273 | favours BNT162b2 ->",
-    y=NULL
+    y=NULL,
+    colour="Matching strategy"
   )+
@@
   scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","), oob=oob_keep)+
-  #scale_fill_brewer(type="qual", palette="Set2")+
-  #scale_colour_brewer(type="qual", palette="Set2")+
+  scale_fill_brewer(type="qual", palette="Set2")+
+  scale_colour_brewer(type="qual", palette="Set2")+
   labs(
     x="12-week hazard ratio\n <- favours mRNA-1273 | favours BNT162b2 ->",
-    y=NULL
+    y=NULL,
+    colour="Matching strategy"
   )+

### 2022-07-06 write-up/report.Rmd
rmds

@@
 ---
-title: "Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: preliminary report"
+title: "Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: a cohort study in OpenSAFELY-TPP"
 output:
-  html_document:
-    self_contained: yes
+  word_document: default
   pdf_document:
@@
     toc: no
-  word_document: default
+  html_document:
+    self_contained: yes
 bibliography: references.bib
@@
       "covidadmitted",
-      "covidadmittedproxy1",
-      "covidcritcare",
+      #"covidadmittedproxy1",
+      #"covidcritcare",
       "coviddeath",
@@
 
+```{r match-A, echo=FALSE, warning=FALSE, message=FALSE}
+
+matchset <- "A"
+
+match_flowchart <-  read_csv(fs::path(output_dir_os, matchset, "match_flowchart.csv"))
+match_coverage <- read_csv(fs::path(output_dir_os, matchset, "match_coverage.csv")) %>%
+   mutate(
+    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
+  )
+
+match_table1 <- read_csv(fs::path(output_dir_os, matchset, "match_table1.csv"))
+match_table1by <- read_csv(fs::path(output_dir_os, matchset, "match_table1by.csv"))
+match_smd <-  read_csv(fs::path(output_dir_os, matchset, "match_smd.csv"))
+
+ci_estimates <- read_csv(fs::path(output_dir_os, matchset, "ci_estimates.csv")) %>% 
+  mutate(
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
+    subgroup_level_descr = replace_na(subgroup_level_descr, "")
+  )
+contrasts_daily <- read_csv(fs::path(output_dir_os, matchset, "ci_contrasts_daily.csv")) %>% 
+  mutate(
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
+    subgroup_level_descr = replace_na(subgroup_level_descr, "")
+  )
+contrasts_cuts <- read_csv(fs::path(output_dir_os, matchset, "ci_contrasts_cuts.csv")) %>%
+  mutate(
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
+    subgroup_level_descr = replace_na(subgroup_level_descr, "")
+  )
+contrasts_overall <- read_csv(fs::path(output_dir_os, matchset, "ci_contrasts_overall.csv")) %>% 
+  mutate(
+    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
+    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
+    subgroup_level_descr = replace_na(subgroup_level_descr, ""),
+
+  )
+
+perN <- 1000
+perN_format <- label_number(1, 1, big.mark=",")(perN)
+
+label_number_n <- label_number(1, 1, big.mark=",")
+label_number_risk <- label_number(0.01, perN, style_negative="minus", big.mark=",")
+label_number_rr <- label_number(0.01, 1, style_negative="minus")
+
+contrasts_overall_table <-
+  contrasts_overall %>%
+  mutate(
+    outcome,
+    outcome_descr,
+    subgroup,
+    subgroup_descr,
+    subgroup_level_descr,
+    n.event = label_number_n(n.event_0+n.event_1),
+    persontime_weeks = label_number_n((persontime_0 + persontime_1) / 7),
+    n.atrisk = label_number_n(n.atrisk_0),
+    persontime_0=label_number_n(persontime_0/7), 
+    persontime_1=label_number_n(persontime_1/7),
+    n.event_0=label_number_n(n.event_0), 
+    n.event_1=label_number_n(n.event_1),
+
+    risk.ll_0 = 1-surv.ul_0,
+    risk.ul_0 = 1-surv.ll_0,
+    risk.ll_1 = 1-surv.ul_1,
+    risk.ul_1 = 1-surv.ll_1,
+    riskCI_0 = glue("{label_number_risk(risk_0)} ({label_number_risk(risk.ll_0)} to {label_number_risk(risk.ul_0)})"),
+    riskCI_0_95 = glue("{label_number_risk(risk_0)} (95%CI {label_number_risk(risk.ll_0)} to {label_number_risk(risk.ul_0)})"),
+    riskCI_1 = glue("{label_number_risk(risk_1)} ({label_number_risk(risk.ll_1)} to {label_number_risk(risk.ul_1)})"),
+    cirdCI = glue("{label_number_risk(cird)} ({label_number_risk(cird.ll)} to {label_number_risk(cird.ul)})"),
+    cirdCI_95 = glue("{label_number_risk(cird)} (95%CI {label_number_risk(cird.ll)} to {label_number_risk(cird.ul)})"),
+    #cirrCI = glue("{label_number_rr(cirr)} ({label_number_rr(cirr.ll)}-{label_number_rr(cirr.ul)})"),
+    #kmirrCI =   glue("{label_number_rr(kmirr)} ({label_number_rr(kmirr.ll)}-{label_number_rr(kmirr.ul)})"),
+    coxhrCI =   glue("{label_number_rr(coxhr)} ({label_number_rr(coxhr.ll)} to {label_number_rr(coxhr.ul)})"),
+    coxhrCI_95 =   glue("{label_number_rr(coxhr)} (95%CI {label_number_rr(coxhr.ll)} to {label_number_rr(coxhr.ul)})"),
+  )
+
+```
+
+William J Hulme^1^, Elsie Horne^3,4^, Ruth Keogh^2^, Edward PK Parker^2^, Venexia Walker^3,5^, Elizabeth J Williamson^2^, Tom Palmer^3,5^, Linda Nab^1^, Helen J Curtis^1^, Ben Goldacre^1^, Miguel A Hernán^8,9^, Jonathan AC Sterne^3,4,7^, and the OpenSAFELY collaborative.
+
+1\. The Bennett Institute for Applied Data Science, Nuffield Department of Primary Care Health Sciences, University of Oxford, OX26GG, UK
+
+2\. London School of Hygiene and Tropical Medicine, Keppel Street, London, WC1E 7HT, UK
+
+3\. Population Health Sciences, University of Bristol, Oakfield House, Oakfield Grove, Bristol, BS8 2BN, UK
+
+4\. NIHR Bristol Biomedical Research Centre, Bristol, UK
+
+5\. MRC Integrative Epidemiology Unit, Bristol Medical School, University of Bristol, Bristol, UK
+
+6\. TPP, TPP House, 129 Low Lane, Horsforth, Leeds, LS18 5PX
+
+7\. Health Data Research UK South West
+
+8\. CAUSALab, Harvard T.H. Chan School of Public Health, Boston, MA 02115
+
+9\. Departments of Epidemiology and Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA 02115
+
 \newpage
 
-## Overview
+## Abstract
+
+*Introduction* The  BNT162b2 and mRNA-1273 vaccines were both widely used for COVID-19 booster vaccination in England, and are both effectiveness against COVID-19 related outcomes, particularly severe outcomes.
+Direct comparisons of the effectiveness of these two vaccine types when used for boosting have not been made in trials or observational data.
+
+*Methods* Using the OpenSAFELY-TPP linked electronic health record database, we matched adult recipients of each vaccine type on date of vaccination, primary vaccine course, prior SARS-CoV-2 infection, and other important demographic and clinical characteristics.
+Recipients were eligible for inclusion if boosted between 29 October 2021 to 31 January 2022, and following up for 12 weeks.
+Outcomes were positive SARS-CoV-2 test, COVID-19 hospitalisation, and COVID-19 death.
+We used the Aalen-johansen estimator to obtain the cumulative incidence of each outcome, and compared these using risk differences (RD) and hazard ratios (HRs).
+
+*Results* `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` people were matched in each group, contributing a total of `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up.
+The 12-week risk per `r perN_format` people of positive SARS-CoV-2 test for those receiving BNT162b2 and Moderna mRNA-1273 was `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` (RD `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)`; HR `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`) respectively.
+
+*Conclusion* COVID-19-related outcomes after booster vaccination with BNT162b2 or mRNA-1273 were rare, and the risk was slightly lower for mRNA-1273 than for BNT162b2.
+These finding were broadly consistent across subgroups.
+
+\newpage
+
+## Background
 
@@
 
-BNT162b2 booster recipients were matched with Moderna booster recipients using two different matching strategies.
-one simpler strategy that will identify more matches but the risk of greater imbalance between treatment groups, and another more complex strategy that will identify fewer matches but with less imbalance.
+BNT162b2 booster recipients were matched 1:1 with Moderna booster recipients without replacement.
+Two different matching strategies were used: one simpler strategy resulting in more matches but with greater imbalance between treatment groups, and another more complex strategy resulting in fewer matches but with less imbalance.
 
-#### Matching strategy A
+#### Matching strategies
 
-Booster recipients were matched on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years and age groups used by JCVI for prioritisation; English Index of Multiple Deprivation, grouped by quintile.
+```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE}
 
-#### Matching strategy B
+tribble(
+  ~Factor, ~Description, ~A, ~B,
 
-Booster recipients were matched on the following characteristics: date of booster dose; primary course vaccine brand; date of second vaccine dose, within 7 days; Sustainability and Transformation Partnership (STP) a geographical grouping of NHS and local authorities; evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation; English Index of Multiple Deprivation, grouped by quintile; sex; and morbidity count (0, 1, 2 or more).
+  "Date of booster dose",  "Matched exactly",  
+    "X", "X",
+  "Primary course vaccine brand", "BNT162b2; ChadOx1-S", 
+    "X", "X",
+  "Date of second vaccine dose", "7 day caliper", 
+    "X", "X",
+  "Age",  "3 year caliper", 
+    "X", "X",
+  "JCVI age group", "18-39; 40-49; 50-54; 55-59; 60-64; 65-69; 70-75; 75-79; 80 and over",
+    "X", "X",
+  "JCVI Clinical risk group", "clinically extremely vulnerable; clinically at-risk; neither",
+    "X", "X",
+  "English Index of Multiple Deprivation", "Grouped by quintile",
+    "X", "X",
+  "NHS region", "East of England; Midlands; London; North East and Yorkshire; North West; South East; South West",
+    "X", "",
+  "Sustainability and Transformation Partnership (STP)", "a geographical grouping of NHS and local authorities",
+    "", "X",
+  "Sex", "Male; Female",
+    "", "X", 
+  "Evidence of prior SARS-CoV-2 infection", 'Any of: positive SARS-CoV-2 test; "probable" infection documented in primary care; COVID-19 hospital admission',
+    "", "X",
+  "Morbidity count", "0, 1, or 2 or more of the following conditions: diabetes; BMI over 40kg/m^2^; chronic heart disease; chronic kidney disease; chronic liver disease; chronic respiratory disease or severe asthma; chronic neurological disease",
+    "", "X"
+) %>%
+gt() %>%
+tab_spanner(
+  label = "Matching strategy",
+  columns = all_of(c("A", "B"))
+) %>%
+fmt_markdown(columns = TRUE)
 
-Morbidity count included the following conditions: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease.
+cat("  \n\n")
+```
+Morbidity count included the following conditions: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease"
 
@@
 
-Each person was followed from the day of boosting (i.e., time zero) until the earliest of outcome, death, practice de-registration, 10 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.
+Each person was followed from the day of boosting (i.e., time zero) until the earliest of outcome, death, practice de-registration, 12 weeks, or `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.
 
@@
 
-We estimated Kaplan-Meier (KM) cumulative incidence curves separately by vaccine brand, and compared brands using the 10-week risk differences and risk ratios.
-The incidence rate ratio (events per person-time-at-risk) in time-period specific intervals will also be reported.
-This is equivalent to a period-specific hazard ratio if a constant hazard within each time period is assumed.
+We estimated the cumulative incidence in each treatment group using the Aalen-Johansen estimator, with COVID-19 and non-COVID-19 treated as competing events (when these were not the event of interest).
+The risk differences (RD) and risk ratios (RRs) of the cumulative incidences were derived.
+We also estimated hazard ratios (HRs) using period-specific Cox models, in 7-day periods up to 28-days then 14-day periods thereafter.
+
+<!-- 95% confidence limits for the AJ estimates were derived from a Greenwood-type formula.
+Confidence limits for the risk difference were derived from the sum of squares of the AJ standard errors, derived from Greenwood’s formula.
+Confidence limits for the survival ratio were based on the standard error of the cumulative hazard.
+Confidence limits for the risk ratio were derived using the Delta method.
+-->
 
 <!-- 95% confidence limits for the KM estimates, risk differences, risk ratios, and incidence rate ratios are derived by bootstrapping the matched pairs. 500 bootstrap replicates will be used. -->
-
-95% confidence limits for the KM estimates were derived from the complementary log-log transformation.
-Confidence limits for the risk difference were derived from the sum of squares of the treatment-specific KM standard errors, derived from Greenwood’s formula.
-Confidence limits for the survival ratio were based on the standard error of the cumulative hazard.
-Confidence limits for the risk ratio were derived using the Delta method.
 
@@
 
-We also estimated comparative effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely .vulnerable or not; evidence of prior SARS-CoV-2 infection or not; dominant SARS-CoV-2 variant era (Delta or Omicron).
+We estimated comparative effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.
 
@@
 
-In this study, we round counts up to the nearest multiple of 7.
-Kaplan-Meier survival estimates are rounded upwards by deferring event-times until there are at least 7 events, so that each KM “step” is based on at least 7 events.
-All other quantities will be derived from these rounded KM estimates.
+To reduce re-identification risk, counts are reported to the nearest $n*6-3$ ($3, 9, 15, 21, ...$).
+Similarly, event times were deferred until there were at least 6 events, so that each "step" in the Aalen-Johansen cumulative incidence estimates is based on at least 6 events.
 
 # Results
-
-```{r match-A, echo=FALSE, warning=FALSE, message=FALSE}
-
-matchset <- "A"
-
-match_flowchart <-  read_csv(fs::path(output_dir_os, matchset, "match_flowchart.csv"))
-match_coverage <- read_csv(fs::path(output_dir_os, matchset, "match_coverage.csv")) %>%
-   mutate(
-    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
-  )
-
-match_table1 <- read_csv(fs::path(output_dir_os, matchset, "match_table1.csv"))
-match_table1by <- read_csv(fs::path(output_dir_os, matchset, "match_table1by.csv"))
-match_smd <-  read_csv(fs::path(output_dir_os, matchset, "match_smd.csv"))
-
-km_estimates <- read_csv(fs::path(output_dir_os, matchset, "km_estimates.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-contrasts_daily <- read_csv(fs::path(output_dir_os, matchset, "contrasts_daily.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-contrasts_overall <- read_csv(fs::path(output_dir_os, matchset, "contrasts_overall.csv")) %>% 
-  mutate(
-    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
-    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
-    subgroup_level_descr = replace_na(subgroup_level_descr, "")
-  )
-
-```
 
@@
 
-## Matching strategy A
+For the simpler matching criteria, `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
 
-For the simpler matching criteria, `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(n))` were matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type=="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).
+The following results are based on matching strategy A, with results for strategy B included in supplementary materials.
+
+Characteristics were well-balanced between vaccine types at the start of follow up (Table 2), and the proportion of people with prior clinical conditions was generally similar between the groups.
+The standardised mean difference was consistently below 0.1.
+
+## Estimated comparative effectiveness
+
+### Main analysis
+
+12 weeks after booster vaccination by either BNT162b2 or Moderna mRNA-1273 there were `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(n.event)`positive SARS-CoV-2 tests, `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths across `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 2).
+
+Figure 2a shows the cumulative incidence curves for each outcome and vaccine type.
+Absolute risks at 12 weeks of positive SARS-CoV-2 test for those receiving BNT162b2 and Moderna mRNA-1273 were `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(riskCI_0_95)` and `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people respectively, representing a risk difference of `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` nad a hazard ratio of `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(coxhrCI_95)`.
+Corresponding 12-week risks for COVID-19 hospitalisation were `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(coxhrCI)`),
+and for COVID-19 death were `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` per `r perN_format` people (RD `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI)`; HR `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(coxhrCI)`).
+
+### Subgroup analyses
+
+Whilst absolute risks differed between subgroups, and therefore so did the risk differences comparing vaccine types, subgroup-specific hazard ratios were broadly similar.
+
+A notable exception was COVID-19 hospitalisation outcomes stratified by clinical vulnerability: the HR was lower amongst those were were clinically at risk or extremely clinically vulnerable, compared with those who were not clinically at-risk (not clinically at-risk HR `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="cev_cv", subgroup_level=="Not clinically at-risk") %>% pull(cirdCI)`; clinically at-risk HR `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="cev_cv", subgroup_level=="Clinically at-risk") %>% pull(cirdCI)`; clinically extremely vulnerable `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="cev_cv", subgroup_level=="Clinically extremely vulnerable") %>% pull(cirdCI)`).
+
+## Discussion
+
+This observational study in over 4 million adults compares the effectiveness of BNT162b2 and Moderna mRNA-1273 vaccines against COVID19-related outcomes in the national booster vaccination programme in England.
+We estimated that Moderna mRNA-1273 provided slightly more protection than BNT162b2 did against a positive SARS-CoV-2 test (RD = `r contrasts_overall_table %>% filter(outcome=="postest", subgroup=="all") %>% pull(cirdCI_95)` per `r perN_format`) and hospitalisation (RD = `r contrasts_overall_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(cirdCI_95)`) in the first 12 weeks after boosting.
+For COVID-19 death, the difference was equivocal (RD = `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(cirdCI_95)`) though there were only `r contrasts_overall_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` events across both vaccine types, limiting statistical power.
+
+Around 1 in 10 people had a positive SARS-CoV-2 test in the 12 weeks following booster vaccination, reflecting both the high infection rates in the study period, particularly from December 2021 onward, and the imperfect protection booster vaccination offers against infection.
+However, rates of severe disease were rare: after 12-weeks, fewer than 1 in 1000 booster recipients were hospitalised with COVID-19, and fewer than 1 in 20,000 died due to COVID-19 disease.
+
+Comparative effectiveness against these severe outcomes appeared similar regardless of age, whether the primary vaccine course was with ChAdOx1-S or BNT162b2, or whether there was evidence of prior infection or not.
+However, the relative advantage of mRNA-1273 against severe COVID-19 was even higher in people classified as clinically at-risk or extremely vulnerable, compared with those not clinically at risk.
+
+### Strengths and weaknesses
+
+The OpenSAFELY-TPP database covers around 40% of the English population and contains rich clinical information which enabled us to closely match BNT162b2 and mRNA-1273 recipients and study a range of clinical outcomes.
+
+We excluded certain groups who were unlikely to undergo boosting, such as those with a recent positive test or in palliative care.
+We also excluded groups where we could not reliably capture high heterogeneity in outcome risk, such as those living in care homes and health and social care workers.
+This may limit the generalisability of out findings.
+
+Outcomes may be under-ascertained.
+Positive SARS-CoV-2 tests only include those reported via the national COVID-19 surveillance system (SGSS), and so many asymptomatic and some symptomatic infections will be missed.
+Positive SARS-CoV-2 testing is therefore not a reliable surrogate for SARS-CoV-2 infection.
+Some COVID-19 hospitalisations resulting in longer hospital stays may have been missed as these events are only reported once the patient has been discharged.
+
+### Findings in context
+
+Our findings are consistent with studies identifying a small benefit of mRNA-1273 compared with BNT162b2 for primary vaccination (DOI 10.1056/NEJMoa2115463).
+
+### Conclusions
+
+COVID-19-related outcomes after booster vaccination with BNT162b2 or mRNA-1273 were rare, though the risk was slightly higher for BNT162b2 and for mRNA-1273.
+These finding were broadly consistent across subgroups.
+These are important differences at a population-level and may help inform vaccine procurement decisions for future booster programmes.
+Individuals should not defer booster vaccination by BNT162b2 in favour of waiting for mRNA-1273 if offered.
+
+## Figures and tables
 
@@
   ) %>%
-  # tab_spanner(
-  #   label = "mRNA-1273",
-  #   columns = ends_with("moderna")
-  # ) %>%
   fmt_percent(
@@
   match_table1 %>%
-  #filter(outcome %in% outcomes) %>%
   group_by(variable) %>%
@@
 
-## Estimated comparative effectiveness
+#### Table 3: Kaplan-Meier 12-week risk estimates
 
-### Main analysis
-
-#### Table 3: Kaplan-Meier 10-week risk estimates
-
-Counts and survival estimates are based on values rounded up to the nearest 7, for disclosure control.
+Counts and survival estimates are based on values rounded up to the nearest n*6-3, for disclosure control.
 
@@
 
-label_number_n <- label_number(1, 1, big.mark=",")
-label_number_risk <- label_number(0.1, 10000, style_negative="minus", big.mark=",")
-label_number_rr <- label_number(0.01, 1, style_negative="minus")
-
-contrasts_overall %>%
-  transmute(
+contrasts_overall_table %>%
+  filter(outcome %in% outcomes) %>%
+  group_by(outcome_descr) %>%
+  select(
     outcome_descr,
@@
     subgroup_level_descr,
-    n.atrisk=label_number_n(n.atrisk_0),
-    persontime_0=label_number_n(persontime_0/7), 
-    persontime_1=label_number_n(persontime_1/7),
-    n.event_0=label_number_n(n.event_0), 
-    n.event_1=label_number_n(n.event_1),
-    riskCI_0 = glue("{label_number_risk(1-surv_0)} ({label_number_risk(1-surv.ul_0)} - {label_number_risk(1-surv.ll_0)})"),
-    riskCI_1 = glue("{label_number_risk(1-surv_1)} ({label_number_risk(1-surv.ul_1)} - {label_number_risk(1-surv.ll_1)})"),
-    kmrdCI = glue("{label_number_risk(kmrd)} ({label_number_risk(kmrd.ll)} - {label_number_risk(kmrd.ul)})"),
-    kmrrCI = glue("{label_number_rr(kmrr)} ({label_number_rr(kmrr.ll)} - {label_number_rr(kmrr.ul)})"),
-    irrCI =   glue("{label_number_rr(irr)} ({label_number_rr(irr.ll)} - {label_number_rr(irr.ul)})"),
-    #coxhrCI =  paste0(label_number(0.01)(kmrr), " (", label_number(0.01)(coxhr.ll), "-", label_number(0.01)(coxhr.ul), ")"),
+
+    n.atrisk,
+
+    persontime_0,
+    persontime_1,
+
+    n.event_0,
+    n.event_1,
+
+    riskCI_0,
+    riskCI_1,
+
+    cirdCI,
+    #cirrCI,
+    #kmirrCI,
+    coxhrCI,
   ) %>%
-  group_by(outcome_descr) %>%
   mutate(
@@
 
-    riskCI_0 = "10-week Risk per 10,000 people (95% CI)",
-    riskCI_1 = "10-week Risk per 10,000 people (95% CI)",
+    riskCI_0 = glue("12-week risk per {perN_format} people (95% CI)"),
+    riskCI_1 = glue("12-week risk per {perN_format} people (95% CI)"),
 
-    kmrdCI = "Risk difference per 10,000 people (95% CI)",
-    kmrrCI = "Risk ratio (95% CI)",
-    irrCI = "Incidence rate ratio (95% CI)",
-    #coxhrCI = "Hazard ratio (95% CI)"
+    cirdCI = glue("Risk difference per {perN_format} people (95% CI)"),
+    #cirrCI = "Risk ratio (95% CI)",
+    #kmirrCI = "Incidence rate ratio (95% CI)",
+    coxhrCI = "Hazard ratio (95% CI)"
   ) %>%
@@
     align = c("right"),
-    columns = ends_with(c("_1", "_0", "CI"))
+    columns = ends_with(c("atrisk", "_1", "_0", "CI"))
   ) %>%
@@
 
-```{r, function-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+```{r, function-ci, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
 
-km_theme <- function(...) {
+ci_theme <- function(...) {
   theme(
@@
       strip.text = element_blank(),
-
+      strip.placement="outside",
       panel.border = element_blank(),
@@
 
-colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
+#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
+colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")
 
-plot_km_subgroup <- function(subgroup){
+plot_ci_subgroup <- function(subgroup){
 
@@
   if(subgroup=="all"){
-    colour_guide <- list(guides(colour="none"))
+    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
   } else{
-    colour_guide <- NULL
+    colour_guide <- list(guides(fill="none", alpha="none"))
   }
 
-  plot_km <- km_estimates %>%
+  plot_ci <- ci_estimates %>%
     filter(
@@
     ) %>%
-    ggplot(aes(group=colgroup, colour=subgroup_level_descr, fill=subgroup_level_descr, alpha=treatment_descr)) +
-    geom_step(aes(x=time, y=1-surv))+
-    geom_rect(aes(xmin=lagtime, xmax=time, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
+    #group_by(colgroup, outcome_descr) %>%
+    ggplot(aes(group=colgroup, colour=treatment_descr, fill=treatment_descr, alpha=treatment_descr)) +
+    geom_step(aes(x=lagtime, y=risk, linetype=treatment_descr))+
+    #geom_segment(aes(x=lagtime, xend=time, y=risk, yend=risk))+
+    #geom_segment(aes(x=lagtime, xend=lagtime, y=lag(risk,1,0), yend=risk))+
+    geom_rect(aes(xmin=lagtime, xmax=time, ymin=risk.ll, ymax=risk.ul), alpha=0.1, colour="transparent")+
     geom_hline(aes(yintercept=0), colour="black")+
-    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
+    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), scales="free_y", switch="y")+
     scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
@@
     scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
-    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, 10000))+
+    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number_risk)+
     scale_alpha_discrete(range=c(0.3,0.9))+
@@
       y=NULL,
-      title="Risk per 10,000",
+      #title=glue("Cumulative incidence per {perN_format}"),
       colour=NULL,
-      alpha=NULL
+      alpha=NULL,
+      linetype=NULL
     )+
-    colour_guide+
+    guides(linetype="none", fill="none", alpha="none")+
     theme_minimal(base_size = 9)+
       theme(
-        legend.position = c(0.05,0.1),
-        legend.justification = c(0,0),
-        legend.direction = "vertical",
+        # legend.position = c(0.05,0.1),
+        # legend.justification = c(0,0),
+        legend.position = "bottom",
+        #legend.direction = "vertical",
         axis.text.x.top=element_text(hjust=0),
@@
 
-  plot_kmrd <- contrasts_daily %>%
-    filter(
-      subgroup==subgroup0,
-      outcome %in% outcomes
-    ) %>%
-    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
-    geom_step(aes(x=period_start, y=kmrd))+
-    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrd.ll, ymax=kmrd.ul), alpha=0.1, colour="transparent")+
-    geom_hline(aes(yintercept=0), colour="black")+
-    facet_grid(rows=vars(outcome_descr), scales="free_y")+
-    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
-    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
-    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, 10000))+
-    coord_cartesian(xlim=c(0, NA))+
-    labs(
-      x="Days",
-      y=NULL,
-      title="Risk difference",
-      colour=NULL
-    )+
-    theme_minimal(base_size = 9)+
-    km_theme()+
-    NULL
+  # plot_cird <- contrasts_daily %>%
+  #   filter(
+  #     subgroup==subgroup0,
+  #     outcome %in% outcomes
+  #   ) %>%
+  #   ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+  #   geom_segment(aes(x=period_start, xend=period_end, y=cird, yend=cird))+
+  #   geom_segment(aes(x=period_start, xend=period_start, y=lag(cird), yend=cird))+
+  #   geom_rect(aes(xmin=period_start, xmax=period_end, ymin=cird.ll, ymax=cird.ul), alpha=0.1, colour="transparent")+
+  #   geom_hline(aes(yintercept=0), colour="black")+
+  #   facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  #   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  #   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  #   scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+  #   scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, 10000))+
+  #   coord_cartesian(xlim=c(0, NA))+
+  #   labs(
+  #     x="Days",
+  #     y=NULL,
+  #     title="Risk difference",
+  #     colour=NULL
+  #   )+
+  #   theme_minimal(base_size = 9)+
+  #   ci_theme()+
+  #   NULL
 
-  plot_kmrr <- contrasts_daily %>%
-    filter(
-      subgroup==subgroup0,
-      outcome %in% outcomes
-    ) %>%
-    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
-    geom_step(aes(x=period_start, y=kmrr))+
-    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmrr.ll, ymax=kmrr.ul), alpha=0.1, colour="transparent")+
-    geom_hline(aes(yintercept=1), colour="black")+
-    facet_grid(rows=vars(outcome_descr), scales="free_y")+
-    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
-    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
-    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
-    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
-    coord_cartesian(xlim=c(0, NA))+
-    labs(
-      x="Days",
-      y=NULL,
-      title="Risk ratio",
-      colour=NULL
-    )+
-    theme_minimal(base_size = 9)+
-    km_theme()+
-    NULL
+  # plot_cirr <- contrasts_daily %>%
+  #   filter(
+  #     subgroup==subgroup0,
+  #     outcome %in% outcomes
+  #   ) %>%
+  #   ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+  #   geom_segment(aes(x=period_start, xend=period_end, y=cirr, yend=cirr))+
+  #   geom_segment(aes(x=period_start, xend=period_start, y=lag(cirr), yend=cirr))+
+  #   geom_rect(aes(xmin=period_start, xmax=period_end, ymin=cirr.ll, ymax=cirr.ul), alpha=0.1, colour="transparent")+
+  #   geom_hline(aes(yintercept=1), colour="black")+
+  #   facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  #   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  #   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  #   scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+  #   scale_y_log10(expand = expansion(mult=c(0,0.01)))+
+  #   coord_cartesian(xlim=c(0, NA))+
+  #   labs(
+  #     x="Days",
+  #     y=NULL,
+  #     title="Risk ratio",
+  #     colour=NULL
+  #   )+
+  #   theme_minimal(base_size = 9)+
+  #   ci_theme()+
+  #   NULL
 
-  plot_irr <- contrasts_daily %>%
+  # plot_kmirr <- contrasts_daily %>%
+  #   # approximately same as the hazard
+  #   filter(
+  #     subgroup==subgroup0,
+  #     outcome %in% outcomes
+  #   ) %>%
+  #   ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
+  #   geom_segment(aes(x=period_start, xend=period_end, y=kmirr, yend=kmirr))+
+  #   geom_segment(aes(x=period_start, xend=period_start, y=lag(kmirr), yend=kmirr))+
+  #   geom_rect(aes(xmin=period_start, xmax=period_end, ymin=kmirr.ll, ymax=kmirr.ul), alpha=0.1, colour="transparent")+
+  #   geom_hline(aes(yintercept=0), colour="black")+
+  #   facet_grid(rows=vars(outcome_descr), scales="free_y")+
+  #   scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  #   scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  #   scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
+  #   scale_y_log10(expand = expansion(mult=c(0,0.01)))+
+  #   coord_cartesian(xlim=c(0, NA))+
+  #   labs(
+  #     x="Days",
+  #     y=NULL,
+  #     title="IRR",
+  #     colour=NULL
+  #   )+
+  #   theme_minimal(base_size = 9)+
+  #   ci_theme()+
+  #   NULL
+
+  plot_coxhr <- contrasts_cuts %>%
     # equivalent to the hazard
@@
     ) %>%
+    group_by(outcome_descr, subgroup_level_descr) %>%
     ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
-    geom_step(aes(x=period_start, y=irr))+
-    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=irr.ll, ymax=irr.ul), alpha=0.1, colour="transparent")+
-    geom_hline(aes(yintercept=0), colour="black")+
+    geom_segment(aes(x=period_start, xend=period_end, y=coxhr, yend=coxhr))+
+    geom_segment(aes(x=period_start, xend=period_start, y=lag(coxhr), yend=coxhr))+
+    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=coxhr.ll, ymax=coxhr.ul), alpha=0.1, colour="transparent")+
+    geom_hline(aes(yintercept=1), colour="black")+
     facet_grid(rows=vars(outcome_descr), scales="free_y")+
-    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
-    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+    # scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+    # scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+    scale_color_manual(values=colourvalues, na.value="grey") +
+    scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
     scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
@@
     coord_cartesian(xlim=c(0, NA))+
+    colour_guide+
     labs(
       x="Days",
-      y=NULL,
+      y="higher risk\n<- BNT162b2 | mRNA 1273 ->",
       title="Hazard ratio",
@@
     theme_minimal(base_size = 9)+
-    km_theme()+
+    ci_theme()+
     NULL
 
+  n_subgroups <-
+    ci_estimates %>%
+    filter(
+      subgroup==subgroup0
+    ) %>% 
+    `[[`("subgroup_level_descr") %>%
+    n_distinct()
+
   plot_combined <- patchwork::wrap_plots(
-    plot_km, 
-    plot_kmrd, 
-    plot_kmrr, 
-    #plot_irr,
-    ncol=3
+    plot_ci, 
+    #plot_cird, 
+    #plot_cirr, 
+    #plot_kmirr,
+    plot_coxhr,
+    ncol=2,
+    widths=c(n_subgroups,1)
   )
@@
   ggsave(
-    filename=glue("km_{subgroup}.png"),
+    filename=glue("ci_{subgroup}.png"),
     path=output_dir_rmd,
-    plot=plot_combined,
+    plot=plot_ci,
   )
 
-  plot_combined
+  plot_ci
 }
@@
 
-#### Figure 3a: Kaplan-Meier cumulative incidence estimates
+#### Figure 2a: Cumulative incidence estimates per `r perN_format` people
 
-```{r, figure-km-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
-plot_km_subgroup("all")
+```{r, figure-ci-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_ci_subgroup("all")
 ```
@@
 
-#### Figure 3b: Kaplan-Meier cumulative incidence estimates, by primary course type
+#### Figure 2b: Cumulative incidence estimates per `r perN_format` people by primary course type
 
-```{r, figure-km-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
-plot_km_subgroup("vax12_type")
+```{r, figure-ci-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_ci_subgroup("vax12_type")
 ```
@@
 
-#### Figure 3c: Kaplan-Meier cumulative incidence estimates, by age group
+#### Figure 2c: Cumulative incidence estimates per `r perN_format` people by age group
 
-```{r, figure-km-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
-plot_km_subgroup("age65plus")
+```{r, figure-ci-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_ci_subgroup("age65plus")
 ```
@@
 
-#### Figure 3d Kaplan-Meier cumulative incidence estimates, by prior infection status
+#### Figure 2d Cumulative incidence estimates per `r perN_format` people by prior infection status
 
-```{r, figure-km-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
-plot_km_subgroup("prior_covid_infection")
+```{r, figure-ci-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_ci_subgroup("prior_covid_infection")
 ```
@@
 
-#### Figure 3e Kaplan-Meier cumulative incidence estimates, by clinical vulnerability
+#### Figure 2e Cumulative incidence estimates per `r perN_format` people by clinical vulnerability
 
-```{r, figure-km-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
-plot_km_subgroup("cev_cv")
+```{r, figure-ci-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+plot_ci_subgroup("cev_cv")
 ```

### 2022-07-06 analysis/study_definition.py
new product names

@@ -237,7 +237,7 @@ study = StudyDefinition(
     name = "covid_vax_az",
     index_date = "1900-01-01",
     n = 4,
-    product_name_matches="COVID-19 Vac AstraZeneca (ChAdOx1 S recomb) 5x10000000000 viral particles/0.5ml dose sol for inj MDV"
+    product_name_matches="COVID-19 Vaccine Vaxzevria 0.5ml inj multidose vials (AstraZeneca)"
   ),
   
   # moderna

### 2022-07-06 analysis/study_definition.py
simplify carehome flag

@@ -442,33 +442,17 @@ study = StudyDefinition(
   # health or social care worker  
   hscworker = patients.with_healthcare_worker_flag_on_covid_vaccine_record(returning="binary_flag"),
   
-  care_home_type=patients.care_home_status_as_of(
+  care_home_tpp=patients.satisfying(
+    "care_home=='1'",
+    
+    care_home = patients.care_home_status_as_of(
       "covid_vax_disease_3_date - 1 day",
       categorised_as={
-          "Carehome": """
-            IsPotentialCareHome
-            AND LocationDoesNotRequireNursing='Y'
-            AND LocationRequiresNursing='N'
-          """,
-          "Nursinghome": """
-            IsPotentialCareHome
-            AND LocationDoesNotRequireNursing='N'
-            AND LocationRequiresNursing='Y'
-          """,
-          "Mixed": "IsPotentialCareHome",
+          "1": "IsPotentialCareHome",
           "": "DEFAULT",  # use empty string
-      },
-      return_expectations={
-          "category": {"ratios": {"Carehome": 0.05, "Nursinghome": 0.05, "Mixed": 0.05, "": 0.85, }, },
-          "incidence": 1,
       },
   ),
   
-  # simple care home flag
-  care_home_tpp=patients.satisfying(
-      """care_home_type""",
-      return_expectations={"incidence": 0.01},
-  ),
   
   # Patients in long-stay nursing and residential care
   care_home_code=patients.with_these_clinical_events(
@@ -477,7 +461,6 @@ study = StudyDefinition(
       returning="binary_flag",
       return_expectations={"incidence": 0.01},
   ),
-  
   
   
 

### 2022-07-06 lib/dummydata/dummydata.R

@@ -119,12 +119,8 @@ sim_list = lst(
     missing_rate = ~ 0.1
   ),
 
-  care_home_type = bn_node(
-    ~rfactor(n=..n, levels=c("Carehome", "Nursinghome", "Mixed", ""), p = c(0.01, 0.01, 0.01, 0.97))
-  ),
-
   care_home_tpp = bn_node(
-    ~care_home_type!=""
+    ~rbernoulli(n=..n, p = 0.01)
   ),
 
   care_home_code = bn_node(

### 2022-07-06 analysis/study_definition.py
simplify carehome flag

@@ -451,6 +451,7 @@ study = StudyDefinition(
           "1": "IsPotentialCareHome",
           "": "DEFAULT",  # use empty string
       },
+    ),
   ),
   
   

### 2022-07-06 analysis/data_selection.R
simplify in-hospital exclusion

@@ -65,7 +65,7 @@ data_criteria <- data_processed %>%
     has_knownvax2 = vax2_type %in% c("pfizer", "az"),
     has_expectedvax3type = vax3_type %in% c("pfizer", "moderna"),
     has_norecentcovid = vax3_date - anycovid_0_date >= 90 | is.na(anycovid_0_date),
-    isnot_inhospitalunplanned = !inhospital_unplanned,
+    isnot_inhospital = !inhospital,
 
     jcvi_group_6orhigher = jcvi_group %in% as.character(1:6),
 
@@ -78,7 +78,7 @@ data_criteria <- data_processed %>%
       isnot_hscworker &
       isnot_carehomeresident & isnot_endoflife & isnot_housebound &
       has_norecentcovid &
-      isnot_inhospitalunplanned
+      isnot_inhospital
 
     ),
   )
@@ -102,7 +102,7 @@ data_inclusioncriteria <- data_criteria %>%
     c3 = c2 & (isnot_hscworker ),
     c4 = c3 & (isnot_carehomeresident & isnot_endoflife & isnot_housebound),
     c5 = c4 & (has_norecentcovid),
-    c6 = c5 & (isnot_inhospitalunplanned),
+    c6 = c5 & (isnot_inhospital),
   ) %>%
   filter(c0)
 
@@ -137,7 +137,7 @@ data_flowchart <-
       crit == "c3" ~ "  and not a HSC worker",
       crit == "c4" ~ "  and not a care/nursing home resident, end-of-life or housebound",
       crit == "c5" ~ "  and no COVID-19-related events within 90 days",
-      crit == "c6" ~ "  and not in hospital at time of booster",
+      crit == "c6" ~ "  and not admitted in hospital at time of booster",
       TRUE ~ NA_character_
     )
   ) #
@@ -172,7 +172,7 @@ data_flowchart_rounded <-
       crit == "c3" ~ "  and not a HSC worker",
       crit == "c4" ~ "  and not a care/nursing home resident, end-of-life or housebound",
       crit == "c5" ~ "  and no COVID-19-related events within 90 days",
-      crit == "c6" ~ "  and not in hospital at time of booster",
+      crit == "c6" ~ "  and not admitted in hospital at time of booster",
       TRUE ~ NA_character_
     )
   ) #

### 2022-07-06 analysis/study_definition.py

@@ -898,39 +898,21 @@ study = StudyDefinition(
     restrict_to_earliest_specimen_date=False,
   ),
   
-  # unplanned hospital admission at time of 3rd / booster dose
-  inhospital_unplanned = patients.satisfying(
+  # overnight hospital admission at time of 3rd / booster dose
+  inhospital = patients.satisfying(
   
-    "discharged_unplanned_0_date >= covid_vax_disease_3_date",
+    "discharged_0_date >= covid_vax_disease_3_date",
     
-    discharged_unplanned_0_date=patients.admitted_to_hospital(
+    discharged_0_date=patients.admitted_to_hospital(
       returning="date_discharged",
-      on_or_before="covid_vax_disease_3_date - 1 day", #FIXME -- need to decide whether to include admissions discharged on the same day as booster dose or not
+      on_or_before="covid_vax_disease_3_date", # this is the admission date
       # see https://github.com/opensafely-core/cohort-extractor/pull/497 for codes
       # see https://docs.opensafely.org/study-def-variables/#sus for more info
-      with_admission_method=["21", "22", "23", "24", "25", "2A", "2B", "2C", "2D", "28"],
+      with_admission_method = ['11', '12', '13', '21', '2A', '22', '23', '24', '25', '2D', '28', '2B', '81'],
       with_patient_classification = ["1"], # ordinary admissions only
       date_format="YYYY-MM-DD",
       find_last_match_in_period=True,
     ), 
-  ),
-  
-  # planned hospital admission at time of 3rd / booster dose
-  inhospital_planned = patients.satisfying(
-  
-    "discharged_planned_0_date >= covid_vax_disease_3_date",
-    
-    discharged_planned_0_date=patients.admitted_to_hospital(
-      returning="date_discharged",
-      on_or_before="covid_vax_disease_3_date - 1 day", #FIXME -- need to decide whether to include admissions discharged on the same day as booster dose or not
-      # see https://github.com/opensafely-core/cohort-extractor/pull/497 for codes
-      # see https://docs.opensafely.org/study-def-variables/#sus for more info
-      with_admission_method=["11", "12", "13", "81"],
-      with_patient_classification = ["1"], # ordinary admissions only
-      date_format="YYYY-MM-DD",
-      find_last_match_in_period=True
-    ), 
-  
   ),
   
 

### 2022-07-06 lib/dummydata/dummydata.R

@@ -221,8 +221,7 @@ sim_list = lst(
     missing_rate = ~0
   ),
 
-  inhospital_planned = bn_node( ~rbernoulli(n=..n, p = 0.05)),
-  inhospital_unplanned = bn_node( ~rbernoulli(n=..n, p = 0.01)),
+  inhospital = bn_node( ~rbernoulli(n=..n, p = 0.01)),
 
   ## pre-baseline events where event date is relevant
 

### 2022-07-06 analysis/codelists.py
add cancer variable

@@ -389,8 +389,6 @@ learndis = codelist_from_csv(
   column="code",
 )
 
-
-
 # Carer codes
 carer = codelist_from_csv(
   "codelists/primis-covid19-vacc-uptake-carer.csv",
@@ -427,6 +425,35 @@ domcareemployee = codelist_from_csv(
 )
 
 
+
+# Cancer
+
+cancer_haem_snomed=codelist_from_csv(
+    "codelists/opensafely-haematological-cancer-snomed.csv",
+    system="snomed",
+    column="id",
+)
+
+cancer_nonhaem_nonlung_snomed=codelist_from_csv(
+    "codelists/opensafely-cancer-excluding-lung-and-haematological-snomed.csv",
+    system="snomed",
+    column="id",
+)
+
+cancer_lung_snomed=codelist_from_csv(
+    "codelists/opensafely-lung-cancer-snomed.csv",
+    system="snomed",
+    column="id",
+)
+
+cancer_nonhaem_snomed=combine_codelists(
+    cancer_nonhaem_nonlung_snomed,
+    cancer_lung_snomed,
+)
+
+
+# end of life
+
 eol = codelist_from_csv(
     "codelists/nhsd-primary-care-domain-refsets-palcare_cod.csv",
     system="snomed",
@@ -455,4 +482,5 @@ discharged_to_hospital = codelist(
     ["306706006", "1066331000000109", "1066391000000105"],
     system="snomed",
 )
+
 

### 2022-07-06 analysis/data_process.R

@@ -173,7 +173,8 @@ data_processed <- data_extract %>%
       (diabetes) +
       (chronic_liver_disease)+
       (chronic_resp_disease | asthma)+
-      (chronic_neuro_disease)#+
+      (chronic_neuro_disease)+
+      (cancer)#+
       #(learndis)+
       #(sev_mental),
     ,

### 2022-07-06 analysis/data_selection.R

@@ -207,6 +207,7 @@ var_labels <- list(
   chronic_resp_disease ~ "Chronic respiratory disease",
   asthma ~ "Asthma",
   chronic_neuro_disease ~ "Chronic neurological disease",
+  cancer ~ "Cancer, within previous 3 years",
 
   #multimorb ~ "Morbidity count",
   immunosuppressed ~ "Immunosuppressed",

### 2022-07-06 analysis/match_report.R

@@ -262,6 +262,7 @@ var_labels <- list(
   chronic_resp_disease ~ "Chronic respiratory disease",
   asthma ~ "Asthma",
   chronic_neuro_disease ~ "Chronic neurological disease",
+  cancer ~ "Cancer, within previous 3 years",
 
   #multimorb ~ "Morbidity count",
   immunosuppressed ~ "Immunosuppressed",

### 2022-07-06 analysis/study_definition.py

@@ -808,7 +808,35 @@ study = StudyDefinition(
   #   on_or_before="covid_vax_disease_3_date - 1 day",
   #   date_format="YYYY-MM-DD",
   # ),
+  
+  cancer = patients.satisfying(
+    
+    "cancer_primary_care",
+    # cancer_hosp=patients.admitted_to_hospital(
+    #   with_these_diagnoses=combine_codelists(
+    #     codelists.cancer_nonhaem_icd10,
+    #     codelists.cancer_haem_icd10,
+    #     codelists.cancer_unspec_icd10,
+    #   ),
+    #   between=["covid_vax_disease_3_date - 3 years", "covid_vax_disease_3_date - 1 day"],
+    #   returning="binary_flag",
+    # ),
+    cancer_primary_care=patients.with_these_clinical_events( 
+      combine_codelists(
+        codelists.cancer_nonhaem_snomed,
+        codelists.cancer_haem_snomed
+      ),
+      between=["covid_vax_disease_3_date - 3 years", "covid_vax_disease_3_date - 1 day"],
+      returning="binary_flag",
+    ), 
+  )
 
+  
+  
+  #####################################
+  # JCVI groups
+  #####################################
+  
   cev_ever = patients.with_these_clinical_events(
     codelists.shield,
     returning="binary_flag",

### 2022-07-06 lib/dummydata/dummydata.R

@@ -206,6 +206,7 @@ sim_list = lst(
   chronic_heart_disease = bn_node( ~rbernoulli(n=..n, p = 0.02)),
   chronic_kidney_disease = bn_node( ~rbernoulli(n=..n, p = 0.02)),
   chronic_liver_disease = bn_node( ~rbernoulli(n=..n, p = 0.02)),
+  cancer = bn_node( ~rbernoulli(n=..n, p = 0.02)),
   immunosuppressed = bn_node( ~rbernoulli(n=..n, p = 0.02)),
   asplenia = bn_node( ~rbernoulli(n=..n, p = 0.02)),
   learndis = bn_node( ~rbernoulli(n=..n, p = 0.02)),

### 2022-07-06 analysis/data_process.R
tidy up IMD

@@ -137,7 +137,7 @@ data_processed <- data_extract %>%
     ),
 
     imd = as.integer(as.character(imd)), # imd is a factor, so convert to character then integer to get underlying values
-    imd = if_else(imd<0, NA_integer_, imd),
+    imd = if_else(imd<0 | msoa=="", NA_integer_, imd),
     imd_Q5 = fct_case_when(
       (imd >=0) & (imd < 32844*1/5) ~ "1 most deprived",
       (imd >= 32844*1/5) & (imd < 32844*2/5) ~ "2",

### 2022-07-06 analysis/data_selection.R

@@ -194,7 +194,7 @@ var_labels <- list(
   ageband ~ "Age",
   sex ~ "Sex",
   ethnicity_combined ~ "Ethnicity",
-  imd_Q5 ~ "IMD",
+  imd_Q5 ~ "Deprivation",
   region ~ "Region",
   cev_cv ~ "JCVI clinical risk group",
 

### 2022-07-06 analysis/match_report.R

@@ -249,7 +249,7 @@ var_labels <- list(
   ageband ~ "Age",
   sex ~ "Sex",
   ethnicity_combined ~ "Ethnicity",
-  imd_Q5 ~ "IMD",
+  imd_Q5 ~ "Deprivation",
   region ~ "Region",
   cev_cv ~ "JCVI clinical risk group",
 

### 2022-07-06 analysis/data_selection.R
tidy up flowchart tables

@@ -131,10 +131,10 @@ data_flowchart <-
     pct_step = n / lag(n),
     crit = str_extract(criteria, "^c\\d+"),
     criteria = fct_case_when(
-      crit == "c0" ~ "Aged 18+ and recieved booster dose of BNT162b2 or Moderna between 29 October 2021 and 31 January 2022", # paste0("Aged 18+\n with 2 doses on or before ", format(study_dates$lastvax2_date, "%d %b %Y")),
+      crit == "c0" ~ "Aged 18+ and recieved booster dose of BNT162b2 or mRNA-1273 between 29 October 2021 and 31 January 2022", # paste0("Aged 18+\n with 2 doses on or before ", format(study_dates$lastvax2_date, "%d %b %Y")),
       crit == "c1" ~ "  with no missing demographic information",
-      crit == "c2" ~ "  with homologous primary vaccination course of pfizer or AZ",
-      crit == "c3" ~ "  and not a HSC worker",
+      crit == "c2" ~ "  with homologous primary vaccination course of BNT162b2 or ChAdOx1",
+      crit == "c3" ~ "  and not a health and social care worker",
       crit == "c4" ~ "  and not a care/nursing home resident, end-of-life or housebound",
       crit == "c5" ~ "  and no COVID-19-related events within 90 days",
       crit == "c6" ~ "  and not admitted in hospital at time of booster",
@@ -166,10 +166,10 @@ data_flowchart_rounded <-
     pct_step = n / lag(n),
     crit = str_extract(criteria, "^c\\d+"),
     criteria = fct_case_when(
-      crit == "c0" ~ "Aged 18+ and recieved booster dose of BNT162b2 or Moderna between 29 October 2021 and 31 January 2022", # paste0("Aged 18+\n with 2 doses on or before ", format(study_dates$lastvax2_date, "%d %b %Y")),
+      crit == "c0" ~ "Aged 18+ and recieved booster dose of BNT162b2 or mRNA-1273 between 29 October 2021 and 31 January 2022", # paste0("Aged 18+\n with 2 doses on or before ", format(study_dates$lastvax2_date, "%d %b %Y")),
       crit == "c1" ~ "  with no missing demographic information",
-      crit == "c2" ~ "  with homologous primary vaccination course of pfizer or AZ",
-      crit == "c3" ~ "  and not a HSC worker",
+      crit == "c2" ~ "  with homologous primary vaccination course of BNT162b2 or ChAdOx1",
[…]

### 2022-07-06 analysis/match_report.R

@@ -272,8 +272,7 @@ var_labels <- list(
 
   prior_tests_cat ~ "Number of SARS-CoV-2 tests",
 
-  prior_covid_infection ~ "Prior documented SARS-CoV-2 infection",
-  inhospital_planned ~ "In hospital (planned admission)"
+  prior_covid_infection ~ "Prior documented SARS-CoV-2 infection"
 ) %>%
   set_names(., map_chr(., all.vars))
 
@@ -402,13 +401,13 @@ data_flowchart_match <-
     pct_step = n / lag(n),
     crit = str_extract(criteria, "^c\\d+"),
     criteria = fct_case_when(
-      crit == "c0" ~ "Aged 18+ and recieved booster dose of BNT162b2 or Moderna between 29 October 2021 and 31 January 2022", # paste0("Aged 18+\n with 2 doses on or before ", format(study_dates$lastvax2_date, "%d %b %Y")),
+      crit == "c0" ~ "Aged 18+ and recieved booster dose of BNT162b2 or mRNA-1273 between 29 October 2021 and 31 January 2022", # paste0("Aged 18+\n with 2 doses on or before ", format(study_dates$lastvax2_date, "%d %b %Y")),
       crit == "c1" ~ "  with no missing demographic information",
-      crit == "c2" ~ "  with homologous primary vaccination course of pfizer or AZ",
-      crit == "c3" ~ "  and not a HSC worker",
+      crit == "c2" ~ "  with homologous primary vaccination course of BNT162b2 or ChAdOx1",
+      crit == "c3" ~ "  and not a health and social care worker",
       crit == "c4" ~ "  and not a care/nursing home resident, end-of-life or housebound",
       crit == "c5" ~ "  and no COVID-19-related events within 90 days",
-      crit == "c6" ~ "  and not in hospital at time of booster",
+      crit == "c6" ~ "  and not admitted in hospital at time of booster",
       crit == "c7" ~ "  and successfully matched",
       TRUE ~ NA_character_
     )
@@ -444,13 +443,13 @@ data_flowchart_match_rounded <-
     pct_step = n / lag(n),
     crit = str_extract(criteria, "^c\\d+"),
     criteria = fct_case_when(
-      crit == "c0" ~ "Aged 18+ and recieved booster dose of BNT162b2 or Moderna between 29 October 2021 and 31 January 2022", # paste0("Aged 18+\n with 2 doses on or before ", format(study_dates$lastvax2_date, "%d %b %Y")),
[…]

### 2022-07-06 analysis/data_process.R
fixed ICU admissions

@@ -248,11 +248,6 @@ data_processed <- data_extract %>%
     # earliest covid event after study start
     anycovid_date = pmin(postest_date, covidemergency_date, covidadmitted_date, coviddeath_date, na.rm=TRUE),
 
-    #covidadmitted_ccdays = as.integer(as.character(covidadmitted_ccdays)),
-    #FIXME ignores any covid-icu episodes subsequent to the first known covid admission, so we need to extract multiple covid admissions to capture them
-    #covidicu_date = if_else(covidadmitted_ccdays>0, covidadmitted_date, Date(NA)),
-
-
     covidcritcare_date = case_when(
       as.numeric(as.character(potentialcovidcritcare_1_ccdays))>0 & !is.na(potentialcovidcritcare_1_date) ~ potentialcovidcritcare_1_date,
       as.numeric(as.character(potentialcovidcritcare_2_ccdays))>0 & !is.na(potentialcovidcritcare_2_date) ~ potentialcovidcritcare_2_date,

### 2022-07-06 analysis/data_process.R
include covid deaths in ICU event definition

@@ -255,6 +255,8 @@ data_processed <- data_extract %>%
       TRUE ~ as.Date(NA_character_)
     ),
 
+    covidcritcare_date = pmin(covidcritcare_date, coviddeath_date, na.rm=TRUE),
+
     noncoviddeath_date = if_else(!is.na(death_date) & is.na(coviddeath_date), death_date, as.Date(NA_character_)),
 
 

### 2022-07-06 analysis/release_objects.R
whitesapce

@@ -18,7 +18,6 @@ for(matchset in c("A", "B")){
 
   ## pre-matching ----
 
-
   fs::file_copy(here("output", "prematch", "table1.csv"), fs::path(output_dir, "prematch_table1.csv"), overwrite = TRUE)
   fs::file_copy(here("output", "prematch", "table1by.csv"), fs::path(output_dir, "prematch_table1by.csv"), overwrite = TRUE)
 

### 2022-11-09 codelists/codelists_to_html.R
add script to create codelist table

@@ -1,0 +1,38 @@
+library('tidyverse')
+library('here')
+library('gt')
+
+# import codelists from json
+codelists <- jsonlite::read_json(
+  path=here("codelists", "codelists.json")
+)
+
+# reformat
+codelists_formatted <- enframe(codelists[[1]]) %>% unnest_wider(value) %>%
+  mutate(
+    file = name,
+    name= str_extract(id, "(?<=/)(.+)(?=/)"),
+    downloaded_at = as.Date(downloaded_at, "%Y-%m-%d")
+  )
+
+# output to html
+codelists_formatted %>%
+  select(name, url, downloaded_at) %>%
+  gt() %>%
+  cols_label(
+    name = "Name",
+    url = "URL",
+    downloaded_at = "Accessed on"
+  ) %>%
+  gtsave(here("codelists", "codelists.docx"))
+
+
+
+
+## convert codelists to tribble constructor ----
+# paste the output into an r script and build on top
+
+codelists_formatted %>%
+  select(name, id, downloaded_at) %>%
+  deparse::deparsec(as_tribble = TRUE)
+

### 2022-11-09 protocol/protocol.Rmd
28 day exclusion period, not 90

@@
 * they are not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care;
-* there is no evidence of SARS-CoV-2 infection or COVID-19 disease within the 90 days prior to booster dose;
+* there is no evidence of SARS-CoV-2 infection or COVID-19 disease within the 28 days prior to booster dose;
 * they are not undergoing an unplanned hospital admission at the time of boosting;

### 2022-11-09 analysis/data_selection.R

@@ -68,7 +68,7 @@ data_criteria <- data_processed %>%
     has_knownvax1 = vax1_type %in% c("pfizer", "az"),
     has_knownvax2 = vax2_type %in% c("pfizer", "az"),
     has_expectedvax3type = vax3_type %in% c("pfizer", "moderna"),
-    has_norecentcovid = ((vax3_date - anycovid_0_date) >= 90) | is.na(anycovid_0_date),
+    has_norecentcovid = ((vax3_date - anycovid_0_date) >= 28) | is.na(anycovid_0_date),
     isnot_inhospital = !inhospital,
 
     jcvi_group_6orhigher = jcvi_group %in% as.character(1:6),
@@ -140,7 +140,7 @@ data_flowchart <-
       crit == "c2" ~ "  with homologous primary vaccination course of BNT162b2 or ChAdOx1",
       crit == "c3" ~ "  and not a health and social care worker",
       crit == "c4" ~ "  and not a care/nursing home resident, end-of-life or housebound",
-      crit == "c5" ~ "  and no COVID-19-related events within 90 days",
+      crit == "c5" ~ "  and no COVID-19-related events within 28 days",
       crit == "c6" ~ "  and not admitted in hospital at time of booster",
       TRUE ~ NA_character_
     )
@@ -175,7 +175,7 @@ data_flowchart_rounded <-
       crit == "c2" ~ "  with homologous primary vaccination course of BNT162b2 or ChAdOx1",
       crit == "c3" ~ "  and not a health and social care worker",
       crit == "c4" ~ "  and not a care/nursing home resident, end-of-life or housebound",
-      crit == "c5" ~ "  and no COVID-19-related events within 90 days",
+      crit == "c5" ~ "  and no COVID-19-related events within 28 days",
       crit == "c6" ~ "  and not admitted in hospital at time of booster",
       TRUE ~ NA_character_
     )

### 2022-11-09 analysis/match_report.R

@@ -403,7 +403,7 @@ data_flowchart_match <-
       crit == "c2" ~ "  with homologous primary vaccination course of BNT162b2 or ChAdOx1",
       crit == "c3" ~ "  and not a health and social care worker",
       crit == "c4" ~ "  and not a care/nursing home resident, end-of-life or housebound",
-      crit == "c5" ~ "  and no COVID-19-related events within 90 days",
+      crit == "c5" ~ "  and no COVID-19-related events within 28 days",
       crit == "c6" ~ "  and not admitted in hospital at time of booster",
       crit == "c7" ~ "  and successfully matched",
       TRUE ~ NA_character_
@@ -445,7 +445,7 @@ data_flowchart_match_rounded <-
       crit == "c2" ~ "  with homologous primary vaccination course of BNT162b2 or ChAdOx1",
       crit == "c3" ~ "  and not a health and social care worker",
       crit == "c4" ~ "  and not a care/nursing home resident, end-of-life or housebound",
-      crit == "c5" ~ "  and no COVID-19-related events within 90 days",
+      crit == "c5" ~ "  and no COVID-19-related events within 28 days",
       crit == "c6" ~ "  and not admitted in hospital at time of booster",
       crit == "c7" ~ "  and successfully matched",
       TRUE ~ NA_character_

### 2022-11-09 analysis/contrasts_combine.R
add post baseline counts of tests and positive tests

@@ -343,6 +343,10 @@ plot_estimates(coxhr, coxhr.ll, coxhr.ul, "coxhr")
 plot_estimates(irr, irr.ll, irr.ul, "irr")
 
 
+## move event counts data ----
+
+eventcounts <- read_rds(here("output", "match", matchset, "eventcounts", "eventcounts.rds"))
+write_csv(eventcounts, fs::path(output_dir, "eventcounts.csv"))
 
 ## follow-up summary statistics ----
 

### 2022-11-09 analysis/release_objects.R

@@ -38,6 +38,7 @@ for(matchset in c("A", "B")){
   fs::file_copy(here("output", "match", matchset, "combined", "contrasts_overall_rounded.csv"), fs::path(output_dir, "contrasts_overall_rounded.csv"), overwrite = TRUE)
   fs::file_copy(here("output", "match", matchset, "combined", "contrasts_20_rounded.csv"), fs::path(output_dir, "contrasts_20_rounded.csv"), overwrite = TRUE)
 
+  fs::file_copy(here("output", "match", matchset, "combined", "eventcounts.csv"), fs::path(output_dir, "eventcounts.csv"), overwrite = TRUE)
   fs::file_copy(here("output", "match", matchset, "combined", "followup_rounded.csv"), fs::path(output_dir, "followup_rounded.csv"), overwrite = TRUE)
   fs::file_copy(here("output", "match", matchset, "combined", "followup_treatment_rounded.csv"), fs::path(output_dir, "followup_treatment_rounded.csv"), overwrite = TRUE)
 

### 2022-11-09 analysis/study_definition.py

@@ -22,6 +22,7 @@ with open("./lib/design/study-dates.json") as f:
 studystart_date = study_dates["studystart_date"] 
 studyend_date = study_dates["studyend_date"]
 followupend_date = study_dates["followupend_date"]
+postestfollowupend_date = study_dates["postestfollowupend_date"]
 firstpfizer_date = study_dates["firstpfizer_date"]
 firstaz_date = study_dates["firstaz_date"]
 firstmoderna_date = study_dates["firstmoderna_date"]
@@ -1005,6 +1006,22 @@ study = StudyDefinition(
   death_date=patients.died_from_any_cause(
     returning="date_of_death",
     date_format="YYYY-MM-DD",
+  ),
+  
+  test_count = patients.with_test_result_in_sgss(
+      pathogen = "SARS-CoV-2",
+      test_result = "any",
+      returning = "number_of_matches_in_period",
+      between = ["anycovidvax_3_date", f"{postestfollowupend_date}"],
+      restrict_to_earliest_specimen_date=False
+    ),
+  
+  postest_count = patients.with_test_result_in_sgss(
+    pathogen = "SARS-CoV-2",
+    test_result = "positive",
+    returning = "number_of_matches_in_period",
+    between = ["anycovidvax_3_date", f"{postestfollowupend_date}"],
+    restrict_to_earliest_specimen_date=False
   ),
 
 

### 2022-11-09 create-project.R

@@ -156,7 +156,20 @@ action_delayedentry_contrasts <- function(
 }
 
 
-
+action_eventcounts <- function(matchset) {
+  action(
+    name = glue("eventcounts_{matchset}"),
+    run = glue("r:latest analysis/eventcounts.R"),
+    arguments = c(matchset),
+    needs = list(
+      glue("match_{matchset}"),
+      "data_selection"
+    ),
+    moderately_sensitive = lst(
+      rds = glue("output/match/{matchset}/eventcounts/*.rds"),
+    )
+  )
+}
 
 
 ## model action function ----
@@ -204,7 +217,8 @@ action_contrasts_combine <- function(
             outcome=outcomes
           ),
           "delayedentry_{matchset}_{subgroup}_{outcome}"
-        ) %>% as.list
+        ) %>% as.list,
+        list(glue("eventcounts_{matchset}"))
       ),
       moderately_sensitive = lst(
         csv = glue("output/match/{matchset}/combined/*.csv"),
@@ -584,6 +598,9 @@ actions_list <- splice(
   action_contrasts("B", "variantera", "coviddeath"),
   action_contrasts("B", "variantera", "noncoviddeath"),
 
+
+  action_eventcounts("A"),
+  action_eventcounts("B"),
 
   comment("# # # # # # # # # # # # # # # # # # #", "Combine KM / CI estimates across outcomes and subgroups", "# # # # # # # # # # # # # # # # # # #"),
 

### 2022-11-09 lib/dummydata/dummydata.R

@@ -362,6 +362,14 @@ sim_list = lst(
     missing_rate = ~0.90
   ),
 
+  test_count = bn_node(
+    ~ rpois(n = ..n, 1)
+  ),
+
+  postest_count = bn_node(
+    ~ rpois(n = ..n, 0.1)
+  )
+
 
 )
 bn <- bn_create(sim_list, known_variables = known_variables)

### 2022-11-09 run-local.R

@@ -33,8 +33,9 @@ action_match <- compose_action(
 )
 
 action_contrasts <- compose_action(
-    #here("analysis", "km.R"),
-    here("analysis", "ci.R")
+    here("analysis", "km.R"),
+    #here("analysis", "ci.R"),
+    here("analysis", "eventcounts.R")
 )
 
 action_combine <- compose_action(

### 2024-01-16 write-up/manuscript/manuscript.qmd
first working maunscript draft
- needs a lot of tidying up!

---
title: "Comparative effectiveness of Pfizer B.A4-5 versus Sanofi during the Spring 2023 COVID-19 booster programme in England: a cohort study in OpenSAFELY-TPP"
format:
  # html:
  #   toc: false
  #   embed-resources: true
  docx
knitr:
  opts_chunk: 
    collapse: true
    echo: false
    fig.path: ./write-up/figures/
    fig.pos: H
    R.options:
      knitr.graphics.auto_pdf: true
bibliography: references.bib
csl: vancouver.csl
link-citations: true
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r}
#| label: setup
#| include: false
#| message: false
#| warnings: false
source(here::here("write-up", "manuscript", "setup.R"))
```

```{r}
#| label: process
#| include: false
#| message: false
source(here("write-up", "manuscript", "import-process.R"))
```

William J Hulme^1^, Edward PK Parker^2^, Em Prestige^2^, Ruth H Keogh^2^, Elizabeth J Williamson^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Colm Andrews^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^6,7^\*, Jonathan AC Sterne^3,5,8^\*, and the OpenSAFELY collaborative.

1\. The Bennett Institute for Applied Data Science, Nuffield Department of Primary Care Health Sciences, University of Oxford, OX26GG, UK

2\. London School of Hygiene and Tropical Medicine, Keppel Street, London, WC1E 7HT, UK

3\. Population Health Sciences, University of Bristol, Oakfield House, Oakfield Grove, Bristol, BS8 2BN, UK

4\. MRC Integrative Epidemiology Unit, Bristol Medical School, University of Bristol, Bristol, UK

5\. NIHR Bristol Biomedical Research Centre, Bristol, UK

6\. CAUSALab, Harvard T.H. Chan School of Public Health, Boston, MA 02115

7\. Departments of Epidemiology and Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA 02115

8\. Health Data Research UK South West

\newpage

## Abstract

*Introduction* The spring 2023 COVID-19 booster vaccination programme in England used both Pfizer Ba45 and Sanofi vaccines. All people aged 75 years and over, or considered to be clinically vulnerable were eligible to receive a booster dose. Direct comparisons of the effectiveness against severe COVID-19 of these two vaccines for boosting have not been made in trials or observational data.

*Methods* On behalf of NHS England, we used the OpenSAFELY-TPP database to compare recipients of each vaccine, investigating the over 75s and the clinically vulnerable groups separately. Recipients were matched on date of vaccination, vaccine history, age, and other characteristics. Recipients were included if boosted between 1 April and 30 June 2023, and followed up for X weeks. Outcomes were COVId-19 emergency department attendance, COVID-19 hospital admission, and COVID-19 death. We estimated the cumulative incidence of each outcome, and quantified comparative effectiveness using risk differences (RD) and incidence rate ratios (IRRs).

*Results* `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4") %>% pull(n) %>% sum())` people were 1-1 matched in the CV cohort, and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4") %>% pull(n) %>% sum())` in the 75+ cohort, contributing a total `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up respectively.

In both CV and 75+ cohorts, effectiveness against all measured COVID-19-related outcomes was similar. This finding was observed overall and in pre-specified subgroups.

In the CV cohort, the X-week risks per `r perN_format` of COVID-19 A&E attendance were `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer B.A4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_1)` for Sanofi: the IRR comparing Sanofi with PfizerB.A4-5 was `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(irrCI)`. For COVID-19 hospitalisation, X-week risks were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)`: IRR `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`. For COVID-19 death: `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)`: IRR `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)`.

In the 75+ cohort, the X-week risks per `r perN_format` for COVID-19 A&E attendance were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer B.A4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_1)` for Sanofi: the IRR comparing Sanofi with PfizerB.A4-5 was `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(irrCI)`. For COVID-19 hospitalisation, X-week risks were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)`: IRR `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`. For COVID-19 death: `r perN_format` were `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)`: IRR `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)`.

*Conclusion* This observational study comparing effectiveness of PFizer and Sanofi vaccine during the Spring 2023 programme in England in the two main eligible cohorts -- people aged 75 and over and in clinically vulnerable people -- found no evidence of substantial differences between the two vaccine products within the first X weeks after vaccination.

\newpage

```{r break1}
2+2
```

## Background

In X 2022, NHS XXX in England announced the Spring 2023 COVID-19 booster vaccination programme, available to those aged 75 years and over, and those considered to be vulnerable to infection or severe COVID-19 disease, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI) @covid19,. The use of the both the Pfizer-BioNTech B.A.4/5 vaccine and the Sanofi vaccine during the spring 2023 programme enables a direct comparison of their effectiveness against severe COVID-19 disease, which has not yet been conducted in randomised trials.

On behalf of NHS England, we used the OpenSAFELY-TPP database, covering around 45% of English primary care practices and linked to hospital and death records, to compare the effectiveness of boosting with Pfizer and Sanofi in the 75+ years group and the clinically vulnerable group who received either vaccine in the spring 2023 programme. During this period and the weeks following, the X variant was dominant.

\newpage

## Methods

### Data source

All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>. With the approval of NHS England, primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). COVID-19 vaccination history and health and social care worker status is available in the GP record directly via the National Immunisation Management System (NIMS).

### Eligibility criteria

We considered two cohorts: all adults aged 75 years or over ("75+"), and all adults aged 18 or over who were classified as clinically vulnerable ("CV"). These groups are not mutually exclusive. Clinical vulnerability encompasses various chronic or immunocompromising conditions, immunosuppressive treatments, and ..., and is defined fully in supplementary materials.

In each group, any person receiving a Pfizer or Sanofi vaccine dose between `r study_dates_format$studystart_date` and `r study_dates_format$studyend_date` inclusive were considered for inclusion. People were eligible if they: were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; were not a health or social care worker, not resident in a care or nursing home and not medically housebound or receiving end-of-life care; had no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; were not hospitalised at the time of boosting; and had complete information on sex, deprivation, and Sustainability and Transformation Partnership (STP, a geographical grouping of NHS and local authorities).

### Matching

Pfizer and Sanofi booster recipients were matched 1-1 without replacement, on the following characteristics: date of vaccination; number of previous COVID-19 vaccine doses; time since previous COVID-19 vaccine dose; sex (male or female); age (3 year caliper and within age groups defined by JVCI risk groups); Index of Multiple Deprivation (IMD, grouped by quintile); STP as a surrogate for geographical region; evidence of prior SARS-CoV-2 infection (any of positive SARS-CoV-2 test, probable infection documented in primary care, or COVID-19 hospital attendance or admission); and morbidity count (grouped as 0, 1, or 2 or more conditions from the following list: diabetes, BMI over 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease, cancer within 3 years). Clinical vulnerability was an additional matchng factor in the 75+ cohort.

There are no missing values for these variables: people with missing values for sex, IMD or STP were excluded, and all other variables were defined by the presence or absence of clinical codes or events.The supplementary materials provide more information on how these characteristics were chosen and defined.

### Outcomes and follow-up

We considered three primary outcomes. COVID-19 emergency department admission was identified using HES emergency department records, with SNOMIED.

COVID-19 hospitalisation was identified using Secondary Use Service (SUS) in-patient hospital records with International Statistical Classification of Diseases and Related Health Problems 10th Revision (ICD-10) codes U07.1, U07.2, or U.109 as the primary or non-primary reason for admission. COVID-19 death was defined as death with U07.1, U07.2, or U.109 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death). We also report on non-COVID-19 deaths.

<!-- We also considered the follows safety outcomes: -->

<!-- * Adverse allergic reaction / Anaphylaxis -->

<!-- * Myocarditis or Pericarditis -->

<!-- *  -->

<!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->

Each person was followed from receipt of vaccination (time zero) until the outcome of interest, with censoring at death, practice de-registration, XX weeks, or `r study_dates_format$followupend_date`.

### Statistical Analysis

Baseline characteristics of each vaccine group were tabulated, with between-group balance examined using standardised mean differences. We estimated the cumulative incidence of each outcome in each vaccine group using the Kaplan-Meier (KM) estimator. We estimated cumulative risk differences and risk ratios (RRs) comparing the vaccine groups for each outcome. These were estimated up to 20 weeks where available, but estimates at X weeks are reported as the primary comparison to ensure potential follow-up in those vaccinated towards the end of the booster programme was not censored. We also estimated Incidence Rate Ratios (IRRs) over the first X weeks, and within period-specific intervals defined by splits on weeks 1, 2, 4, 8, 12, 16, and X, where available.

<!-- days 7, 14, 21, 28, 42, 56, 70, 84, 112, 140, 168 and 196. -->

<!-- 95% confidence limits for the AJ estimates were derived from a Greenwood-type formula. Confidence limits for the risk difference were derived from the sum of squares of the AJ standard errors, derived from Greenwood’s formula. Confidence limits for the survival ratio were based on the standard error of the cumulative hazard. Confidence limits for the risk ratio were derived using the Delta method. -->

95% confidence limits for the cumulative incidences were derived using standard errors on the log-KM scale. Confidence limits for the risk differences were derived from the sum of squares of the standard errors on the KM scale, using Greenwood's formula. Confidence limits for the risk ratio were derived from the sum of squares of the standard errors on the log-KM scale. Confidence limits for the IRRs were derived from the sum of squares of the standard errors on the log-KM scale. <!-- 95% confidence limits for the Kaplan-Meier estimates, risk differences, risk ratios, and incidence rate ratios are derived by bootstrapping the matched pairs. 500 bootstrap replicates will be used. -->

### Subgroup and secondary analyses

We estimated comparative effectiveness separately in the following subgroups: age band in the CV group only (18-39, 40-54, 55-64, 65-74, 75+); clinical vulnerability in the 75+ group only; prior vaccination history; evidence of prior SARS-CoV-2 infection or not. We used $\chi^2$ tests to examine evidence for heterogeneity in comparative effectiveness between subgroups.

We reported the rates of SARS-CoV-2 tests undertaken from baseline up to 31 March 2022 in each vaccine group, as lower testing rates in one group might lead to under-ascertainment of outcomes, particularly positive SARS-CoV-2 tests.

We reported outcomes at one week as a negative control outcome @Lipsitch2010, during which time the vaccine is not expected to have any protective immunological effect.

### Disclosure control

To satisfy strict re-identification minimisation requirements for statistical outputs from OpenSAFELY's Trusted Research Environment, counts were rounded to the nearest 3, 9, 15, and so on. Plots of cumulative event counts and the Kaplan-Meier cumulative incidence estimates were rounded such that each increment is based on at least 6 events. Event rates, risk differences, and risk ratios were derived from these rounded estimates.

### Software, code, and reproducibility

Data management and analyses were conducted using OpenSAFELY tools, Python version 3.8.10 and R version 4.0.2. All code is shared openly for review and re-use under MIT open license at <https://github.com/opensafely/comparative-booster-spring2023>. Codelists are available at <https://www.opencodelists.org/>. The supplementary materials provide further details of codelists and data sources used for all variables in the study. Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.

\newpage

## Results

```{r break2}
2+2
```

### Study population and matching

\[\[fix total identified in data_process script then re-write\]\]

<!-- We identified `r label_number(1, big.mark=",")(filter(rst$cv$total_flowchart, boost_type=="Any") %>% pull(n) %>% sum)` clinically vulnerable people and `r label_number(1, big.mark=",")(filter(rst$age75plus$total_flowchart, boost_type=="Any") %>% pull(n) %>% sum)` people aged 75 or over who were recorded as having received a COVID-19 vaccine between `r study_dates_format$studystart_date` and `r study_dates_format$studyend_date` whilst registered at a TPP practice. This included `r label_number(1, big.mark=",")(filter(rst$age75plus$total_flowchart, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$total_flowchart, boost_type=="pfizerBA45") %>% pull(pct))`) Pfizer B.A4/5 and `r label_number(1, big.mark=",")(filter(rst$cv$total_flowchart, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$total_flowchart, boost_type=="sanofi") %>% pull(pct))`) Sanofi. -->

<!-- `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c0") %>% pull(n) %>% sum())` were in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c1", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c1", boost_type=="pfizerBA45") %>% pull(pct_all))`) received Pfizer BA45 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c1", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c1", boost_type=="sanofi") %>% pull(pct_all))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1). -->

`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c0") %>% pull(n) %>% sum())` were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c1", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c1", boost_type=="pfizerBA45") %>% pull(pct_all))`) received Pfizer BA45 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c1", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c1", boost_type=="sanofi") %>% pull(pct_all))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).

Eligible Pfizer BA45 recipients were on average older, more deprived and had higher rates of prior clinical conditions than Sanofi recipients (supplementary Table S1). They were also more likely to have received more/less/X prior COVID-19 vaccine doses...

In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1).

In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1).

Characteristics at the start of follow-up were well-balanced between groups (Table 1) with standardised mean differences consistently below 0.05 (supplementary Figure S2 and Table S1).

In particular, prior clinical conditions, which were not directly matched on, were well-balanced between the groups after, though not before, matching. The median follow-up was X weeks for COVID-19 death outcomes (Table S3).

SARS-CoV-2 testing rates after baseline were slightly less frequent for Pfizer BA45 than for Sanofi: `r rst$cv$eventcounts %>% filter(subgroup=="all", treatment==0) %>% pull(test_rate) %>% {label_number(0.001, 7)(.)}` versus `r rst$cv$eventcounts %>% filter(subgroup=="all", treatment==1) %>% pull(test_rate) %>% {label_number(0.001, 7)(.)}` per week for the CV cohort; and `r rst$age75plus$eventcounts %>% filter(subgroup=="all", treatment==0) %>% pull(test_rate) %>% {label_number(0.001, 7)(.)}` versus `r rst$age75plus$eventcounts %>% filter(subgroup=="all", treatment==1) %>% pull(test_rate) %>% {label_number(0.001, 7)(.)}` per week for the 75+ cohort (supplementary Table S4).

### Estimated comparative effectiveness

#### CV cohort

```{r break3}
2+2
```

By X weeks there were `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(n.event)` COVID-19 A&E attendances, `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths (Table 2).

<!-- across `r contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 2). -->

The X-week risks (cumulative incidence) per `r perN_format` people of COVID-19 A&E attendances were `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA45 and `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 2). The corresponding X-week risks per `r perN_format` people for COVID-19 hospitalisation were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI)`). The X-week risks per `r perN_format` people of COVID-19 death were `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`). For non-COVID-19 death, these were `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(rdCI)`).

<!-- For COVID-19 A&E attendance, the cumulative risk ratio comparing Sanofi with Pfizer BA45 reached a minimum of 0.91 (0.90 to 0.92) at 4 weeks, then attenuated with increasing follow up (Figure 2). -->

<!-- For COVID-19 hospitalisation the minimum risk ratio was 0.67 (0.59 to 0.76) at around 11 weeks, attenuating to 0.89 (0.83 to 0.96) by 28 weeks. -->

<!-- The number of COVID-19 deaths was too small to reliably examine patterns of cumulative incidence over time. -->

<!-- For non-COVID-19 death the risk ratio was imprecisely estimated in the first few weeks, but stabilised to 0.92 (0.86 to 0.99) at 28 weeks. -->

IRRs (where less than one favours Sanofi) in the X-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 A&E attendances, `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)` for COVID-19 hospitalisation, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).

Period-specific IRRs are displayed in Supplementary Figure S4a. During the first week after vaccination event rates were similar between vaccine groups: the estimated HR was `r rst$cv$contrasts_cuts_table %>% filter(outcome=="covidemergency", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI_95)` for COVID-19 A&E attendances, `r rst$cv$contrasts_cuts_table %>% filter(outcome=="covidadmitted", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 hospitalisation, `r rst$cv$contrasts_cuts_table %>% filter(outcome=="coviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 death, `r rst$cv$contrasts_cuts_table %>% filter(outcome=="noncoviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for non-COVID-19 death. After this initial period, the rate for A&E was lower for Sanofi than Pfizer BA45 until week X, after which they were higher for Sanofi. For COVID-19 hospitalisation, the minimum IRR was X (X to X) during weeks 5-8, with IRRs between X and X weeks consistent with no difference between vaccine groups. For non-COVID-19 death IRRs were consistently close to 1.

#### 75+ cohort

```{r break4}
2+2
```

By X weeks there were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(n.event)` COVID-19 A&E attendances, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths (Table 2).

<!-- across `r contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 2). -->

The X-week risks (cumulative incidence) per `r perN_format` people of COVID-19 A&E attendances were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA45 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 2). The corresponding X-week risks per `r perN_format` people for COVID-19 hospitalisation were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI)`). The X-week risks per `r perN_format` people of COVID-19 death were `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`). For non-COVID-19 death, these were `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(rdCI)`).

<!-- For COVID-19 A&E attendance, the cumulative risk ratio comparing Sanofi with Pfizer BA45 reached a minimum of 0.91 (0.90 to 0.92) at 4 weeks, then attenuated with increasing follow up (Figure 2). -->

<!-- For COVID-19 hospitalisation the minimum risk ratio was 0.67 (0.59 to 0.76) at around 11 weeks, attenuating to 0.89 (0.83 to 0.96) by 28 weeks. -->

<!-- The number of COVID-19 deaths was too small to reliably examine patterns of cumulative incidence over time. -->

<!-- For non-COVID-19 death the risk ratio was imprecisely estimated in the first few weeks, but stabilised to 0.92 (0.86 to 0.99) at 28 weeks. -->

IRRs (where less than one favours Sanofi) in the X-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 A&E attendance, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)` for COVID-19 hospitalisation, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).

Period-specific IRRs are displayed in Supplementary Figure S4a. During the first week after vaccination event rates were similar between vaccine groups: the estimated HR was `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidemergency", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI_95)` for COVID-19 A&E attendances, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidadmitted", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 hospitalisation, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="coviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 death, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="noncoviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for non-COVID-19 death. After this initial period, the rate for A&E was lower for Sanofi than Pfizer BA45 until week X, after which they were higher for Sanofi. For COVID-19 hospitalisation, the minimum IRR was X (X to X) during weeks 5-8, with IRRs between X and X weeks consistent with no difference between vaccine groups. For non-COVID-19 death IRRs were consistently close to 1.

### Subgroup analyses

Findings in the main analyses were similar when examining subgroups...

<!-- For COVID-19 A&E attendances the estimated relative benefit of Sanofi compared with Pfizer BA45 was greater/lower in those with fewer prior vaccines (IRR `r contrasts_table %>% filter(outcome=="covidemergency", subgroup=="covid_vax_history", subgroup_level=="pfizer-pfizer") %>% pull(irrCI_95)`) than in those who received ChAdOx1 (IRR `r contrasts_table %>% filter(outcome=="covidemergency", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(irrCI_95)`, heterogeneity p=`r contrasts_table %>% filter(outcome=="covidemergency", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhr.ln.p) %>% first()`) (Figure 3). -->

<!-- Similarly, the estimated relative benefit of mRNA-1273 compared with BNT162b2 was greater in those with prior evidence of infection (IRR `r contrasts_table %>% filter(outcome=="covidemergency", subgroup_level=="TRUE", subgroup=="prior_covid_infection") %>% pull(irrCI_95)`) versus those without (IRR `r contrasts_table %>% filter(outcome=="covidemergency", subgroup_level=="FALSE", subgroup=="prior_covid_infection") %>% pull(irrCI_95)`, heterogeneity p=`r contrasts_table %>% filter(outcome=="covidemergency", subgroup_level=="FALSE", subgroup=="prior_covid_infection") %>% pull(coxhr.ln.p) %>% first()`). -->

<!-- For COVID-19 hospitalisation the estimated relative benefit of mRNA-1273 compared with BNT162b2 was similar by type of first two vaccine doses (IRR `r contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="vax12_type", subgroup_level=="pfizer-pfizer") %>% pull(irrCI_95)` for BNT162b2 and `r contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(irrCI_95)` for ChAdOx1; heterogeneity p=`r contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhr.ln.p) %>% first()`). -->

<!-- The estimated relative benefit of mRNA-1273 was greater in those with prior evidence of infection (IRR `r contrasts_table %>% filter(outcome=="covidadmitted", subgroup_level=="TRUE", subgroup=="prior_covid_infection") %>% pull(irrCI_95)`) than those without (IRR `r contrasts_table %>% filter(outcome=="covidadmitted", subgroup_level=="FALSE", subgroup=="prior_covid_infection") %>% pull(irrCI_95)`, heterogeneity p=`r contrasts_table %>% filter(outcome=="covidadmitted", subgroup_level=="FALSE", subgroup=="prior_covid_infection") %>% pull(coxhr.ln.p) %>% first()`). -->

<!-- For COVID-19 death, IRRs were estimated imprecisely due to low event rates so no strong conclusions about variation between subgroups could be made. -->

<!-- For Non-COVID-19 death, the relative benefit of mRNA-1273 was generally greater in older and more clinically vulnerable people. -->

Supplementary materials provide cumulative incidence plots by subgroup (Figures S3a-e), the corresponding cumulative risk differences, cumulative risk ratios, and period-specific IRRs (Figures S4a-e), and subgroup-specific heterogeneity tests for risk differences and risk ratios (Table S2).

\newpage

## Discussion

This observational study in over X million adults receiving a COVID-19 vaccine during the Spring 2023 booster programme in England compared the effectiveness of Sanofi versus Pfizer BA45 vaccines and is, to our knowledge, the first to do so against severe COVID-19 outcomes. We estimated that ... there were no substantial differences between either vaccine on any of the studies effectiveness outcomes during the first X weeks after vaccination. Estimates appeared broadly similar regardless of age, clinical vulnerability, and ...

Rates of COVID-19-related events were higher (the cumulative incidence curve is stepper) during the immune induction period up to 2 weeks, during which time the booster dose has yet to take full effect, than subsequently. Consistent with this delayed immunity, rates of the outcomes were similar for the two vaccines in the first week of follow-up, though rates also remained similar after this period.

Only `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` people aged 75 or over died with COVID-19 recorded as an underlying or contributory cause across both vaccine groups within X weeks (1 in 38,500) so the IRR was estimated imprecisely. <!-- The risk of non-COVID-19 death was around 1 in 1,600 people, and was marginally lower/higher for Sanofi than Pfizer BA45. --> <!-- This could indicate differences in vaccine safety, under-ascertainment of COVID-19 related deaths (for example post-COVID cardiovascular deaths not directly attributed to SARS-CoV-2 infection), or sub-optimal confounding control. -->

### Strengths and limitations

The OpenSAFELY-TPP database covers around 45% of the English population and contains rich clinical information which enabled us to closely match Sanofi and Pfizer BA45 recipients to control for potential confounding, to study a range of clinical outcomes, including severe COVID-19, and compare comparative safety and effectiveness between important clinical subgroups.

To make fair comparisons between the two vaccine types, we exploited the concurrent roll-out of both vaccines across the same eligible population during the same time period. The type of vaccine administered was based largely on local supply and availability, and clinical characteristics did not inform the type of vaccine offered on-site \[except for age??\]. It remained important to control for any potential differences in the distribution of prognostic factors between the two vaccine groups, and these were well-balanced after matching. We cannot rule out residual confounding as we did not match on or adjust for all potential confounders, for example some clinical subgroups where relative protection between vaccines may differ such as immunosuppressive conditions. However, unmeasured potential confounders, particularly relating to unmeasured health-seeking behaviours, are less problematic in comparative effectiveness studies than in studies comparing vaccinated with unvaccinated people, as all people entering the study have sought and received at three vaccine doses. We found no evidence that rates of other outcomes during the first week after vaccination differed between the vaccine groups, although some of these comparisons were estimated imprecisely.

<!--# [People were usually not offered a choice of vaccine type (although they could refuse boosting if they were not offered their preference) and clinical characteristics did not inform the type of vaccine offered. Anecdotally, we understand that occasionally some people were offered a choice in centres using both vaccines, though we are not aware of any specific policy that allowed this.] -->

We excluded groups who were unlikely to undergo boosting such as those in palliative care. We also excluded those identified as health and social care workers due to these groups being much more likely to receive their vaccines at their place of work, rather than at walk-in or bookable community vaccination centres, which may introduce bias. For example, occupation type for health care workers may influence both vaccine administration setting (and therefore vaccine type) and risk of infection.

<!-- The later introduction of the mRNA-1273 vaccine for the UK booster campaign meant that the matched cohort contained younger and healthier people than the whole population receiving a third dose, due to the prioritisation of older and higher risk people. -->

<!-- For example, the number of people with immunosuppressive conditions in the matched cohort was just 1.5%. -->

<!-- The applicability of our findings to the most vulnerable groups may therefore be limited. -->

<!-- However, numbers remained sufficient to compare subgroups defined by age, clinical vulnerability, and prior infection status, including for COVID-19 hospitalisation. -->

To control for spatio-temporal heterogeneity in infection risk during the study period we matched on the date of third dose receipt and local health administration (STP). Some residual heterogeneity within STP regions is possible, but more precise geographical matching was not feasible due to poor matching success at this level, partly due to time periods during which only one vaccine type was available within a small geographical area.

Post-baseline SARS-CoV-2 testing rates were slightly higher in X than in X recipients, which largely rule out differences in testing behaviours between vaccine groups.

This study does not include anyone who received the Moderna mRNA-1273 vaccine which was also available during the spring 2023 booster programme but was administered much more infrequently.

### Findings in context

<!-- The COV-BOOST phase II randomised trial assessed the safety and immunogenicity of seven COVID-19 vaccines for boosting, including BNT162b2 and mRNA-1273 against a menACWY control in people with no prior history of infection @munro2021. -->

<!-- It found higher immunogenicity of mRNA-1273 compared with BNT162b2 in subgroups with a two-dose BNT162b2 primary course (geometric mean ratio (GMR) of SARS-CoV-2 anti-spike IgG versus control = 11.49 (95%CI 9.36 to 14.12) versus (6.78 (5.51 to 8.35)) and a two-dose ChAdOx1 primary course (GMR = 32.30 (24.84 to 42.01) versus 16.80 (12.97 to 21.76)). -->

<!-- Likewise, the MixNMatch phase I-II non-randomised trial documented higher antibody levels after mRNA-1273 boost than BNT162b2 boost among individuals who received a primary course of BNT162b2 @atmar2022. -->

<!-- These trials did not evaluate efficacy endpoints, and we are aware of no other planned or published randomised trials making direct comparisons of boosting with BNT162b2 versus mRNA-1273. -->

<!-- A study in Spanish registry data of people aged 40 years or over with no previous positive test for SARS-CoV-2 examined effectiveness of mRNA vaccine boosting, including a comparison of BNT162b2 with mRNA-1273 @monge2022. -->

<!-- For positive SARS-CoV-2 tests, it estimated a 34-day risk difference comparing mRNA-1273 with BNT162b2 of −2.2 (−2.7 to −1.6) per 1,000 people (risk ratio 0.87 (0.84, 0.90)). -->

<!-- There was no assessment of effectiveness against hospitalisation or other severe outcomes. -->

<!-- A study in US Veteran Affairs health care system in the Alpha and Delta eras using a similar matching approach @dickermanComparativeEffectivenessThird2023, with 65,196 matched recipients in each vaccine group, reported 16-week risk ratio of 0.87 (0.77 to 0.94) for documented SARS-CoV-2 infection, 0.61 (0.36 to 0.79) for COVID-19 hospitalisation and 0.92 (0.16 to 2.17) for COVID-19 death. -->

<!-- Comparisons of vaccines for primary vaccination in observational data have also been made. -->

<!-- Another US Veteran Affairs study by the same group during the Alpha period @dickerman2022, identified an additional 1.23 (95% CI 0.72 to 1.81) documented infections per 1,000 people in those with a BNT162b2 first dose compared with a mRNA-1273 first dose at 24 weeks, 0.55 (95%CI 0.36 to 0.83) COVID-19 hospitalisations, 0.10 (95%CI 0.00 to 0.26) COVID-19 ICU admissions, and 0.02 (95%CI −0.06 to 0.12) COVID-19 deaths. -->

<!-- Another study using Veterans Affairs data with exact- and propensity-matching in a similar time period @ioannou2022a, estimated similar differences at 24 weeks for documented infection (1.73, 95%CI 1.50 to 1.96 additional events per 1,000 people in those with a BNT162b2 first dose compared with a mRNA-1273), COVID-19 hospitalisation (0.56, 95%CI 0.45 to 0.67) and COVID-19 death (0.03, 95%CI -0.00 to 0.07). -->

<!-- Accounting for the longer follow-up period and statistical uncertainty, these estimates are consistent with our findings for the comparative effectiveness of mRNA-1273 versus BNT162b2 boosting. -->

<!-- A third Veteran Affairs study compared three doses of BNT162b2 with three doses of mRNA-1273 in the Omicron era in a secondary analysis. https://doi.org/10.1093/cid/ciac328 -->

<!-- A study using English data to estimate the effectiveness of BNT162b2 boosting versus no boosting and mRNA-1273 boosting versus no boosting separately using a test-negative-control design @andrews2022, suggests marginally higher effectiveness against symptomatic disease in mRNA-1273 than in BNT162b2, but this assumes that the respective control populations in each analysis were similar. -->

<!-- The evidence therefore strongly points to a benefit of mRNA-1273 over BNT162b2 for primary vaccination and subsequent booster doses. -->

<!-- This is relevant to vaccine procurement decisions for future booster programmes. -->

<!-- However, both vaccines are safe [@barda2021a; @klein2021; @diaz2021; @dickerman2022a] and strongly effective [@hulme2022; @barda2021; @andrews2022; @monge2022] against infection and COVID-19 disease, compared to no boosting. -->

<!-- Findings from this study should not discourage individuals from receiving BNT162b2 booster vaccination if offered. -->

<!-- Though small, an additional 0.22 hospitalisations per 1,000 people becomes 100s or over 1,000 hospitalisations in England at a population level, depending on booster uptake and infection risk, contributing substantially to healthcare resource use both during the acute phase of disease and by those with longer-term illness -->

### Conclusions

COVID-19-related outcomes after vaccination with Sanofi or Pfizer BA45 during the Spring 2023 booster programme in England were rare, though risks were estimated to be similar between vaccine types. These finding were broadly consistent across subgroups. Given the similarly in effectiveness of the two vaccines compared in this study, future COVID-19 booster programmes may focus on other characteristics of the vaccines, for instance cost effectiveness or logistical constraints, to determine which to recommend for roll-out.

\newpage

## Contributions

Conceptualization: WJH, EPKP, RHK, EJW, VW, TP, BG, MAH, and JACS. Data curation: WJH, HJC, AW, AM, BM, CA, and SCJB. Formal analysis: WJH, EPKP, and VW. Funding acquisition: JM, BG, and JACS. Methodology: WJH, EPKP, RHK, EJW, VW, TP, MAH, and JACS. Project administration: WJH, AM, JM, BG, CA, and JACS. Resources: WJH, HJC, AW, AM, JM, BM, SCJB, and BG. Software: WJH, EPKP, and SCJB. Supervision: MAH and JACS. Validation: WJH, EPKP, CA, and VW. Visualization: WJH, and EPKP. Writing - original draft: WJH. Writing - review & editing: WJH, EPKP, RHK, EJW, VW, TP, HJC, AW, CA, AM, JM, BM, SCJB, BG, MAH, and JACS.

## Information governance and ethical approval

NHS England is the data controller for OpenSAFELY-TPP; TPP is the data processor; and study authors using OpenSAFELY have the approval of NHS England. This implementation of OpenSAFELY is hosted within the TPP environment which is accredited to the ISO 27001 information security standard and is NHS IG Toolkit compliant @datasec; Patient data has been pseudonymised for analysis and linkage using industry standard cryptographic hashing techniques; all pseudonymised datasets transmitted for linkage onto OpenSAFELY are encrypted; access to the platform is via a virtual private network (VPN) connection, restricted to a small group of researchers; the researchers hold contracts with NHS England and only access the platform to initiate database queries and statistical models; all database activity is logged; only aggregate statistical outputs leave the platform environment following best practice for anonymisation of results such as statistical disclosure control for low cell counts @isb1523 . The OpenSAFELY research platform adheres to the obligations of the UK General Data Protection Regulation (GDPR) and the Data Protection Act 2018. In March 2020, the Secretary of State for Health and Social Care used powers under the UK Health Service (Control of Patient Information) Regulations 2002 (COPI) to require organisations to process confidential patient information for the purposes of protecting public health, providing healthcare services to the public and monitoring and managing the COVID-19 outbreak and incidents of exposure; this sets aside the requirement for patient consent @coronavi2020 . Taken together, these provide the legal bases to link patient datasets on the OpenSAFELY platform. General practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.

This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by the LSHTM Ethics Board (reference 21863).

JACS is the guarantor.

## Funding

This work was jointly funded by UKRI \[COV0076;MR/V015737/1\], the Longitudinal Health and Wellbeing strand of the National Core Studies programme (MC_PC_20030; MC_PC_20059; COV-LT-0009), NIHR and Asthma UK-BLF. The OpenSAFELY data science platform is funded by the Wellcome Trust (222097/Z/20/Z).

BG's work on better use of data in healthcare more broadly is currently funded in part by: the Bennett Foundation, the Wellcome Trust, NIHR Oxford Biomedical Research Centre, NIHR Applied Research Collaboration Oxford and Thames Valley, the Mohn-Westlake Foundation; all Bennett Institute staff are supported by BG's grants on this work. EW holds grants from MRC. RHK was funded by UK Research and Innovation (Future Leaders Fellowship MR/S017968/1).

The views expressed are those of the authors and not necessarily those of the NIHR, NHS England, UK Health Security Agency (UKHSA) or the Department of Health and Social Care.

Funders had no role in the study design, collection, analysis, and interpretation of data; in the writing of the report; and in the decision to submit the article for publication.

For the purpose of Open Access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission.

## Data access and verification

Access to the underlying identifiable and potentially re-identifiable pseudonymised electronic health record data is tightly governed by various legislative and regulatory frameworks, and restricted by best practice. The data in OpenSAFELY is drawn from General Practice data across England where TPP is the data processor.

TPP developers initiate an automated process to create pseudonymised records in the core OpenSAFELY database, which are copies of key structured data tables in the identifiable records. These pseudonymised records are linked onto key external data resources that have also been pseudonymised via SHA-512 one-way hashing of NHS numbers using a shared salt. Bennett Institute for Applied Data Science developers and PIs holding contracts with NHS England have access to the OpenSAFELY pseudonymised data tables as needed to develop the OpenSAFELY tools.

These tools in turn enable researchers with OpenSAFELY data access agreements to write and execute code for data management and data analysis without direct access to the underlying raw pseudonymised patient data, and to review the outputs of this code. All code for the full data management pipeline---from raw data to completed results for this analysis---and for the OpenSAFELY platform as a whole is available for review at github.com/OpenSAFELY.

\newpage

## Figures and tables

#### Table 1: Baseline characteristics before and after matching

```{r table-table1}
#| echo: false
#| warning: false
#| message: false

# tab1 <- rst$cv$match_table1_wide %>%
#   filter(
#     strategy=="post",
#     !(variable  %in% c("N", "immunosuppressed", "region"))
#   ) %>%
#   mutate(
#     smd=if_else(variable %in% c(matching_variables$B$exact, "ageband"), NA_character_, smd),
#     var_label = fct_recode(var_label, "Any immunosuppressive condition"= "Immunosuppressed (all)"),
#     order_level = row_number(),
#     order_var = cumsum(lag(variable, 1, first(variable))!=variable),
#     var_label = if_else(lag(as.character(var_label),n=1,default="")==var_label, "", as.character(var_label)),
#   ) %>%
#   select(var_label, variable_levels, stat_0, stat_1, smd, order_var, level) %>%
#   gt(
#     #groupname_col = "var_label"
#   ) %>%
#   cols_label(
#     var_label = "Variable",
#     variable_levels = " ",
#     `stat_0` = glue("Pfizer BA45 (N={filter(rst$cv$match_table1, strategy=='post', variable=='N', by==0) %>% pull(n)})"),
#     `stat_1` = glue("Sanofi (N={filter(rst$cv$match_table1, strategy=='post', variable=='N', by==1) %>% pull(n)})"),
#     `smd` = "Standardised mean difference"
#   ) %>%
#   cols_align(
#     align = c("right"),
#     columns =  c("stat_0", "stat_1")
#   ) %>%
#   cols_align(
#     align = c("left"),
#     columns =  c("var_label", "variable_levels")
#   ) %>%
#   sub_missing(
#       columns = "smd",
#       missing_text = "-"
#   ) %>%
#   tab_style(
#     style = list(
#       cell_text(weight = "bold")
#     ),
#     locations = cells_column_labels(everything())
#   ) %>%
#   tab_style(
#     style = list(
#       cell_fill(color = "gray96")
#     ),
#     locations = cells_body(
#       rows = order_var%%2 == 0
#     )
#   ) %>%
#   cols_hide(c("level", "order_var")) %>%
#   tab_options(
#     #row_group.as_column = TRUE,
#     table.font.size = 8
#   )

match_table1_wide_wide <-
  match_table1_wide %>%
  filter(
    strategy=="post",
    !(variable  %in% c("N", "immunosuppressed", "region"))
  ) %>%
  mutate(
    smd = if_else(variable %in% c(matching_variables$B$exact), NA_character_, smd)
  ) %>%
  pivot_wider(
    id_cols = c(var_label, variable, variable_levels, level),
    names_from = cohort,
    values_from = c(stat_0, stat_1, smd)
  ) %>%
  mutate(
    order_var = cumsum(lag(variable, 1, first(variable))!=variable)
  )

tab1 <-
  match_table1_wide_wide %>%
  gt(
    groupname_col = "var_label"
  ) %>%
  fmt_markdown(columns = "var_label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order_var%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
    var_label = "Variable",
    variable_levels = "",
    `stat_0_cv` = glue("Pfizer BA45 (N={filter(match_table1, strategy=='post', cohort=='cv', variable=='N', by==0) %>% pull(n)})"),
    `stat_1_cv` = glue("Sanofi (N={filter(match_table1, strategy=='post', cohort=='cv', variable=='N', by==1) %>% pull(n)})"),
   `smd_cv` = "SMD",
    `stat_0_age75plus` = glue("Pfizer BA45 (N={filter(match_table1, strategy=='post', cohort=='age75plus', variable=='N', by==0) %>% pull(n)})"),
    `stat_1_age75plus` = glue("Sanofi (N={filter(match_table1, strategy=='post', cohort=='age75plus', variable=='N', by==1) %>% pull(n)})"),
    `smd_age75plus` = "SMD"
  ) %>%
  sub_missing(
      columns = starts_with("smd_"),
      missing_text = "-"
  ) %>%
  tab_spanner(
    label = "CV",
    columns = ends_with("_cv")
  ) %>%
  tab_spanner(
    label = "75+",
    columns = ends_with("_age75plus")
  ) %>%
  cols_align(
    align = c("right"),
    columns =  ends_with(c("_cv", "_age75plus"))
  ) %>%
  cols_hide(c(
    "variable", "level", "order_var",
    #"smd_cv", "smd_age75plus",
    NULL
    )) %>% 
  tab_options(
    table.font.size = 8,
    row_group.as_column = TRUE
  )

gtsave(tab1, fs::path(output_dir_qmd, "table1.html"))

tab1

cat("  \n\n")
```

Standardised mean differences reported as "-" indicate that the variable was exactly matched on and is therefore zero by design.

All values are counts and percentages, except where "\*" indicates means and standard deviations.

#### Table 2: Estimates at X-week follow-up of risk per `r perN_format` people in each vaccine group, and corresponding risk differences, overall and within subgroups

Counts and survival estimates are based on values rounded up to the nearest $6n-3$ for disclosure control.

```{r table-contrasts}
#| echo: false
#| warning: false
#| message: false

tabcontrasts <- 
  contrasts_table %>%
  filter(
    outcome %in% outcomes_primary,
    subgroup == "all"
  ) %>%
  group_by(outcome_descr) %>%
  select(
    cohort_descr,
    outcome_descr,

    #n.atrisk,

    persontime_0,
    persontime_1,

    n.event_0,
    n.event_1,

    riskCI_0,
    riskCI_1,

    rdCI,
    #rrCI,
    #kmirrCI,
    #coxhrCI,
  ) %>%
  mutate(
    # add indicator for row shading
    shade = cumsum((lag(as.character(outcome_descr), 1, "")!=outcome_descr)*1L),
    # remove repeat values of subgroup_level
    #subgroup_descr = if_else(lag(as.character(subgroup_descr),n=1,default="")==subgroup_descr, "", as.character(subgroup_descr)),
  ) %>%
  ungroup() %>%
  gt(groupname_col="cohort_descr") %>%
  cols_label(
    cohort_descr = "Cohort",
    outcome_descr = "Outcome",

    #n.atrisk = "N per group",

    persontime_0 = "Person-weeks", 
    persontime_1 = "Person-weeks",

    n.event_0 = "Events",
    n.event_1 = "Events",

    riskCI_0 = glue("Risk per {perN_format} people (95% CI)"),
    riskCI_1 = glue("Risk per {perN_format} people (95% CI)"),

    rdCI = glue("Risk difference per {perN_format} people (95% CI)"),
    #rrCI = "Risk ratio (95% CI)",
    #kmirrCI = "Incidence rate ratio (95% CI)",
    #coxhrCI = "Hazard ratio (95% CI)"
  ) %>%
  tab_spanner(
    label = "Pfizer BA45",
    columns = ends_with("_0")
  ) %>%
   tab_spanner(
    label = "Sanofi",
    columns = ends_with("_1")
  )  %>%
   tab_spanner(
    label = "Comparison (reference=Pfizer BA45)",
    columns = ends_with("CI")
  ) %>%
  cols_align(
    align = c("right"),
    columns = ends_with(c("atrisk", "_1", "_0", "CI"))
  ) %>%
  tab_options(
    table.font.size = 8
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = shade%%2 != 0
    )
  ) %>%
  cols_hide(c("shade"))

gtsave(tabcontrasts, fs::path(output_dir_qmd, "tabcontrasts.html"))

tabcontrasts

cat("  \n\n")
```

#### Figure 1: Flow of participants into the study

```{r table-flowchart}
#| echo: false
#| warning: false
#| message: false

flowchart %>%
  select(-crit) %>%
  mutate(
    pct_step1= 1-pct_exclude
  ) %>%
  pivot_wider(
    id_cols = c(criteria, cohort_descr),
    names_from = c(boost_type),
    values_from = c(n, n_exclude, pct_all, pct_step, pct_step1)
  ) %>%
  gt(groupname_col="cohort_descr") %>%
  cols_label(
    criteria = "",

    n_pfizerBA45 = "N",
    n_sanofi = "N",

    n_exclude_pfizerBA45 = "N excluded",
    n_exclude_sanofi = "N excluded",

    pct_all_pfizerBA45 = "% (overall)",
    pct_all_sanofi = "% (overall)",
    pct_step_pfizerBA45 = "% (incremental)",
    pct_step_sanofi = "% (incremental)",
    pct_step1_pfizerBA45 = "% (incremental 2)",
    pct_step1_sanofi = "% (incremental 2)",
  ) %>%
  tab_spanner(
    label = "Pfizer BA45",
    columns = ends_with("pfizerBA45")
  ) %>%
   tab_spanner(
    label = "Sanofi",
    columns = ends_with("sanofi")
  ) %>%
  fmt_percent(
    columns = starts_with(c("pct_")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("n_")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "criteria")
  ) %>%
  tab_options(
    table.font.size = 8
  )

cat("  \n\n")
```

```{r, function-ci, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_subgroup <- function(cohort, subgroup){

  subgroup0 <- subgroup

  if(subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_ci <- rst[[cohort]]$estimates %>%
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes_primary,
      outcome != "noncoviddeath"
    ) %>%
    mutate(
      treatment_descr = fct_recode(treatment_descr, "pfizerBA45"="Pfizer BA45", "sanofi"="Sanofi"),
      colgroup = interaction(treatment_descr, subgroup_level_descr, sep=", "),
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10)
    ) %>%
    group_by(colgroup, outcome_descr) %>%
    mutate(
      risk = if_else(outcome=="covidemergency" & time>20*7, NA_real_, risk),
      risk.ll = if_else(outcome=="covidemergency" & time>20*7, NA_real_, risk.ll),
      risk.ul = if_else(outcome=="covidemergency" & time>20*7, NA_real_, risk.ul),
      lagrisk=lag(risk,1,0),
    ) %>%
    ggplot(aes(group=colgroup, colour=treatment_descr, fill=treatment_descr, alpha=treatment_descr)) +
    geom_hline(aes(yintercept=0), colour="darkgrey")+
    #geom_step(aes(x=lagtime, y=risk*perN, linetype=treatment_descr))+
    geom_segment(aes(x=lagtime, xend=time, y=risk*perN, yend=risk*perN))+
    geom_segment(aes(x=lagtime, xend=lagtime, y=lagrisk*perN, yend=risk*perN))+
    geom_rect(aes(xmin=lagtime, xmax=time, ymin=risk.ll*perN, ymax=risk.ul*perN), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set2", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set2", guide="none", na.value="grey") +
    #scale_color_manual(values=colourvalues, na.value="grey") +
    #scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,28), expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_continuous(
      expand = expansion(mult=c(0,0.01)), 
      #label=label_number_risk
    )+
    scale_alpha_discrete(range=c(0.3,0.9))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Weeks",
      y=NULL,
      #title=glue("Cumulative incidence per {perN_format}"),
      colour=NULL,
      alpha=NULL,
      linetype=NULL
    )+
    guides(linetype="none", fill="none", alpha="none")+
    theme_minimal(base_size = 9)+
      theme(
        # legend.position = c(0.05,0.1),
        # legend.justification = c(0,0),
        legend.position = "bottom",
        #legend.direction = "vertical",
        axis.text.x.top=element_text(hjust=0),

        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),
        strip.placement="outside",
        strip.text.y.left = element_text(angle=0),

        panel.border = element_blank(),
        panel.spacing = unit(0.8, "lines"),

      )+
      NULL

  plot_contrasts_data <- rst[[cohort]]$contrasts_daily %>%
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes_primary
    ) %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    mutate(
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
      # remove values before 7 days to avoid infinite ratios before both groups have events
      #rr = if_else(fup_start<7, NA_real_, rr),
      #rr.ll = if_else(fup_start<7, NA_real_, rr.ll),
      #rr.ul = if_else(fup_start<7, NA_real_, rr.ul),
      lagrd=lag(rd,1,0),
      lagrr=lag(rr,1,1),

    )

  plot_rr <- 
    plot_contrasts_data %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    ggplot(aes(group=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="darkgrey")+
    geom_segment(aes(x=fup_start, xend=fup_end, y=rr, yend=rr))+
    geom_segment(aes(x=fup_start, xend=fup_start, y=lagrr, yend=rr))+
    geom_rect(aes(xmin=fup_start, xmax=fup_end, ymin=rr.ll, ymax=rr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    #scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    #scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,28), expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_log10(
      expand = expansion(mult=c(0,0.01)),
      label=label_number_rr,
      sec.axis = dup_axis(breaks=NULL, labels=NULL, name = "higher risk\n<- BNT162b2 | mRNA 1273 ->")
    )+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Weeks",
      y=NULL,
      title="Cumulative risk ratio",
      colour=NULL
    )+
    #colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      #axis.line.y = element_line(colour = "darkgrey"),
      #axis.line.x.top = element_line(colour = "darkgrey"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines")
    )+
    NULL

  plot_irr <- rst[[cohort]]$contrasts_cuts %>%
    # equivalent to the hazard
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes_primary
    ) %>%
    group_by(outcome_descr, subgroup_level_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="darkgrey")+
    geom_segment(aes(x=fup_start, xend=fup_end, y=irr, yend=irr))+
    geom_segment(aes(x=fup_start, xend=fup_start, y=lag(irr), yend=irr))+
    geom_rect(aes(xmin=fup_start, xmax=fup_end, ymin=irr.ll, ymax=irr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y")+
    # scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    # scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    #scale_color_manual(values=colourvalues, na.value="grey") +
    #scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,28), expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_log10(
      expand = expansion(mult=c(0,0.01)),
      label=label_number_rr
    )+
    coord_cartesian(xlim=c(0, NA))+
    colour_guide+
    labs(
      x="Weeks",
      y="higher risk\n<- Pfizer BA45 | Sanofi ->",
      title="Incidence rate ratio",
      colour=NULL
    )+
    theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      #axis.line.y = element_line(colour = "darkgrey"),
      #axis.line.x.top = element_line(colour = "darkgrey"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines")
    )+
    NULL

  n_subgroups <-
    rst[[cohort]]$estimates %>%
    filter(
      subgroup==subgroup0
    ) %>% 
    `[[`("subgroup_level_descr") %>%
    n_distinct()

  plot_combined <- patchwork::wrap_plots(
    plot_ci, 
    #plot_rd, 
    #plot_rr, 
    #plot_kmirr,
    #plot_coxhr,
    ncol=2,
    widths=c(n_subgroups,1)
  )

  ggsave(
    filename=glue("ci_{cohort}_{subgroup}.png"),
    path=output_dir_qmd,
    plot=plot_ci,
    unit="cm",
    width=13,
    height=13
  )

  plot_ci
}

```

#### Figure 2: Estimates of cumulative risk per `r perN_format` people (left plots) and cumulative risk ratios (right plots).

```{r, figure-ci-all}
#| echo: false
#| message: false
#| warning: false
#| out-width: '100%'
#| out-height: '50%'
#| output: 'asis'
#| out-extra: ''

plot_subgroup("cv", "all")
```

\newpage

#### Figure 3: Estimated incidence rate ratios at X-week follow-up comparing the effectiveness of Sanofi and Pfizer BA45, overall and within subgroups, together with p-values for between-subgroup heterogeneity.

<!-- ```{r figure-contrasts-irr} -->

<!-- #| echo: false -->

<!-- #| message: false -->

<!-- #| warning: false -->

<!-- #| fig-height: 10 -->

<!-- #| out-extra: '' -->

<!-- contrasts_plot <-  -->

<!--   contrasts_overall %>% -->

<!--   filter( -->

<!--     outcome %in% outcomes, -->

<!--     subgroup %in% c("all", "ageband", "cv"), -->

<!--   ) %>% -->

<!--   group_by(cohort_descr, outcome_descr) %>% -->

<!--   mutate( -->

<!--    # outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10), -->

<!--     subgroup_level_descr = fct_rev(subgroup_level_descr), -->

<!--     subgroup_level_descr = fct_recode(subgroup_level_descr, "All"=""), -->

<!--     irrCI =   glue("{label_number_rr(irr)} ({label_number_rr(irr.ll)} - {label_number_rr(irr.ul)})"), -->

<!--     order = cumsum(lag(subgroup_descr, 1, first(subgroup_descr))!=subgroup_descr), -->

<!--     patch = (order %% 2) == 1, -->

<!--     rownum = n() - row_number() + 1 -->

<!--   ) %>% -->

<!--   group_by(cohort_descr, outcome_descr, subgroup_descr) %>% -->

<!--   mutate( -->

<!--     midgroup = mean(rownum), -->

<!--     irr.ln.p = if_else(row_number()==1 & (subgroup!="all"), irr.ln.p, NA_real_), -->

<!--     irr.ln.p = if_else(irr.ln.p<0.001, "<0.001", label_number(0.001, trim=FALSE)(irr.ln.p)) -->

<!--   ) -->

<!-- plot_irr_subgroup <- contrasts_plot %>% -->

<!--   ggplot(aes(y=subgroup_level_descr)) + -->

<!--   geom_vline(aes(xintercept=1), linetype="dotted", colour="darkgrey")+ -->

<!--   geom_point(aes(x=irr), position=position_dodge(width=-0.3))+ -->

<!--   geom_linerange(aes(xmin=irr.ll, xmax=irr.ul), position=position_dodge(width=-0.3))+ -->

<!--   geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=min(irr.ll), xmax=max(irr.ul), fill=patch), alpha=0.2)+ -->

<!--   #facet_grid(rows=vars(outcome_descr), scales="free", space="free_y", switch="y")+ -->

<!--   facet_wrap(vars(outcome_descr), scales="free_y", ncol=1, strip.position = "top")+ -->

<!--   scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","), oob=~oob_squish(.,range=log(c(0.4,2), 10)), expand=expansion())+ -->

<!--   #scale_fill_brewer(type="qual", palette="Set2")+ -->

<!--   scale_fill_manual(values=c("transparent", "darkgrey"))+ -->

<!--   labs( -->

<!--     subtitle="\n\n\nX-week Hazard Ratio", -->

<!--     x="<- favours Sanofi  |  favours Pfizer BA45 ->", -->

<!--     y=NULL -->

<!--   )+ -->

<!--   theme_minimal()+ -->

<!--   theme( -->

<!--     plot.margin=grid::unit(c(0,0,0,0), "mm"), -->

<!--     plot.subtitle=element_text(hjust=0.5, vjust=0), -->

<!--     legend.position="none", -->

<!--     #panel.grid.minor.x = element_blank(), -->

<!--     panel.grid.major.y = element_blank(), -->

<!--     panel.grid.minor.y = element_blank(), -->

<!--     #strip.background = element_blank(), -->

<!--     #strip.placement="outside", -->

<!--     strip.text = element_text(hjust=0), -->

<!--     #strip.text.y.left = element_text(angle=0), -->

<!--     #strip.text.y.left = element_blank(), -->

<!--     panel.border = element_blank(), -->

<!--     panel.spacing = unit(0.5, "lines"), -->

<!--   ) -->

<!-- plot_irr_subgroup_table <- contrasts_plot %>% -->

<!--   ggplot(aes(y=subgroup_level_descr)) + -->

<!--   geom_text(aes(x=0.5, label=irrCI), hjust=0.5, vjust=0.5, size=9/.pt)+ -->

<!--   geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=0.2, xmax=0.8, fill=patch), alpha=0.2)+ -->

<!--   facet_wrap(vars(outcome_descr), scales="free_y", ncol=1, strip.position = "top")+ -->

<!--   scale_fill_manual(values=c("transparent", "darkgrey"))+ -->

<!--   scale_x_continuous(expand=expansion())+ -->

<!--   labs(subtitle="\nHazard\nRatio\n(95% CI)")+ -->

<!--   theme_minimal()+ -->

<!--   theme( -->

<!--     plot.margin=grid::unit(c(0,0,0,0), "mm"), -->

<!--     plot.subtitle=element_text(hjust=0.5, vjust=0), -->

<!--     legend.position="none", -->

<!--     panel.grid = element_blank(), -->

<!--     strip.text =element_text(colour="transparent"), -->

<!--     axis.text.y = element_blank(), -->

<!--     axis.title.y = element_blank(), -->

<!--     axis.text.x = element_blank(), -->

<!--     axis.title.x=element_blank(), -->

<!--     panel.border = element_blank(), -->

<!--     panel.spacing = unit(0.5, "lines"), -->

<!--   )+ -->

<!--   NULL -->

<!-- contrasts_plot %>% select(outcome_descr, subgroup_level_descr, irrCI, irr.ln.p) %>% write_csv(here("table3.csv")) -->

<!-- plot_p_subgroup_table <- contrasts_plot %>% -->

<!--   ggplot() + -->

<!--   geom_text(aes(x=0.5, y=subgroup_level_descr, label=irr.ln.p), hjust=0.5, vjust=0.5, size=9/.pt, colour="transparent")+ -->

<!--   geom_text(aes(x=0.5, y=midgroup, label=irr.ln.p), hjust=0.5, vjust=0.55, size=9/.pt)+ -->

<!--   geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=0, xmax=1, fill=patch), alpha=0.2)+ -->

<!--   facet_wrap(vars(outcome_descr), scales="free_y", ncol=1, strip.position = "top")+ -->

<!--   scale_fill_manual(values=c("transparent", "darkgrey"))+ -->

<!--   scale_x_continuous(expand=expansion())+ -->

<!--   labs(subtitle="P for\nbetween\nsubgroup\ndifferences")+ -->

<!--   theme_minimal()+ -->

<!--   theme( -->

<!--     plot.margin=grid::unit(c(0,0,0,0), "mm"), -->

<!--     plot.subtitle=element_text(hjust=0.5, vjust=0), -->

<!--     legend.position="none", -->

<!--     panel.grid = element_blank(), -->

<!--     strip.text =element_text(colour="transparent"), -->

<!--     axis.text.y = element_blank(), -->

<!--     axis.title.y = element_blank(), -->

<!--     axis.text.x = element_blank(), -->

<!--     axis.title.x=element_blank(), -->

<!--     panel.border = element_blank(), -->

<!--     panel.spacing = unit(0.5, "lines"), -->

<!--   )+ -->

<!--   NULL -->

<!--   plot_combined <- patchwork::wrap_plots( -->

<!--     plot_irr_subgroup,  -->

<!--     plot_irr_subgroup_table,  -->

<!--     plot_p_subgroup_table, -->

<!--     ncol=3, -->

<!--     widths=c(2,0.7,0.5) -->

<!--   ) -->

<!--   ggsave( -->

<!--     filename=glue("irr_pvals.jpeg"), -->

<!--     path=output_dir_qmd, -->

<!--     plot=plot_combined, -->

<!--     width=20, height=20, units="cm" -->

<!--   ) -->

<!-- plot_combined -->

<!-- cat("  \n\n") -->

<!-- ``` -->

\newpage

## References

### 2024-01-16 write-up/manuscript/import-process.R

@@ -1,0 +1,359 @@
+
+library("fs")
+
+
+
+## ollection of formatting functions that may be used ----
+
+perN <- 1000
+perN_format <- label_number(1, 1, big.mark=",")(perN)
+
+label_number_n <- label_number(1, 1, big.mark=",")
+
+label_number_risk <- function(x, scale=perN){
+  x_scale <- x*scale
+  x_abs <- abs(x_scale)
+  case_when(
+    x_abs>=100 ~ label_number(0.1, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<100 & x_abs>=10 ~ label_number(0.1, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<10 & x_abs>=1 ~ label_number(0.01, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<1 & x_abs>=0.1 ~ label_number(0.01, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<0.1 ~ label_number(0.001, scale, style_negative="minus", big.mark=",")(x),
+    NA ~ NA_character_
+  )
+}
+
+label_number_rr <- function(x, scale=1){
+  x_scale <- x*scale
+  x_abs <- abs(x_scale)
+  case_when(
+    x_abs>=100 ~ label_number(0.01, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<100 & x_abs>=10 ~ label_number(0.01, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<10 & x_abs>=1 ~ label_number(0.01, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<1 & x_abs>=0.1 ~ label_number(0.01, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<0.1 & x_abs>=0.01 ~ label_number(0.01, scale, style_negative="minus", big.mark=",")(x),
+    x_abs<0.01 ~ label_number(0.0001, scale, style_negative="minus", big.mark=",")(x),
+    NA ~ NA_character_
+  )
+}
+
+
+
+# store results for each cohort in a list,
+# accessed for example as follows:  rst$cv$contrasts
+
+# initialise list
+rst <- lst(cv=NULL, age75plus=NULL)
+
+for(cohort in c("age75plus", "cv")){
+
+  # new temporary environment
+  env <- new.env()
+
+  # import data within new environment
+  evalq({
+
+
+    cohort_dir <- path(output_dir_os, cohort)
[…]

### 2024-01-16 write-up/manuscript/setup.R

@@ -1,0 +1,66 @@
+
+library('tidyverse')
+library('here')
+library('glue')
+library('lubridate')
+library('gt')
+library('patchwork')
+library('scales')
+
+# where are the outputs (ie the inputs for this manuscript!) saved?
+output_dir_os <- here("output", "release-objects")
+#output_dir_os <- here("released-output", "release-objects-v2")
+
+
+## import functions and design elements
+source(here("analysis", "functions", "utility.R"))
+source(here("analysis", "design", "design.R"))
+
+study_dates_format <- map(study_dates, ~format(as.Date(.), "%e %B %Y"))
+
+# where should we put the objects created within this rmd script?
+output_dir_qmd <- here("write-up", "manuscript", "figures")
+fs::dir_create(output_dir_qmd)
+
+# only applicable if `self-contained: yes` option is enables in yaml header
+# knitr::opts_chunk$set(
+#   echo = TRUE,
+#   fig.path = paste0(output_dir_rmd, "/"),
+#   fig.pos = 'H' #to stop figures floating around when rendering to pdf
+# )
+
+
+## define outcome sets
+
+outcomes_primary <- c(
+  "covidemergency",
+  "covidadmitted",
+  #"covidadmittedproxy1",
+  #"covidcritcare",
+  "coviddeath",
+  "noncoviddeath",
+  NULL
+) %>%
+  set_names(., .)
+
+# outcomes_negative <- c(
+#   #"noncoviddeath",
+#   #"fracture",
+#   NULL
+# ) %>%
+#   set_names(., .)
+
+outcomes <- outcomes_primary
+
+subgroups <-
+  c(
+    "all",
+    "vax_prior_history",
+    "ageband",
+    "prior_covid_infection",
+    "cv",
+    NULL
+  ) %>%
+  set_names(., .)
+
+

### 2024-01-16 analysis/design/design.R
extend follow up; add cohort description

@@ -105,6 +105,10 @@ threshold <- 6
 
 recoder <-
   lst(
+    cohort = c(
+      `cv` = "Clinically vulnerable",
+      `age75plus` = "Aged 75 years or over"
+    ),
     subgroups = c(
       `Main` = "all",
       `Age` = "ageband",

### 2024-01-16 analysis/design/design.R
update from main

@@ -107,7 +107,8 @@ recoder <-
   lst(
     cohort = c(
       `cv` = "Clinically vulnerable",
-      `age75plus` = "Aged 75 years or over"
+      `Clinically Vulnerable`  = "cv",
+      `Aged 75 years or over` = "age75plus"
     ),
     subgroups = c(
       `Main` = "all",

### 2024-01-16 analysis/design/study-dates.R

@@ -1,7 +1,9 @@
 # # # # # # # # # # # # # # # # # # # # #
 # Purpose: Define the study dates that are used throughout the rest of the project
 # Notes:
-# This script is separate from the design.R script as the dates are used by the study definition as well as analysis R scripts.
+# This script is separate from the design.R script as
+#  the dates are also used by the dataset definition (as well as analysis R scripts),
+#  where they cannot be "sourced" as usual
 # # # # # # # # # # # # # # # # # # # # #
 
 ## create output directories ----

### 2024-01-16 analysis/design/design.R
Update design.R

@@ -106,7 +106,6 @@ threshold <- 6
 recoder <-
   lst(
     cohort = c(
-      `cv` = "Clinically vulnerable",
       `Clinically Vulnerable`  = "cv",
       `Aged 75 years or over` = "age75plus"
     ),

### 2024-01-16 analysis/design/design.R
Update design.R

@@ -106,8 +106,8 @@ threshold <- 6
 recoder <-
   lst(
     cohort = c(
-      `cv` = "Clinically vulnerable",
-      `age75plus` = "Aged 75 years or over"
+      `Clinically vulnerable` = "cv",
+      `Aged 75 years or over` = "age75plus"
     ),
     subgroups = c(
       `Main` = "all",

### 2024-01-31 write-up/manuscript/manuscript.qmd
update manuscript

@@
 ---
-title: "Comparative effectiveness of Pfizer B.A4-5 versus Sanofi during the Spring 2023 COVID-19 booster programme in England: a cohort study in OpenSAFELY-TPP"
+title: "Comparative safety and effectiveness of Pfizer BA.4-5 versus Sanofi during the spring 2023 COVID-19 booster vaccination programme in England: a matched cohort study in OpenSAFELY-TPP"
 format:
@@
   #   embed-resources: true
-  docx
+  rtf
+output-file: manuscript_draft.rtf
+keep-md: false
+execute:
+  echo: false
+  output: true
+  #warning: true
 knitr:
   opts_chunk: 
-    collapse: true
-    echo: false
-    fig.path: ./write-up/figures/
+    #fig.path: ./
     fig.pos: H
@@
 link-citations: true
-zotero: yes
-header-includes:
-- \usepackage{float} 
-- \floatplacement{figure}{H}
-- \usepackage{caption}
-- \captionsetup[figure]{labelformat=empty}
-- \captionsetup[table]{labelformat=empty}
+zotero: true
 ---
@@
 #| warnings: false
+#| cache: false
 source(here::here("write-up", "manuscript", "setup.R"))
@@
 #| message: false
+#| cache: false
 source(here("write-up", "manuscript", "import-process.R"))
@@
 
-William J Hulme^1^, Edward PK Parker^2^, Em Prestige^2^, Ruth H Keogh^2^, Elizabeth J Williamson^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Colm Andrews^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^6,7^\*, Jonathan AC Sterne^3,5,8^\*, and the OpenSAFELY collaborative.
+TBD: Colm Andrews^1^, Edward PK Parker^2^, Em Prestige^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^5,6^\*, Jonathan AC Sterne^3,7,8^\*, William J Hulme^1^, and the OpenSAFELY collaborative.
 
@@
 
-5\. NIHR Bristol Biomedical Research Centre, Bristol, UK
-
-6\. CAUSALab, Harvard T.H. Chan School of Public Health, Boston, MA 02115
-
-7\. Departments of Epidemiology and Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA 02115
+5\. CAUSALab, Harvard T.H. Chan School of Public Health, Boston, MA 02115
+
+6\. Departments of Epidemiology and Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA 02115
+
+7\. NIHR Bristol Biomedical Research Centre, Bristol, UK
 
@@
 
-*Introduction* The spring 2023 COVID-19 booster vaccination programme in England used both Pfizer Ba45 and Sanofi vaccines.
-All people aged 75 years and over, or considered to be clinically vulnerable were eligible to receive a booster dose.
-Direct comparisons of the effectiveness against severe COVID-19 of these two vaccines for boosting have not been made in trials or observational data.
-
-*Methods* On behalf of NHS England, we used the OpenSAFELY-TPP database to compare recipients of each vaccine, investigating the over 75s and the clinically vulnerable groups separately.
-Recipients were matched on date of vaccination, vaccine history, age, and other characteristics.
-Recipients were included if boosted between 1 April and 30 June 2023, and followed up for X weeks.
-Outcomes were COVId-19 emergency department attendance, COVID-19 hospital admission, and COVID-19 death.
-We estimated the cumulative incidence of each outcome, and quantified comparative effectiveness using risk differences (RD) and incidence rate ratios (IRRs).
+*Introduction* The spring 2023 COVID-19 booster vaccination programme in England used both Pfizer BA.4-5 and Sanofi vaccines.
+All people aged 75 years or over and the clinically vulnerable were eligible to receive a booster dose.
+Direct comparisons of effectiveness against severe COVID-19 events of these two vaccines for boosting have not been made in trials or observational data.
+
+*Methods* With the approval of NHS England, we used the OpenSAFELY-TPP database to compare effectiveness of the Pfizer BA.4-5 and Sanofi vaccines during the spring 2023 booster programme, between 1 April and 30 June 2023.
+We investigated two cohorts separately: those aged 75 or over (75+); those aged 50 or over and clinically vulnerable (CV).
+In each cohort, vaccine recipients were matched on date of vaccination, COVID-19 vaccine history, age, and other characteristics.
+Effectiveness outcomes were COVID-19 Emergency Department (ED) attendance, COVID-19 hospital admission, and COVID-19 death up to 16 weeks after vaccination.
+Safety outcomes were Pericarditis and Myocarditis up to 4 weeks after vaccination.
+We report the cumulative incidence of each outcome, and compare safety and effectiveness using risk differences (RD), relative risks (RR), and incidence rate ratios (IRRs).
 
 *Results* `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4") %>% pull(n) %>% sum())` people were 1-1 matched in the CV cohort, and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4") %>% pull(n) %>% sum())` in the 75+ cohort, contributing a total `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up respectively.
-
-In both CV and 75+ cohorts, effectiveness against all measured COVID-19-related outcomes was similar. This finding was observed overall and in pre-specified subgroups.
-
-In the CV cohort, the X-week risks per `r perN_format` of COVID-19 A&E attendance were `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer B.A4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_1)` for Sanofi: the IRR comparing Sanofi with PfizerB.A4-5 was `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(irrCI)`.
-For COVID-19 hospitalisation, X-week risks were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)`: IRR `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`.
-For COVID-19 death: `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)`: IRR `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)`.
-
-In the 75+ cohort, the X-week risks per `r perN_format` for COVID-19 A&E attendance were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer B.A4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_1)` for Sanofi: the IRR comparing Sanofi with PfizerB.A4-5 was `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(irrCI)`.
-For COVID-19 hospitalisation, X-week risks were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)`: IRR `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`.
-For COVID-19 death: `r perN_format` were `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` and `r contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)`: IRR `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)`.
-
-*Conclusion* This observational study comparing effectiveness of PFizer and Sanofi vaccine during the Spring 2023 programme in England in the two main eligible cohorts -- people aged 75 and over and in clinically vulnerable people -- found no evidence of substantial differences between the two vaccine products within the first X weeks after vaccination.
+The incidence of COVID-19 hospital admission was higher for Sanofi than for Pfizer BA.4-5.
+In the CV cohort, 16-week risks per `r perN_format` people were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)`.
+In the 75+ cohort, these were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`.
+These finding were similar across all pre-specified subgroups.
+More severe COVID-19 related outcomes (critical care admission and death), and safety outcomes at 4 weeks, were rare in both vaccines so we could not reliably estimate superiority of either vaccine.
+
+*Conclusion* This observational study comparing effectiveness of Pfizer BA.4-5 and Sanofi vaccine during the spring 2023 programme in England in the two main eligible cohorts -- people aged 75 and over and in clinically vulnerable people -- found some evidence of superior effectiveness against COVID-19 hospital admission for Pfizer BA.4-5 compared with Sanofi within 16 weeks after vaccination.
+Differences were largely observed during weeks 2 to 4 after vaccination, suggesting the Pfizer BA.4-5 offers earlier protection via more rapid immunogenicity, as shown in RCTs.
 
@@
 
-```{r break1}
-2+2
-```
+````{comment}
+
+````
 
@@
 
-In X 2022, NHS XXX in England announced the Spring 2023 COVID-19 booster vaccination programme, available to those aged 75 years and over, and those considered to be vulnerable to infection or severe COVID-19 disease, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI) @covid19,.
-The use of the both the Pfizer-BioNTech B.A.4/5 vaccine and the Sanofi vaccine during the spring 2023 programme enables a direct comparison of their effectiveness against severe COVID-19 disease, which has not yet been conducted in randomised trials.
-
-On behalf of NHS England, we used the OpenSAFELY-TPP database, covering around 45% of English primary care practices and linked to hospital and death records, to compare the effectiveness of boosting with Pfizer and Sanofi in the 75+ years group and the clinically vulnerable group who received either vaccine in the spring 2023 programme.
-During this period and the weeks following, the X variant was dominant.
+The spring 2023 COVID-19 booster vaccination programme in England offered COVID-19 vaccination to those aged 75 years or over, resident in a care home, and those who are immunosuppressed, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI).[@NHSEnglandNHSc; @JCVIStatementCOVID19; @JCVIStatementSpring] Both the Pfizer BA.4/BA.5 vaccine and the Sanofi vaccine were used during this programme.
+This allows direct comparisons of their safety and effectiveness against severe COVID-19 disease, something not investigated in randomised trials.
+
+We used the OpenSAFELY-TPP database, covering around 45% of English primary care practices and linked to hospital and death records, to compare safety and effectiveness of the Pfizer BA.4-5 and Sanofi vaccines received during the spring 2023 booster programme.
 
@@
 
-All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>.
-With the approval of NHS England, primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
+All data were linked, stored and analysed securely using the OpenSAFELY platform: <https://opensafely.org/>, as part of the NHS England OpenSAFELY COVID-19 service.
+Primary care records managed by the GP software provider, TPP were linked, using NHS numbers, to emergency department attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
 COVID-19 vaccination history and health and social care worker status is available in the GP record directly via the National Immunisation Management System (NIMS).
@@
 
-We considered two cohorts: all adults aged 75 years or over ("75+"), and all adults aged 18 or over who were classified as clinically vulnerable ("CV").
+We considered two cohorts: all adults aged 75 years or over ("75+"), and all adults aged 50 or over who were considered clinically vulnerable ("CV").
 These groups are not mutually exclusive.
-Clinical vulnerability encompasses various chronic or immunocompromising conditions, immunosuppressive treatments, and ..., and is defined fully in supplementary materials.
-
-In each group, any person receiving a Pfizer or Sanofi vaccine dose between `r study_dates_format$studystart_date` and `r study_dates_format$studyend_date` inclusive were considered for inclusion.
-People were eligible if they: were registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting; were not a health or social care worker, not resident in a care or nursing home and not medically housebound or receiving end-of-life care; had no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; were not hospitalised at the time of boosting; and had complete information on sex, deprivation, and Sustainability and Transformation Partnership (STP, a geographical grouping of NHS and local authorities).
+We only considered those aged 50 years or over because the Sanofi vaccine was rarely offered to younger people.
+Clinical vulnerability encompasses various chronic or immunosuppressive conditions and medications as set out in the Green Book chapter 14a @COVID19GreenBook, and defined in supplementary materials.
+
+In each group, any person receiving a Pfizer BA.4-5 or Sanofi vaccine dose between `r study_dates_format$studystart_date` and `r study_dates_format$studyend_date` inclusive was considered.
+People were eligible if they: were registered at a GP practice using TPP's SystmOne clinical information system at the time of vaccination; had received at least two prior COVID-19 vaccine doses; were not a health or social care worker and not receiving end-of-life care; had no evidence of SARS-CoV-2 infection or COVID-19 disease during the 28 days prior to vaccination; were not a hospital in-patient at the time of vaccination; and had complete information on sex, deprivation, and Sustainability and Transformation Partnership (STP, a geographical grouping of NHS and Local Authorities).
 
@@
 
-Pfizer and Sanofi booster recipients were matched 1-1 without replacement, on the following characteristics: date of vaccination; number of previous COVID-19 vaccine doses; time since previous COVID-19 vaccine dose; sex (male or female); age (3 year caliper and within age groups defined by JVCI risk groups); Index of Multiple Deprivation (IMD, grouped by quintile); STP as a surrogate for geographical region; evidence of prior SARS-CoV-2 infection (any of positive SARS-CoV-2 test, probable infection documented in primary care, or COVID-19 hospital attendance or admission); and morbidity count (grouped as 0, 1, or 2 or more conditions from the following list: diabetes, BMI over 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease, cancer within 3 years).
-Clinical vulnerability was an additional matchng factor in the 75+ cohort.
-
-There are no missing values for these variables: people with missing values for sex, IMD or STP were excluded, and all other variables were defined by the presence or absence of clinical codes or events.The supplementary materials provide more information on how these characteristics were chosen and defined.
+Pfizer BA.4-5 and Sanofi vaccine recipients were matched 1-1 without replacement, on the following characteristics: date of vaccination (3-day caliper); number of previous COVID-19 vaccine doses (grouped as 2 to 4, 5, or 6 or more); time since previous COVID-19 vaccine dose (28-day caliper); sex (male or female); age (3-year caliper and within age groups 50-64, 65-75, 75-79, 80-84, 85+); Index of Multiple Deprivation (IMD, within a 5000 IMD-rank caliper); STP as a surrogate for geographical region; any evidence of prior SARS-CoV-2 infection (including any of: positive SARS-CoV-2 test, infection as documented in primary care, COVID-19 hospital attendance or admission); and morbidity count (grouped as 0, 1, or 2 or more conditions from the following list: diabetes, BMI over 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease, cancer within 3 years).
+Clinical vulnerability was an additional matching factor in the 75+ cohort.
+
+Age was calculated as at 30 June 2023, to allow for a degree of operational flexibility in who was offered vaccination during the programme.
+All other characteristics were calculated as on the day before vaccination.
+In the analysis cohorts there are no missing values for any characteristics as people with missing values for sex, IMD or STP were excluded, and remaining characteristics were defined by the presence or absence of clinical codes or events in the health record.
+The supplementary materials provide more information on how these characteristics were defined.
 
@@
 
-We considered three primary outcomes. COVID-19 emergency department admission was identified using HES emergency department records, with SNOMIED.
-
-COVID-19 hospitalisation was identified using Secondary Use Service (SUS) in-patient hospital records with International Statistical Classification of Diseases and Related Health Problems 10th Revision (ICD-10) codes U07.1, U07.2, or U.109 as the primary or non-primary reason for admission.
-COVID-19 death was defined as death with U07.1, U07.2, or U.109 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
+We considered three effectiveness outcomes.
+COVID-19 hospital admission was identified using Secondary Use Service (SUS) in-patient hospital spell records with International Statistical Classification of Diseases and Related Health Problems 10th Revision (ICD-10) codes U07.1, U07.2, or U.109 as the primary or non-primary reason for admission.
+COVID-19 critical care admissions were any COVID-19 hospital spell with at least one day spent in ciritcal care (including any such spells occurring subsequent to the first spell after vaccination).
+COVID-19 death was defined as death COVID-19 ICD-10 codes (as before) mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
 We also report on non-COVID-19 deaths.
 
-<!-- We also considered the follows safety outcomes: -->
-
-<!-- * Adverse allergic reaction / Anaphylaxis -->
-
-<!-- * Myocarditis or Pericarditis -->
-
-<!-- *  -->
-
-<!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->
-
-Each person was followed from receipt of vaccination (time zero) until the outcome of interest, with censoring at death, practice de-registration, XX weeks, or `r study_dates_format$followupend_date`.
+We considered two safety outcomes.
+Pericarditis within 28 days of vaccination was identified using SUS emergency department records with pericarditis listed as the discharge diagnosis (SNOMED), and using ICD-10 coded hospital admission and death records.
+Myocarditis was identified similarly.
+Details of the codelists used to define these outcomes are available in supplementary materials.
+
+<!--# COVID-19 related attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->
+
+Each person was followed from receipt of vaccination (time zero) until the outcome of interest, with censoring at death, practice de-registration, `r study_dates_format$followupend_date`.
+Outcomes were observed for up to 16 weeks for effectiveness outcomes, and up to 4 weeks for safety outcomes.
 
@@
 We estimated cumulative risk differences and risk ratios (RRs) comparing the vaccine groups for each outcome.
-These were estimated up to 20 weeks where available, but estimates at X weeks are reported as the primary comparison to ensure potential follow-up in those vaccinated towards the end of the booster programme was not censored.
-We also estimated Incidence Rate Ratios (IRRs) over the first X weeks, and within period-specific intervals defined by splits on weeks 1, 2, 4, 8, 12, 16, and X, where available.
-
-<!-- days 7, 14, 21, 28, 42, 56, 70, 84, 112, 140, 168 and 196. -->
+We also estimated Incidence Rate Ratios (IRRs) over the full follow-up duration, and within period-specific intervals defined by splits on weeks 1, 2, 4, 8, 12, and 16.
 
@@
 Confidence limits for the risk differences were derived from the sum of squares of the standard errors on the KM scale, using Greenwood's formula.
-Confidence limits for the risk ratio were derived from the sum of squares of the standard errors on the log-KM scale.
-Confidence limits for the IRRs were derived from the sum of squares of the standard errors on the log-KM scale.
+Confidence limits for the risk ratios and incidence rate ratios were derived from the sum of squares of the standard errors on the log-KM scale.
 <!-- 95% confidence limits for the Kaplan-Meier estimates, risk differences, risk ratios, and incidence rate ratios are derived by bootstrapping the matched pairs.
@@
 
-We estimated comparative effectiveness separately in the following subgroups: age band in the CV group only (18-39, 40-54, 55-64, 65-74, 75+); clinical vulnerability in the 75+ group only; prior vaccination history; evidence of prior SARS-CoV-2 infection or not.
-We used $\chi^2$ tests to examine evidence for heterogeneity in comparative effectiveness between subgroups.
-
-We reported the rates of SARS-CoV-2 tests undertaken from baseline up to 31 March 2022 in each vaccine group, as lower testing rates in one group might lead to under-ascertainment of outcomes, particularly positive SARS-CoV-2 tests.
-
-We reported outcomes at one week as a negative control outcome @Lipsitch2010, during which time the vaccine is not expected to have any protective immunological effect.
+We estimated comparative effectiveness separately in the following subgroups: age band in the CV group only (50-64, 65-75, 75-79, 80-84, 85+); clinical vulnerability in the 75+ group only; prior COVID-19 vaccine count.
+We used $\chi^2$ tests to examine evidence for heterogeneity in estimates between subgroups.
+
+<!-- We reported outcomes at one week as a negative control outcome @Lipsitch2010, during which time the vaccine is not expected to have any protective immunological effect. -->
 
@@
 
-```{r break2}
-2+2
+````{comment}
+
+````
+
+### Study population and matching
+
+We identified `r label_number(1, big.mark=",")(filter(flowchart_totals_allcohorts, boost_type=="any") %>% pull(n))` people aged 50 years or over who were recorded as having received a COVID-19 vaccine between `r study_dates_format$studystart_date` and `r study_dates_format$studyend_date` whilst registered at a TPP practice.
+This included `r label_number(1, big.mark=",")(filter(flowchart_totals_allcohorts, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(flowchart_totals_allcohorts, boost_type=="pfizerBA45") %>% pull(pct))`) Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(flowchart_totals_allcohorts, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(flowchart_totals_allcohorts, boost_type=="sanofi") %>% pull(pct))`) Sanofi recipients.
+
+We identified `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).
+
+`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).
+
+Sanofi was more likely to be administered in April 2023, and Pfizer BA.4-5 more likely in May and June 2023.
+Prior to matching, Sanofi recipients were on average older, more likely to be recorded as resident in a care or nursing home, and less likely to have received fewer than 5 prior COVID-19 vaccine doses compared with Pfizer BA.4-5 recipients.
+Sanofi recipients were less likely that Pfizer BA.4-5 recipients to have previously received the Pfizer/BA.1 vaccine during an earlier booster programme and more likely to have received Moderna/Omicron.
+Sex, deprivation levels, and ethnicity were similar in both groups prior to matching (supplementary Tables S1a and S1b).
+
+In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1a).
+In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1b).
+
+After matching, characteristics at the start of follow-up were well-balanced between groups (Table 1) with standardised mean differences consistently under 0.05 (supplementary Figures S2a and S2b), including for those characteristics not used for matching.
+
+### Estimated comparative effectiveness
+
+#### CV cohort
+
+```{r break3}
+
 ```
 
-### Study population and matching
-
-\[\[fix total identified in data_process script then re-write\]\]
-
-<!-- We identified `r label_number(1, big.mark=",")(filter(rst$cv$total_flowchart, boost_type=="Any") %>% pull(n) %>% sum)` clinically vulnerable people and `r label_number(1, big.mark=",")(filter(rst$age75plus$total_flowchart, boost_type=="Any") %>% pull(n) %>% sum)` people aged 75 or over who were recorded as having received a COVID-19 vaccine between `r study_dates_format$studystart_date` and `r study_dates_format$studyend_date` whilst registered at a TPP practice.
-This included `r label_number(1, big.mark=",")(filter(rst$age75plus$total_flowchart, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$total_flowchart, boost_type=="pfizerBA45") %>% pull(pct))`) Pfizer B.A4/5 and `r label_number(1, big.mark=",")(filter(rst$cv$total_flowchart, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$total_flowchart, boost_type=="sanofi") %>% pull(pct))`) Sanofi.
--->
-
-<!-- `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c0") %>% pull(n) %>% sum())` were in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c1", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c1", boost_type=="pfizerBA45") %>% pull(pct_all))`) received Pfizer BA45 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c1", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c1", boost_type=="sanofi") %>% pull(pct_all))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).
--->
-
-`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c0") %>% pull(n) %>% sum())` were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c1", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c1", boost_type=="pfizerBA45") %>% pull(pct_all))`) received Pfizer BA45 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c1", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c1", boost_type=="sanofi") %>% pull(pct_all))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).
-
-Eligible Pfizer BA45 recipients were on average older, more deprived and had higher rates of prior clinical conditions than Sanofi recipients (supplementary Table S1).
-They were also more likely to have received more/less/X prior COVID-19 vaccine doses...
-
-In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1).
-
-In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1).
-
-Characteristics at the start of follow-up were well-balanced between groups (Table 1) with standardised mean differences consistently below 0.05 (supplementary Figure S2 and Table S1).
-
-In particular, prior clinical conditions, which were not directly matched on, were well-balanced between the groups after, though not before, matching.
-The median follow-up was X weeks for COVID-19 death outcomes (Table S3).
-
-SARS-CoV-2 testing rates after baseline were slightly less frequent for Pfizer BA45 than for Sanofi: `r rst$cv$eventcounts %>% filter(subgroup=="all", treatment==0) %>% pull(test_rate) %>% {label_number(0.001, 7)(.)}` versus `r rst$cv$eventcounts %>% filter(subgroup=="all", treatment==1) %>% pull(test_rate) %>% {label_number(0.001, 7)(.)}` per week for the CV cohort; and `r rst$age75plus$eventcounts %>% filter(subgroup=="all", treatment==0) %>% pull(test_rate) %>% {label_number(0.001, 7)(.)}` versus `r rst$age75plus$eventcounts %>% filter(subgroup=="all", treatment==1) %>% pull(test_rate) %>% {label_number(0.001, 7)(.)}` per week for the 75+ cohort (supplementary Table S4).
-
-### Estimated comparative effectiveness
-
-#### CV cohort
-
-```{r break3}
-2+2
+At 16 weeks there were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(n.event)` COVID-19 critical care admissions, and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths (Table 2).
+
+The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 2).
+The corresponding 16-week risks per `r perN_format` people for COVID-19 critical care admission were `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(rdCI)`).
+The 16-week risks per `r perN_format` people of COVID-19 death were `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`).
+For non-COVID-19 death, these were `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(rdCI)`).
+
+IRRs (where less than one favours Sanofi) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
+
+#### 75+ cohort
+
+```{r break4}
+
 ```
 
-By X weeks there were `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(n.event)` COVID-19 A&E attendances, `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths (Table 2).
-
-<!-- across `r contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 2). -->
-
-The X-week risks (cumulative incidence) per `r perN_format` people of COVID-19 A&E attendances were `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA45 and `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 2).
-The corresponding X-week risks per `r perN_format` people for COVID-19 hospitalisation were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI)`).
-The X-week risks per `r perN_format` people of COVID-19 death were `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`).
-For non-COVID-19 death, these were `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(rdCI)`).
-
-<!-- For COVID-19 A&E attendance, the cumulative risk ratio comparing Sanofi with Pfizer BA45 reached a minimum of 0.91 (0.90 to 0.92) at 4 weeks, then attenuated with increasing follow up (Figure 2).
--->
-
-<!-- For COVID-19 hospitalisation the minimum risk ratio was 0.67 (0.59 to 0.76) at around 11 weeks, attenuating to 0.89 (0.83 to 0.96) by 28 weeks. -->
-
-<!-- The number of COVID-19 deaths was too small to reliably examine patterns of cumulative incidence over time. -->
-
-<!-- For non-COVID-19 death the risk ratio was imprecisely estimated in the first few weeks, but stabilised to 0.92 (0.86 to 0.99) at 28 weeks. -->
-
-IRRs (where less than one favours Sanofi) in the X-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 A&E attendances, `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)` for COVID-19 hospitalisation, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
-
-Period-specific IRRs are displayed in Supplementary Figure S4a.
-During the first week after vaccination event rates were similar between vaccine groups: the estimated HR was `r rst$cv$contrasts_cuts_table %>% filter(outcome=="covidemergency", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI_95)` for COVID-19 A&E attendances, `r rst$cv$contrasts_cuts_table %>% filter(outcome=="covidadmitted", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 hospitalisation, `r rst$cv$contrasts_cuts_table %>% filter(outcome=="coviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 death, `r rst$cv$contrasts_cuts_table %>% filter(outcome=="noncoviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for non-COVID-19 death.
-After this initial period, the rate for A&E was lower for Sanofi than Pfizer BA45 until week X, after which they were higher for Sanofi.
-For COVID-19 hospitalisation, the minimum IRR was X (X to X) during weeks 5-8, with IRRs between X and X weeks consistent with no difference between vaccine groups.
-For non-COVID-19 death IRRs were consistently close to 1.
-
-#### 75+ cohort
-
-```{r break4}
-2+2
-```
-
-By X weeks there were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(n.event)` COVID-19 A&E attendances, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospitalisations, and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths (Table 2).
-
-<!-- across `r contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up (Table 2). -->
-
-The X-week risks (cumulative incidence) per `r perN_format` people of COVID-19 A&E attendances were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA45 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 2).
-The corresponding X-week risks per `r perN_format` people for COVID-19 hospitalisation were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI)`).
-The X-week risks per `r perN_format` people of COVID-19 death were `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`).
-For non-COVID-19 death, these were `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA45 and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(rdCI)`).
-
-<!-- For COVID-19 A&E attendance, the cumulative risk ratio comparing Sanofi with Pfizer BA45 reached a minimum of 0.91 (0.90 to 0.92) at 4 weeks, then attenuated with increasing follow up (Figure 2).
--->
-
-<!-- For COVID-19 hospitalisation the minimum risk ratio was 0.67 (0.59 to 0.76) at around 11 weeks, attenuating to 0.89 (0.83 to 0.96) by 28 weeks. -->
-
-<!-- The number of COVID-19 deaths was too small to reliably examine patterns of cumulative incidence over time. -->
-
-<!-- For non-COVID-19 death the risk ratio was imprecisely estimated in the first few weeks, but stabilised to 0.92 (0.86 to 0.99) at 28 weeks. -->
-
-IRRs (where less than one favours Sanofi) in the X-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidemergency", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 A&E attendance, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)` for COVID-19 hospitalisation, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
-
-Period-specific IRRs are displayed in Supplementary Figure S4a.
-During the first week after vaccination event rates were similar between vaccine groups: the estimated HR was `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidemergency", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI_95)` for COVID-19 A&E attendances, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidadmitted", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 hospitalisation, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="coviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 death, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="noncoviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for non-COVID-19 death.
-After this initial period, the rate for A&E was lower for Sanofi than Pfizer BA45 until week X, after which they were higher for Sanofi.
-For COVID-19 hospitalisation, the minimum IRR was X (X to X) during weeks 5-8, with IRRs between X and X weeks consistent with no difference between vaccine groups.
-For non-COVID-19 death IRRs were consistently close to 1.
+By 16 weeks there were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(n.event)` COVID-19 hospital admissions, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(n.event)` COVID-19 critical care admissions, and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` COVID-19 deaths (Table 2).
+
+The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 2).
+The corresponding 16-week risks per `r perN_format` people for COVID-19 critical care admission were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(rdCI)`).
+The 16-week risks per `r perN_format` people of COVID-19 death were `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`).
+For non-COVID-19 death, these were `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(rdCI)`).
+
+IRRs in the 16-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
+
+During the first week after vaccination event rates were similar between vaccine groups: the estimated HR was `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidadmitted", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidcritcare", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="coviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 death, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="noncoviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for non-COVID-19 death.
+
+Period-specific IRRs show that, in both cohorts during the first week after vaccination, event rates were similar between vaccine groups across all effectiveness outcomes (Supplementary Figure S4).
+
+### Estimated comparative safety
+
+In the CV cohort, there were `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients respectively.
+This corresponded to a cumulative incidence per `r perN_format` people of `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
+For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events, with a cumulative incidence of `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+
+In the 75+ cohort, `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients respectively, with a cumulative incidence per `r perN_format` people of `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
+For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events, with a cumulative incidence of `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
 
@@
 
-Findings in the main analyses were similar when examining subgroups...
-
-<!-- For COVID-19 A&E attendances the estimated relative benefit of Sanofi compared with Pfizer BA45 was greater/lower in those with fewer prior vaccines (IRR `r contrasts_table %>% filter(outcome=="covidemergency", subgroup=="covid_vax_history", subgroup_level=="pfizer-pfizer") %>% pull(irrCI_95)`) than in those who received ChAdOx1 (IRR `r contrasts_table %>% filter(outcome=="covidemergency", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(irrCI_95)`, heterogeneity p=`r contrasts_table %>% filter(outcome=="covidemergency", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhr.ln.p) %>% first()`) (Figure 3).
--->
-
-<!-- Similarly, the estimated relative benefit of mRNA-1273 compared with BNT162b2 was greater in those with prior evidence of infection (IRR `r contrasts_table %>% filter(outcome=="covidemergency", subgroup_level=="TRUE", subgroup=="prior_covid_infection") %>% pull(irrCI_95)`) versus those without (IRR `r contrasts_table %>% filter(outcome=="covidemergency", subgroup_level=="FALSE", subgroup=="prior_covid_infection") %>% pull(irrCI_95)`, heterogeneity p=`r contrasts_table %>% filter(outcome=="covidemergency", subgroup_level=="FALSE", subgroup=="prior_covid_infection") %>% pull(coxhr.ln.p) %>% first()`).
--->
-
-<!-- For COVID-19 hospitalisation the estimated relative benefit of mRNA-1273 compared with BNT162b2 was similar by type of first two vaccine doses (IRR `r contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="vax12_type", subgroup_level=="pfizer-pfizer") %>% pull(irrCI_95)` for BNT162b2 and `r contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(irrCI_95)` for ChAdOx1; heterogeneity p=`r contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="vax12_type", subgroup_level=="az-az") %>% pull(coxhr.ln.p) %>% first()`).
--->
-
-<!-- The estimated relative benefit of mRNA-1273 was greater in those with prior evidence of infection (IRR `r contrasts_table %>% filter(outcome=="covidadmitted", subgroup_level=="TRUE", subgroup=="prior_covid_infection") %>% pull(irrCI_95)`) than those without (IRR `r contrasts_table %>% filter(outcome=="covidadmitted", subgroup_level=="FALSE", subgroup=="prior_covid_infection") %>% pull(irrCI_95)`, heterogeneity p=`r contrasts_table %>% filter(outcome=="covidadmitted", subgroup_level=="FALSE", subgroup=="prior_covid_infection") %>% pull(coxhr.ln.p) %>% first()`).
--->
-
-<!-- For COVID-19 death, IRRs were estimated imprecisely due to low event rates so no strong conclusions about variation between subgroups could be made. -->
-
-<!-- For Non-COVID-19 death, the relative benefit of mRNA-1273 was generally greater in older and more clinically vulnerable people. -->
-
-Supplementary materials provide cumulative incidence plots by subgroup (Figures S3a-e), the corresponding cumulative risk differences, cumulative risk ratios, and period-specific IRRs (Figures S4a-e), and subgroup-specific heterogeneity tests for risk differences and risk ratios (Table S2).
+Findings in the main analyses were similar when examining subgroups, those estimated with reduced precision.
+Supplementary materials provide cumulative incidence plots by subgroup (Figures S3), the corresponding cumulative risk differences, cumulative risk ratios, and period-specific IRRs (Figures S4), and subgroup-specific heterogeneity tests for risk differences and risk ratios (Table S2).
 
@@
 
-This observational study in over X million adults receiving a COVID-19 vaccine during the Spring 2023 booster programme in England compared the effectiveness of Sanofi versus Pfizer BA45 vaccines and is, to our knowledge, the first to do so against severe COVID-19 outcomes.
-We estimated that ...
-there were no substantial differences between either vaccine on any of the studies effectiveness outcomes during the first X weeks after vaccination.
-Estimates appeared broadly similar regardless of age, clinical vulnerability, and ...
-
-Rates of COVID-19-related events were higher (the cumulative incidence curve is stepper) during the immune induction period up to 2 weeks, during which time the booster dose has yet to take full effect, than subsequently.
-Consistent with this delayed immunity, rates of the outcomes were similar for the two vaccines in the first week of follow-up, though rates also remained similar after this period.
-
-Only `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(n.event)` people aged 75 or over died with COVID-19 recorded as an underlying or contributory cause across both vaccine groups within X weeks (1 in 38,500) so the IRR was estimated imprecisely.
-<!-- The risk of non-COVID-19 death was around 1 in 1,600 people, and was marginally lower/higher for Sanofi than Pfizer BA45.
---> <!-- This could indicate differences in vaccine safety, under-ascertainment of COVID-19 related deaths (for example post-COVID cardiovascular deaths not directly attributed to SARS-CoV-2 infection), or sub-optimal confounding control.
--->
+This observational study of COVID-19 vaccine recipients during the spring 2023 booster programme in England is, to our knowledge, the first to directly compare the Pfizer BA.4-5 and Sanofi vaccines for safety and effectiveness against severe COVID-19 outcomes.
+We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Sanofi vaccine compared with Pfizer BA.4-5 during the first 16 weeks after vaccination.
+For the more severe COVID-19 outcomes of critical care admission and death, there were very few events and insufficient power to identify differences between vaccines.
 
@@
 
-The OpenSAFELY-TPP database covers around 45% of the English population and contains rich clinical information which enabled us to closely match Sanofi and Pfizer BA45 recipients to control for potential confounding, to study a range of clinical outcomes, including severe COVID-19, and compare comparative safety and effectiveness between important clinical subgroups.
+The OpenSAFELY-TPP database covers around 45% of the English population and contains rich clinical information which enabled us to closely match Pfizer BA.4-5 and Sanofi recipients to control for potential confounding and compare comparative safety and effectiveness.
 
 To make fair comparisons between the two vaccine types, we exploited the concurrent roll-out of both vaccines across the same eligible population during the same time period.
-The type of vaccine administered was based largely on local supply and availability, and clinical characteristics did not inform the type of vaccine offered on-site \[except for age??\].
-It remained important to control for any potential differences in the distribution of prognostic factors between the two vaccine groups, and these were well-balanced after matching.
-We cannot rule out residual confounding as we did not match on or adjust for all potential confounders, for example some clinical subgroups where relative protection between vaccines may differ such as immunosuppressive conditions.
-However, unmeasured potential confounders, particularly relating to unmeasured health-seeking behaviours, are less problematic in comparative effectiveness studies than in studies comparing vaccinated with unvaccinated people, as all people entering the study have sought and received at three vaccine doses.
-We found no evidence that rates of other outcomes during the first week after vaccination differed between the vaccine groups, although some of these comparisons were estimated imprecisely.
-
-<!--# [People were usually not offered a choice of vaccine type (although they could refuse boosting if they were not offered their preference) and clinical characteristics did not inform the type of vaccine offered.
-Anecdotally, we understand that occasionally some people were offered a choice in centres using both vaccines, though we are not aware of any specific policy that allowed this.] -->
+We used 1-1 matching to control for potential differences in the distribution of prognostic factors between the two vaccine groups, and although this led to the omission of a substantial number of people from the final analysis, characteristics were very well-balanced after matching.
+
+As an observational study, we cannot rule out residual confounding, though this is less problematic in comparative effectiveness studies than in studies comparing vaccinated with unvaccinated people, as all people entering the study have sought and received a booster vaccine.
 
@@
 
-<!-- The later introduction of the mRNA-1273 vaccine for the UK booster campaign meant that the matched cohort contained younger and healthier people than the whole population receiving a third dose, due to the prioritisation of older and higher risk people.
--->
-
-<!-- For example, the number of people with immunosuppressive conditions in the matched cohort was just 1.5%. -->
-
-<!-- The applicability of our findings to the most vulnerable groups may therefore be limited. -->
-
-<!-- However, numbers remained sufficient to compare subgroups defined by age, clinical vulnerability, and prior infection status, including for COVID-19 hospitalisation. -->
-
-To control for spatio-temporal heterogeneity in infection risk during the study period we matched on the date of third dose receipt and local health administration (STP).
-Some residual heterogeneity within STP regions is possible, but more precise geographical matching was not feasible due to poor matching success at this level, partly due to time periods during which only one vaccine type was available within a small geographical area.
-
-Post-baseline SARS-CoV-2 testing rates were slightly higher in X than in X recipients, which largely rule out differences in testing behaviours between vaccine groups.
-
 This study does not include anyone who received the Moderna mRNA-1273 vaccine which was also available during the spring 2023 booster programme but was administered much more infrequently.
@@
 
-<!-- The COV-BOOST phase II randomised trial assessed the safety and immunogenicity of seven COVID-19 vaccines for boosting, including BNT162b2 and mRNA-1273 against a menACWY control in people with no prior history of infection @munro2021.
--->
-
-<!-- It found higher immunogenicity of mRNA-1273 compared with BNT162b2 in subgroups with a two-dose BNT162b2 primary course (geometric mean ratio (GMR) of SARS-CoV-2 anti-spike IgG versus control = 11.49 (95%CI 9.36 to 14.12) versus (6.78 (5.51 to 8.35)) and a two-dose ChAdOx1 primary course (GMR = 32.30 (24.84 to 42.01) versus 16.80 (12.97 to 21.76)).
--->
-
-<!-- Likewise, the MixNMatch phase I-II non-randomised trial documented higher antibody levels after mRNA-1273 boost than BNT162b2 boost among individuals who received a primary course of BNT162b2 @atmar2022.
--->
-
-<!-- These trials did not evaluate efficacy endpoints, and we are aware of no other planned or published randomised trials making direct comparisons of boosting with BNT162b2 versus mRNA-1273. -->
-
-<!-- A study in Spanish registry data of people aged 40 years or over with no previous positive test for SARS-CoV-2 examined effectiveness of mRNA vaccine boosting, including a comparison of BNT162b2 with mRNA-1273 @monge2022.
--->
-
-<!-- For positive SARS-CoV-2 tests, it estimated a 34-day risk difference comparing mRNA-1273 with BNT162b2 of −2.2 (−2.7 to −1.6) per 1,000 people (risk ratio 0.87 (0.84, 0.90)). -->
-
-<!-- There was no assessment of effectiveness against hospitalisation or other severe outcomes. -->
-
-<!-- A study in US Veteran Affairs health care system in the Alpha and Delta eras using a similar matching approach @dickermanComparativeEffectivenessThird2023, with 65,196 matched recipients in each vaccine group, reported 16-week risk ratio of 0.87 (0.77 to 0.94) for documented SARS-CoV-2 infection, 0.61 (0.36 to 0.79) for COVID-19 hospitalisation and 0.92 (0.16 to 2.17) for COVID-19 death.
--->
-
-<!-- Comparisons of vaccines for primary vaccination in observational data have also been made. -->
-
-<!-- Another US Veteran Affairs study by the same group during the Alpha period @dickerman2022, identified an additional 1.23 (95% CI 0.72 to 1.81) documented infections per 1,000 people in those with a BNT162b2 first dose compared with a mRNA-1273 first dose at 24 weeks, 0.55 (95%CI 0.36 to 0.83) COVID-19 hospitalisations, 0.10 (95%CI 0.00 to 0.26) COVID-19 ICU admissions, and 0.02 (95%CI −0.06 to 0.12) COVID-19 deaths.
--->
-
-<!-- Another study using Veterans Affairs data with exact- and propensity-matching in a similar time period @ioannou2022a, estimated similar differences at 24 weeks for documented infection (1.73, 95%CI 1.50 to 1.96 additional events per 1,000 people in those with a BNT162b2 first dose compared with a mRNA-1273), COVID-19 hospitalisation (0.56, 95%CI 0.45 to 0.67) and COVID-19 death (0.03, 95%CI -0.00 to 0.07).
--->
-
-<!-- Accounting for the longer follow-up period and statistical uncertainty, these estimates are consistent with our findings for the comparative effectiveness of mRNA-1273 versus BNT162b2 boosting.
--->
-
-<!-- A third Veteran Affairs study compared three doses of BNT162b2 with three doses of mRNA-1273 in the Omicron era in a secondary analysis. https://doi.org/10.1093/cid/ciac328 -->
-
-<!-- A study using English data to estimate the effectiveness of BNT162b2 boosting versus no boosting and mRNA-1273 boosting versus no boosting separately using a test-negative-control design @andrews2022, suggests marginally higher effectiveness against symptomatic disease in mRNA-1273 than in BNT162b2, but this assumes that the respective control populations in each analysis were similar.
--->
-
-<!-- The evidence therefore strongly points to a benefit of mRNA-1273 over BNT162b2 for primary vaccination and subsequent booster doses. -->
-
-<!-- This is relevant to vaccine procurement decisions for future booster programmes. -->
-
-<!-- However, both vaccines are safe [@barda2021a; @klein2021; @diaz2021; @dickerman2022a] and strongly effective [@hulme2022; @barda2021; @andrews2022; @monge2022] against infection and COVID-19 disease, compared to no boosting.
--->
-
-<!-- Findings from this study should not discourage individuals from receiving BNT162b2 booster vaccination if offered. -->
-
-<!-- Though small, an additional 0.22 hospitalisations per 1,000 people becomes 100s or over 1,000 hospitalisations in England at a population level, depending on booster uptake and infection risk, contributing substantially to healthcare resource use both during the acute phase of disease and by those with longer-term illness -->
+...
 
@@
 
-COVID-19-related outcomes after vaccination with Sanofi or Pfizer BA45 during the Spring 2023 booster programme in England were rare, though risks were estimated to be similar between vaccine types.
-These finding were broadly consistent across subgroups.
-Given the similarly in effectiveness of the two vaccines compared in this study, future COVID-19 booster programmes may focus on other characteristics of the vaccines, for instance cost effectiveness or logistical constraints, to determine which to recommend for roll-out.
+During the spring 2023 COVID-19 booster programme in England, COVID-19-related hospital admissions were slightly more frequent following receipt of the Pfizer BA.4-5 vaccine than the Sanofi vaccine after matching on a number of important potential confounders.
+These finding were broadly consistent across subgroups, though estimated with less precision.
 
@@
 
-Conceptualization: WJH, EPKP, RHK, EJW, VW, TP, BG, MAH, and JACS.
+Conceptualization: WJH, EPKP, VW, TP, BG, MAH, and JACS.
 Data curation: WJH, HJC, AW, AM, BM, CA, and SCJB.
@@
 Funding acquisition: JM, BG, and JACS.
-Methodology: WJH, EPKP, RHK, EJW, VW, TP, MAH, and JACS.
+Methodology: WJH, EPKP, VW, TP, MAH, and JACS.
 Project administration: WJH, AM, JM, BG, CA, and JACS.
@@
 Writing - original draft: WJH.
-Writing - review & editing: WJH, EPKP, RHK, EJW, VW, TP, HJC, AW, CA, AM, JM, BM, SCJB, BG, MAH, and JACS.
+Writing - review & editing: WJH, EPKP, VW, TP, HJC, AW, CA, AM, JM, BM, SCJB, BG, MAH, and JACS.
 
@@
 
-NHS England is the data controller for OpenSAFELY-TPP; TPP is the data processor; and study authors using OpenSAFELY have the approval of NHS England.
-This implementation of OpenSAFELY is hosted within the TPP environment which is accredited to the ISO 27001 information security standard and is NHS IG Toolkit compliant @datasec; Patient data has been pseudonymised for analysis and linkage using industry standard cryptographic hashing techniques; all pseudonymised datasets transmitted for linkage onto OpenSAFELY are encrypted; access to the platform is via a virtual private network (VPN) connection, restricted to a small group of researchers; the researchers hold contracts with NHS England and only access the platform to initiate database queries and statistical models; all database activity is logged; only aggregate statistical outputs leave the platform environment following best practice for anonymisation of results such as statistical disclosure control for low cell counts @isb1523 .
-The OpenSAFELY research platform adheres to the obligations of the UK General Data Protection Regulation (GDPR) and the Data Protection Act 2018.
-In March 2020, the Secretary of State for Health and Social Care used powers under the UK Health Service (Control of Patient Information) Regulations 2002 (COPI) to require organisations to process confidential patient information for the purposes of protecting public health, providing healthcare services to the public and monitoring and managing the COVID-19 outbreak and incidents of exposure; this sets aside the requirement for patient consent @coronavi2020 .
-Taken together, these provide the legal bases to link patient datasets on the OpenSAFELY platform.
-General practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.
-
-This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by the LSHTM Ethics Board (reference 21863).
-
-JACS is the guarantor.
+NHS England is the data controller for the NHS England OpenSAFELY COVID-19 Service; TPP is the data processor; all study authors using OpenSAFELY have the approval of NHS England @NHSEnglandOpenSAFELY.
+This implementation of OpenSAFELY is hosted within the TPP environment which is accredited to the ISO 27001 information security standard and is NHS IG Toolkit compliant @datasec.
+
+Patient data has been pseudonymised for analysis and linkage using industry standard cryptographic hashing techniques; all pseudonymised datasets transmitted for linkage onto OpenSAFELY are encrypted; access to the NHS England OpenSAFELY COVID-19 service is via a virtual private network (VPN) connection; the researchers hold contracts with NHS England and only access the platform to initiate database queries and statistical models; all database activity is logged; only aggregate statistical outputs leave the platform environment following best practice for anonymisation of results such as statistical disclosure control for low cell counts @isb1523.
+
+The service adheres to the obligations of the UK General Data Protection Regulation (UK GDPR) and the Data Protection Act 2018.
+The service previously operated under notices initially issued in February 2020 by the the Secretary of State under Regulation 3(4) of the Health Service (Control of Patient Information) Regulations 2002 (COPI Regulations), which required organisations to process confidential patient information for COVID-19 purposes; this set aside the requirement for patient consent @WithdrawnWithdrawnCoronavirus2022.
+As of 1 July 2023, the Secretary of State has requested that NHS England continue to operate the Service under the COVID-19 Directions 2020 @COVID19PublicHealth.
+In some cases of data sharing, the common law duty of confidence is met using, for example, patient consent or support from the Health Research Authority Confidentiality Advisory Group @ConfidentialityAdvisoryGroup.
+
+Taken together, these provide the legal bases to link patient datasets using the service.
+GP practices, which provide access to the primary care data, are required to share relevant health information to support the public health response to the pandemic, and have been informed of how the service operates.
+
+This study was approved by the Health Research Authority (REC reference 20/LO/0651).
+
+## Data access and verification
+
+Access to the underlying identifiable and potentially re-identifiable pseudonymised electronic health record data is tightly governed by various legislative and regulatory frameworks, and restricted by best practice.
+The data in the NHS England OpenSAFELY COVID-19 service is drawn from General Practice data across England where TPP is the data processor.
+
+TPP developers initiate an automated process to create pseudonymised records in the core OpenSAFELY database, which are copies of key structured data tables in the identifiable records.
+These pseudonymised records are linked onto key external data resources that have also been pseudonymised via SHA-512 one-way hashing of NHS numbers using a shared salt.
+University of Oxford, Bennett Institute for Applied Data Science developers and PIs, who hold contracts with NHS England, have access to the OpenSAFELY pseudonymised data tables to develop the OpenSAFELY tools.
+
+These tools in turn enable researchers with OpenSAFELY data access agreements to write and execute code for data management and data analysis without direct access to the underlying raw pseudonymised patient data, and to review the outputs of this code.
+All code for the full data management pipeline --- from raw data to completed results for this analysis --- and for the OpenSAFELY platform as a whole is available for review at www.github.com/OpenSAFELY.
 
@@
 
-## Data access and verification
-
-Access to the underlying identifiable and potentially re-identifiable pseudonymised electronic health record data is tightly governed by various legislative and regulatory frameworks, and restricted by best practice.
-The data in OpenSAFELY is drawn from General Practice data across England where TPP is the data processor.
-
-TPP developers initiate an automated process to create pseudonymised records in the core OpenSAFELY database, which are copies of key structured data tables in the identifiable records.
-These pseudonymised records are linked onto key external data resources that have also been pseudonymised via SHA-512 one-way hashing of NHS numbers using a shared salt.
-Bennett Institute for Applied Data Science developers and PIs holding contracts with NHS England have access to the OpenSAFELY pseudonymised data tables as needed to develop the OpenSAFELY tools.
-
-These tools in turn enable researchers with OpenSAFELY data access agreements to write and execute code for data management and data analysis without direct access to the underlying raw pseudonymised patient data, and to review the outputs of this code.
-All code for the full data management pipeline---from raw data to completed results for this analysis---and for the OpenSAFELY platform as a whole is available for review at github.com/OpenSAFELY.
-
-\newpage
+## Acknowledgements
+
+We are very grateful for all the support received from the TPP Technical Operations team throughout this work, and for generous assistance from the information governance and database teams at NHS England and the NHS England Transformation Directorate.
 
@@
 
-# tab1 <- rst$cv$match_table1_wide %>%
-#   filter(
-#     strategy=="post",
-#     !(variable  %in% c("N", "immunosuppressed", "region"))
-#   ) %>%
-#   mutate(
-#     smd=if_else(variable %in% c(matching_variables$B$exact, "ageband"), NA_character_, smd),
-#     var_label = fct_recode(var_label, "Any immunosuppressive condition"= "Immunosuppressed (all)"),
-#     order_level = row_number(),
-#     order_var = cumsum(lag(variable, 1, first(variable))!=variable),
-#     var_label = if_else(lag(as.character(var_label),n=1,default="")==var_label, "", as.character(var_label)),
-#   ) %>%
-#   select(var_label, variable_levels, stat_0, stat_1, smd, order_var, level) %>%
-#   gt(
-#     #groupname_col = "var_label"
-#   ) %>%
-#   cols_label(
-#     var_label = "Variable",
-#     variable_levels = " ",
-#     `stat_0` = glue("Pfizer BA45 (N={filter(rst$cv$match_table1, strategy=='post', variable=='N', by==0) %>% pull(n)})"),
-#     `stat_1` = glue("Sanofi (N={filter(rst$cv$match_table1, strategy=='post', variable=='N', by==1) %>% pull(n)})"),
-#     `smd` = "Standardised mean difference"
-#   ) %>%
-#   cols_align(
-#     align = c("right"),
-#     columns =  c("stat_0", "stat_1")
-#   ) %>%
-#   cols_align(
-#     align = c("left"),
-#     columns =  c("var_label", "variable_levels")
-#   ) %>%
-#   sub_missing(
-#       columns = "smd",
-#       missing_text = "-"
-#   ) %>%
-#   tab_style(
-#     style = list(
-#       cell_text(weight = "bold")
-#     ),
-#     locations = cells_column_labels(everything())
-#   ) %>%
-#   tab_style(
-#     style = list(
-#       cell_fill(color = "gray96")
-#     ),
-#     locations = cells_body(
-#       rows = order_var%%2 == 0
-#     )
-#   ) %>%
-#   cols_hide(c("level", "order_var")) %>%
-#   tab_options(
-#     #row_group.as_column = TRUE,
-#     table.font.size = 8
-#   )
-
 match_table1_wide_wide <-
@@
   filter(
-    strategy=="post",
+    matchset!="prematch",
     !(variable  %in% c("N", "immunosuppressed", "region"))
@@
     variable_levels = "",
-    `stat_0_cv` = glue("Pfizer BA45 (N={filter(match_table1, strategy=='post', cohort=='cv', variable=='N', by==0) %>% pull(n)})"),
-    `stat_1_cv` = glue("Sanofi (N={filter(match_table1, strategy=='post', cohort=='cv', variable=='N', by==1) %>% pull(n)})"),
+    `stat_0_cv` = glue("Pfizer BA.4-5 (N={filter(match_table1, matchset=='postmatch', cohort=='cv', variable=='N', by==0) %>% pull(n)})"),
+    `stat_1_cv` = glue("Sanofi (N={filter(match_table1, matchset=='postmatch', cohort=='cv', variable=='N', by==1) %>% pull(n)})"),
    `smd_cv` = "SMD",
-    `stat_0_age75plus` = glue("Pfizer BA45 (N={filter(match_table1, strategy=='post', cohort=='age75plus', variable=='N', by==0) %>% pull(n)})"),
-    `stat_1_age75plus` = glue("Sanofi (N={filter(match_table1, strategy=='post', cohort=='age75plus', variable=='N', by==1) %>% pull(n)})"),
+    `stat_0_age75plus` = glue("Pfizer BA.4-5 (N={filter(match_table1, matchset=='postmatch', cohort=='age75plus', variable=='N', by==0) %>% pull(n)})"),
+    `stat_1_age75plus` = glue("Sanofi (N={filter(match_table1, matchset=='postmatch', cohort=='age75plus', variable=='N', by==1) %>% pull(n)})"),
     `smd_age75plus` = "SMD"
@@
 
-#### Table 2: Estimates at X-week follow-up of risk per `r perN_format` people in each vaccine group, and corresponding risk differences, overall and within subgroups
+#### Table 2: Estimates at 16-week follow-up of risk per `r perN_format` people in each vaccine group, and corresponding risk differences, overall and within subgroups
 
@@
   filter(
-    outcome %in% outcomes_primary,
+    outcome %in% outcomes,
     subgroup == "all"
@@
     rdCI,
-    #rrCI,
-    #kmirrCI,
-    #coxhrCI,
+    rrCI,
+    irrCI
+
   ) %>%
@@
     rdCI = glue("Risk difference per {perN_format} people (95% CI)"),
-    #rrCI = "Risk ratio (95% CI)",
-    #kmirrCI = "Incidence rate ratio (95% CI)",
-    #coxhrCI = "Hazard ratio (95% CI)"
+    rrCI = "Risk ratio (95% CI)",
+    irrCI = "Incidence rate ratio (95% CI)"
   ) %>%
   tab_spanner(
-    label = "Pfizer BA45",
+    label = "Pfizer BA.4-5",
     columns = ends_with("_0")
@@
    tab_spanner(
-    label = "Comparison (reference=Pfizer BA45)",
+    label = "Comparison (reference = Pfizer BA.4-5)",
     columns = ends_with("CI")
@@
   tab_spanner(
-    label = "Pfizer BA45",
+    label = "Pfizer BA.4-5",
     columns = ends_with("pfizerBA45")
@@
 
-```{r, function-ci, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
+```{r}
+#| label: function-ci
+#| out.width: '100%'
+#| out.height: '50%'
+#| output: 'asis'
+#| out-extra: ''
 
@@
       subgroup==subgroup0,
-      outcome %in% outcomes_primary,
+      outcome %in% outcomes,
       outcome != "noncoviddeath"
@@
     mutate(
-      treatment_descr = fct_recode(treatment_descr, "pfizerBA45"="Pfizer BA45", "sanofi"="Sanofi"),
+      treatment_descr = fct_recode(treatment_descr, "pfizerBA45"="Pfizer BA.4-5", "sanofi"="Sanofi"),
       colgroup = interaction(treatment_descr, subgroup_level_descr, sep=", "),
@@
     mutate(
-      risk = if_else(outcome=="covidemergency" & time>20*7, NA_real_, risk),
-      risk.ll = if_else(outcome=="covidemergency" & time>20*7, NA_real_, risk.ll),
-      risk.ul = if_else(outcome=="covidemergency" & time>20*7, NA_real_, risk.ul),
-      lagrisk=lag(risk,1,0),
+      risk = if_else(outcome=="covidadmitted" & time>20*7, NA_real_, risk),
+      risk.ll = if_else(outcome=="covidadmitted" & time>20*7, NA_real_, risk.ll),
+      risk.ul = if_else(outcome=="covidadmitted" & time>20*7, NA_real_, risk.ul),
+      lagrisk = lag(risk,1,0),
     ) %>%
@@
       subgroup==subgroup0,
-      outcome %in% outcomes_primary
+      outcome %in% outcomes
     ) %>%
@@
   plot_irr <- rst[[cohort]]$contrasts_cuts %>%
-    # equivalent to the hazard
+    # equivalent(ish) to the hazard
     filter(
       subgroup==subgroup0,
-      outcome %in% outcomes_primary
+      outcome %in% outcomes
     ) %>%
@@
       x="Weeks",
-      y="higher risk\n<- Pfizer BA45 | Sanofi ->",
+      y="higher risk\n<- Pfizer BA.4-5 | Sanofi ->",
       title="Incidence rate ratio",
@@
     #plot_rr, 
-    #plot_kmirr,
-    #plot_coxhr,
+    plot_irr,
     ncol=2,
@@
     path=output_dir_qmd,
-    plot=plot_ci,
+    plot=plot_combined,
     unit="cm",
@@
 
-  plot_ci
+  plot_combined
 }
@@
 
-#### Figure 2: Estimates of cumulative risk per `r perN_format` people (left plots) and cumulative risk ratios (right plots).
-
-```{r, figure-ci-all}
+#### Figure 2a: Estimates in the CV cohort of cumulative risk per `r perN_format` people (left plots) and incidence rate ratios (right plots).
+
+```{r figure-ci-all}
 #| echo: false
@@
 
+#### Figure 2b: Estimates in the 75+ cohort of cumulative risk per `r perN_format` people (left plots) and incidence rate ratios (right plots).
+
+```{r figure-ci-age75plus-all}
+#| echo: false
+#| message: false
+#| warning: false
+#| out-width: '100%'
+#| out-height: '50%'
+#| output: 'asis'
+#| out-extra: ''
+
+plot_subgroup("age75plus", "all")
+```
+
 \newpage
 
-#### Figure 3: Estimated incidence rate ratios at X-week follow-up comparing the effectiveness of Sanofi and Pfizer BA45, overall and within subgroups, together with p-values for between-subgroup heterogeneity.
-
-<!-- ```{r figure-contrasts-irr} -->
-
-<!-- #| echo: false -->
-
-<!-- #| message: false -->
-
-<!-- #| warning: false -->
-
-<!-- #| fig-height: 10 -->
-
-<!-- #| out-extra: '' -->
-
-<!-- contrasts_plot <-  -->
-
-<!--   contrasts_overall %>% -->
-
-<!--   filter( -->
-
-<!--     outcome %in% outcomes, -->
-
-<!--     subgroup %in% c("all", "ageband", "cv"), -->
-
-<!--   ) %>% -->
-
-<!--   group_by(cohort_descr, outcome_descr) %>% -->
-
-<!--   mutate( -->
-
-<!--    # outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10), -->
-
-<!--     subgroup_level_descr = fct_rev(subgroup_level_descr), -->
-
-<!--     subgroup_level_descr = fct_recode(subgroup_level_descr, "All"=""), -->
-
-<!--     irrCI =   glue("{label_number_rr(irr)} ({label_number_rr(irr.ll)} - {label_number_rr(irr.ul)})"), -->
-
-<!--     order = cumsum(lag(subgroup_descr, 1, first(subgroup_descr))!=subgroup_descr), -->
-
-<!--     patch = (order %% 2) == 1, -->
-
-<!--     rownum = n() - row_number() + 1 -->
-
-<!--   ) %>% -->
-
-<!--   group_by(cohort_descr, outcome_descr, subgroup_descr) %>% -->
-
-<!--   mutate( -->
-
-<!--     midgroup = mean(rownum), -->
-
-<!--     irr.ln.p = if_else(row_number()==1 & (subgroup!="all"), irr.ln.p, NA_real_), -->
-
-<!--     irr.ln.p = if_else(irr.ln.p<0.001, "<0.001", label_number(0.001, trim=FALSE)(irr.ln.p)) -->
-
-<!--   ) -->
-
-<!-- plot_irr_subgroup <- contrasts_plot %>% -->
-
-<!--   ggplot(aes(y=subgroup_level_descr)) + -->
-
-<!--   geom_vline(aes(xintercept=1), linetype="dotted", colour="darkgrey")+ -->
-
-<!--   geom_point(aes(x=irr), position=position_dodge(width=-0.3))+ -->
-
-<!--   geom_linerange(aes(xmin=irr.ll, xmax=irr.ul), position=position_dodge(width=-0.3))+ -->
-
-<!--   geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=min(irr.ll), xmax=max(irr.ul), fill=patch), alpha=0.2)+ -->
-
-<!--   #facet_grid(rows=vars(outcome_descr), scales="free", space="free_y", switch="y")+ -->
-
-<!--   facet_wrap(vars(outcome_descr), scales="free_y", ncol=1, strip.position = "top")+ -->
-
-<!--   scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","), oob=~oob_squish(.,range=log(c(0.4,2), 10)), expand=expansion())+ -->
-
-<!--   #scale_fill_brewer(type="qual", palette="Set2")+ -->
-
-<!--   scale_fill_manual(values=c("transparent", "darkgrey"))+ -->
-
-<!--   labs( -->
-
-<!--     subtitle="\n\n\nX-week Hazard Ratio", -->
-
-<!--     x="<- favours Sanofi  |  favours Pfizer BA45 ->", -->
-
-<!--     y=NULL -->
-
-<!--   )+ -->
-
-<!--   theme_minimal()+ -->
-
-<!--   theme( -->
-
-<!--     plot.margin=grid::unit(c(0,0,0,0), "mm"), -->
-
-<!--     plot.subtitle=element_text(hjust=0.5, vjust=0), -->
-
-<!--     legend.position="none", -->
-
-<!--     #panel.grid.minor.x = element_blank(), -->
-
-<!--     panel.grid.major.y = element_blank(), -->
-
-<!--     panel.grid.minor.y = element_blank(), -->
-
-<!--     #strip.background = element_blank(), -->
-
-<!--     #strip.placement="outside", -->
-
-<!--     strip.text = element_text(hjust=0), -->
-
-<!--     #strip.text.y.left = element_text(angle=0), -->
-
-<!--     #strip.text.y.left = element_blank(), -->
-
-<!--     panel.border = element_blank(), -->
-
-<!--     panel.spacing = unit(0.5, "lines"), -->
-
-<!--   ) -->
-
-<!-- plot_irr_subgroup_table <- contrasts_plot %>% -->
-
-<!--   ggplot(aes(y=subgroup_level_descr)) + -->
-
-<!--   geom_text(aes(x=0.5, label=irrCI), hjust=0.5, vjust=0.5, size=9/.pt)+ -->
-
-<!--   geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=0.2, xmax=0.8, fill=patch), alpha=0.2)+ -->
-
-<!--   facet_wrap(vars(outcome_descr), scales="free_y", ncol=1, strip.position = "top")+ -->
-
-<!--   scale_fill_manual(values=c("transparent", "darkgrey"))+ -->
-
-<!--   scale_x_continuous(expand=expansion())+ -->
-
-<!--   labs(subtitle="\nHazard\nRatio\n(95% CI)")+ -->
-
-<!--   theme_minimal()+ -->
-
-<!--   theme( -->
-
-<!--     plot.margin=grid::unit(c(0,0,0,0), "mm"), -->
-
-<!--     plot.subtitle=element_text(hjust=0.5, vjust=0), -->
-
-<!--     legend.position="none", -->
-
-<!--     panel.grid = element_blank(), -->
-
-<!--     strip.text =element_text(colour="transparent"), -->
-
-<!--     axis.text.y = element_blank(), -->
-
-<!--     axis.title.y = element_blank(), -->
-
-<!--     axis.text.x = element_blank(), -->
-
-<!--     axis.title.x=element_blank(), -->
-
-<!--     panel.border = element_blank(), -->
-
-<!--     panel.spacing = unit(0.5, "lines"), -->
-
-<!--   )+ -->
-
-<!--   NULL -->
-
-<!-- contrasts_plot %>% select(outcome_descr, subgroup_level_descr, irrCI, irr.ln.p) %>% write_csv(here("table3.csv")) -->
-
-<!-- plot_p_subgroup_table <- contrasts_plot %>% -->
-
-<!--   ggplot() + -->
-
-<!--   geom_text(aes(x=0.5, y=subgroup_level_descr, label=irr.ln.p), hjust=0.5, vjust=0.5, size=9/.pt, colour="transparent")+ -->
-
-<!--   geom_text(aes(x=0.5, y=midgroup, label=irr.ln.p), hjust=0.5, vjust=0.55, size=9/.pt)+ -->
-
-<!--   geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=0, xmax=1, fill=patch), alpha=0.2)+ -->
-
-<!--   facet_wrap(vars(outcome_descr), scales="free_y", ncol=1, strip.position = "top")+ -->
-
-<!--   scale_fill_manual(values=c("transparent", "darkgrey"))+ -->
-
-<!--   scale_x_continuous(expand=expansion())+ -->
-
-<!--   labs(subtitle="P for\nbetween\nsubgroup\ndifferences")+ -->
-
-<!--   theme_minimal()+ -->
-
-<!--   theme( -->
-
-<!--     plot.margin=grid::unit(c(0,0,0,0), "mm"), -->
-
-<!--     plot.subtitle=element_text(hjust=0.5, vjust=0), -->
-
-<!--     legend.position="none", -->
-
-<!--     panel.grid = element_blank(), -->
-
-<!--     strip.text =element_text(colour="transparent"), -->
-
-<!--     axis.text.y = element_blank(), -->
-
-<!--     axis.title.y = element_blank(), -->
-
-<!--     axis.text.x = element_blank(), -->
-
-<!--     axis.title.x=element_blank(), -->
-
-<!--     panel.border = element_blank(), -->
-
-<!--     panel.spacing = unit(0.5, "lines"), -->
-
-<!--   )+ -->
-
-<!--   NULL -->
-
-<!--   plot_combined <- patchwork::wrap_plots( -->
-
-<!--     plot_irr_subgroup,  -->
-
-<!--     plot_irr_subgroup_table,  -->
-
-<!--     plot_p_subgroup_table, -->
-
-<!--     ncol=3, -->
-
-<!--     widths=c(2,0.7,0.5) -->
-
-<!--   ) -->
-
-<!--   ggsave( -->
-
-<!--     filename=glue("irr_pvals.jpeg"), -->
-
-<!--     path=output_dir_qmd, -->
-
-<!--     plot=plot_combined, -->
-
-<!--     width=20, height=20, units="cm" -->
-
-<!--   ) -->
-
-<!-- plot_combined -->
-
-<!-- cat("  \n\n") -->
-
-<!-- ``` -->
+#### Figure 3: Estimated incidence rate ratios at comparing the safety (4-weeks) and effectiveness (16-weeks) of Sanofi and Pfizer BA.4-5, overall and within subgroups, together with p-values for between-subgroup heterogeneity.
+
+```{r figure-contrasts-irr}
+#| echo: false
+#| message: false
+#| warning: false
+#| fig-height: 10
+#| out-extra: ''
+
+contrasts_plot <-
+  contrasts_overall %>%
+  filter(
+    outcome %in% outcomes,
+    subgroup %in% c("all"),
+  ) %>%
+  group_by(outcome_descr) %>%
+  mutate(
+    outcome_descr = fct_rev(outcome_descr),
+    #outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
+    #subgroup_level_descr = fct_rev(subgroup_level_descr),
+    #subgroup_level_descr = fct_recode(subgroup_level_descr, "All"=""),
+    irrCI =   glue("{label_number_rr(irr)} ({label_number_rr(irr.ll)} - {label_number_rr(irr.ul)})"),
+  ) %>%
+  group_by(cohort_descr, outcome_descr) %>%
+  mutate(
+    irr.ln.p = if_else(row_number()==1 & (subgroup!="all"), irr.ln.p, NA_real_),
+    irr.ln.p = if_else(irr.ln.p<0.001, "<0.001", label_number(0.001, trim=FALSE)(irr.ln.p))
+  )
+
+plot_irr_lines <- contrasts_plot %>%
+  ggplot(aes(y=outcome_descr)) +
+  geom_vline(aes(xintercept=1), linetype="dotted", colour="darkgrey")+
+  geom_point(aes(x=irr), position=position_dodge(width=0.3))+
+  geom_linerange(aes(xmin=irr.ll, xmax=irr.ul), position=position_dodge(width=0.3))+
+  #facet_grid(rows=vars(outcome_descr), scales="free", space="free_y", switch="y")+
+  facet_wrap(vars(cohort_descr), scales="free_y", ncol=1, strip.position = "top")+
+  scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","), oob=~oob_squish(.,range=log(c(0.4,2), 10)), expand=expansion())+
+  #scale_fill_brewer(type="qual", palette="Set2")+
+  scale_fill_manual(values=c("transparent", "darkgrey"))+
+  labs(
+    subtitle="\n\n\nIncidence Rate Ratios",
+    x="<- favours Sanofi  |  favours Pfizer BA.4-5 ->",
+    y=NULL
+  )+
+  theme_minimal()+
+  theme(
+    plot.margin=grid::unit(c(0,0,0,0), "mm"),
+    plot.subtitle=element_text(hjust=0.5, vjust=0),
+    legend.position="none",
+    #panel.grid.minor.x = element_blank(),
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    #strip.background = element_blank(),
+    #strip.placement="outside",
+    strip.text = element_text(hjust=0),
+    #strip.text.y.left = element_text(angle=0),
+    #strip.text.y.left = element_blank(),
+    panel.border = element_blank(),
+    panel.spacing = unit(0.5, "lines"),
+  )
+
+plot_irr_table <- contrasts_plot %>%
+  ggplot(aes(y=outcome_descr)) +
+  geom_text(aes(x=0.5, label=irrCI), hjust=0.5, vjust=0.5, size=9/.pt)+
+  #geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=0.2, xmax=0.8, fill=patch), alpha=0.2)+
+  facet_wrap(vars(cohort_descr), scales="free_y", ncol=1, strip.position = "top")+
+  scale_fill_manual(values=c("transparent", "darkgrey"))+
+  scale_x_continuous(expand=expansion())+
+  #labs(subtitle="\nIncidence\nRate Ratio\n(95% CI)")+
+  theme_minimal()+
+  theme(
+    plot.margin=grid::unit(c(0,0,0,0), "mm"),
+    plot.subtitle=element_text(hjust=0.5, vjust=0),
+    legend.position="none",
+    panel.grid = element_blank(),
+    strip.text =element_text(colour="transparent"),
+    axis.text.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.x = element_blank(),
+    axis.title.x=element_blank(),
+    panel.border = element_blank(),
+    panel.spacing = unit(0.5, "lines"),
+  )+
+  NULL
+
+# contrasts_plot %>% select(cohort_descr, subgroup_level_descr, outcome_descr, irrCI, irr.ln.p) %>% write_csv(here("table3.csv"))
+# 
+# plot_p_table <- contrasts_plot %>%
+#   ggplot() +
+#   geom_text(aes(x=0.5, y=subgroup_level_descr, label=irr.ln.p), hjust=0.5, vjust=0.5, size=9/.pt, colour="transparent")+
+#   geom_text(aes(x=0.5, y=midgroup, label=irr.ln.p), hjust=0.5, vjust=0.55, size=9/.pt)+
+#   geom_rect(aes(ymin=rownum-0.5, ymax=rownum+0.5, xmin=0, xmax=1, fill=patch), alpha=0.2)+
+#   facet_wrap(vars(outcome_descr), scales="free_y", ncol=1, strip.position = "top")+
+#   scale_fill_manual(values=c("transparent", "darkgrey"))+
+#   scale_x_continuous(expand=expansion())+
+#   labs(subtitle="P for\nbetween\nsubgroup\ndifferences")+
+#   theme_minimal()+
+#   theme(
+#     plot.margin=grid::unit(c(0,0,0,0), "mm"),
+#     plot.subtitle=element_text(hjust=0.5, vjust=0),
+#     legend.position="none",
+#     panel.grid = element_blank(),
+#     strip.text =element_text(colour="transparent"),
+#     axis.text.y = element_blank(),
+#     axis.title.y = element_blank(),
+#     axis.text.x = element_blank(),
+#     axis.title.x=element_blank(),
+#     panel.border = element_blank(),
+#     panel.spacing = unit(0.5, "lines"),
+#   )+
+#   NULL
+
+  plot_combined <- patchwork::wrap_plots(
+    plot_irr_lines,
+    plot_irr_table,
+    #plot_p_subgroup_table,
+    ncol=3,
+    widths=c(2,0.7,0.5)
+  )
+
+  ggsave(
+    filename=glue("irr_pvals.jpeg"),
+    path=output_dir_qmd,
+    plot=plot_combined,
+    width=20, height=20, units="cm"
+  )
+
+plot_combined
+
+cat("  \n\n")
+
+```
 

### 2024-01-31 write-up/manuscript/supplement.qmd
update manuscript

---
title: Supplementary materials for Comparative safety and effectiveness of Pfizer BA.4-5 versus Sanofi during the Spring 2023 COVID-19 booster vaccination programme in England
format:
  # html:
  #   toc: false
  #   embed-resources: true
  docx
output-file: supplementary_draft.docx
keep-md: false
execute:
  echo: false
  output: true
  warning: false
  #message: true
knitr:
  opts_chunk: 
    #fig.path: ./
    fig.pos: H
    R.options:
      knitr.graphics.auto_pdf: true
bibliography: references.bib
csl: vancouver.csl
link-citations: true
zotero: true
# header-includes:
# - \usepackage{float} 
# - \floatplacement{figure}{H}
# - \usepackage{caption}
# - \captionsetup[figure]{labelformat=empty}
# - \captionsetup[table]{labelformat=empty}
---

```{r}
#| label: setup

source(here::here("write-up", "manuscript", "setup.R"))
```

```{r}
#| label: process

source(here("write-up", "manuscript", "import-process.R"))
```

```{r}
#| label: figure-function

show_coverage <- function(.cohort){

  xmin <- min(match_coverage$boost_date )
  xmax <- max(match_coverage$boost_date )+1

  data_annotate <- tibble(
      boost_date = date(c("2023-05-01", "2023-05-01", "2023-05-20", "2023-06-07")),
      cumuln = c(1000000, -100000, 0, -10000),
      dailyn = c(50000, -25000, 0, -12000),
      label = c("Sanofi", "Pfizer BA.4-5", "Matched", "Unmatched"),
      colour = c(brewer_pal(type="qual", palette="Set2")(2)[c(2,1)], "grey50", "grey70"),
      vjust = c(-1, 1, 0, 0.5)
    )

  plot_coverage_cumuln <-
    rst[[.cohort]]$match_coverage %>%
    mutate(
     # treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
      #status_descr = fct_rev(fct_recoderelevel(as.character(status), recoder$status)),
      #status_descr = fct_rev(status_descr),
      cumuln = cumuln*((treatment*2) - 1),
      dailyn = n*((treatment*2) - 1)
    ) %>%
    ggplot()+
    geom_col(
      aes(
        x=boost_date+0.5,
        y=dailyn,
        group=paste0(treatment,status_descr),
        fill=treatment_descr,
        alpha=status_descr,
        colour=NULL
      ),
      position=position_stack(reverse=TRUE),
      width=1
    )+
    geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
    geom_text(data=data_annotate, aes(x=boost_date, y=dailyn, label=label), colour=data_annotate$colour, vjust=data_annotate$vjust, hjust=1.05, fontface="bold")+
    scale_x_date(
      breaks = seq(xmin, xmax+1, by=14),
      #breaks = unique(lubridate::ceiling_date(match_coverage$boost_date, "1 month")),
      limits = c(xmin-1, NA),
      labels = scales::label_date("%d/%m"),
      expand = expansion(add=1),
      sec.axis = sec_axis(
        trans = ~as.Date(.),
        breaks=as.Date(seq(floor_date(xmin, "month"), ceiling_date(xmax, "month"),by="month")),
        labels = scales::label_date("%B %y")
      )
    )+
    scale_y_continuous(
      labels = ~scales::label_number(accuracy = 1, big.mark=",")(abs(.)),
      expand = expansion(c(0, NA))
    )+
    scale_fill_brewer(type="qual", palette="Set2")+
    scale_colour_brewer(type="qual", palette="Set2")+
    scale_alpha_discrete(range= c(0.8,0.4))+
    labs(
      x="Date",
      y="Booster recipients",
      colour=NULL,
      fill=NULL,
      alpha=NULL
    ) +
    theme_minimal()+
    theme(
      axis.line.x.bottom = element_blank(),
      axis.text.x.top=element_text(hjust=0),
      strip.text.y.right = element_text(angle = 0),
      axis.ticks.x=element_line(),
      legend.position = "none",
      legend.box = "horizontal"
    )+
    NULL

  ggsave(
      filename=glue("coverage_{.cohort}.png"),
      path=output_dir_qmd,
      plot=plot_coverage_cumuln,
      unit="cm",
      width=15,
      height=10
    )

  return(plot_coverage_cumuln)

}
```

#### Figure S1a: Cumulative numbers included in the matched sample for the CV cohort
```{r}
#| label: figure-coverage-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_coverage("cv")
```

#### Figure S1b: Cumulative numbers included in the matched sample for the 75+ cohort

```{r}
#| label: figure-coverage-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_coverage("age75plus")
```
&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

```{r}
#| label: table-table1-function

show_table1 <- function(.cohort){
  match_table1_wide_wide <-
    rst[[.cohort]]$match_table1_wide %>%
    filter(!(variable %in% c("N","immunossuppressed"))) %>%
    mutate(
      smd = if_else((variable %in% matching_variables$B$exact) & (matchset == "postmatch"), "-", smd)
    ) %>%
    pivot_wider(
      id_cols = c(var_label, variable, variable_levels, level),
      names_from=matchset,
      values_from = c(stat_0, stat_1, smd)
    ) %>%
    mutate(
      order_var = cumsum(lag(variable, 1, first(variable))!=variable),
      var_label = if_else(lag(as.character(var_label), n=1, default="") == as.character(var_label), "", as.character(var_label)),
    )

  match_table1_wide_wide %>%
    gt() %>%
    fmt_markdown(columns = "var_label") %>%
    tab_style(
      style = list(
        cell_fill(color = "gray96")
      ),
      locations = cells_body(
        rows = order_var%%2 == 0
      )
    ) %>%
    tab_style(
      style = list(
        cell_text(weight = "bold")
      ),
      locations = cells_column_labels(everything())
    ) %>%
    cols_label(
      var_label = "Variable",
      variable_levels = "",
      `stat_0_prematch` = glue("Pfizer BA.4-5 (N={filter(match_table1, cohort==.cohort, matchset=='prematch', variable=='N', by==0) %>% pull(n)})"),
      `stat_1_prematch` = glue("Sanofi (N={filter(match_table1, cohort==.cohort, matchset=='prematch', variable=='N', by==1) %>% pull(n)})"),
      `smd_prematch` = "SMD",
      `stat_0_postmatch` = glue("Pfizer BA.4-5 (N={filter(match_table1, cohort==.cohort, matchset=='postmatch', variable=='N', by==0) %>% pull(n)})"),
      `stat_1_postmatch` = glue("Sanofi (N={filter(match_table1, cohort==.cohort, matchset=='postmatch', variable=='N', by==1) %>% pull(n)})"),
      `smd_postmatch` = "SMD"
    ) %>%
    tab_spanner(
      label = "Before matching",
      columns = ends_with("_prematch")
    ) %>%
    tab_spanner(
      label = "After matching",
      columns = ends_with("_postmatch")
    ) %>%
    cols_align(
      align = c("right"),
      columns =  ends_with(c("_prematch", "_postmatch"))
    ) %>%
    cols_hide(c("variable", "level", "order_var")) %>%
    tab_options(
      table.font.size = 8
    )

}
```

#### Table S1a: Baseline characteristics in the CV cohort between Pfizer BA.4-5 and Sanofi groups, before and after matching

```{r}
#| label: table-table1-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_table1("cv")
```
For variables that were matched exactly the SMD is zero by design, and so are not shown.

#### Table S1b: Baseline characteristics in the 75+ cohort between Pfizer BA.4-5 and Sanofi groups, before and after matching

```{r}
#| label: table-table1-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_table1("age75plus")
```
For variables that were matched exactly the SMD is zero by design, and so are not shown.

```{r}
#| label: figure-smd-function

show_smd <- function(.cohort) {
  plot_smd <-
    rst[[.cohort]]$data_smd %>%
    filter(
      matchset %in% c("prematch", "postmatch")
    ) %>%
    filter(
      #!(variable %in% matching_variables$B$exact)
    ) %>%
    mutate(
      matchset=fct_recoderelevel(matchset, c("Pre-match"="prematch", "Post-match"="postmatch")),
    ) %>%
    droplevels() %>%
    group_by(matchset) %>%
    arrange(matchset, variable, level) %>%
    mutate(
      variable_levels = replace_na(as.character(variable_levels), ""),
    ) %>%
    mutate(
      level=fct_rev(fct_inorder(as.character(level))),
      cardn = row_number()
    ) %>%
    ggplot()+
    geom_point(aes(x=smd, y=level, group=matchset, colour=matchset), position=position_dodge(width=0.3))+
    geom_linerange(aes(xmax=smd, xmin=0, y=level, group=matchset, colour=matchset), position=position_dodge(width=0.3))+
    geom_rect(aes(alpha = variable_card, ymin = rev(cardn)-0.5, ymax =rev(cardn+0.5), group=matchset), xmin = -Inf, xmax = Inf, fill='grey', colour="transparent") +
    scale_alpha_continuous(range=c(0,0.1), guide='none')+
    scale_colour_brewer(type="qual", palette="Set1")+
    labs(
      x="Standardised mean difference\n <- more in Pfizer BA.4-5  |  more in Sanofi -> ",
      y=NULL,
      alpha=NULL,
      colour= NULL
    )+
    theme_minimal(base_size=10) +
    theme(
      legend.position = "bottom",
      strip.placement = "outside",
      strip.background = element_rect(fill="transparent", colour="transparent"),
      strip.text.y.left = element_text(angle = 0, hjust=1),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0, "lines")
    )

  plot_smd
}

```

#### Figure S2a: Standardised Mean Differences for CV cohort between Pfizer BA.4-5 and Sanofi groups, before and after matching
```{r}
#| label: figure-smd-cv
#| out-width: '90%'
#| out-height: '100%'
#| output: asis
#| out-extra: ''

show_smd("cv")
```

#### Figure S2b: Standardised Mean Differences for 75+ cohort between Pfizer BA.4-5 and Sanofi groups, before and after matching
```{r}
#| label: figure-smd-age75plus
#| out-width: '90%'
#| out-height: '100%'
#| output: asis
#| out-extra: ''

show_smd("age75plus")
```

\newpage

```{r}
#| label: table-contrasts-function

show_contrasts <- function(.cohort){

  rst[[.cohort]]$contrasts_table %>%
    filter(
      outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    group_by(outcome_descr, subgroup_descr) %>%
    mutate(
      rd.p = if_else(row_number()==1 & (subgroup!="all"), rd.p, NA_character_),
      rr.ln.p = if_else(row_number()==1 & (subgroup!="all"), rr.ln.p, NA_character_),
      irr.ln.p = if_else(row_number()==1 & (subgroup!="all"), irr.ln.p, NA_character_)
    ) %>%
    group_by(outcome_descr) %>%
    select(
      outcome_descr,
      subgroup_descr,
      subgroup_level_descr,

      rdCI,
      rd.p,
      rrCI,
      rr.ln.p,
      irrCI,
      irr.ln.p,
    ) %>%
    mutate(
      # add indicator for row shading
      shade = cumsum((lag(as.character(subgroup_descr), 1, "")!=subgroup_descr)*1L),
      # remove repeat values of subgroup_level
      subgroup_descr = if_else(lag(as.character(subgroup_descr), n=1, default="")==subgroup_descr & row_number()!=1, "", as.character(subgroup_descr)),
    ) %>%
    ungroup() %>%
    gt(groupname_col="outcome_descr") %>%
    cols_label(

      outcome_descr = "Outcome",
      subgroup_descr = "Sub-group",
      subgroup_level_descr = "",

      rdCI = glue("Risk difference per {perN_format} people (95% CI)"),
      rd.p = "P-value",
      rrCI = "Risk ratio (95% CI)",
      rr.ln.p = "P-value",
      irrCI = "IRR (95% CI)",
      irr.ln.p = "P-value",
    ) %>%
    cols_align(
      align = c("right"),
      columns = ends_with(c("CI", ".p"))
    ) %>%
    sub_missing(
      columns = ends_with(c("CI", ".p")),
      missing_text = ""
    ) %>%
    tab_options(
      table.font.size = 8
    ) %>%
    tab_style(
      style = list(
        cell_fill(color = "gray96")
      ),
      locations = cells_body(
        rows = shade%%2 != 0
      )
    ) %>%
    cols_hide(c("shade"))

}

```

#### Table S2a: Risk differences, risk ratios, and incidence rate ratios in the CV cohort, with p-values for subgroup heterogeneity

```{r}
#| label: table-contrasts-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_contrasts("cv")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Table S2b: Risk differences, risk ratios, and incidence rate ratios in the age75plus cohort, with p-values for subgroup heterogeneity

```{r}
#| label: table-contrasts-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_contrasts("age75plus")
```

\newpage

```{r}
#| label: table-fup-function

show_fup <- function(.cohort){
  rst[[.cohort]]$followup_table %>%
    filter(
      outcome %in% outcomes,
      subgroup %in% "all",
    ) %>%
    pivot_longer(
      cols=ends_with("weeks"),
      names_to = "stat",
      values_to = "value"
    ) %>%
    pivot_wider(
      id_cols = c(outcome_descr, subgroup_descr, subgroup_level_descr, treatment_descr),
      names_from = "stat",
      values_from = "value"
    ) %>%
    group_by(outcome_descr) %>%
    select(
      outcome_descr,
      #subgroup_descr,
      #subgroup_level_descr,
      treatment_descr,
      ends_with("weeks")
    ) %>%
    gt(groupname_col="outcome_descr") %>%
    cols_label(

      outcome_descr = "Outcome",
      #subgroup_descr = "Sub-group",
      #subgroup_level_descr = "",
      treatment_descr = "Vaccine",
      total_weeks = "Total",
      mean_weeks = "Mean",
      median_weeks = "Median",
      Q1_weeks = "Q1",
      Q3_weeks = "Q3",
    ) %>%
    tab_spanner(
      label = "Weeks of follow-up",
      columns = ends_with(c("weeks"))
    ) %>%
    cols_align(
      align = c("right"),
      columns = ends_with(c("weeks"))
    ) %>%
    tab_options(
      table.font.size = 8
    )
}
cat("  \n\n")
```

#### Table S3a: Summary of follow-up time for CV cohort

```{r}
#| label: table-fup-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_fup("cv")
```

#### Table S3b: Summary of follow-up time for 75+ cohort

```{r}
#| label: table-fup-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_fup("age75plus")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

```{r}
#| label: table-testrates-function

show_testingrates <- function(.cohort){

  rst[[.cohort]]$eventcounts %>%
    filter(
      #outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    transmute(
      subgroup_descr,
      subgroup_level_descr,
      treatment,
      persontime = label_number(1, 1/7)(persontime),
      test_rate = label_number(0.001, 7)(test_rate), 
    ) %>%
    pivot_wider(
      id_cols = c(subgroup_descr, subgroup_level_descr),
      names_from = treatment,
      values_from = c(persontime, test_rate)
    ) %>%
    mutate(
      # add indicator for row shading
      shade = cumsum((lag(subgroup_descr, 1, "")!=subgroup_descr)*1L),
    ) %>%
    group_by(subgroup_descr) %>%
    mutate(
      # remove repeat values of subgroup_level
      subgroup_descr = if_else(lag(subgroup_descr,n=1, default="")==subgroup_descr, "", as.character(subgroup_descr)),
      subgroup_level_descr = coalesce(subgroup_level_descr, ""),
    ) %>%
    ungroup() %>%
    gt() %>%
    cols_label(
      subgroup_descr = "Sub-group",
      subgroup_level_descr = "",
      persontime_0 = "Person-weeks",
      persontime_1 = "Person-weeks",
      test_rate_0 = "All tests",
      test_rate_1 = "All tests"
    ) %>%
    tab_spanner(
      label = "Pfizer BA.4-5",
      columns = c(persontime_0, test_rate_0)
    ) %>%
    tab_spanner(
      label = "Sanofi",
      columns = c(persontime_1, test_rate_1)
    ) %>%
    cols_align(
      align = c("right"),
      columns = c(persontime_0, persontime_1, test_rate_0, test_rate_1)
    ) %>%
    tab_style(
      style = list(
        cell_fill(color = "gray96")
      ),
      locations = cells_body(
        rows = shade%%2 != 0
      )
    ) %>%
    cols_hide(c("shade")) %>%
    tab_options(
      table.font.size = 8
    )
}

```

#### Table S4a: Testing rates after boosting for the CV cohort

```{r}
#| label: table-testingrates-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: true
#| out-extra: ''

show_testingrates("cv")
```

#### Table S4b: Testing rates after boosting for the 75+ cohort

```{r}
#| label: table-testingrates-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_testingrates("age75plus")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

### Cumulative incidences by subgroup

```{r}
#| label: ci-function

ci_theme <- function(...) {
  theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      #panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
      ...
    )
}

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_ci_subgroup <- function(.cohort, .subgroup){

  if(.subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_ci <- 
    rst[[.cohort]]$estimates %>%
    filter(
      subgroup==.subgroup,
      outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    mutate(
      colgroup = interaction(treatment_descr, subgroup_level_descr, sep=", "),
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
      subgroup_level_descr = fct_inorder(subgroup_level_descr)
    ) %>%
    group_by(treatment_descr, subgroup_level_descr, outcome_descr) %>%
    mutate(
      lagrisk=lag(risk,1,0),
    ) %>%
    #group_by(colgroup, outcome_descr) %>%
    ggplot(aes(group=colgroup, colour=treatment_descr, fill=treatment_descr, alpha=treatment_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    #geom_step(aes(x=lagtime, y=risk*perN, linetype=treatment_descr))+
    geom_segment(aes(x=lagtime, xend=time, y=risk*perN, yend=risk*perN, linetype=treatment_descr))+
    geom_segment(aes(x=lagtime, xend=lagtime, y=lagrisk*perN, yend=risk*perN, linetype=treatment_descr))+
    geom_rect(aes(xmin=lagtime, xmax=time, ymin=risk.ll*perN, ymax=risk.ul*perN), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set2", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set2", guide="none", na.value="grey") +
    #scale_color_manual(values=colourvalues, na.value="grey") +
    #scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,28), expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_continuous(
      expand = expansion(mult=c(0,0.01)), 
      #label=label_number_risk
    )+
    scale_alpha_discrete(range=c(0.3,0.9))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Weeks",
      y=NULL,
      #title=glue("Cumulative incidence per {perN_format}"),
      colour=NULL,
      alpha=NULL,
      linetype=NULL
    )+
    guides(linetype="none", fill="none", alpha="none")+
    theme_minimal(base_size = 9)+
      theme(
        # legend.position = c(0.05,0.1),
        # legend.justification = c(0,0),
        legend.position = "bottom",
        #legend.direction = "vertical",
        axis.text.x.top=element_text(hjust=0),

        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),
        strip.placement="outside",
        strip.text.y.left = element_text(angle=0),

        panel.border = element_blank(),
        panel.spacing = unit(0.8, "lines"),

      )+
      NULL

  n_subgroups <-
    rst[[.cohort]]$estimates %>%
    filter(
      subgroup==.subgroup
    ) %>% 
    `[[`("subgroup_level_descr") %>%
    n_distinct()

  # plot_combined <- patchwork::wrap_plots(
  #   plot_ci,
  #   #plot_rd,
  #   #plot_rr,
  #   #plot_irr,
  #   ncol=2,
  #   widths=c(n_subgroups,1)
  # )

  ggsave(
    filename=glue("ci_{.cohort}_{.subgroup}.png"),
    path=output_dir_qmd,
    plot=plot_ci,
  )

  plot_ci

}

```

#### Figure S3a.1: Cumulative incidence estimates per `r perN_format` people for the CV cohort

```{r}
#| label: figure-ci-cv-all
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("cv", "all")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3a.2: Cumulative incidence estimates per `r perN_format` people by age group for the CV cohort

```{r}
#| label: figure-ci-cv-ageband
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("cv", "ageband")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3a.3: Cumulative incidence estimates per `r perN_format` people by prior vaccine count for the CV cohort

```{r}
#| label: figure-ci-cv-vax_previous_group
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("cv", "vax_previous_group")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3b.1: Cumulative incidence estimates per `r perN_format` people for the 75+ cohort

```{r}
#| label:  figure-ci-age75plus-all
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("age75plus", "all")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3b.2: Cumulative incidence estimates per `r perN_format` people by age group for the 75+ cohort

```{r}
#| label: figure-ci-age75plus-ageband
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("age75plus", "ageband")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3b.2: Cumulative incidence estimates per `r perN_format` people by prior vaccine count for the 75+ cohort

```{r}
#| label: figure-ci-age75plus-vax_previous_group
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("age75plus", "vax_previous_group")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3b.4: Cumulative incidence estimates per `r perN_format` people by clinical vulnerability for the 75+ cohort

```{r}
#| label: figure-ci-age75plus-cv
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("age75plus", "cv")
```

<!--       -->

<!--       -->

<!--       -->

```{r}
#| label: plot-contrasts-function

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_contrast_subgroup <- function(.cohort, .subgroup, coltype){

  scale_color <- 
    if(coltype=="qual"){
      list(
        scale_color_brewer(type="qual", palette="Set1", na.value="grey"),
        scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")
      )
    } else if(coltype=="seq"){
      list(
        scale_color_viridis_d(na.value="grey"),
        scale_fill_viridis_d(guide="none", na.value="grey")
      )
    }

  if(.subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_data <- 
    rst[[.cohort]]$contrasts_daily %>%
    filter(
      subgroup==.subgroup,
      outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    mutate(
      subgroup_level_descr = fct_inorder(subgroup_level_descr)
    ) %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    mutate(
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
      rr = if_else(fup_start<7, NA_real_, rr),
      rr.ll = if_else(fup_start<7, NA_real_, rr.ll),
      rr.ul = if_else(fup_start<7, NA_real_, rr.ul),
      lagrd=lag(rd,1,0),
      lagrr=lag(rr,1,1)
    )

  plot_rd <- 
    plot_data %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    geom_segment(aes(x=fup_start, xend=fup_end, y=rd*perN, yend=rd*perN))+
    geom_segment(aes(x=fup_start, xend=fup_start, y=lagrd*perN, yend=rd*perN))+
    geom_rect(aes(xmin=fup_start, xmax=fup_end, ymin=rd.ll*perN, ymax=rd.ul*perN), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    scale_color +
    scale_x_continuous(breaks = seq(0,600,28), minor_breaks = c(0,1,2,4)*7, expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_continuous(
      expand = expansion(mult=c(0,0.01)), 
      #=label_number(0.1, 10000)
    )+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Weeks",
      y=NULL,
      title="Cumulative\nrisk difference",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      # legend.position = c(0.05,0.1),
      # legend.justification = c(0,0),
      legend.position = "bottom",
      #legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement="outside",
      strip.text.y.left = element_text(angle=0),

      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
    )+
    NULL

  plot_rr <- 
    plot_data %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    geom_segment(aes(x=fup_start, xend=fup_end, y=rr, yend=rr))+
    geom_segment(aes(x=fup_start, xend=fup_start, y=lagrr, yend=rr))+
    geom_rect(aes(xmin=fup_start, xmax=fup_end, ymin=rr.ll, ymax=rr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    scale_color +
    scale_x_continuous(breaks = seq(0,600,28), minor_breaks = c(0,1,2,4)*7, expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Weeks",
      y=NULL,
      title="Cumulative\nrisk ratio",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      #panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines")
    )+
    NULL

  plot_irr <- 
    rst[[.cohort]]$contrasts_cuts  %>%
    filter(
      subgroup==.subgroup,
      outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    group_by(outcome_descr, subgroup_level_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    # geom_segment(aes(x=fup_start, xend=fup_end, y=coxhr, yend=coxhr))+
    # geom_segment(aes(x=fup_start, xend=fup_start, y=lag(coxhr), yend=coxhr))+
    # geom_rect(aes(xmin=fup_start, xmax=fup_end, ymin=coxhr.ll, ymax=coxhr.ul), alpha=0.1, colour="transparent")+
    geom_point(aes(x = (fup_start+fup_end)/2, y=irr), alpha=0.5, position=position_dodge(width=5))+
    geom_linerange(aes(x = (fup_start+fup_end)/2, ymin=irr.ll, ymax=irr.ul), alpha=0.9, position=position_dodge(width=5))+
    facet_grid(rows=vars(outcome_descr), scales="free_y")+
    scale_color +
    scale_x_continuous(breaks = seq(0,600,28), minor_breaks = c(0,1,2,4)*7, expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)), 
                  oob=~oob_squish(.,range=log(c(1/10,10), 10)),
                  sec.axis = dup_axis(breaks=NULL, labels=NULL, name = "higher risk\n<- Pfizer BA.4-5 | Sanofi ->"))+
    coord_cartesian(xlim=c(0, 16*7))+
    colour_guide+
    labs(
      x="Weeks",
      y=NULL,
      title="Period-specific\nhazard ratio",
      colour=NULL
    )+
    theme_minimal(base_size = 9)+
    ci_theme()+
    NULL

  plot_combined <- patchwork::wrap_plots(
    plot_rd, 
    plot_rr, 
    plot_irr,
    ncol=3
  )

  ggsave(
    filename=glue("contrasts_{.cohort}_{.subgroup}.png"),
    path=output_dir_qmd,
    plot=plot_combined,
  )

  plot_combined
}

```

### Comparisons by subgroup

#### Figure S4a.1: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios for the CV cohort

```{r}
#| label: figure-contrasts-cv-all
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("cv", "all", "qual")
```

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

#### Figure S4a.2: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by age group for the CV cohort

```{r}
#| label: figure-contrasts-cv-ageband 
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("cv", "ageband", "qual")
```

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

#### Figure S4a.3: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by prior vaccine count for the CV cohort

```{r}
#| label: figure-contrasts-cv-vax_previous_group 
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("cv", "vax_previous_group", "qual")
```

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

#### Figure S4b.1: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios for the 75+ cohort

```{r}
#| label: figure-contrasts-age7plus-all
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("age75plus", "all","qual")
```

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

#### Figure S4b.2: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by age group for the 75+ cohort

```{r}
#| label: figure-contrasts-age75plus-ageband 
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("age75plus", "ageband","qual")
```

#### Figure S4b.3: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by prior vaccine count for the 75+ cohort

```{r}
#| label: figure-contrasts-age75plus-vax_previous_group
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("age75plus", "vax_previous_group","qual")
```

#### Figure S4b.4: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by clinical vulnerability for the 75+ cohort

```{r}
#| label: figure-contrasts-age75plus-cv
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("age75plus", "cv","qual")
```

### 2024-01-31 write-up/manuscript/import-process.R

@@ -1,7 +1,4 @@
 
-library("fs")
-
-
 
 ## ollection of formatting functions that may be used ----
 
@@ -39,6 +36,8 @@ label_number_rr <- function(x, scale=1){
 
 
 
+flowchart_totals_allcohorts <- read_csv(path(output_dir_os, "flowchart_totals_allcohorts_rounded.csv"))
+
 # store results for each cohort in a list,
 # accessed for example as follows:  rst$cv$contrasts
 
@@ -58,22 +57,31 @@ for(cohort in c("age75plus", "cv")){
 
     ## study flowchart ----
 
-    total_flowchart <-  read_csv(path(cohort_dir, "total_flowchart.csv"))
-    flowchart <- read_csv(path(cohort_dir, "B", "match_flowchart.csv"))
+    flowchart_totals <-
+      read_csv(path(cohort_dir, "flowchart_totals_rounded.csv"))
 
+    flowchart <-
+      read_csv(path(cohort_dir, "flowchart_rounded.csv")) %>%
+      filter(matchset =="B")
 
     ## matching coverage over time ----
 
-    match_coverage <- read_csv(path(cohort_dir, "B", "match_coverage.csv")) %>%
+    match_coverage <-
+      read_csv(path(cohort_dir, "coverage_rounded.csv")) %>%
+      filter(matchset =="B") %>%
       mutate(
         treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
       )
 
-
     ## follow-up summary data ----
 
-    followup <- read_csv(path(cohort_dir, "B", "followup_rounded.csv"))
-    followup_treatment <- read_csv(path(cohort_dir, "B", "followup_treatment_rounded.csv"))
+    followup <-
+      read_csv(path(cohort_dir, "followup_rounded.csv"))  %>%
+      filter(matchset =="B")
+
+    followup_treatment <-
+      read_csv(path(cohort_dir, "followup_treatment_rounded.csv")) %>%
+      filter(matchset =="B")
 
     followup_table <-
       bind_rows(
@@ -81,15 +89,12 @@ for(cohort in c("age75plus", "cv")){
         followup %>% mutate(treatment=2)
       ) %>%
       mutate(
[…]

### 2024-01-31 write-up/manuscript/setup.R

@@ -6,10 +6,11 @@ library('lubridate')
 library('gt')
 library('patchwork')
 library('scales')
+library("fs")
 
 # where are the outputs (ie the inputs for this manuscript!) saved?
-output_dir_os <- here("output", "release-objects")
-#output_dir_os <- here("released-output", "release-objects-v2")
+#output_dir_os <- here("output", "release-objects")
+output_dir_os <- here("released-output", "release1")
 
 
 ## import functions and design elements
@@ -19,8 +20,8 @@ source(here("analysis", "design", "design.R"))
 study_dates_format <- map(study_dates, ~format(as.Date(.), "%e %B %Y"))
 
 # where should we put the objects created within this rmd script?
-output_dir_qmd <- here("write-up", "manuscript", "figures")
-fs::dir_create(output_dir_qmd)
+output_dir_qmd <- here("write-up", "manuscript", "figures_manual")
+dir_create(output_dir_qmd)
 
 # only applicable if `self-contained: yes` option is enables in yaml header
 # knitr::opts_chunk$set(
@@ -32,32 +33,31 @@ fs::dir_create(output_dir_qmd)
 
 ## define outcome sets
 
-outcomes_primary <- c(
-  "covidemergency",
+outcomes_effectiveness <- c(
+  #"covidemergency",
+  "covidcritcare",
   "covidadmitted",
-  #"covidadmittedproxy1",
-  #"covidcritcare",
   "coviddeath",
   "noncoviddeath",
   NULL
 ) %>%
   set_names(., .)
 
-# outcomes_negative <- c(
-#   #"noncoviddeath",
-#   #"fracture",
-#   NULL
-# ) %>%
-#   set_names(., .)
+ outcomes_safety <- c(
+  "pericarditis",
+  "myocarditis",
+  NULL
+) %>%
+  set_names(., .)
 
-outcomes <- outcomes_primary
+
+outcomes <- c(outcomes_effectiveness, outcomes_safety)
 
 subgroups <-
   c(
     "all",
-    "vax_prior_history",
+    "vax_previous_group",
     "ageband",
-    "prior_covid_infection",
     "cv",
     NULL
   ) %>%

### 2024-01-31 analysis/design/design.R
typo

@@ -86,7 +86,7 @@ vax_product_lookup = c(
 # )
 
 
-treatement_lookup <-
+treatment_lookup <-
   tribble(
     ~treatment, ~treatment_descr,
     "pfizerBA45", "pfizer/BA.4-5",

### 2024-02-09 write-up/manuscript/import-process.R
Minor comment correction

@@ -1,6 +1,6 @@
 
 
-## ollection of formatting functions that may be used ----
+## Collection of formatting functions that may be used ----
 
 perN <- 1000
 perN_format <- label_number(1, 1, big.mark=",")(perN)

### 2024-02-09 write-up/manuscript/manuscript.qmd
manuscript edits

@@
 
-TBD: Colm Andrews^1^, Edward PK Parker^2^, Em Prestige^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^5,6^\*, Jonathan AC Sterne^3,7,8^\*, William J Hulme^1^, and the OpenSAFELY collaborative.
+TBD: Colm D Andrews^1^, Edward PK Parker^2^, Em Prestige^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^5,6^\*, Jonathan AC Sterne^3,7,8^\*, William J Hulme^1^, and the OpenSAFELY collaborative.
 
@@
 
-*Results* `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4") %>% pull(n) %>% sum())` people were 1-1 matched in the CV cohort, and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4") %>% pull(n) %>% sum())` in the 75+ cohort, contributing a total `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up respectively.
+*Results* `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5") %>% pull(n) %>% sum())` people were 1-1 matched in the CV cohort, and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5") %>% pull(n) %>% sum())` in the 75+ cohort, contributing a total `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up respectively.
 The incidence of COVID-19 hospital admission was higher for Sanofi than for Pfizer BA.4-5.
-In the CV cohort, 16-week risks per `r perN_format` people were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)`.
-In the 75+ cohort, these were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`.
+In the CV cohort, 16-week risks per `r perN_format` people were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)`.
+In the 75+ cohort, these were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi, with an IRR of `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI)`.
 These finding were similar across all pre-specified subgroups.
@@
 
-The spring 2023 COVID-19 booster vaccination programme in England offered COVID-19 vaccination to those aged 75 years or over, resident in a care home, and those who are immunosuppressed, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI).[@NHSEnglandNHSc; @JCVIStatementCOVID19; @JCVIStatementSpring] Both the Pfizer BA.4/BA.5 vaccine and the Sanofi vaccine were used during this programme.
+The spring 2023 COVID-19 booster vaccination programme in England offered COVID-19 vaccination to those aged 75 years or over, resident in a care home, and those who are immunosuppressed, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI).[@NHSEnglandNHSc; @JCVIStatementCOVID19; @JCVIStatementSpring] Both the Pfizer BA.4-5 vaccine and the Sanofi vaccine were used during this programme.
 This allows direct comparisons of their safety and effectiveness against severe COVID-19 disease, something not investigated in randomised trials.
@@
 All data were linked, stored and analysed securely using the OpenSAFELY platform: <https://opensafely.org/>, as part of the NHS England OpenSAFELY COVID-19 service.
+Data include pseudonymised data such as coded diagnoses, medications and physiological parameters.
+No free text data are included.
+All code is shared openly for review and re-use under MIT open license <https://github.com/opensafely/comparative-booster-spring2023>.
+Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.
 Primary care records managed by the GP software provider, TPP were linked, using NHS numbers, to emergency department attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
@@
 COVID-19 critical care admissions were any COVID-19 hospital spell with at least one day spent in ciritcal care (including any such spells occurring subsequent to the first spell after vaccination).
-COVID-19 death was defined as death COVID-19 ICD-10 codes (as before) mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
+COVID-19 death was defined as COVID-19 ICD-10 codes (as before) mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
 We also report on non-COVID-19 deaths.
@@
 
-Each person was followed from receipt of vaccination (time zero) until the outcome of interest, with censoring at death, practice de-registration, `r study_dates_format$followupend_date`.
+Each person was followed from receipt of vaccination (time zero) until the outcome of interest, with censoring at death, practice de-registration, or `r study_dates_format$followupend_date`.
 Outcomes were observed for up to 16 weeks for effectiveness outcomes, and up to 4 weeks for safety outcomes.
@@
 ### Software, code, and reproducibility
-
 Data management and analyses were conducted using OpenSAFELY tools, Python version 3.8.10 and R version 4.0.2.
-All code is shared openly for review and re-use under MIT open license at <https://github.com/opensafely/comparative-booster-spring2023>.
+Code for data management and analysis, as well as codelists, are archived online <https://github.com/opensafely/comparative-booster-spring2023>.
 Codelists are available at <https://www.opencodelists.org/>.
@@
 Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.
-
 \newpage
@@
 
-We identified `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).
+We identified `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).
 
-`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c3", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).
+`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 2).
 
@@
 
-In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1a).
-In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1b).
+In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1a).
+In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 2, Figure S1b).
 
@@
 
-The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 2).
+The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 3).
 The corresponding 16-week risks per `r perN_format` people for COVID-19 critical care admission were `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(rdCI)`).
@@
 
-IRRs (where less than one favours Sanofi) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
+IRRs (where less than one favours Sanofi) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 4).
 
@@
 
-The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 2).
+The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 3).
 The corresponding 16-week risks per `r perN_format` people for COVID-19 critical care admission were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(rdCI)`).
@@
 
-IRRs in the 16-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
+IRRs in the 16-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 4).
 
-During the first week after vaccination event rates were similar between vaccine groups: the estimated HR was `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidadmitted", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidcritcare", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="coviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for COVID-19 death, `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="noncoviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for non-COVID-19 death.
+During the first week after vaccination event rates were similar between vaccine groups: the estimated HR was `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="covidadmitted", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI_95)` for COVID-19 hospital admissions, and `r rst$age75plus$contrasts_cuts_table %>% filter(outcome=="noncoviddeath", subgroup=="all", fup_start==0, fup_end==7) %>% pull(irrCI)` for non-COVID-19 death.
+COVID-19 critical care admission and COVID-19 deaths were rare so Hazard Ratios could not be reliably estimated.
 
@@
 This corresponded to a cumulative incidence per `r perN_format` people of `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
-For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events, with a cumulative incidence of `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events, with a cumulative incidence of `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(risk_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` (0 to 0) for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
 
 In the 75+ cohort, `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients respectively, with a cumulative incidence per `r perN_format` people of `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
-For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events, with a cumulative incidence of `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events, with a cumulative incidence of `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(risk_0)` (95%CI 0-0) for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
 
@@
 This observational study of COVID-19 vaccine recipients during the spring 2023 booster programme in England is, to our knowledge, the first to directly compare the Pfizer BA.4-5 and Sanofi vaccines for safety and effectiveness against severe COVID-19 outcomes.
-We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Sanofi vaccine compared with Pfizer BA.4-5 during the first 16 weeks after vaccination.
+We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Pfizer BA.4-5 vaccine compared with Sanofi during the first 16 weeks after vaccination.
 For the more severe COVID-19 outcomes of critical care admission and death, there were very few events and insufficient power to identify differences between vaccines.
@@
 
-...
+The MHRA authorisation for Sanofi was based on clinical data from about 800 individuals previously immunised with an mRNA or viral-vectored vaccine [@Healthcare_products_Regulatory_Agency2022-us] and authorisation for Pfizer  BA.4-5 was based on the C4591001a trial, a placebo-controlled efficacy trial, finding efficacy of 86 to 100% against laboratory-confirmed COVID-19.
+[@Thomas2021-zm]
+Real world evidence of the effectiveness of Sanofi and Pfizer BA.4-5 against hospitalisation amongst adults aged 75 years and older in England from a test-negative case control study involving 14,174 eligible tests found similar effectiveness of both Pfizer  BA.4-5 and Sanofi of 30-50%.[@Kirsebom2023-fq]
+We are aware of no other planned or published randomised trials making direct efficacy endpoint comparisons of boosting with Pfizer BA.4-5 and Sanofi.
+The results of our study are relevant to vaccine procurement decisions for future booster programmes.
+However, both vaccines are safe and strongly effective against COVID-19 hospitalisation.
+Findings from this study should not discourage people from receiving Sanofi booster vaccination if offered.
 
@@
 
-During the spring 2023 COVID-19 booster programme in England, COVID-19-related hospital admissions were slightly more frequent following receipt of the Pfizer BA.4-5 vaccine than the Sanofi vaccine after matching on a number of important potential confounders.
+During the spring 2023 COVID-19 booster programme in England, COVID-19-related hospital admissions were slightly more frequent following receipt of the Sanofi vaccine than the Pfizer BA.4-5 vaccine after matching on a number of important potential confounders.
 These finding were broadly consistent across subgroups, though estimated with less precision.
@@
 Conceptualization: WJH, EPKP, VW, TP, BG, MAH, and JACS.
-Data curation: WJH, HJC, AW, AM, BM, CA, and SCJB.
+Data curation: WJH, HJC, AW, AM, BM, CDA, and SCJB.
 Formal analysis: WJH, EPKP, and VW.
@@
 Methodology: WJH, EPKP, VW, TP, MAH, and JACS.
-Project administration: WJH, AM, JM, BG, CA, and JACS.
+Project administration: WJH, AM, JM, BG, CDA, and JACS.
 Resources: WJH, HJC, AW, AM, JM, BM, SCJB, and BG.
@@
 Supervision: MAH and JACS.
-Validation: WJH, EPKP, CA, and VW.
+Validation: WJH, EPKP, CDA, and VW.
 Visualization: WJH, and EPKP.
 Writing - original draft: WJH.
-Writing - review & editing: WJH, EPKP, VW, TP, HJC, AW, CA, AM, JM, BM, SCJB, BG, MAH, and JACS.
+Writing - review & editing: WJH, EPKP, VW, TP, HJC, AW, CDA, AM, JM, BM, SCJB, BG, MAH, and JACS.
 
@@
 
-This study was approved by the Health Research Authority (REC reference 20/LO/0651).
+This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by London School of Hygeine and Tropical Medicine Ethics Board (reference 21863).
 
@@
 This work was jointly funded by UKRI \[COV0076;MR/V015737/1\], the Longitudinal Health and Wellbeing strand of the National Core Studies programme (MC_PC_20030; MC_PC_20059; COV-LT-0009), NIHR and Asthma UK-BLF.
-The OpenSAFELY data science platform is funded by the Wellcome Trust (222097/Z/20/Z).
+The OpenSAFELY data science platform is funded by the Wellcome Trust (222097/Z/20/Z) and MRC (MR/V015737/1, MC_PC_20059, MR/W016729/1).
+In addition, development of OpenSAFELY has been funded by the the NIHR funded CONVALESCENCE programme (COV-LT-0009), NIHR (NIHR135559, COV-LT2-0073), and the Data and Connectivity National Core Study funded by UK Research and Innovation (MC_PC_20058) and Health Data Research UK (HDRUK2021.000).
 
@@
 For the purpose of Open Access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission.
+
+## Conflicts of Interest
+
+BG has received research funding from the Bennett Foundation, the Laura and John Arnold Foundation, the NHS National Institute for Health Research (NIHR), the NIHR School of Primary Care Research, NHS England, the NIHR Oxford Biomedical Research Centre, the Mohn-Westlake Foundation, NIHR Applied Research Collaboration Oxford and Thames Valley, the Wellcome Trust, the Good Thinking Foundation, Health Data Research UK, the Health Foundation, the World Health Organisation, UKRI MRC, Asthma UK, the British Lung Foundation, and the Longitudinal Health and Wellbeing strand of the National Core Studies programme; he has previously been a Non-Executive Director at NHS Digital; he also receives personal income from speaking and writing for lay audiences on the misuse of science.
+BMK is also employed by NHS England working on medicines policy and clinical lead for primary care medicines data.
+
+## Patient and Public Involvement and Engagement (PPIE)
+
+OpenSAFELY has developed a publicly available website https://www.opensafely.org/ through which they invite any patient or member of the public to make contact regarding the broader OpenSAFELY project.
 
@@
 
-#### Figure 1: Flow of participants into the study
+#### Figure 1: Flow of participants into the CV cohort
 
-```{r table-flowchart}
-#| echo: false
-#| warning: false
-#| message: false
+![figure1](figures_manual/flowchart_cv.png){width=60%}
 
-flowchart %>%
-  select(-crit) %>%
-  mutate(
-    pct_step1= 1-pct_exclude
-  ) %>%
-  pivot_wider(
-    id_cols = c(criteria, cohort_descr),
-    names_from = c(boost_type),
-    values_from = c(n, n_exclude, pct_all, pct_step, pct_step1)
-  ) %>%
-  gt(groupname_col="cohort_descr") %>%
-  cols_label(
-    criteria = "",
+#### Figure 2: Flow of participants into the 75+ cohort
 
-    n_pfizerBA45 = "N",
-    n_sanofi = "N",
-
-    n_exclude_pfizerBA45 = "N excluded",
-    n_exclude_sanofi = "N excluded",
-
-    pct_all_pfizerBA45 = "% (overall)",
-    pct_all_sanofi = "% (overall)",
-    pct_step_pfizerBA45 = "% (incremental)",
-    pct_step_sanofi = "% (incremental)",
-    pct_step1_pfizerBA45 = "% (incremental 2)",
-    pct_step1_sanofi = "% (incremental 2)",
-  ) %>%
-  tab_spanner(
-    label = "Pfizer BA.4-5",
-    columns = ends_with("pfizerBA45")
-  ) %>%
-   tab_spanner(
-    label = "Sanofi",
-    columns = ends_with("sanofi")
-  ) %>%
-  fmt_percent(
-    columns = starts_with(c("pct_")),
-    decimals = 1
-  ) %>%
-  fmt_number(
-    columns = starts_with(c("n_")),
-    decimals= 0
-  ) %>%
-  fmt_missing(
-    everything(),
-    missing_text="-"
-  ) %>%
-  tab_style(
-    style=cell_text(whitespace ="pre-wrap"),
-    #style=cell_text(style ="italic"),
-    locations = cells_body(columns = "criteria")
-  ) %>%
-  tab_options(
-    table.font.size = 8
-  )
-
-cat("  \n\n")
-```
+![figure1](figures_manual/flowchart_age75plus.png){width=40%}
 
@@
       outcome %in% outcomes,
-      outcome != "noncoviddeath"
     ) %>%

### 2024-02-12 write-up/manuscript/manuscript.qmd
amend IRR plots

@@
 
+# create csv in form of Jasper plot
+contrasts_jasper_plot <- contrasts_plot %>%
+  ungroup()%>%
+  summarise(
+            Heading=outcome_descr,
+            HR1=irr,
+            LCI1=irr.ll,
+            UCI1=irr.ul,
+            ) %>%
+  replace(is.na(.),Inf) %>%
+  add_row( Heading ="",.before = 0) %>%
+  add_row(Heading = "Clinically Vulnerable", .before = 1) %>%
+  add_row(Heading ="", .after = length(outcomes) +2) %>%
+  add_row( Heading="", .after = length(outcomes) +2) %>%
+  add_row(Heading = "Aged 75 years or over", .after = length(outcomes) +3)
+
+write_csv(contrasts_jasper_plot,fs::path(output_dir_qmd,"irr_plots.csv"),na="")
+
+# Run Jasper IRR plot script (requires https://github.com/arnhew99/Jasper?tab=readme-ov-file#jasper)
+source(here::here("write-up", "manuscript", "irr_plots_jasper_code.r"))
+
 plot_irr_lines <- contrasts_plot %>%
@@
   facet_wrap(vars(cohort_descr), scales="free_y", ncol=1, strip.position = "top")+
-  scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","), oob=~oob_squish(.,range=log(c(0.4,2), 10)), expand=expansion())+
+  scale_x_log10(label=label_number(0.1, style_negative="minus", big.mark=","),n.breaks =8, oob=~oob_squish(.,range=log(c(0.2,8), 10)), expand=expansion())+
   #scale_fill_brewer(type="qual", palette="Set2")+

### 2024-02-13 write-up/manuscript/manuscript.qmd
minor updates as discussed

@@
 #| include: false
-#| message: false
 #| warnings: false
@@
 #| include: false
-#| message: false
 #| cache: false
@@
 
-The spring 2023 COVID-19 booster vaccination programme in England offered COVID-19 vaccination to those aged 75 years or over, resident in a care home, and those who are immunosuppressed, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI).[@NHSEnglandNHSc; @JCVIStatementCOVID19; @JCVIStatementSpring] Both the Pfizer BA.4-5 vaccine and the Sanofi vaccine were used during this programme.
+The spring 2023 COVID-19 booster vaccination programme in England offered COVID-19 vaccination to those aged 75 years or over, resident in a care home, and those who are immunosuppressed, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI).[@NHSEnglandNHSc; @JCVIStatementCOVID19; @JCVIStatementSpring] Both the Pfizer BA.4-5 vaccine (authorised November 2022 [@SecondPfizerBioNTech]) and the Sanofi vaccine (authorised December 2022 [@SanofiPasteurCOVID19]) were used during this programme.
 This allows direct comparisons of their safety and effectiveness against severe COVID-19 disease, something not investigated in randomised trials.
@@
 COVID-19 hospital admission was identified using Secondary Use Service (SUS) in-patient hospital spell records with International Statistical Classification of Diseases and Related Health Problems 10th Revision (ICD-10) codes U07.1, U07.2, or U.109 as the primary or non-primary reason for admission.
-COVID-19 critical care admissions were any COVID-19 hospital spell with at least one day spent in ciritcal care (including any such spells occurring subsequent to the first spell after vaccination).
+COVID-19 critical care admissions were any COVID-19 hospital spell with at least one day spent in critical care (including any such spells occurring subsequent to the first spell after vaccination).
 COVID-19 death was defined as COVID-19 ICD-10 codes (as before) mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
@@
 ### Software, code, and reproducibility
+
 Data management and analyses were conducted using OpenSAFELY tools, Python version 3.8.10 and R version 4.0.2.
@@
 
-We identified `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1).
+We identified `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1a).
 
-`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 2).
+`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1b).
 
@@
 
-In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1, Figure S1a).
-In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 2, Figure S1b).
+In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1a, Figure S1a).
+In the 75+ cohort, `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` were matched in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`), representing `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1b, Figure S1b).
 
@@
 
-```{r break3}
+````{comment}
 
-```
+````
 
@@
 
-IRRs (where less than one favours Sanofi) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 4).
+IRRs (where less than one favours Sanofi) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
 
@@
 
-```{r break4}
+````{comment}
 
-```
+````
 
@@
 
-IRRs in the 16-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 4).
+IRRs in the 16-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
 
@@
 This corresponded to a cumulative incidence per `r perN_format` people of `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
-For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events, with a cumulative incidence of `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(risk_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` (0 to 0) for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` events for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi, (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`)
+
+<!-- with a cumulative incidence of `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(risk_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` (0 to 0) for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+-->
 
 In the 75+ cohort, `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients respectively, with a cumulative incidence per `r perN_format` people of `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
-For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events, with a cumulative incidence of `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(risk_0)` (95%CI 0-0) for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` for Pfizer BA.4/5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi, (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+
+<!-- with a cumulative incidence of `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(risk_0)` (95%CI 0-0) for Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+-->
 
@@
 
-The MHRA authorisation for Sanofi was based on clinical data from about 800 individuals previously immunised with an mRNA or viral-vectored vaccine [@Healthcare_products_Regulatory_Agency2022-us] and authorisation for Pfizer  BA.4-5 was based on the C4591001a trial, a placebo-controlled efficacy trial, finding efficacy of 86 to 100% against laboratory-confirmed COVID-19.
-[@Thomas2021-zm]
-Real world evidence of the effectiveness of Sanofi and Pfizer BA.4-5 against hospitalisation amongst adults aged 75 years and older in England from a test-negative case control study involving 14,174 eligible tests found similar effectiveness of both Pfizer  BA.4-5 and Sanofi of 30-50%.[@Kirsebom2023-fq]
-We are aware of no other planned or published randomised trials making direct efficacy endpoint comparisons of boosting with Pfizer BA.4-5 and Sanofi.
-The results of our study are relevant to vaccine procurement decisions for future booster programmes.
-However, both vaccines are safe and strongly effective against COVID-19 hospitalisation.
-Findings from this study should not discourage people from receiving Sanofi booster vaccination if offered.
+<!-- The MHRA authorisation for Sanofi was based on clinical data from about 800 individuals previously immunised with an mRNA or viral-vectored vaccine \[@SanofiPasteurCOVID19\] and authorisation for Pfizer BA.4/BA.5 was based on a placebo-controlled efficacy trial, finding efficacy of 86 to 100% against laboratory-confirmed COVID-19.
+[@zouNeutralizationBABA2023] -->
+
+Out-of-trial evidence of the effectiveness of Sanofi and Pfizer BA.4-5 vaccines against hospitalisation amongst adults aged 75 years and older in England from a test-negative case control study involving 14,174 eligible tests found similar effectiveness of both Pfizer BA.4-5 and Sanofi of 30-50%.[@kirsebomEffectivenessAdjuvantedSanofi2023] We are aware of no other planned or published randomised trials making direct endpoint comparisons of boosting with Pfizer BA.4-5 and Sanofi.
 
@@
 #| warning: false
-#| message: false
 
@@
 
-#### Figure 1: Flow of participants into the CV cohort
+#### Figure 1a: Flow of participants into the CV cohort
 
-![figure1](figures_manual/flowchart_cv.png){width=60%}
+![figure1a](figures_manual/flowchart_cv.png){width="60%"}
 
-#### Figure 2: Flow of participants into the 75+ cohort
+#### Figure 1b: Flow of participants into the 75+ cohort
 
-![figure1](figures_manual/flowchart_age75plus.png){width=40%}
+![figure1b](figures_manual/flowchart_age75plus.png){width="40%"}
 
@@
 #| label: function-ci
+#| warning: false
 #| out.width: '100%'
 #| out.height: '50%'
-#| output: 'asis'
-#| out-extra: ''
 
@@
   plot_combined <- patchwork::wrap_plots(
-    plot_ci, 
+    plot_ci#, 
     #plot_rd, 
     #plot_rr, 
-    plot_irr,
-    ncol=2,
-    widths=c(n_subgroups,1)
+    #plot_irr,
+    #ncol=2,
+    #widths=c(n_subgroups,1)
   )
@@
 
-```{r figure-ci-all}
+```{r figure-ci-cv-all}
 #| echo: false
-#| message: false
 #| warning: false
 #| out-width: '100%'
-#| out-height: '50%'
-#| output: 'asis'
-#| out-extra: ''
+#| out-height: '80%'
 
@@
 #| echo: false
-#| message: false
 #| warning: false
 #| out-width: '100%'
-#| out-height: '50%'
-#| output: 'asis'
-#| out-extra: ''
+#| out-height: '80%'
 
@@
 #| echo: false
-#| message: false
 #| warning: false
-#| fig-height: 10
-#| out-extra: ''
+#| out-height: '50%'
 
@@
 
-write_csv(contrasts_jasper_plot,fs::path(output_dir_qmd,"irr_plots.csv"),na="")
+write_csv(contrasts_jasper_plot, fs::path(output_dir_qmd,"irr_plots.csv"),na="")
 
 # Run Jasper IRR plot script (requires https://github.com/arnhew99/Jasper?tab=readme-ov-file#jasper)
-source(here::here("write-up", "manuscript", "irr_plots_jasper_code.r"))
+#source(here::here("write-up", "manuscript", "irr_plots_jasper_code.r"))
 
@@
   geom_vline(aes(xintercept=1), linetype="dotted", colour="darkgrey")+
-  geom_point(aes(x=irr), position=position_dodge(width=0.3))+
-  geom_linerange(aes(xmin=irr.ll, xmax=irr.ul), position=position_dodge(width=0.3))+
+  geom_point(aes(x=irr), position=position_dodge(width=0.1))+
+  geom_linerange(aes(xmin=irr.ll, xmax=irr.ul), position=position_dodge(width=0.1))+
   #facet_grid(rows=vars(outcome_descr), scales="free", space="free_y", switch="y")+
@@
   labs(
-    subtitle="\n\n\nIncidence Rate Ratios",
+    #subtitle="\n\n\nIncidence Rate Ratios",
     x="<- favours Sanofi  |  favours Pfizer BA.4-5 ->",
@@
     panel.border = element_blank(),
-    panel.spacing = unit(0.5, "lines"),
+    panel.spacing = unit(0.4, "lines"),
   )
@@
     panel.border = element_blank(),
-    panel.spacing = unit(0.5, "lines"),
+    panel.spacing = unit(0.4, "lines"),
   )+
@@
     plot_irr_table,
-    #plot_p_subgroup_table,
-    ncol=3,
-    widths=c(2,0.7,0.5)
+    ncol=2,
+    widths=c(2,0.9)
   )
@@
     plot=plot_combined,
-    width=20, height=20, units="cm"
+    width=20, height=15, units="cm"
   )

### 2024-02-13 write-up/manuscript/manuscript.qmd
typos; per 10000 not 1000

@@
 
-\newpage
-
 ## Discussion
@@
 During the spring 2023 COVID-19 booster programme in England, COVID-19-related hospital admissions were slightly more frequent following receipt of the Sanofi vaccine than the Pfizer BA.4-5 vaccine after matching on a number of important potential confounders.
-These finding were broadly consistent across subgroups, though estimated with less precision.
+These findings were broadly consistent across subgroups, though estimated with less precision.
 
@@
 The service adheres to the obligations of the UK General Data Protection Regulation (UK GDPR) and the Data Protection Act 2018.
-The service previously operated under notices initially issued in February 2020 by the the Secretary of State under Regulation 3(4) of the Health Service (Control of Patient Information) Regulations 2002 (COPI Regulations), which required organisations to process confidential patient information for COVID-19 purposes; this set aside the requirement for patient consent @WithdrawnWithdrawnCoronavirus2022.
+The service previously operated under notices initially issued in February 2020 by the Secretary of State under Regulation 3(4) of the Health Service (Control of Patient Information) Regulations 2002 (COPI Regulations), which required organisations to process confidential patient information for COVID-19 purposes; this set aside the requirement for patient consent @WithdrawnWithdrawnCoronavirus2022.
 As of 1 July 2023, the Secretary of State has requested that NHS England continue to operate the Service under the COVID-19 Directions 2020 @COVID19PublicHealth.
@@
 
-This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by London School of Hygeine and Tropical Medicine Ethics Board (reference 21863).
+This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by London School of Hygiene and Tropical Medicine Ethics Board (reference 21863).
 

### 2024-02-21 write-up/manuscript/manuscript.qmd
Add manuscript edits

--- 2024-02-12 write-up/manuscript/manuscript.qmd
@@
 In each cohort, vaccine recipients were matched on date of vaccination, COVID-19 vaccine history, age, and other characteristics.
-Effectiveness outcomes were COVID-19 Emergency Department (ED) attendance, COVID-19 hospital admission, and COVID-19 death up to 16 weeks after vaccination.
+Effectiveness outcomes were COVID-19 hospital admission, COVID-19 critical care admission, and COVID-19 death up to 16 weeks after vaccination.
 Safety outcomes were Pericarditis and Myocarditis up to 4 weeks after vaccination.
@@
 Conceptualization: WJH, EPKP, VW, TP, BG, MAH, and JACS.
-Data curation: WJH, HJC, AW, AM, BM, CDA, and SCJB.
-Formal analysis: WJH, EPKP, and VW.
-Funding acquisition: JM, BG, and JACS.
+Data curation: WJH, CDA, EP, CW, BS, HJC, AW, AM, BM, and SCJB.
+Formal analysis: WJH, CDA, EPKP, and VW.
+Funding acquisition: BG, and JACS.
 Methodology: WJH, EPKP, VW, TP, MAH, and JACS.
-Project administration: WJH, AM, JM, BG, CDA, and JACS.
-Resources: WJH, HJC, AW, AM, JM, BM, SCJB, and BG.
-Software: WJH, EPKP, and SCJB.
-Supervision: MAH and JACS.
-Validation: WJH, EPKP, CDA, and VW.
-Visualization: WJH, and EPKP.
+Project administration: WJH, CDA, AM, BG,, and JACS.
+Resources: WJH, CDA, HJC, AW, AM, BM, BS, SCJB, and BG.
+Software: WJH, CDA, EP,  EPKP, BS, and SCJB.
+Supervision: WJH and JACS.
+Validation: WJH, EP, EPKP, CDA, CW, and VW.
+Visualization: WJH, CDA, and EPKP.
 Writing - original draft: WJH.
-Writing - review & editing: WJH, EPKP, VW, TP, HJC, AW, CDA, AM, JM, BM, SCJB, BG, MAH, and JACS.
+Writing - review & editing: WJH, CDA, EPKP, VW, TP, HJC, AW, AM, BM, SCJB, BG, MAH, and JACS.
 
@@
 
-This work was jointly funded by UKRI \[COV0076;MR/V015737/1\], the Longitudinal Health and Wellbeing strand of the National Core Studies programme (MC_PC_20030; MC_PC_20059; COV-LT-0009), NIHR and Asthma UK-BLF.
+This work was jointly funded by UKRI \(COV0076;MR/V015737/1\), the Longitudinal Health and Wellbeing strand of the National Core Studies programme (MC_PC_20030; MC_PC_20059; COV-LT-0009), NIHR and Asthma UK-BLF.
 The OpenSAFELY data science platform is funded by the Wellcome Trust (222097/Z/20/Z) and MRC (MR/V015737/1, MC_PC_20059, MR/W016729/1).
@@
 
-#### Figure 3: Estimated incidence rate ratios at comparing the safety (4-weeks) and effectiveness (16-weeks) of Sanofi and Pfizer BA.4-5, overall and within subgroups, together with p-values for between-subgroup heterogeneity.
+#### Figure 3: Estimated incidence rate ratios at comparing the safety (4-weeks) and effectiveness (16-weeks) of Sanofi and Pfizer BA.4-5, overall and within subgroups.
 

### 2024-02-21 write-up/manuscript/irr_plots_jasper_code.r
remove dead space from irr plot

@@ -25,7 +25,7 @@ print.headings <- parseColHeadings(rawdata$ColHeadings, rd=rawdata)
 
 # set up the Jasper page
 type <- "JPG"
-SetPage(orient="PORTRAIT", perpage=1, type=type, filestem=path(output_dir_qmd, "irr_jasp_plots"), append_datetime=FALSE, suppress.date=TRUE, titlespace=0.05, footerspace=0.42)
+SetPage(orient="PORTRAIT", perpage=1, type=type, filestem=path(output_dir_qmd, "irr_jasp_plots"),titlespace = 0.005,blank.bottom.percent = 0, append_datetime=FALSE, suppress.date=TRUE,page_height = 7)
 
 # Set page margins: format is c(bottom, left, top, right)
 par(mar=c(8.09424320667184, 2, 8.09424320667184, 2))
@@ -77,7 +77,7 @@ text(x=0, y=locs$BlankLocs+ ylabels.offset, labels=left.labels[blank.rows], adj=
 
 # adding other columns
 # add the main title
-mainTitle("Incidence Rate Ratios", mainfont=1)
+# mainTitle("Incidence Rate Ratios", mainfont=1)
 
 # stop writing to file
 closeFile(type)

### 2024-02-27 write-up/manuscript/manuscript.qmd
manuscript edits

@@
 
-TBD: Colm D Andrews^1^, Edward PK Parker^2^, Em Prestige^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^5,6^\*, Jonathan AC Sterne^3,7,8^\*, William J Hulme^1^, and the OpenSAFELY collaborative.
+TBD: Colm D Andrews^1^, Em Prestige^2^, Edward PK Parker^2^, Venexia Walker^3,4^, Tom Palmer^3,4^, Helen J Curtis^1^, Alex Walker^1^, Amir Mehrkar^1^, Brian MacKenna^1^, Sebastian CJ Bacon^1^, Ben Goldacre^1^, Miguel A Hernán^5,6^\*, Jonathan AC Sterne^3,7,8^\*, William J Hulme^1^, and the OpenSAFELY collaborative.
 
@@
 All people aged 75 years or over and the clinically vulnerable were eligible to receive a booster dose.
-Direct comparisons of effectiveness against severe COVID-19 events of these two vaccines for boosting have not been made in trials or observational data.
+Direct comparisons of these two vaccines in boosting protection against severe COVID-19 events have not been made in trials or observational data.
 
 *Methods* With the approval of NHS England, we used the OpenSAFELY-TPP database to compare effectiveness of the Pfizer BA.4-5 and Sanofi vaccines during the spring 2023 booster programme, between 1 April and 30 June 2023.
-We investigated two cohorts separately: those aged 75 or over (75+); those aged 50 or over and clinically vulnerable (CV).
+We investigated two cohorts separately: those aged 75 or over (75+); and those aged 50 or over and clinically vulnerable (CV).
 In each cohort, vaccine recipients were matched on date of vaccination, COVID-19 vaccine history, age, and other characteristics.
 Effectiveness outcomes were COVID-19 hospital admission, COVID-19 critical care admission, and COVID-19 death up to 16 weeks after vaccination.
-Safety outcomes were Pericarditis and Myocarditis up to 4 weeks after vaccination.
+Safety outcomes were pericarditis and myocarditis up to 4 weeks after vaccination.
 We report the cumulative incidence of each outcome, and compare safety and effectiveness using risk differences (RD), relative risks (RR), and incidence rate ratios (IRRs).
 
-*Results* `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5") %>% pull(n) %>% sum())` people were 1-1 matched in the CV cohort, and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5") %>% pull(n) %>% sum())` in the 75+ cohort, contributing a total `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up respectively.
+*Results* `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5") %>% pull(n) %>% sum())` people were 1-1 matched in the CV cohort, and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5") %>% pull(n) %>% sum())` in the 75+ cohort, contributing a total `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(persontime_weeks)` person-weeks of follow-up, respectively.
 The incidence of COVID-19 hospital admission was higher for Sanofi than for Pfizer BA.4-5.
@@
 These finding were similar across all pre-specified subgroups.
-More severe COVID-19 related outcomes (critical care admission and death), and safety outcomes at 4 weeks, were rare in both vaccines so we could not reliably estimate superiority of either vaccine.
-
+More severe COVID-19 related outcomes (critical care admission and death), and safety outcomes at 4 weeks, were rare in both vaccines so we could not reliably compare effectiveness of the two vaccines.
 *Conclusion* This observational study comparing effectiveness of Pfizer BA.4-5 and Sanofi vaccine during the spring 2023 programme in England in the two main eligible cohorts -- people aged 75 and over and in clinically vulnerable people -- found some evidence of superior effectiveness against COVID-19 hospital admission for Pfizer BA.4-5 compared with Sanofi within 16 weeks after vaccination.
@@
 The spring 2023 COVID-19 booster vaccination programme in England offered COVID-19 vaccination to those aged 75 years or over, resident in a care home, and those who are immunosuppressed, based on guidance from the Joint Committee for Vaccination and Immunisation (JCVI).[@NHSEnglandNHSc; @JCVIStatementCOVID19; @JCVIStatementSpring] Both the Pfizer BA.4-5 vaccine (authorised November 2022 [@SecondPfizerBioNTech]) and the Sanofi vaccine (authorised December 2022 [@SanofiPasteurCOVID19]) were used during this programme.
-This allows direct comparisons of their safety and effectiveness against severe COVID-19 disease, something not investigated in randomised trials.
+This allows direct comparisons of their safety and effectiveness against severe COVID-19 disease - something not investigated in randomised trials.
 
@@
 No free text data are included.
-All code is shared openly for review and re-use under MIT open license <https://github.com/opensafely/comparative-booster-spring2023>.
+All code is shared openly for review and re-use under MIT open license (<https://github.com/opensafely/comparative-booster-spring2023>).
 Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.
@@
 We only considered those aged 50 years or over because the Sanofi vaccine was rarely offered to younger people.
-Clinical vulnerability encompasses various chronic or immunosuppressive conditions and medications as set out in the Green Book chapter 14a @COVID19GreenBook, and defined in supplementary materials.
+Clinical vulnerability encompasses various chronic or immunosuppressive conditions and medications as set out in the Green Book chapter 14a @COVID19GreenBook, and defined in the supplementary materials.
 
@@
 
-Age was calculated as at 30 June 2023, to allow for a degree of operational flexibility in who was offered vaccination during the programme.
-All other characteristics were calculated as on the day before vaccination.
+Age was calculated as of 30 June 2023, to allow for a degree of operational flexibility in who was offered vaccination during the programme.
+All other characteristics were calculated as of the day before vaccination.
 In the analysis cohorts there are no missing values for any characteristics as people with missing values for sex, IMD or STP were excluded, and remaining characteristics were defined by the presence or absence of clinical codes or events in the health record.
@@
 Myocarditis was identified similarly.
-Details of the codelists used to define these outcomes are available in supplementary materials.
+Details of the codelists used to define these outcomes are available in the supplementary materials.
 
@@
 We estimated cumulative risk differences and risk ratios (RRs) comparing the vaccine groups for each outcome.
-We also estimated Incidence Rate Ratios (IRRs) over the full follow-up duration, and within period-specific intervals defined by splits on weeks 1, 2, 4, 8, 12, and 16.
+We also estimated Incidence Rate Ratios (IRRs) for Sanofi versus Pfizer BA.4-5 recipients over the full follow-up duration, and within period-specific intervals defined by splits on weeks 1, 2, 4, 8, 12, and 16.
 
@@
 ### Software, code, and reproducibility
-
 Data management and analyses were conducted using OpenSAFELY tools, Python version 3.8.10 and R version 4.0.2.
-Code for data management and analysis, as well as codelists, are archived online <https://github.com/opensafely/comparative-booster-spring2023>.
+Code for data management and analysis, as well as codelists, are archived online (<https://github.com/opensafely/comparative-booster-spring2023>).
 Codelists are available at <https://www.opencodelists.org/>.
@@
 \newpage
-
 ## Results
@@
 Sanofi was more likely to be administered in April 2023, and Pfizer BA.4-5 more likely in May and June 2023.
-Prior to matching, Sanofi recipients were on average older, more likely to be recorded as resident in a care or nursing home, and less likely to have received fewer than 5 prior COVID-19 vaccine doses compared with Pfizer BA.4-5 recipients.
+Prior to matching, Sanofi recipients were on average older, more likely to be recorded as resident in a care or nursing home, and less likely to have received fewer than 5 prior COVID-19 vaccine doses compared with Pfizer BA.4-5 recipients, in both cohorts.
 Sanofi recipients were less likely that Pfizer BA.4-5 recipients to have previously received the Pfizer/BA.1 vaccine during an earlier booster programme and more likely to have received Moderna/Omicron.
 Sex, deprivation levels, and ethnicity were similar in both groups prior to matching (supplementary Tables S1a and S1b).
-
 In the CV cohort, `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(n))` people in each vaccine group (total `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())`) were matched, representing `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="pfizerBA45") %>% pull(pct_step))` and `r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c5", boost_type=="sanofi") %>% pull(pct_step))` of eligible Pfizer and Sanofi recipients respectively (Figure 1a, Figure S1a).
@@
 The 16-week risks (cumulative incidence) per `r perN_format` people of COVID-19 hospital admissions were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(rdCI_95)`) (Table 2; Figure 3).
-The corresponding 16-week risks per `r perN_format` people for COVID-19 critical care admission were `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(rdCI)`).
+The corresponding 16-week risks per `r perN_format` people for COVID-19 critical care admission were `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(rdCI)` where values greater than zero favour Pfizer).
 The 16-week risks per `r perN_format` people of COVID-19 death were `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_0)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(rdCI)`).
@@
 
-IRRs (where less than one favours Sanofi) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
+IRRs (where values greater than one favour Pfizer) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 4).
 
@@
 
-````{comment}
+```{r break4}
 
-````
+```
 
@@
 
-In the CV cohort, there were `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients respectively.
+In the CV cohort, there were `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients, respectively.
 This corresponded to a cumulative incidence per `r perN_format` people of `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
-For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` events for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi, (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`)
+For myocarditis, there were `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` events for Pfizer BA.4-5 and `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi (risk difference `r rst$cv$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`)
 
@@
 
-In the 75+ cohort, `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients respectively, with a cumulative incidence per `r perN_format` people of `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
-For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` for Pfizer BA.4/5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi, (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
+In the 75+ cohort, `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_0)` and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(n.event_1)` pericarditis events within 4 weeks after vaccination for Pfizer BA.4-5 and Sanofi recipients, respectively, with a cumulative incidence per `r perN_format` people of `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_0_95)` for those receiving Pfizer BA.4-5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(riskCI_1)` for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="pericarditis", subgroup=="all") %>% pull(rdCI_95)`).
+For myocarditis, there were `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_0)` events for Pfizer BA.4/5 and `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(n.event_1)` events for Sanofi (risk difference `r rst$age75plus$contrasts_table %>% filter(outcome=="myocarditis", subgroup=="all") %>% pull(rdCI_95)`).
 
@@
 
-Findings in the main analyses were similar when examining subgroups, those estimated with reduced precision.
-Supplementary materials provide cumulative incidence plots by subgroup (Figures S3), the corresponding cumulative risk differences, cumulative risk ratios, and period-specific IRRs (Figures S4), and subgroup-specific heterogeneity tests for risk differences and risk ratios (Table S2).
+Findings in the main analyses were similar when examining subgroups, although estimated with reduced precision.
+The Supplementary materials provide cumulative incidence plots by subgroup (Figures S3), the corresponding cumulative risk differences, cumulative risk ratios, and period-specific IRRs (Figures S4), and subgroup-specific heterogeneity tests for risk differences and risk ratios (Table S2).
+
+\newpage
 
@@
 We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Pfizer BA.4-5 vaccine compared with Sanofi during the first 16 weeks after vaccination.
-For the more severe COVID-19 outcomes of critical care admission and death, there were very few events and insufficient power to identify differences between vaccines.
+For the more severe COVID-19 outcomes of critical care admission and death, there were very few events resultng in reduced power to identify differences between vaccines.
 
@@
 
-The OpenSAFELY-TPP database covers around 45% of the English population and contains rich clinical information which enabled us to closely match Pfizer BA.4-5 and Sanofi recipients to control for potential confounding and compare comparative safety and effectiveness.
+The OpenSAFELY-TPP database covers around 45% of the English population and contains rich clinical information which enabled us to closely match Pfizer BA.4-5 and Sanofi recipients to control for potential confounding and compare safety and effectiveness.
 
@@
 [@zouNeutralizationBABA2023] -->
-
 Out-of-trial evidence of the effectiveness of Sanofi and Pfizer BA.4-5 vaccines against hospitalisation amongst adults aged 75 years and older in England from a test-negative case control study involving 14,174 eligible tests found similar effectiveness of both Pfizer BA.4-5 and Sanofi of 30-50%.[@kirsebomEffectivenessAdjuvantedSanofi2023] We are aware of no other planned or published randomised trials making direct endpoint comparisons of boosting with Pfizer BA.4-5 and Sanofi.
-
 ### Conclusions
@@
 During the spring 2023 COVID-19 booster programme in England, COVID-19-related hospital admissions were slightly more frequent following receipt of the Sanofi vaccine than the Pfizer BA.4-5 vaccine after matching on a number of important potential confounders.
-These findings were broadly consistent across subgroups, though estimated with less precision.
+These finding were broadly consistent across subgroups, though estimated with less precision.
 
@@
 Data curation: WJH, CDA, EP, CW, BS, HJC, AW, AM, BM, and SCJB.
-Formal analysis: WJH, CDA, EPKP, and VW.
+Formal analysis: WJH, CDA, and VW.
 Funding acquisition: BG, and JACS.
@@
 The service adheres to the obligations of the UK General Data Protection Regulation (UK GDPR) and the Data Protection Act 2018.
-The service previously operated under notices initially issued in February 2020 by the Secretary of State under Regulation 3(4) of the Health Service (Control of Patient Information) Regulations 2002 (COPI Regulations), which required organisations to process confidential patient information for COVID-19 purposes; this set aside the requirement for patient consent @WithdrawnWithdrawnCoronavirus2022.
+The service previously operated under notices initially issued in February 2020 by the the Secretary of State under Regulation 3(4) of the Health Service (Control of Patient Information) Regulations 2002 (COPI Regulations), which required organisations to process confidential patient information for COVID-19 purposes; this set aside the requirement for patient consent @WithdrawnWithdrawnCoronavirus2022.
 As of 1 July 2023, the Secretary of State has requested that NHS England continue to operate the Service under the COVID-19 Directions 2020 @COVID19PublicHealth.
@@
 
-This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by London School of Hygiene and Tropical Medicine Ethics Board (reference 21863).
+This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by London School of Hygeine and Tropical Medicine Ethics Board (reference 21863).
 
@@
 #| warning: false
+#| message: false
 
@@
 
-```{r figure-ci-cv-all}
+```{r figure-ci-all}
 #| echo: false
+
 #| warning: false
@@
 #| warning: false
-#| out-height: '50%'
+#| out-height: 50%
 
@@
     panel.border = element_blank(),
-    panel.spacing = unit(0.4, "lines"),
+    panel.spacing = unit(0.5, "lines"),
   )+

### 2024-02-27 write-up/manuscript/manuscript.qmd

@@
 We also report on non-COVID-19 deaths.
+We considered three effectiveness outcomes.
+COVID-19 hospital admission was identified using Secondary Use Service (SUS) in-patient hospital spell records with International Statistical Classification of Diseases and Related Health Problems 10th Revision (ICD-10) codes U07.1, U07.2, or U.109 as the primary or non-primary reason for admission.
+COVID-19 critical care admissions were any COVID-19 hospital spell with at least one day spent in critical care (including any such spells occurring subsequent to the first spell after vaccination).
+COVID-19 death was defined as COVID-19 ICD-10 codes (as before) mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
+We also report on non-COVID-19 deaths.
 
@@
 We identified `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1a).
+We identified `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people in the CV cohort, and of these `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$cv$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1a).
 
+`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1b).
 `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="any") %>% pull(n) %>% sum())` people were in the 75+ cohort, and of these `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="pfizerBA45") %>% pull(pct))`) received Pfizer BA.4-5 and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart_totals, boost_type=="sanofi") %>% pull(pct))`) received Sanofi, with `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="pfizerBA45") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(n))` (`r label_percent(0.1)(filter(rst$age75plus$flowchart, crit=="c4", boost_type=="sanofi") %>% pull(pct_all))`) eligible for matching respectively (Figure 1b).
@@
 ````{comment}
+````{comment}
 
+````
 ````
@@
 
+IRRs in the 16-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
 IRRs in the 16-week period after vaccination were `r rst$age75plus$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$age75plus$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$age75plus$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
@@
 The Supplementary materials provide cumulative incidence plots by subgroup (Figures S3), the corresponding cumulative risk differences, cumulative risk ratios, and period-specific IRRs (Figures S4), and subgroup-specific heterogeneity tests for risk differences and risk ratios (Table S2).
-
-\newpage
 
@@
 #### Figure 1a: Flow of participants into the CV cohort
+#### Figure 1a: Flow of participants into the CV cohort
 
 ![figure1a](figures_manual/flowchart_cv.png){width="60%"}
+![figure1a](figures_manual/flowchart_cv.png){width="60%"}
 
+#### Figure 1b: Flow of participants into the 75+ cohort
 #### Figure 1b: Flow of participants into the 75+ cohort
@@
 #| label: function-ci
+#| warning: false
 #| warning: false
@@
     plot_ci#, 
+    plot_ci#, 
     #plot_rd, 
     #plot_rr, 
+    #plot_irr,
+    #ncol=2,
+    #widths=c(n_subgroups,1)
     #plot_irr,
@@
 
-```{r figure-ci-all}
+```{r figure-ci-cv-all}
 #| echo: false
@@
 #| out-width: '100%'
+#| out-height: '80%'
 #| out-height: '80%'
@@
 write_csv(contrasts_jasper_plot, fs::path(output_dir_qmd,"irr_plots.csv"),na="")
+write_csv(contrasts_jasper_plot, fs::path(output_dir_qmd,"irr_plots.csv"),na="")
 
 # Run Jasper IRR plot script (requires https://github.com/arnhew99/Jasper?tab=readme-ov-file#jasper)
+#source(here::here("write-up", "manuscript", "irr_plots_jasper_code.r"))
 #source(here::here("write-up", "manuscript", "irr_plots_jasper_code.r"))
@@
   labs(
+    #subtitle="\n\n\nIncidence Rate Ratios",
     #subtitle="\n\n\nIncidence Rate Ratios",
@@
     panel.border = element_blank(),
+    panel.spacing = unit(0.4, "lines"),
     panel.spacing = unit(0.4, "lines"),
@@
     widths=c(2,0.9)
+    ncol=2,
+    widths=c(2,0.9)
   )
@@
     plot=plot_combined,
+    width=20, height=15, units="cm"
     width=20, height=15, units="cm"

### 2024-02-27 write-up/manuscript/supplement.qmd

@@
   plot_smd
+
+  ggsave(
+    filename=glue("smd_{.cohort}.jpeg"),
+    path=output_dir_qmd,
+    plot=plot_smd,
+    width=20, height=20, units="cm"
+  )
 }
@@
 #| out-width: '90%'
-#| out-height: '100%'
+#| out-height: '200%'
 #| output: asis

### 2024-02-27 write-up/manuscript/manuscript.qmd
fix merge errors

--- 2024-02-27 write-up/manuscript/manuscript.qmd
@@
 
-IRRs (where values greater than one favour Pfizer) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 4).
+IRRs (where values greater than one favour Pfizer) in the 16-week period after vaccination were `r rst$cv$contrasts_table %>% filter(outcome=="covidadmitted", subgroup=="all") %>% pull(irrCI_95)` for COVID-19 hospital admissions, `r rst$cv$contrasts_table %>% filter(outcome=="covidcritcare", subgroup=="all") %>% pull(irrCI)` for COVID-19 critical care admission, `r rst$cv$contrasts_table %>% filter(outcome=="coviddeath", subgroup=="all") %>% pull(irrCI)` for COVID-19 death, and `r rst$cv$contrasts_table %>% filter(outcome=="noncoviddeath", subgroup=="all") %>% pull(irrCI)` for non-COVID-19 death (Figure 3).
 
@@
 We estimated that, in both CV (N=`r label_number(1, big.mark=",")(filter(rst$cv$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) and 75+ (N=`r label_number(1, big.mark=",")(filter(rst$age75plus$flowchart, crit=="c5", boost_type %in% c("pfizerBA45", "sanofi")) %>% pull(n) %>% sum())` people) cohorts, COVID-19 related hospital admissions were less frequent in those receiving the Pfizer BA.4-5 vaccine compared with Sanofi during the first 16 weeks after vaccination.
-For the more severe COVID-19 outcomes of critical care admission and death, there were very few events resultng in reduced power to identify differences between vaccines.
+For the more severe COVID-19 outcomes of critical care admission and death, there were very few events resulting in reduced power to identify differences between vaccines.
 
@@
 
+![figure1a](figures_manual/flowchart_cv.png){width="60%"}
 ![figure1a](figures_manual/flowchart_cv.png){width="60%"}
@@
 #| warning: false
-#| out-height: 50%
+#| out-height: '50%'
 
